### 2013-01-22 paper.tex
More work on dpl Initial tex

\documentclass[12pt]{article}

\usepackage[utf8]{inputenc}
\usepackage{amsthm,amssymb,amsmath,dsfont,mathrsfs,nicefrac}

\let\stdsection\section
\renewcommand\section{\newpage\stdsection}

\title{Title of my paper}
\author{Thomas Dybdahl Ahle}
\date{March 9, 2013}

\begin{document}
\maketitle

\begin{abstract}
Your abstract goes here...
...
\end{abstract}

\tableofcontents

\section{Introduction}
5 pagish
Donkey example to follow through

\section{Formal Stuff}
10 pagish
This subsection's content...
\subsection{DPL}
This subsection's content...
\subsection{DRA}
This subsection's content...

\section{Implementation}
8pagish
This subsubsection's content...

\section{Perspectives}
2pagish
What I learned

\section*{Acknowledgements}
Thanks to Samson
This section's content...\cite{visser1999donkey}
the PCFG parser\cite{klein2003accurate}

\subsection{Checklist}
Things learned from courses

\bibliographystyle{plain}
\bibliography{refs}

\end{document}

### 2013-01-22 donkey.hs

@@ -1,0 +1,43 @@
+-- http://stackoverflow.com/questions/3228856/predicate-logic-in-haskell
+
+import Data.List
+
+data Prop =
+	  Not   Prop
+    | And   Prop Prop
+    | Or    Prop Prop
+    | Impl  Prop Prop
+    | Equiv Prop Prop
+    | Equals Obj Obj
+    | ForAll (Obj -> Prop)
+    | Exists (Obj -> Prop)
+    deriving (Eq, Ord)
+
+data Dra =
+	  DraComp Dra Dra
+	| DraId
+	| DraBottom
+	| DraImpl Dra Dra
+	| Set
+
+class Pol p where
+	plus :: 
+
+data Pol = 1 | -1
+type Dlp = (Prop, Prop, Pol)
+
+true :: Dlp
+true = (true, true, 1)
+
+false :: Dlp
+false = (true, false, 1)
+
+switch :: Dlp
+switch = (true, true, -1)
+
+mul :: Dlp -> Dlp -> Dlp
+mul (qm, qp, 1) (rm, rp, b) = (pmul qm rm, pmul qp rp, b)
+mul (qm, qp, -1) (rm, rp, b) = (pmul qm rp, pmul qp rm, -b)
+
+test :: Dlp -> Prop
+test (qm, qp, a) = Impl qm qp

### 2013-01-22 dra.hs

@@ -1,0 +1,66 @@
+import Data.List
+
+type Var = Integer
+type Rels = Var -> Var -> Bool
+type M = [Var] -> Rels
+
+-- monaden skal have constanter, værdier, variable, funktioner...
+
+create :: [(Var,Var)] -> Rels
+create pairs x y = elem (x,y) pairs
+
+result :: Rels -> M
+result r vs = r
+
+new = result . create
+
+-- ($>) :: M -> (Rels -> M) -> M
+-- (r $> f) vs = f (filter r (zip vs' vs')) where vs' = r vs
+
+draComp :: M -> M -> M
+draComp r s vs x y = any (\z -> r vs x z && s vs z y) vs
+
+draId :: M
+draId vs x y = x == y
+
+draBot :: M
+draBot vs x y = False
+
+impl a b = not a || b
+
+draImpl :: M -> M -> M
+draImpl r s vs x y = x == y && all (\z -> r vs x z `impl` any (\u -> s vs z u) vs) vs
+
+draNot :: M -> M
+draNot r = r `draImpl` draBot
+
+cross :: [a] -> [b] -> [(a, b)]
+cross [] ys = []
+cross (x:xs) ys = [(x, y) | y <- ys] ++ cross xs ys
+
+draDiag :: [Var] -> M
+draDiag xs vs x y = x == y && elem x xs
+
+draShow :: M -> [Var] -> [(Var,Var)]
+draShow r vs = filter (uncurry (r vs)) (cross vs vs)
+
+draDom :: M -> [Var] -> [Var]
+draDom r vs = map fst (draShow r vs)
+
+test :: [(Var,Var)] -> [Var] -> Bool
+test pairs vs =
+    draShow (draNot r) vs == draShow (draDiag (vs \\ draDom r vs)) vs
+    && draShow ((draNot.draNot) r) vs == draShow (draId `draImpl` r) vs
+    && draShow (draId `draImpl` r) vs == draShow (draDiag (draDom r vs)) vs
+    where r = (result . create) pairs
+
+testAll :: [Var] -> Bool
+testAll vs = and [test pairs vs | pairs <- subsequences (cross vs vs)]
+
+-- Der må være forskel på værdier og variable
+-- exists skal tage en variabel, og assigne den til alle forskellige værdier
+-- indtil det, den bliver kædet med, virker.
+-- Så faktisk skal der være noget effekt at et element, nemlig at det kan ændre
+-- assigmenten på variablerne
+-- Kræver måske renaming...
+exists 

### 2013-02-26 paper.tex
Work on paper New ideas for semantics

@@
 \usepackage{amsthm,amssymb,amsmath,dsfont,mathrsfs,nicefrac}
+\usepackage{listings}
 
@@
 \section{Introduction}
-5 pagish
-Donkey example to follow through
+5 pagish Donkey example to follow through.
+There are many reasons why it is interesting to represent information in a logical language.
+You can do things such as unification, resolution, and question answering.
+
+% It is not possible
+It is clear that not all sentences can be unambigiously translated from natural language to logic. A sentence like 	
+\begin{quotation}
+A farmer beats a donkey
+\end{quotation}
+Might be translated as 
+\begin{equation}
+\exists x ( farmer(x) \wedge \exists y ( donkey(y) \wedge beats(x,y)))
+\end{equation}
+Indicating that ``a farmer'' refers to some specific individual. Or we can might see it as a general rule, and translate it as
+\begin{equation}
+\forall x ( farmer(x) \rightarrow \exists y ( donkey(y) \wedge beats(x,y)))
+\end{equation}
+Indicating that every farmer has some donkey he beats. Or we could even read it as
+\begin{equation}
+\exists y ( donkey(y) \wedge \forall x ( farmer(x) \rightarrow beats(x,y)))
+\end{equation}
+Indicating that the poor donkey is common between all farmers.
+
+In this paper we will not try to solve the above problem.
+(Though it could be interesting to generate multiple possibilities and perhaps use context to filter out the correct one).
+Instead we will try to show how some old method for doing this conversion can be enhanced with some new tools
+
+\begin{equation}
+\forall x . farmer(x) \rightarrow (\forall y . donkey(y) \wedge owns(x,y) \rightarrow beats(x,y))
+\end{equation}
+
+lksadjf lakdf jlsak dfjlask dfjlask dfjlsa kdfjasld kfjsald kfjsalkd fjslakd fjlsakd fjlkasd fjlaskd jflsak dfj
+
+If a farmer owns a donkey he beats it
+\begin{equation}
+\forall x . farmer(x) \rightarrow (\forall y . donkey(y) \wedge owns(x,y) \rightarrow beats(x,y))
+\end{equation}
+
+There are a lot of problems with this representation
+\begin{itemize}
+  \item It is not compositional. 
+  dfdf
+  \item It doesn't very well follow the order of the original sentence
+  \item It is just not very easy to get right
+\end{itemize}
 
@@
 \subsection{DPL}
+Dynamic semantics: http://plato.stanford.edu/entries/dynamic-semantics/
 This subsection's content...
+\subsection{Using the monad}
+This might be what I call DRA underneath
 \subsection{DRA}
+This might be just the same as DPL
+\subsection{Dependencies}
+This subsection's content...
+\subsection{Montague semantic}
+This subsection's content...
+\subsection{Semantics via dependencies}
 This subsection's content...
@@
 This subsubsection's content...
+
+\subsection{My implementaiton of DPL semantics}
+Maybe this should be in the formal stuff. Nah, there are two versions
+
+Below are some direct implementation of the semantics for DPL. We use the types
+\begin{lstlisting}
+type Assignment = [(Ref, Val)]
+type Relation = [Assignment] -> [Assignment]
+\end{lstlisting}
+This differs from the indirect set of types are used in the earlier formalisms:
+\begin{lstlisting}
+type Assignment = [(Ref, Val)]
+type Relation = [Assignment] -> [Assignment] -> Bool
+\end{lstlisting}
+The later is very difficult to implement, since the semantics require quantification over all assignments. Like for composition.
+
+\begin{lstlisting}
+:: Relation
+true = id
+false = const []
+
+predi1 :: [Val] -> Ref -> Rel
+predi1 f k = filter (\as -> elem (get k as) f)
+
+predi2 :: [(Val,Val)] -> Ref -> Ref -> Rel
+predi2 f k l = filter (\as -> elem (get k as, get l as) f)
+
+:: Relation -> Relation
+neg r = filter (\as -> null (r [as]))
+
+:: Relation -> Relation -> Relation
+comp = flip (.)
+impl s r = rnot (s `comp` (rnot r))
+\end{lstlisting}
+Then comes the most interesting one. To ``randomly reset'' a variable, we have to make sure that we have all assignments in the cross product of new values for the variable and what was there before.
+
+Bum bum. Maybe we should just use a set, then we could overwrite variable assignments
+\begin{lstlisting}
+exist :: Ref -> [Val] -> Rel
+exist k vs ass = [a : as | a <- news, as <- olds]
+	where olds = map (filter ((/=k) . fst)) ass
+	      news = [(k,v) | v <- vs]
+\end{lstlisting}
+
+\subsection{Using machine learning}
+The $\bowtie$ and $\triangle$ notation makes logic look close enough to natural language that it starts looking tempting to use machine learning methods such as those used in translation to translate.
+
+My main problem here is that such translation requires large ammounts of dualingo data. It was out of scope for my project to try create such.
+
+\subsection{Handcoding}
+This is not as bad as it sounds. It is similar to montague and what the stanfordparser does for dependencies.
 

### 2013-02-26 Dpl.hs

@@ -1,9 +1,7 @@
-module Dpl ( parse, deps, corefs, Prop ) where
+module Dpl (dpl2prop, Type2, s2dpl,
+	sT, sF, sSwitch, sScope, sPredi, sExist, sComp, sTest0, sTest1) where
 
-import Data.List
-import Text.XML.Light
-import Control.Monad
-import Data.Maybe
+import Prop (Ref, Prop(..))
 
 -- Type 1: Basic dynamic predicate logic
 ----------------------------------------
@@ -17,19 +15,20 @@ data DPL =
 	  | DImpl DPL DPL
 	  deriving (Eq, Show)
 
--- TODO: Move run code here
+-- TODO: Move exec code here
 
-dpl2prop' dpl = dpl2prop PT [dpl]
+dpl2prop :: DPL -> Prop
+dpl2prop dpl = addDpl T [dpl]
 
-dpl2prop :: Prop -> [DPL] -> Prop
-dpl2prop p [] = p
-dpl2prop p (DF:ds) = PF
-dpl2prop p (DT:ds) = dpl2prop p ds
-dpl2prop p ((DPredi f ks):ds) = dpl2prop (PAnd (PPredi f ks) p) ds
-dpl2prop p ((DExist k):ds) = dpl2prop (PExist k p) ds
-dpl2prop p ((DComp d1 d2):ds) = dpl2prop p (d2:d1:ds)
-dpl2prop p ((DNot d):ds) = dpl2prop (PAnd (PNot (dpl2prop PT [d])) p) ds
-dpl2prop p ((DImpl d1 d2):ds) = dpl2prop p ((DNot (DComp d1 (DNot d2))):ds)
+addDpl :: Prop -> [DPL] -> Prop
+addDpl p [] = p
+addDpl p (DF:ds) = F
+addDpl p (DT:ds) = addDpl p ds
+addDpl p ((DPredi f ks):ds) = addDpl (And (Predi f ks) p) ds
+addDpl p ((DExist k):ds) = addDpl (Exist k p) ds
+addDpl p ((DComp d1 d2):ds) = addDpl p (d2:d1:ds)
+addDpl p ((DNot d):ds) = addDpl (And (Not (addDpl T [d])) p) ds
+addDpl p ((DImpl d1 d2):ds) = addDpl p ((DNot (DComp d1 (DNot d2))):ds)
 
 -- If there's a farmer and a donkey, the farmer beats the donkey
[…]

### 2013-02-26 Main.hs

@@ -1,6 +1,8 @@
 module Main (main) where
 
-import Stanford (run)
+import Stanford (run, runDry)
+import Prop (Prop)
+import ToLogic (logify)
 
 import Data.List
 import Text.XML.Light
@@ -9,23 +11,36 @@ import Data.Maybe
 import System.Process
 
 super :: [String] -> IO [Prop]
-super sentences = run sentences >>= (return . logify)
+--super sentences = run sentences >>= (return . map logify)
+super sentences = runDry sentences >>= (return . map logify)
 
 main :: IO ()
 main = do
 		putStrLn "Test1"
 		props <- super [
 			"If a farmer owns a donkey, he beats it",
-			"If a farmer owns it and he beats it , it is a donkey", -- fails because 'donkey' becomes head somehow. (It even has a nsubj, huh?)
-			"If a farmer owns it and beats it , it is a donkey", -- fails because co-ref doesn't handle single occourences
+			"If a farmer owns it and he beats it, it is a donkey", -- fails because 'donkey' becomes head somehow. (It even has a nsubj, huh?)
+			"If a farmer owns it and beats it, it is a donkey", -- fails because co-ref doesn't handle single occourences
 			"She is a mother of five", -- mother becomes the root, because is is not a 'real' verb
-			"A farmer . A Donkey . He owns it . He beats it", -- fails because we assume all sentences have a VP root
-			"If you like me , you will hate my cousin",
+			"A farmer. A Donkey. He owns it. He beats it", -- fails because we assume all sentences have a VP root
+			"If you like me, you will hate my cousin",
 			"Do you still beat your wife?",
 			-- "The man could not make his donkey walk.", fails because SP things walk is a noun
 			"The man couldn't make his donkey see.",
 			"If there's a farmer and a donkey, the farmer beats the donkey.",
-			"If there's a farmer who owns a donkey, the farmer beats the donkey."]
+			"If there's a farmer who owns a donkey, the farmer beats the donkey.",
+			"James won a Nobel prize", -- commitment
+			"The King of France is bold", -- unique, what if there's no king?
+			"John regrets that his dog died.", -- what is the object?
+			"When did you stop beating your wife",
+			"A man comes in. He sees a dog. He smiles." -- very sequential
+			]
 		putStrLn "Test2"
 		sequence (map print props)
 		return ()
+
+{-
+These look hard (Existential non-commitment):
+	Jim wants to marry a norweigan girl
+	Fred was looking for a unicorn
+-}

### 2013-02-26 Prop.hs

@@ -1,4 +1,4 @@
-module Prop (Prop(..), simplify) where
+module Prop (Ref, Prop(..), simplify) where
 
 type Ref = String
 

### 2013-02-26 Stanford.hs

@@ -1,5 +1,6 @@
+module Stanford (DepTree(..), Word, Sentence, run, runDry, runOnFile, PosTree(..), toMapping, postreeS) where
 
-module Stanford (DepTree, Ref, run, PosTree(..), toMapping, postreeS) where
+import Prop (Ref)
 
 import Text.ParserCombinators.Parsec
 import Data.Map hiding (map, (\\))
@@ -10,14 +11,13 @@ import Control.Monad
 import Data.Maybe
 import System.Process
 
-type Ref = String
 type Word = String
 type Sentence = [Word]
 
 type PosTag = String
-data PosTree = Phrase PosTag [PosTree] | Leaf PosTag Word deriving (Show)
+data PosTree a = Phrase PosTag [PosTree a] | Leaf PosTag a deriving (Show)
 
-data DepTree = Dep Int [(Ref, DepTree)] deriving Show
+data DepTree = Dep Int [(Ref, DepTree)] deriving (Show, Eq)
 
 ------------------
 
@@ -27,16 +27,16 @@ data DepTree = Dep Int [(Ref, DepTree)] deriving Show
 -- (did -> do, n't -> not)
 
 lemmas :: Element -> Sentence
-lemmas doc = map strContent (findElements (unqual "word") doc)
+lemmas doc = map strContent (findElements (unqual "lemma") doc)
 
 ------------------
 
-postree :: Element -> PosTree
+postree :: Element -> PosTree Word
 postree doc = postreeS dat
 	where
 		dat = (strContent . fromJust) (findElement (unqual "parse") doc)
 
-postreeS :: String -> PosTree
+postreeS :: String -> PosTree Word
 postreeS dat = tree
 	where
 		Right tree = parse posEither "(unknown)" dat
@@ -60,7 +60,7 @@ posLeaf =
 		word <- many (noneOf " ()")
 		return (Leaf tag word)
 
-toMapping :: [Ref] -> PosTree -> ([Ref], Map Int Ref)
+toMapping :: [Ref] -> PosTree Word -> ([Ref], Map Int Ref)
 toMapping vs (Leaf _ _) = (vs, empty)
 
 toMapping vs tr@(Phrase "NP" subs) = (vs', submaps `union` np)
@@ -73,7 +73,7 @@ toMapping vs (Phrase _ (sub:subs)) = (vs'', p `union` shifted)
 	      shifted = mapKeys (+tsize sub) subm
 	      (vs'', p) = toMapping vs' sub
 
[…]

### 2013-02-26 Tests.hs

@@ -1,5 +1,3 @@
-import Stanford
-
 import Test.Framework (defaultMain, testGroup)
 import Test.Framework.Providers.HUnit
 import Test.Framework.Providers.QuickCheck (testProperty)
@@ -7,9 +5,14 @@ import Test.Framework.Providers.QuickCheck (testProperty)
 import Test.QuickCheck
 import Test.HUnit
 
+import System.IO.Unsafe
+
 import Data.List
 import Data.Map hiding (map)
 import Data.Char (chr, ord)
+
+import Stanford
+import Prop(Ref)
 
 main = defaultMain tests
 
@@ -17,7 +20,8 @@ tests = [
     testGroup "Stanford tests" [
         testCase "test1" test_tomapping1,
         testCase "test2" test_tomapping2,
-        testProperty "test3" prop_tomapping
+        testProperty "test3" prop_tomapping,
+        testCase "test4" test_stanford
      ],
     testGroup "Dpl tests" [
 
@@ -58,7 +62,13 @@ test_tomapping2 = (toMapping2 . postreeS) inp @?= out
           out = fromList [(0, "b"), (1, "b"), (5, "a"), (6, "a")]
 
 prop_tomapping :: PosTree -> Bool
-prop_tomapping postree = cntNp postree <= (length . nub . elems . toMapping2) postree
+prop_tomapping postree = cntNp postree >= (length . nub . elems . toMapping2) postree
   where cntNp (Leaf _ _) = 0
         cntNp (Phrase "NP" subs) = 1 + sum (map cntNp subs)
         cntNp (Phrase _ subs) = sum (map cntNp subs)
+
+test_stanford = (sentence, refs, deps) @?= (sout, rout, dout)
+    where (sentence, refs, deps) = unsafePerformIO (runOnFile "test.xml")
+          sout = ["if", "a", "farmer", "own", "a", "donkey", ",", "he", "beat", "it"]
+          rout = fromList [(1, "e"), (2, "e"), (4, "f"), (5, "f"), (7, "e"), (9, "f")]
+          dout = Dep 8 [("advcl",Dep 3 [("mark",Dep 0 []),("nsubj",Dep 2 [("det",Dep 1 [])]),("dobj",Dep 5 [("det",Dep 4 [])])]),("nsubj",Dep 7 []),("dobj",Dep 9 [])]

### 2013-02-26 ToLogic.hs

@@ -1,9 +1,13 @@
 module ToLogic (logify) where
 
-import Dpl
+import Dpl (dpl2prop, s2dpl, Type2, sT, sF, sSwitch, sScope, sPredi, sExist, sComp, sTest0, sTest1)
 import Prop
+import Stanford (DepTree(..), Word, Sentence)
 
-parseVP :: Sentence -> Map Int Ref -> DepTree -> Scop
+import Data.Map hiding (lookup, mapMaybe)
+import Data.Maybe
+
+parseVP :: Sentence -> Map Int Ref -> DepTree -> Type2
 parseVP ws refs (Dep n ds) = foldl sComp sT parts `sComp` predi
 	where
 		parts = mapMaybe (\(name, dep) -> (case name of
@@ -15,18 +19,34 @@ parseVP ws refs (Dep n ds) = foldl sComp sT parts `sComp` predi
 		Just (Dep obj _) = lookup "dobj" ds
 		predi = sPredi (ws !! n) [refs ! sub, refs ! obj]
 
-parseAdvcl :: Sentence -> Map Int Ref -> DepTree -> Scop
+parseAdvcl :: Sentence -> Map Int Ref -> DepTree -> Type2
 parseAdvcl ws refs dep = sSwitch `sComp` (parseVP ws refs dep) `sComp` sSwitch
 
 -- Check if we are the primary mention, if true then add an exists and a test
-parseNP :: Sentence -> Map Int Ref -> DepTree -> Scop
+parseNP :: Sentence -> Map Int Ref -> DepTree -> Type2
 parseNP ws refs (Dep n ds) = case isJust (lookup "det" ds) of
 	 True -> sExist var `sComp` sPredi (ws !! n) [var]
 	 False -> sT
 	where var = refs ! n
 
-logify :: (Sentence, Map Int Refs, DepTree) -> Prop
-logify (sentence, refs, deps) = (simplify . dpl2prop' . s2dpl) logic
+logify :: (Sentence, Map Int Ref, DepTree) -> Prop
+logify (sentence, refs, deps) = (simplify . dpl2prop . s2dpl) logic
 	where
 		logic = parseVP sentence refs deps
 
+monte :: PosTree (Int, Word) -> Map Int Ref -> Type2
+monte (Phrase "ROOT" ts) = monte (head ts)
[…]

### 2013-02-26 exec2.hs

@@ -1,4 +1,5 @@
 
+-- These direct semantics don't work, since they require quantification over all assignments
 
 import Data.List
 import Data.Maybe
@@ -6,53 +7,35 @@ import Data.Maybe
 type Val = Integer
 type Ref = String
 type Ass = [(Ref, Val)]
-type Rel = [Ass] -> [Ass]
+type Rel = Ass -> Ass -> Bool
 
--- Rel a = [b in ASS | Ez in a (z R b)]
-
-sigma :: [Ref]
 sigma = ["x","y"]
-
-domain :: [Val]
 domain = [0,1,2]
-
-everything :: [Ass]
 everything = sequence [[(k,v) | v <- domain] | k <- sigma]
 
-
-
 true :: Rel
-true = id
+true x y = x == y
 
 false :: Rel
-false = const []
+false x y = False
 
--- like and
 comp :: Rel -> Rel -> Rel
-comp = flip (.)
-
-test :: Rel -> Ass -> Bool
-test s ass = s [ass] /= []
+comp r s x y = or [x `r` z && z `s` y | z <- everything]
 
 rnot :: Rel -> Rel
-rnot s = filter (\as -> not (test s as))
+rnot r x y = x == y && not (r x y)
 
 impl :: Rel -> Rel -> Rel
 impl s r = rnot (s `comp` (rnot r))
 
-exist :: Ref -> [Val] -> Rel
-exist k vs ass = nub [set (k,v) as | v <- vs, as <- ass]
+exist :: Ref -> Rel
+exist key x y = and [lookup k x == lookup k y | k <- sigma \\ [key]]
 
 predi1 :: [Val] -> Ref -> Rel
-predi1 f k = filter (\as -> elem (get k as) f)
+predi1 f k x y = x == y && elem (get k x) f
 
 predi2 :: [(Val,Val)] -> Ref -> Ref -> Rel
-predi2 f k l = filter (\as -> elem (get k as, get l as) f)
-
--- like or
-union :: Rel -> Rel -> Rel
-union s r ass = (s ass) ++ (r ass)
-
+predi2 f k l x y = x == y && elem ((get k x), (get l x)) f
 
 -- Will fail if ref not in ass
 get :: Ref -> Ass -> Val
@@ -61,19 +44,16 @@ get k = fromJust . (lookup k)
 set :: (Ref, Val) -> Ass -> Ass
[…]

### 2013-02-26 exec3.hs

@@ -1,0 +1,84 @@
+
+
+import Data.List
+import Data.Maybe
+
+type Val = Integer
+type Ref = String
+type Ass = [(Ref, Val)]
+type Rel = [Ass] -> [Ass]
+
+-- "The correct way to go is undoubtly to replace relations by indexed sets of relations read disjunctively"
+
+sigma = ["x","y"]
+domain = [0,1,2]
+everything = sequence [[(k,v) | v <- domain] | k <- sigma]
+
+true :: [Rel]
+true = [id]
+
+false :: [Rel]
+false = [const []]
+
+-- like and, conjunction, cross product
+-- (a -> b eller c -> d) og (d -> e eller f -> g)
+comp :: [Rel] -> [Rel] -> [Rel]
+comp rs ss = [s.r | r <- rs, s <- ss]
+
+disj :: [Rel] -> [Rel] -> [Rel]
+disj = (++)
+
+test :: [Rel] -> Ass -> Bool
+test rs as = any (not.null) [r [as] | r <- rs]
+
+rtest :: Rel -> Ass -> Bool
+rtest r as = (not.null) (r [as])
+
+-- not in dpl works by filtering out exactly those assignments that would before have been accepted
+-- (not rel) ass = ass - (rel ass)
+rnot :: Rel -> Rel
+rnot r = filter (\as -> not (rtest r as))
+
+-- not (a -> b eller c -> d)
+-- we must push the not through the disjunction with de-morgan
+dnot :: [Rel] -> [Rel]
+dnot rs = foldl comp true [[rnot r] | r <- rs]
+
+impl :: [Rel] -> [Rel] -> [Rel]
+impl rs ss = dnot (rs `comp` (dnot ss))
+--impl rs ss = (dnot rs) `disj` ss
+-- This bottom one somehow doesn't allow left E's to be used in right. Hm
+-- So di morgan er ikke rigtig paa plads
+
+exist :: Ref -> [Rel]
+exist k = [\ass -> [set (k,v) as | v <- domain, as <- ass]]
+
+predi1 :: [Val] -> Ref -> [Rel]
+predi1 f k = [filter (\as -> elem (get k as) f)]
+
+predi2 :: [(Val,Val)] -> Ref -> Ref -> [Rel]
+predi2 f k l = [filter (\as -> elem (get k as, get l as) f)]
+
+-- Will fail if ref not in ass
[…]

### 2013-03-02 paper.tex
more

@@
 \usepackage[utf8]{inputenc}
-\usepackage{amsthm,amssymb,amsmath,dsfont,mathrsfs,nicefrac}
+\usepackage{amsthm,amssymb,amsmath,dsfont,mathrsfs,nicefrac,wasysym}
 \usepackage{listings}
-
 \let\stdsection\section
@@
 Indicating that the poor donkey is common between all farmers.
+% This seams wrong to me
 
@@
 Dynamic semantics: http://plato.stanford.edu/entries/dynamic-semantics/
-This subsection's content...
-\subsection{Using the monad}
+
+growth of information in time
+Inspired by programming language
+not philosofically, states can be anything. I don't want to get into that
+isomorphic with normal logic (visser siger generalization/projection)
+doen't promise to be a perfect dual of natural language, but it's closer, and has been rewarding
+Dynamic semantics aims to model meaning and interpretation. You can do that without answering broader philosophical questions
+takes the meaning of concepts for given
+discourse representation theory, file change, update semantics, 
+Hans Kamp, Veltman, Groenendijk
+
+compoitionality. A man comes in. He sees a dog. He smiles.
+
+Vi har brug for en logik, der ikke behøver at vide alting fra starten
+
+direct translation not possible
+`He saw the man with the telescope'
+and even if we understand the syntax
+`The students all attended one class'
+has at least two meanings.
+interpretation of discourse is influenced and guided by the common ground that exists between speaker and hearer
+
+relation between sets of assignments
+$\{\overline{PQR}, \overline{PQ}R, \overline{P}Q\overline{R}, \overline{P}QR, P\overline{QR}, P\overline{Q}R, PQ\overline{R}, PQR\}$
+
+example
+
+real world example
+
+something about more information, less models
+the interpretation of false
+
+Something about kleiski combinator?
+
+the thing about quantifiers
+how they would work in a programming language
+$\exists x; A$
+$\exists x; x = 4$ hvis $x=4$ skal virke og reducere mængden af modeller, så må exists gøre dem alle sammen mulige
+
+A man comes in. En fri variabel vi ikke ved noget om.
+A man comes in. It's Thomas..
+
+we take as DPL-meanings binary relations between assignments
+
+later we will use an isomorphism and take them as functions from assignments to sets of assignments.
+
+\begin{equation}
+\phi ::= \bot \mid \top \mid \pi \mid \epsilon \mid \phi \cdot \phi \mid \lnot(\phi)
+\end{equation}
+\begin{equation}
+\alpha [\bot] \beta :\equiv False
+\end{equation}
+\begin{equation}
+\alpha [\top] \beta :\equiv \alpha = \beta
+\end{equation}
+\begin{equation}
+\alpha[P(x_0, ..., xn-1)]\beta :\equiv \alpha = \beta and (\alpha x0, ..., \alpha x n-1 ) \in I(P)
+\end{equation}
+
+\subsection{The problem with disjunction}
+
+a farmer doesn't own a horse or a donkey
+
+Ex.farmer(x) . ~(Ey.horse(y).owns(x,y) v Ey.donkey(y).owns(x,y))
+Ex.farmer(x) . ~Ey.horse(y).owns(x,y) . ~Ey.donkey(y).owns(x,y)
+
+a farmer beats a horse or a donkey
+
+Ex.farmer(x).(Ey.horse(y) v Ez.donkey(z)).beats(x,y) = {Nothing, Ex.farmer(x).EyJust .horse(y).beats(x,y)}
+
+Ex.farmer(x).~((Ey.horse(y) v Ez.donkey(z)).beats(x,y)) =
+Ex.farmer(x).(~Ey.horse(y).~Ez.donkey(z) v ~beats(x,y)) = {Nothing, Ex.farmer(x).~Ey.horse(y).~Ez.donkey(z)}
+
+Ex.farmer(x).~(tall(x) v blue(x)) = Ex.famer(x).~tall(x).~blue(x)
+
+There is a farmer who maybe has a donkey. He beats it.
+Ex.farmer(x).(Ey.donkey(y) v false).beats(x,y)
+With union semantics for v this would be equal to
+Ex.farmer(x).Ey.donkey(y).beats(x,y)
+which is stupid.
+Intead with sets of relations, this would be equal to
+Ex.farmer(x).Ey.donkey(y).beats(x,y) v Ex.farmer(x).false.beats(x,y)
+the right side being an error? no
+
+If a farmer is not poor he owns a donkey. He beats it.
+Ex.farmer(x).~poor(x) -> Ey.donkey(y).owns(x,y).beats(x,y)
+
+A farmer is either poor or he owns a donkey. He beats it.
+Ex.farmer(x).(poor(x) v Ey.donkey(y).owns(x,y)).beats(x,y) =
+
+(Ex v false) . f(x)
+(exist x `union` false) [] = [[("x",0)],[("x",1)],[("x",2)]
+[Nothing, Just [[("x",0)],[("x",1)],[("x",2)]]]
+
+(Ex v false) = [Just [[("x",0)],[("x",1)],[("x",2)], Just []]
+not (Ex v false) =
+
+not (Ex v f(x))
+= ~Ex.~f(x)
+
+A farmer owns a horse or a donkey
+Ex.farmer(x).~(~(Ey.horse(y)).~(Ey.donkey(y))).owns(x,y)   this surely doesn't work
+
+nothing survives out of a not (or impl) it's normal. Don't worry about it
+
+Let's just say that if some of 
+maybe not (bla v bla v bla v ERROR) should be ERROR
+it could also be something more involved, like ERROR(bla v bla v bla v ERROR), but let's not do that
+fmap
+
+\subsection{Monadic variants}
 This might be what I call DRA underneath
-\subsection{DRA}
-This might be just the same as DPL
+In the previous subsection we saw that DPL allowed us to rewrite the donkey sentence as
+
+\begin{equation}
+\exists x \cdot farmer(x) \cdot \exists y \cdot donkey(y) \cdot owns(x,y) \rightarrow beats(x,y)
+\end{equation}
+
+($\cdot$ binds stronger than $\rightarrow$)
+
+But think about the slight modification to the sentence:
+
+\begin{quotation}
+A farmer owns a donkey, if he beats it
+\end{quotation}
+
+We would have to write that as
+
+\begin{equation}
+\exists x \cdot farmer(x) \cdot \exists y \cdot donkey(y) \cdot beats(x,y) \rightarrow owns(x,y)
+\end{equation}
+
+The problem here is that reorderings like this work against our sought property of linearity: of being able to add new information as we go.
+We cannot compose `A farmer owns a donkey' and `if he beats it' in a meaningful way.
+
+To resque us from this problem Albert Visser suggests constructing a special e-monoid over the dynamic relational algebra.
+The idea is that we are going to simultaniously work on two `streams' of dpl, and switch easily between them with a special element $\Bowtie$.
+
+Our sentence from above will then be expressible as
+
+\begin{equation}
+\Bowtie \cdot \exists x \cdot farmer(x) \cdot \exists y \cdot donkey(y) \cdot \Bowtie \cdot owns(x,y) \Bowtie \cdot beats(x,y)
+\end{equation}
+
+This is however still not quite as close to natural language as we would like. Specially
+
 \subsection{Dependencies}

### 2013-03-02 Dpl.hs

@@ -11,6 +11,7 @@ data DPL =
 	  | DPredi Ref [Ref]
 	  | DExist Ref
 	  | DComp DPL DPL
+	  | DOr DPL DPL
 	  | DNot DPL
 	  | DImpl DPL DPL
 	  deriving (Eq, Show)
@@ -26,9 +27,10 @@ addDpl p (DF:ds) = F
 addDpl p (DT:ds) = addDpl p ds
 addDpl p ((DPredi f ks):ds) = addDpl (And (Predi f ks) p) ds
 addDpl p ((DExist k):ds) = addDpl (Exist k p) ds
-addDpl p ((DComp d1 d2):ds) = addDpl p (d2:d1:ds)
+addDpl p ((DComp d1 d2):ds) = addDpl p (d2:d1:ds) -- backwards
 addDpl p ((DNot d):ds) = addDpl (And (Not (addDpl T [d])) p) ds
 addDpl p ((DImpl d1 d2):ds) = addDpl p ((DNot (DComp d1 (DNot d2))):ds)
+addDpl p ((DOr d1 d2):ds) = addDpl (And (Or (addDpl T [d1]) (addDpl T [d2])) p) ds
 
 -- If there's a farmer and a donkey, the farmer beats the donkey
 r1 = DImpl (DComp (DComp (DComp (DExist "x") (DPredi "farmer" ["x"])) (DExist "y")) (DPredi "donkey" ["y"])) (DPredi "beats" ["x", "y"])

### 2013-03-02 Prop.hs

@@ -7,6 +7,7 @@ data Prop =
 	| F
 	| Not   Prop
     | And   Prop Prop
+    | Or    Prop Prop
     | Predi Ref [Ref]
     | Exist Ref Prop
     | All Ref Prop

### 2013-03-02 exec3.hs

@@ -1,84 +1,110 @@
 
 
-import Data.List
 import Data.Maybe
+import Data.Either
+import Control.Monad ((>=>), liftM, liftM2)
 
 type Val = Integer
 type Ref = String
 type Ass = [(Ref, Val)]
-type Rel = [Ass] -> [Ass]
+type Rel = Ass -> [Ass]
 
--- "The correct way to go is undoubtly to replace relations by indexed sets of relations read disjunctively"
+-- We don't have to used indexed sets and disjoint union, since we don't eliminate duplicates
+type DRel = [Maybe Rel]
 
+-- Rel a = [b in ASS | Ez in a (z R b)]
+-- tanke: den initielle assignment er til de fri variable.
+
+--
+
+sigma :: [Ref]
 sigma = ["x","y"]
+
+domain :: [Val]
 domain = [0,1,2]
+
+everything :: [Ass]
 everything = sequence [[(k,v) | v <- domain] | k <- sigma]
 
-true :: [Rel]
-true = [id]
 
-false :: [Rel]
-false = [const []]
+true :: DRel
+true = [Just return]
 
--- like and, conjunction, cross product
--- (a -> b eller c -> d) og (d -> e eller f -> g)
-comp :: [Rel] -> [Rel] -> [Rel]
-comp rs ss = [s.r | r <- rs, s <- ss]
+false :: DRel
+false = [Just (const [])]
 
-disj :: [Rel] -> [Rel] -> [Rel]
+err :: DRel
+err = [Nothing]
+
+
+-- like and
+comp :: DRel -> DRel -> DRel
+comp rs ss = [liftM2 (>=>) r s | r <- rs, s <- ss]
+
+test :: DRel -> Ass -> [Maybe [Ass]]
+test rs a = map (liftM ($a)) rs
+
+--test :: Rel -> Ass -> Rel
+--test r a = if r a == [] then [] else [[]]
+-- test = (rnot . rnot)
+-- test = impl true
+
+-- notice 'not' cannot bring anybody back 'from the dead'
+rnot :: Rel -> Rel
+rnot r a = if r a == [] then [a] else []
+
+-- just use 'all' and 'test'. Doh
+dnot :: DRel -> DRel
+dnot rs = [foldl (liftM2 (\a b -> a >=> (rnot b))) (Just return) rs]
+
+impl :: DRel -> DRel -> DRel
[…]

### 2013-03-02 exec4.hs

@@ -6,7 +6,7 @@ import Control.Monad ((>=>))
 type Val = Integer
 type Ref = String
 type Ass = [(Ref, Val)]
-type Rel = Ass -> [Ass]
+type Rel = Maybe Ass -> [Maybe Ass]
 
 -- Rel a = [b in ASS | Ez in a (z R b)]
 -- tanke: den initielle assignment er til de fri variable.
@@ -27,43 +27,50 @@ true = return
 false :: Rel
 false = const []
 
+panic :: Rel
+panic = const [Nothing]
+
 -- like and
 comp :: Rel -> Rel -> Rel
 comp = (>=>)
 
-test :: Rel -> Rel
-test r a = if r a == [] then [] else [[]]
+test :: Rel -> Ass -> Bool
+test r a = case sequence (r (Just a)) of
+    Just [] -> False
+    Just _ -> True
+    Nothing -> undefined
 -- test = (rnot . rnot)
 -- test = impl true
 
--- notice 'not' cannot bring anybody back 'from the dead'
 rnot :: Rel -> Rel
-rnot r a = if r a == [] then [a] else []
+rnot r ma = case sequence (r ma) of
+    Just [] -> [Just []]
+    Just _ -> []
+    Nothing -> [Nothing]
 
 impl :: Rel -> Rel -> Rel
--- impl r s = rnot (r `comp` (rnot s))
-impl r s a = if all (\z -> s z /= []) (r a) then [a] else []
+impl r s = rnot (r `comp` (rnot s))
+-- impl r s a = if all (\z -> s z /= []) (r a) then [a] else []
 
 exist :: Ref -> Rel
-exist k a = [set k v a | v <- domain]
+exist k ma = fromMaybe [Nothing] (do
+    a <- ma
+    return [Just (set k v a) | v <- domain])
 
 predi1 :: [Val] -> Ref -> Rel
-predi1 f k a = if elem (get k a) f then [a] else []
+predi1 f k ma = fromMaybe [Nothing] (do
+    a <- ma
+    e <- lookup k a
+    if elem e f then return [Just a] else return [])
 
 predi2 :: [(Val,Val)] -> Ref -> Ref -> Rel
-predi2 f k l a = if elem (get k a, get l a) f then [a] else []
-
--- like or
-union :: Rel -> Rel -> Rel
-union s r a = s a ++ r a
[…]

### 2013-03-02 exec5.hs

@@ -1,15 +1,17 @@
 
 
 import Data.Maybe
-import Control.Monad ((>=>))
+import Control.Monad ((>=>), liftM, liftM2)
 
 type Val = Integer
 type Ref = String
 type Ass = [(Ref, Val)]
-type Rel = Ass -> [Ass]
+type Rel = Maybe Ass -> Maybe [Ass]
 
 -- Rel a = [b in ASS | Ez in a (z R b)]
 -- tanke: den initielle assignment er til de fri variable.
+
+--ma (|>)
 
 sigma :: [Ref]
 sigma = ["x","y"]
@@ -20,46 +22,43 @@ domain = [0,1,2]
 everything :: [Ass]
 everything = sequence [[(k,v) | v <- domain] | k <- sigma]
 
+unlift :: (Maybe a -> Maybe b) -> a -> b
+unlift f x = fromJust (f (Just x))
 
 true :: Rel
-true = return
+true = liftM return
 
 false :: Rel
-false = const []
+false = liftM (const [])
+
+panic :: Rel
+panic = const Nothing
 
 -- like and
 comp :: Rel -> Rel -> Rel
-comp = (>=>)
+comp r s ma = r ma >>= (\as -> liftM concat $ sequence $ map (s . return) as)
 
-test :: Rel -> Rel
-test r a = if r a == [] then [] else [[]]
+--test :: Rel -> Rel
+--test r a = if r a == [] then [] else [[]]
 -- test = (rnot . rnot)
 -- test = impl true
 
--- notice 'not' cannot bring anybody back 'from the dead'
+-- hm, this will make not (error) = [], should be error
 rnot :: Rel -> Rel
-rnot r a = if r a == [] then [a] else []
+rnot r = liftM (\a -> if r a == [] then [a] else [])
 
 impl :: Rel -> Rel -> Rel
 -- impl r s = rnot (r `comp` (rnot s))
-impl r s a = if all (\z -> s z /= []) (r a) then [a] else []
+impl r s = liftM (\a -> if all (\z -> s z /= []) (r a) then [a] else [])
 
 exist :: Ref -> Rel
 exist k a = [set k v a | v <- domain]
 
 predi1 :: [Val] -> Ref -> Rel
-predi1 f k a = if elem (get k a) f then [a] else []
+predi1 f k = liftM (\a -> if elem (get k a) f then [a] else [])
 
[…]

### 2013-03-03 paper.tex
wrote more

@@
 \renewcommand\section{\newpage\stdsection}
+
+\renewcommand{\arraystretch}{1.1}
 
@@
 You can do things such as unification, resolution, and question answering.
+
+In this paper we won't do anything about animal violence
 
@@
 But think about the slight modification to the sentence:
-
+%
 \begin{quotation}
@@
 \end{quotation}
-
+%
 We would have to write that as
-
+%
 \begin{equation}
@@
 \end{equation}
-
+%
 The problem here is that reorderings like this work against our sought property of linearity: of being able to add new information as we go.
@@
 Our sentence from above will then be expressible as
+%
+\begin{equation}
+\Bowtie \cdot \exists x \cdot farmer(x) \cdot \exists y \cdot donkey(y) \cdot \Bowtie \cdot owns(x,y) \cdot \Bowtie \cdot beats(x,y)
+\end{equation}
+%
+which can be written in `stream form' as
+%
+\begin{equation}
+\begin{tabular}{|clll|}
+    \hline
+    $(-)$ & $\exists x \cdot farmer(x) \cdot \exists y \cdot donkey(y)$ & ~ & $beats(x,y)$ \\\hline
+    $(+)$ & ~ & $owns(x,y)$ & ~ \\\hline
+\end{tabular}
+\end{equation}
+%
+The meaning of the above is $(-) \rightarrow (+)$
 
+Should we have one more example here?
+
+So how do we define this nifty $\Bowtie$? The trick is to move the entire dynamic relation algebra from before into an e-monoid that will help us keep track of some details.
+
+If $\phi$ from before is the basic algebra, we specify the construction of our monoid $\Phi$ as follows.
+%
+\begin{flalign}
+&\top := \langle \top, \top, + \rangle & \\
+&\bot := \langle \top, \bot, + \rangle & \\
+&\Bowtie := \langle \top, \top, - \rangle & \\
+&\langle q_-, q_+, \alpha \rangle \cdot \langle r_-, r_+, \beta \rangle := \langle q_- \cdot r_{-\alpha}, q_+ \cdot r_{+\alpha}, \alpha\beta \rangle&
+\end{flalign}
+%
+Notice how we keep the negative stream in the first position of the tuple, and the positive stream at the second position.
+The third position idicates to what stream we are currently writing.
+For example, if we are writing to the negative stream, and we apply an element that itself asks us to continue at its negative stream, we go back to the positive stream.
+
+More examples on the utilities of the above monad can be found in Vissers paper.
+
+An even more interesting element than $\Bowtie$ is $\triangle$. $\triangle$ is a scope modifier, and it allows us to write:
+%
+\begin{quote}
+A farmer owns a donkey. He beats it.
+\end{quote}
+%
+(that is just some single farmer and donkey) as
+%
 \begin{equation}
-\Bowtie \cdot \exists x \cdot farmer(x) \cdot \exists y \cdot donkey(y) \cdot \Bowtie \cdot owns(x,y) \Bowtie \cdot beats(x,y)
+\triangle \cdot \exists x \cdot farmer(x) \cdot \triangle \cdot owns(x,y) \cdot \triangle \cdot \exists y \cdot donkey(y) \cdot \triangle \cdot beats(x,y)
 \end{equation}
+%
+That is pretty close to the order of the natural language!
 
-This is however still not quite as close to natural language as we would like. Specially
+And combining the two monads we can do our original sentence as
+%
+\begin{quote}
+If a farmer owns a donkey, he beats it
+\end{quote}
+%
+as
+\begin{equation}
+\Bowtie \cdot \triangle \cdot \exists x \cdot farmer(x) \cdot \triangle \cdot owns(x,y) \cdot \triangle \cdot \exists y \cdot donkey(y) \cdot \triangle \cdot \Bowtie \cdot beats(x,y)
+\end{equation}
+%
+In stream form, that is:
+%
+\begin{equation}
+\begin{tabular}{|cllll|}
+    \hline
+    $(-,1)$ & $\exists x \cdot farmer(x)$ & ~ & ~ $ \exists y \cdot donkey(y)$ & ~ \\\hline
+    $(-,0)$ & ~ & $owns(x,y)$ & ~ & ~ \\\hline
+    $(+,1)$ & ~ & ~ & ~ & ~ \\\hline
+    $(+,0)$ & ~ & ~ & ~ & $beats(x,y)$ \\\hline
+\end{tabular}
+\end{equation}
+%
+With the meaning being $(-,1) \cdot (-,0) \rightarrow (+,1) \cdot (+,0)$.
+Notice that the negative relations are again put on the left side of the implication, and the relations with larger scope value are put before those of lower value.
+
+Notice $\Bowtie$ and $\triangle$ are commutative, since they work entirely on their own indicator.
+It's an interesting question to consider what sentences might require three or more levels of scoping, but those seam to be so rare that it doesn't now make much sense to clutter our algebra with an extra infinity of streams.
+
+We define the monad as follows:
+%
+\begin{flalign}
+&\top := \langle \top, \top, \top, \top, +, 0 \rangle & \\
+&\bot := \langle \top, \top, \top, \bot, +, 0 \rangle & \\
+&\Bowtie := \langle \top, \top, \top, \top -, 0 \rangle & \\
+&\triangle := \langle \top, \top, \top, \top +, 1 \rangle & \\
+&\langle q_{-,1}, q_{-,0}, q_{+,1}, q_{+,0}, \alpha, i \rangle \cdot \langle r_{-,1}, r_{-,0}, r_{+,1}, r_{+,0}, \beta, j \rangle := &\\
+&\hspace{1cm}\langle q_{-,1} \cdot r_{-\alpha,1+i}, q_{-,0} \cdot r_{-\alpha,i}, q_{+,1} \cdot r_{\alpha,1+i}, q_{+,0} \cdot r_{\alpha,i}, \alpha\beta, i+j \rangle& \nonumber
+\end{flalign}
+
+In his paper Visser now goes on to define $\triangleleft$ and $\triangleright$ which work similarly to $\Bowtie$ except they work in different directions. 
+%
+\begin{equation}
+\Bowtie \cdot \triangle \cdot \exists x \cdot farmer(x) \cdot \triangle \cdot owns(x,y) \cdot \triangle \cdot \exists y \cdot donkey(y) \cdot \triangle \cdot \triangleleft \cdot beats(x,y)
+\end{equation}
+%
+This is quite interesting for certain centences containing `surprises', like
+%
+\begin{quotation}
+He sees... no donkey
+\end{quotation}
+%
+In this case we would compositionally assume to start out with
+$sees(x,y)$ and finish with $\cdot \exists y \cdot donkey(y)\triangle$ hence we want to write something like:
+%
+\begin{equation}
+sees(x,y) \cdot \triangleleft \cdot \bot \cdot \triangleright \cdot \triangle \cdot \exists y \cdot donkey(y) \cdot \triangle
+\end{equation}
+%
+Notice that that means the word `no'/`not' must be $\triangleleft\cdot\bot\cdot\triangleright$.
+Unfortunately these retrospective triangles tend to mess sentence enough up, that they require lots of `stream unification brackets', and don't generally give very nice formulas.
+In the semantic parsing in the next section, they wont be used at all.
 

### 2013-03-22 paper.tex
More equations and nice stuff

@@
 \usepackage[utf8]{inputenc}
-\usepackage{amsthm,amssymb,amsmath,dsfont,mathrsfs,nicefrac,wasysym}
-\usepackage{listings}
+\usepackage{amsthm,amssymb,amsmath,dsfont,mathrsfs,nicefrac,wasysym,listings}
+\usepackage{tikz}
+\usepackage{enumitem}
+
+\usetikzlibrary{shapes,arrows}
+
 \let\stdsection\section
@@
 \renewcommand{\arraystretch}{1.1}
+
+\lstloadlanguages{Haskell}
+\lstnewenvironment{haskell}
+    {\lstset{}%
+      \csname lst@SetFirstLabel\endcsname}
+    {\csname lst@SaveFirstLabel\endcsname}
+    \lstset{
+      basicstyle=\small\ttfamily,
+      escapeinside={/+}{+/},
+      flexiblecolumns=false,
+      xleftmargin=0.05\textwidth,
+      basewidth={0.5em,0.45em},
+      literate={+}{{$+$}}1
+               {/}{{$/$}}1
+               {*}{{$*$}}1
+               {=}{{$=$}}1
+               {>}{{$>$}}1
+               {<}{{$<$}}1
+               {\\}{{$\lambda$}}1
+               {++}{{$+\!\!\!+$}}1
+               {::}{{$:\!\!\!:$}}1
+               {\\\\}{{\char`\\\char`\\}}1
+               {->}{{$\rightarrow$}}2
+               {>=}{{$\geq$}}2
+               {<-}{{$\leftarrow$}}2
+               {<=}{{$\leq$}}2
+               {=>}{{$\Rightarrow$}}2
+               {/=}{{$\neq$}}2
+               {.}{{$\circ$}}2
+               {<<}{{$\ll$}}2
+               {>>}{{$\gg$}}2
+               {>>=}{{$\gg\!=$}}2
+               {>=>}{{$>\!=\!>$}}2
+               {|}{{$\mid$}}1
+               {undefined}{{$\bot$}}1
+               {`elem`}{{$\in$}}1
+    }
 
@@
 Your abstract goes here...
-...
+
+Érica is super hot
+
+‘a brief description of what you did; no more than a page.’
+The abstract provides a summary of the project report. It may contain some or all of the following moves, 
+often, but not always, in the following order. Only use one or two sentences for each move. Keep the 
+language and sentence structure simple. The reader should be able to read the abstract and obtain the 
+necessary information quickly and efficiently.
+Move 1:  Background to the project
+Move 2: Purpose of the project
+Move 3: Problem tackled
+Move 4: Work carried out
+Move 5: Results 
+Move 6: Conclusions or implications
+Move 7: Achievements of the project
+
 \end{abstract}
@@
 
+Move 1 Background 
+Possible Steps 1. Why the area is important
+2. Giving background information
+3. Reviewing previous research
+Move 2 Indicating a Problem or Need
+Move 3 Presenting the Project
+Possible Steps 1. Purposes, aims, objectives
+2. Work carried out
+3. Justification or importance of the project
+4. Outline of the structure of the report
+
 In this paper we won't do anything about animal violence
+
+% Define block styles
+\tikzstyle{decision} = [diamond, draw, fill=blue!20,
+    text width=4.5em, text badly centered, node distance=2.5cm, inner sep=0pt]
+\tikzstyle{block} = [rectangle, draw, fill=blue!20,
+    text width=5em, text centered, rounded corners, minimum height=4em]
+\tikzstyle{line} = [draw, -latex']
+
+\begin{tikzpicture}[node distance = 3cm, auto]
+    % Place nodes
+    \node [block] (nl) {Natural Language};
+    \node [block, right of=nl] (dr) {Dependency Representation};
+    \node [block, right of=dr] (monad) {Enhanced DPL Representation};
+    \node [block, right of=monad] (dpl) {Dynamic Propositional Logic};
+    \node [block, right of=dpl] (prop) {First Order Logic};
+    % Draw edges
+    \path [line] (nl) -- (dr);
+    \path [line] (dr) -- (monad);
+    \path [line] (monad) -- (dpl);
+    \path [line] [dotted] (dpl) -- (prop);
+\end{tikzpicture}
 
@@
 
+Additionally, the indefinite article `a' is normally understood as an existential quantifier
+
+Week and strong readings, check this out \cite{kanazawa1994weak}
+
 In this paper we will not try to solve the above problem.
@@
 
-lksadjf lakdf jlsak dfjlask dfjlask dfjlsa kdfjasld kfjsald kfjsalkd fjslakd fjlsakd fjlkasd fjlaskd jflsak dfj
+This one has problems with scope
+\begin{equation}
+\forall x (farmer(x) \wedge \exists y (donkey(y) \wedge owns(x,y) \rightarrow beats(x,y)))
+\end{equation}
+
+Moving the scope doesn't solve it, since now if there are two donkeys and he only owns one of them, the sentence will be true nomatter if he beats them or not.
+Also if he owns say a donkey and a horse, the left part of the implication would be vacously true and hence the beats part wouldn't matter.
+\begin{equation}
+\forall x \exists y (farmer(x) \wedge donkey(y) \wedge owns(x,y) \rightarrow beats(x,y))
+\end{equation}
 
@@
 
+More stuff from wikipedia:
+
+Unfortunately, this translation leads to a serious problem of inconsistency.
+Indefinites must sometimes be interpreted as existential quantifiers, and other times as universal quantifiers, without any apparent regularity
+
+More questions:
+
+Should those three have the same meaning?
+[1a]  Jane likes a painting.
+[1b]  It is not the case that Jane doesn't like a painting.
+[1c]  Jane likes a painting and either it is raining or it isn't.
+
 \section{Formal Stuff}
 10 pagish
+
 This subsection's content...
+
+Background should introduce any information that is necessary for the examiner to understand your project. 
+The information here is generally more specific than the background given in the Introduction.
+Requirements gives the program requirements. These chapters prepare the ground for the Design chapter.
+That is: What do we need to handle to be able to solve this problem
+
+In particular, anaphora resolution is part of DRT, but in dynamic semantic systems anaphora resolution is assumed to be performed by a separate component of the theory
+
 \subsection{DPL}
@@
 takes the meaning of concepts for given
-discourse representation theory, file change, update semantics, 
-Hans Kamp, Veltman, Groenendijk
+dynamic assignment language, discourse representation theory, file change, update semantics, 
+Jan van Eijck, Hans Kamp, Veltman, Groenendijk
 
@@
 real world example
+
+hvad med sudoku. I started er alle assignments mulige. Nae.. Men. hm
 
@@
 \end{equation}
+
+\subsection{Connection with first order logic}
+
+normal logik kan også laves om til dpl. Nemt, bare brug not.not til at scope E.
+
+An alternative way of defining the semantics
+
+Shows that we aren't doing anything `odd' or `fancy'
+
+shows sound and complete wrt.
+
+Preconditions
+weakest phi sa <phi> pi <T> is still true
+
+Like hoare stuff
+
+inspired Dynamic assignment language \cite{eijck1992dynamic}
+Jan van Eijck 
+Fer-Jan de Vries
+
+like using a stack
+
+$\alpha \models \langle\psi\rangle\phi$ iff there is an assignment $\beta$ with $\alpha[\psi]\beta$, and $\beta \models \phi$
+
+Definition of true and false $\alpha\models\phi :\leftrightarrow \exists\beta(alpha\phi\beta)$.
+
+%{
+
+%Det her skal med for at vi kan forsvare ligningerne
+
+%α[⊥]β :⇔ 0 ≠ 0.
+%α[⊤]β :⇔ α = β.
+%α[P(x0, …, xn−1)]β :⇔ α = β and ⟨αx0, …, αxn−1 ⟩ ∈ I(P), 
+%where P is a predicate symbol of Σ with arity n.
+%α[∃v]β :⇔ α[v]β, where α[v]β iff αw = βw, for all variables w ≢ v.
+%α[φ · ψ]β :⇔ ∃γ  α[φ]γ[ψ]β.
+%α[~(φ)]β :⇔ α = β and ∀γ  ¬ α[φ]γ.
+%
+
+Det giver ikke mening at skrive $\not\models[\psi]$ da typen af $[\psi]$ er en relation. Eller hvad?
+
+\begin{align}
+\alpha[\psi_1\cdot\psi_2]\beta &:\leftrightarrow \exists\gamma (\alpha[\psi_1]\gamma \wedge \gamma[\psi_2]\beta) \label{sem_and} \\
+\alpha[\psi_1\cup\psi_2]\beta &:\leftrightarrow \alpha[\psi_1]\beta \vee \alpha[\psi_2]\beta \label{sem_or} \\
+\alpha[P(x_1,\dots,x_n)]\beta &:\leftrightarrow \alpha = \beta \wedge P(\alpha_{x_1},\dots,\alpha_{x_n}) \label{sem_pred} \\
+\alpha[\neg\psi]\beta &:\leftrightarrow \alpha = \beta \wedge \alpha\not\models[\psi] \label{sem_neg}\\
+\alpha[\exists x]\beta &:\leftrightarrow \forall\omega (\omega \neq x \rightarrow \alpha_{\omega} = \beta_{\omega}) \label{sem_exists} \\
+\intertext{Using these we can derive the following:}
+\alpha[\bot]\beta &\leftrightarrow \alpha = \beta \wedge \bot \nonumber\\
+                  &\leftrightarrow \bot \label{sem_bot}\\
+\alpha[\top]\beta &\leftrightarrow \alpha = \beta \wedge \top \nonumber\\
+                  &\leftrightarrow \alpha = \beta \label{sem_top}\\
+\intertext{Taking $\psi_1\rightarrow\psi_2$ to mean $\neg(\psi_1\wedge\neg\psi_2)$ we also get the following neat equation, originally by Groenendijk and Stokhof\cite{groenendijk1991dynamic}:}
+\alpha[\psi_1\rightarrow\psi_2]\beta
+ &:\leftrightarrow \alpha[\neg(\psi_1\wedge\neg\psi_2)]\beta \nonumber\\
+ &\leftrightarrow \alpha = \beta \wedge \alpha\not\models[\psi_1\wedge\neg\psi_2] \nonumber\\
+ &\leftrightarrow \alpha = \beta \wedge \neg\exists\gamma(\alpha[\psi_1\wedge\neg\psi_2]\gamma) \nonumber\\
+ &\leftrightarrow \alpha = \beta \wedge \neg\exists\gamma(\exists\delta(\alpha[\psi_1]\delta\wedge\delta[\neg\psi_2]\gamma)) \nonumber\\
+ &\leftrightarrow \alpha = \beta \wedge \neg\exists\gamma(\exists\delta(\alpha[\psi_1]\delta\wedge\delta=\gamma\wedge\delta\not\models[\psi_2])) \nonumber\\
+ &\leftrightarrow \alpha = \beta \wedge \forall\gamma(\forall\delta(\delta=\gamma\rightarrow(\alpha[\psi_1]\delta\rightarrow\delta\models[\psi_2]))) \nonumber\\
+ &\leftrightarrow \alpha = \beta \wedge \forall\gamma(\alpha[\psi_1]\gamma\rightarrow\gamma\models[\psi_2]) \label{sem_impl}
+\end{align}
+
+Hence we can read implication as a the filter, that if you can get `through' $\psi_1$ then you must also be able to get through $\psi_2$.
+Also notice that $\gamma$ doesn't have to equal $\alpha$.
+Hence variables introduced in $\psi_1$ can bind variables in $\psi_2$.
+This will show to be extreamly important in the treatment of donkey sentences later.
+
+Talking about binding, also notice that our union or `or' construct is a bit special.
+$\psi_1$ and $\psi_2$ cannot bind variables in each other, but in $(\psi_1\cup\psi_2)\cdot\psi_3$ they can both bind variables in $\psi_3$.
+This is going to introduce a lot of weird situations later.
+An alternative definition of `or': $\neg(\neg\psi_1\cdot\neg\psi_2)$ will also be considered.
+
+For \eqref{sem_pred} $P$ is a predicate symbol of $\Sigma$ with arity n.
+
+static negation. DOesn't allow new things. LIke a filter
+
+some people have a closed/static or: $\neg(\neg\psi_1\wedge\neg\psi_2)$ \cite{groenendijk1991dynamic}
+
+%Notice that \eqref{sem_bot} and \eqref{sem_top} are special cases of \eqref{sem_pred}.
+
+We are going to introduce formuals that allow us to recursively convert any DRA formula into PL.
+
+To smoothen the following proofs, we will introduce a special syntax `$\langle\psi\rangle\phi$' where $\psi$ is a DRA formula, $\phi$ is a PL formula and the whole thing is a PL formula.
+For some assignment $\alpha$, we define `$\alpha\models\langle\psi\rangle\phi$' to mean $\exists\beta(\alpha[\psi]\beta \wedge \beta\models\phi)$.
+
+\begin{align}
+\alpha\models\langle\exists x\rangle\phi
+ & \leftrightarrow \exists\beta (\alpha[\exists x]\beta \wedge \beta\models\phi) \nonumber\\
+ & \leftrightarrow \exists\beta (\forall\omega (\omega \neq x \rightarrow \alpha_{\omega} = \beta_{\omega}) \wedge \beta\models\phi) \nonumber\\
+ & \leftrightarrow \exists\beta (\exists\nu (\forall\omega (\alpha[x:=\nu]_{\omega} = \beta_{\omega})) \wedge \beta\models\phi) \nonumber\\
+ & \leftrightarrow \exists\beta (\exists\nu (\alpha[x:=\nu] = \beta) \wedge \beta\models\phi) \nonumber\\
+% & \leftrightarrow \exists\nu (\exists\beta (\alpha[x:=\nu] = \beta \wedge \beta\models\phi)) \nonumber\\
+ & \leftrightarrow \exists\nu (\alpha[x:=\nu]\models\phi) \nonumber\\
+ & \leftrightarrow \alpha\models\exists x (\phi) \label{conv_exists}\\\nonumber\\
+%
+\alpha\models\langle\psi_1\cdot\psi_2\rangle\phi
+ & \leftrightarrow \exists\beta (\alpha[\psi_1\cdot\psi_2]\beta \wedge \beta\models\phi) \nonumber\\
+ & \leftrightarrow \exists\beta(\exists\gamma (\alpha[\psi_1]\gamma \wedge \gamma[\psi_2]\beta) \wedge \beta\models\phi) \nonumber\\
+ & \leftrightarrow \exists\gamma (\alpha[\psi_1]\gamma \wedge \exists\beta(\gamma[\psi_2]\beta \wedge \beta\models\phi)) \nonumber\\
+ & \leftrightarrow \exists\gamma (\alpha[\psi_1]\gamma \wedge \gamma\models\langle\psi_2\rangle\phi) \nonumber\\
+ & \leftrightarrow \alpha\models\langle\psi_1\rangle\langle\psi_2\rangle\phi \label{conv_and}\\\nonumber\\
+%
+\alpha\models\langle\psi_1 \cup \psi_2\rangle\phi
+ & \leftrightarrow \exists\beta (\alpha[\psi_1 \cup \psi_2]\beta \wedge \beta\models\phi) \nonumber\\
+ & \leftrightarrow \exists\beta ((\alpha[\psi_1]\beta \vee \alpha[\psi_2]\beta) \wedge \beta\models\phi) \nonumber\\
+ & \leftrightarrow \exists\beta (\alpha[\psi_1]\beta \wedge \beta\models\phi) \vee \exists\beta (\alpha[\psi_2]\beta \wedge \beta\models\phi) \nonumber\\
+ & \leftrightarrow (\alpha\models\langle\psi_1\rangle\phi) \vee (\alpha\models\langle\psi_2\rangle\phi) \nonumber\\
+ & \leftrightarrow \alpha\models\langle\psi_1\rangle\phi \vee \langle\psi_2\rangle\phi \label{conv_or}\\\nonumber\\
+%
+\alpha\models\langle P(x_1,\dots,x_n)\rangle\phi
+ & \leftrightarrow \exists\beta (\alpha[P(x_1,\dots,x_n)]\beta \wedge \beta\models\phi) \nonumber\\
+ & \leftrightarrow \exists\beta (\alpha = \beta \wedge P(\alpha_{x_1},\dots,\alpha_{x_n}) \wedge \beta\models\phi) \nonumber\\
+ & \leftrightarrow P(\alpha_{x_1},\dots,\alpha_{x_n}) \wedge \alpha\models\phi \nonumber\\
+ & \leftrightarrow \alpha\models P(x_1,\dots,x_n) \wedge \phi \label{conv_pred}\\\nonumber\\
+%
+\alpha\models\langle\neg\psi\rangle\phi
+ & \leftrightarrow \exists\beta (\alpha[\neg\psi]\beta \wedge \beta\models\phi) \nonumber\\
+ & \leftrightarrow \exists\beta (\alpha=\beta \wedge \neg\exists\gamma(\alpha[\psi]\gamma) \wedge \beta\models\phi) \nonumber\\
+ & \leftrightarrow \neg\exists\gamma(\alpha[\psi]\gamma) \wedge \alpha\models\phi \nonumber\\
+ & \leftrightarrow \neg\exists\gamma(\alpha[\psi]\gamma \wedge \gamma\models\top) \wedge \alpha\models\phi \nonumber\\
+ & \leftrightarrow \alpha\not\models\langle\psi\rangle\top \wedge \alpha\models\phi \nonumber\\
+ & \leftrightarrow \alpha\models \neg\langle\psi\rangle\top \wedge \phi \label{conv_neg}
+\end{align}
+
+Notice that $\alpha\models\langle\bot\rangle\phi$ and $\alpha\models\langle\top\rangle\phi$ follow trivially from \eqref{conv_pred}.
+
+\begin{haskell}
+addDpl :: Prop -> [DPL] -> Prop
+addDpl p [] = p
+addDpl p (DF:ds) = F
+addDpl p (DT:ds) = addDpl p ds
+addDpl p ((DPredi f ks):ds) = addDpl (And (Predi f ks) p) ds
+addDpl p ((DExists k):ds) = addDpl (Exist k p) ds
+addDpl p ((DComp d1 d2):ds) = addDpl p (d2:d1:ds) -- backwards
+addDpl p ((DNot d):ds) = addDpl (And (Not (addDpl T [d])) p) ds
+addDpl p ((DImpl d1 d2):ds) = addDpl p ((DNot (DComp d1 (DNot d2))):ds)
+addDpl p ((DOr d1 d2):ds) = addDpl (And (Or (addDpl T [d1]) (addDpl T [d2])) p) ds
+\end{haskell}
 
@@
 fmap
+
+We can also try to define semantics using first order logic. This totally doesn't work and convert reasonable statements into mess. In particular it is hard to have or not being closed scopewise.
+
+The difference between those two:
+
+$addDpl p ((DOr d1 d2):ds) = addDpl (And (Or (addDpl T [d1]) (addDpl T [d2])) p) ds$
+
+$addDpl p ((DOr d1 d2):ds) = Or (addDpl p (d1:ds)) (addDpl p (d2:ds))$
+
+Fra Dynamic Montague Grammar:
+This means that the extended dynamic semantics is also able to deal with a donkey
+disjunction such as (49), and in such a way that it is equivalent with the donkey
+conditional (50). It can further be noted that this notion of dynamic disjunction is
+like dynamic implication in this respect that sentences which follow it, are in fact
+conjoined with its second argument:
+Fact 24 [either p or q] ; r = either p or [q ; r]
+This is needed for a treatment of examples such as:
+(54) Either there is no bathroom here, or it is in a funny place. In any case, it is
+not on the ground floor.
+
+Den ene giver mulighed for at binde fra a til b. Den anden giver mulighed for at begge dele binder res
+
+\subsubsection{Weird disjunctive subjects}
+
+problem wouldn't work in Spanish
+A farmer or their donkey was arrested this morning.
+
+\begin{equation}
+Ex(farmer(x)) ^ Ey(donkey(y)) ^ (a-t-m(x) v a-t-m(y))
+ExEy(farmer(x) ^ donkey(y) ^ owns(x,y) ^ (a-t-m(x) v a-t-m(y)))
+Ex(farmer(x)) v Ey(donkey(y))
+((Ex . farmer(x) . Es.s=x) v (Ey . donkey(y) . owns(x,y) . Es.s=y)) . a-t-m(s)
+((Ex . farmer(x) . (Es.s=x v (Ey . donkey(y) . owns(x,y) . Es.s=y)) v (Ey . donkey(y) . owns(x,y) . Es.s=y)) . a-t-m(s)
+\end{equation}
+
+If a farmer is Spanish or a Japanese she beats her donkey
+It is a donkey or a horse
+You beat it or I do
+Either a farmer or his donkey is happy
+
+either = xor?
+
+Vissers dynamic bracketing
+
+\subsection{Errors}
+
+Or should this be in implementation? Hard for me to say
 
@@
 
+\subsection{Montague semantic}
+Montague grammars are an approach to combinatorical semantics based on types and lambda calculus. It is interesting because it 
+Types
+Quantification
+
+montague can make any logic compositional
+
+Types of:
+  One-place predicate constant (farmer, sleeps, walks): Entity -> Bool
+  Transitive verb (owns, beats): Entity -> Entity -> Bool
+  Attributive adjective (good, inteligent, former): (Entity -> Bool) -> (Entity -> Bool)
+  Noun phrase: (Entity -> Bool) -> Bool
+  Proper noun: Entity
+  Quantifiers/Determiners: a, the, every, no: (Entity -> Bool) -> (Entity -> Bool) -> Bool
+  Sentence: Bool
+
+\subsection{Weak and strong readings}
+
+Let's return multiple intepretations
+
+drt
+
+The upshot of the foregoing observations is that, apparently, indefinites are neither quantifiers nor referential terms, and this problem entrains another one, for as long as it unclear what indefinites mean, it will also remain obscure how they can serve as antecedents to pronouns
+
+Is this the same as the problem of multiple `scope bearing operators?'
+Sentences with multiple scope bearing operators - e.g.,
+quantified noun phrases or negations - are often ambiguous.
+
+%$• “Every student reads a book”
+%- ∀x(student(x) ⇒ ∃y(book(y) ∧ read(y)(x)))
+%- ∃y(book(y) ∧ ∀x(student(x) ⇒ read(y)(x)))
+
 \subsection{Dependencies}
-This subsection's content...
-\subsection{Montague semantic}
-This subsection's content...
+
+Extracted from parse trees using tregex\cite{de2006generating}
+
+\tikzstyle{level 1}=[level distance=2.5cm, sibling distance=3.5cm]
+\begin{figure}
+\centering
+\begin{tikzpicture}[grow=down, sloped]
+\node {root}
+    child {
+        node {owns}
+        child {
+            node {If}
+            edge from parent
+            node[above] {mark}
+        }
+        child {
+            node {farmer}
+            child {
+                node {a}
+                edge from parent
+                node[above] {det}
+            }
+            edge from parent
+            node[above] {nsubj}
+        }
+        child {
+            node {beats}
+            child {
+                node {he}
+                edge from parent
+                node[above] {nsubj}
+            }
+            child {
+                node {it}
+                edge from parent
+                node[above] {dobj}
+            }
+            edge from parent
+            node[above] {advcl}
+        }
+        child {
+            node {donkey}
+            child {
+                node {a}
+                edge from parent
+                node[above] {det}
+            }
+            edge from parent
+            node[above] {dobj}
+        }
+        edge from parent
+        node[above] {root}
+    };
+\end{tikzpicture}
+\caption{M1} \label{fig:M1}
+\end{figure}
+
+\begin{equation}
+NP < NN \$ VP
+\end{equation}
+
+An NP over an NN and w/ sister VP
+
+\begin{equation}
+NP < (NN < dog) \$ (VP <<\# (barks > VBZ))
+\end{equation}
+
+An NP both over an NN over `dog' and with
+a sister VP headed by `barks' under VBZ
+
+\subsection{Corefferences}
+Corefferences (or anaphora) is essential in order to create advanced discourse, since it is what allows us maintain a topic beond the first mention.
+Without corefferences we couldn't even express simple logical statements such as transitivity: `a man's brother's brother is his brother' etc.
+
+Identifying corefferences (or resolving anaphora) is hence an important part of converting natural language into logic.
+The issue is not touched upon in Visser's paper, which casually sidesteps the problem by assigning variables to all mentioned entities.
+
+For the subtask of resolving simple pronouns a solution could be to work in terms of subjects, objects, indirect objects and so on.
+However if not earlier, this certainly fails for pronouns seperated from their antecedent by sentence bounderies.
+Also we want to resolve other kinds of corefferences, such as `John, the farmer, owns a donkey', where `the farmer' and `John' reffers to the same entity.
+
+It quickly becomes clear that the problem is not only syntactical.
+In the sentence `John and his wife had a farm.
+He took care of the donkeys' we could substitute `He' with `She' and to change the refferant of the pronoun.
+In this case we need semantic knowledge of genders to proceed.
+
+In my project I have taken advantage of the corefference tagger built into Stanford CoreNLP\cite{lee2013deterministic}\cite{lee2011stanford}\cite{raghunathan2010multi}.
+In the following section I will summerize the workings of this machine, made to work across large texts as well as simple sentences.
+
+Correference tagging is one part of computational linguistics where machine learning algorithms still haven't been able to outperform large itterative applications of linguistic heuristics.
+Of course the procedure assumes (most likely) machine learned postagging and named entity recognization, but the main algorithm is transparent and deterministic.
+
+The idea is to iteratively go over the text with different `sieves'. The below example explains each step with an example I have adopted from \cite{lee2013deterministic}:
+
+\begin{quotation}
+John is a farmer. He owned a stubborn donkey.
+A girl was looking at the donkey.
+``It was my favorite,'' John said to her.
+\end{quotation}
+
+\begin{description}
+\item[Mention detection]
+The first step is to identify all noun phrases (NP) and personal (PRP) and possesive (PRP\$) pronouns. This information is easily pulled from the postree. Notice we also identify nested mentions:
+
+[John]$_1^1$ is [a farmer]$_2^2$. [He]$_3^3$ owned [a stubborn donkey]$_4^4$.\newline
+[A girl]$_5^5$ was looking at [the donkey]$_6^6$.\newline
+``[It]$_7^7$ was [[my]$_9^9$ favorite]$_8^8$,'' [John]$_{10}^{10}$ said to [her]$_{11}^{11}$.
+
+All mentions are initially assigned to seperate entities (entity ids).
+The entities are also attached information we can easily pull from the postree, such as `a girl: {number: singular, gender: female}' which are used for the following rules.
+\item[Speaker Sieve]
+The purpose of this first sieve is to link self referring pronouns to speakers.
+Speakers are identified simply by their connection to verbs such as `say' and proximity to the quotations.
+In our case the pronoun `my' gets linked to `John':
+
+[John]$_1^1$ is [a farmer]$_2^2$. [He]$_3^3$ owned [a stubborn donkey]$_4^4$.\newline
+[A girl]$_5^5$ was looking at [the donkey]$_6^6$.\newline
+``[It]$_7^7$ was [\textbf{[my]$_9^1$} favorite]$_8^8$,'' \textbf{[John]$_{10}^{9}$} said to [her]$_{11}^{11}$.
+
+In conversational text we would store information about the speakers in our entities. Thus we could e.g. match mentions of `you' to the speaker of the previous quote we saw.
+
+We also note restrictions for future use, so we don't risk corefferencing an `I' with a `he' etc.
+\item[String Match]
+The second step is to merge entities reffered to by exactly the same string. In our case we have two mentions of `John':
+
+\textbf{[John]$_1^1$} is [a farmer]$_2^2$. [He]$_3^3$ owned [a stubborn donkey]$_4^4$.\newline
+[A girl]$_5^5$ was looking at [the donkey]$_6^6$.\newline
+``[It]$_7^7$ was [[my]$_9^1$ favorite]$_8^8$,'' \textbf{[John]$_{10}^{1}$} said to [her]$_{11}^{11}$.
+\item[Relaxed String Match]
+This third sieve doesn't get activated by our example.
+It will try to remove relative clauses and other text following the head word of a mention, and do an exact match of the result.
+For example [John] and [John, whose donkey loves him] will correctly be merged.
+
+Like the `String Match' sieve, we are not doing anything fancy here, and we will incorrectly corefference the two mentions of `John' in: `[John, who loves donkeys] and [John, who hates donkeys] were brothers.'
+\item[Precise Constructs]
+The `Precise Constructs' sieve uses a lot of common patterns that link metions. In our case two `X is Y' patterns are found:
+
+\textbf{[John]$_1^1$} is \textbf{[a farmer]$_2^1$}. [He]$_3^3$ owned [a stubborn donkey]$_4^4$.\newline
+[A girl]$_5^5$ was looking at [the donkey]$_6^6$.\newline
+``\textbf{[It]$_7^7$} was \textbf{[}[my]$_9^1$ \textbf{favorite]$_8^7$},'' [John]$_{10}^1$ said to [her]$_{11}^{11}$.
+
+Other patterns include acronyms and appositives.
+For example [Canadian Donkey \& Mule Association] and [CDMA] are linked, because the second is tagged as an acronym and it matches the upper case letters of the first one.
+An appositive is usually a pattern `X, Y, ...' where Y is another description of X.
+\item[Strict Head Match A, B, C] are simply rules strip mentions down to their head word and try to find exact matches.
+This is similar to `Relaxed String Match', but more radical.
+In our case we correctly get a link between `a stubborn donkey' and `the donkey':
+
+[John]$_1^1$ is [a farmer]$_2^1$. [He]$_3^3$ owned \textbf{[a stubborn donkey]$_4^4$}.\newline
+[A girl]$_5^5$ was looking at \textbf{[the donkey]$_6^4$}.\newline
+``[It]$_7^7$ was [[my]$_9^1$ favorite]$_8^7$,'' [John]$_{10}^1$ said to [her]$_{11}^{11}$.
+
+Linking mentions based on their head word is dangerous.
+e.g.
+`University of Oxford' and `University of Cambridge' both have `University' as their head word, but are (obviously?) different entities.
+To combat this a lot of strict rules must be observed.
+These are however gradually weakened in sieve B and C.
+\item[Proper Head Noun Match] could be called `Strict Head Match D'.
+It is the weakest form of the sieve, as it has only the three contraints: two mentions of the same entity cannot have one included in the other, they cannot contain words reffering to conflicting geographical locations, and they cannot have different numbers.
+e.g.
+[donkeys] and [at least 5 donkeys].
+\item[Pronoun Match]
+This last sieve is perhaps the most important one.
+It is done in the end so that we may have as much data in our entities as possible for matching.
+Using guesses on gender, number and animacy we can make our final merges:
+
+\textbf{[John]$_1^1$} is [a farmer]$_2^1$. \textbf{[He]$_3^1$} owned [a stubborn donkey]$_4^4$.\newline
+\textbf{[A girl]$_5^5$} was looking at \textbf{[the donkey]$_6^4$}.\newline
+``\textbf{[It]$_7^4$} was [[my]$_9^1$ favorite]$_8^4$,'' [John]$_{10}^1$ said to \textbf{[her]$_{11}^{5}$}.
+
+As always there are a few contraints put on what can be merged. One rule is that the distance between pronun and antecedent must be maximum 3 sentences.
+\end{description}
+
+After the final step different post processing options are usually performed. For example singleton mentions are often discarded. That would mean we got rid of `a musician' and `my favorite'.
+
+The reason the rules are applied in the order above is to give the most certain rules the highest priority.
+
+The Stanford procedure above is currently the strongest competitor in various correference competitions.
+
 \subsection{Semantics via dependencies}
@@
 
+There is not a big first order logic component in my project, but in the following it will often be useful to refer to the following structure
+
+\begin{haskell}
+data Prop =
+    T | F
+  | Not   Prop
+  | And   Prop Prop
+  | Or    Prop Prop
+  | Predi Ref [Ref]
+  | Exist Ref Prop
+  | All Ref Prop
+  | Impl Prop Prop
+\end{haskell}
+
+And since we are going to deal with some large computer generated formulas, I've implemented a simple simplification function, which simlply applies basic identities for $\top$ and $\bot$, and a comnbination of De Morgan's laws and the implication identity to eliminate negations.
+
+One interesting part is that since the simplifyer has no look ahead, it will sometimes not find all applicable simplifications in one pass. For example
+%
+\begin{haskell}
+simplify (Not (Impl p F)) = Not (simplify (Impl p F))
+                          = Not (Not simplify p)
+                          = Not (Not q)
+\end{haskell}
+%
+Hence we need to run it a couple of times to get a good result. We can run it until the output doesn't change, or even simpler, just use a function like
+%
+\begin{haskell}
+fexp :: (a -> a) -> Int -> (a -> a)
+fexp f e = (iterate (f.) id) !! e
+\end{haskell}
+
 \subsection{My implementaiton of DPL semantics}
-Maybe this should be in the formal stuff. Nah, there are two versions
+
+Syntactically we will be using the following structure
+
+\begin{lstlisting}
+data DPL =
+    DT | DF
+  | DComp DPL DPL
+  | DOr DPL DPL
+  | DNot DPL
+  | DImpl DPL DPL
+  | DExists Ref
+  | DPredi Ref [Ref]
+\end{lstlisting}
 
@@
 type Assignment = [(Ref, Val)]
-type Relation = [Assignment] -> [Assignment]
+type Relation = Assignment -> [Assignment]
 \end{lstlisting}
@@
 type Assignment = [(Ref, Val)]
-type Relation = [Assignment] -> [Assignment] -> Bool
+type Relation = Assignment -> Assignment -> Bool
 \end{lstlisting}
+
+intuitive
+
+uses well known isomorphism
+
+more direct
+
 The later is very difficult to implement, since the semantics require quantification over all assignments. Like for composition.
@@
 
+\subsection{Handling variables}
+One thing I haven't found much \cite{visser1999donkey}
+Stanford: \cite{lee2013deterministic}
+\cite{lee2011stanford}
+\cite{raghunathan2010multi}
+Coreferences
+Using the parse tree and NPs
+
+Something that's very interesting is when to use free variables and when to use bound variables.
+We can think of free variables as actors known by the context and bound variables as actors introduced in the sentence.
+So if we had `The farmer had a friend.
+His name was Anders', we could write the first part as $\exists y \cdot friends(x,y)$ and the second part as $name(y, `Anders')$.
+This is important to keep compositionality, but it does of course raise the interesting question of how to make sure that the variable for `Anders' in the first and the second sentence is the same.
+
+In addition to being essential for composition, free variables are also very useful for information extraction type sentences.
+We can immagine writing a curious researcher writing into a search engine `farmers who like donkeys' which translates into $farmer(x) \cdot likes(x,y) \cdot \triangle \cdot \exists y \cdot donkey(y)$.
+
+So how do we know when to include an $\exists$ and when not to?
+
+The coreference parser does however not find entities that are only mentioned once. Hence to make the later processing simpler, we assign those entitie to free variable now as well.
+
+The task is slightly complicated by the fact that NP's can be nested, so `The farmer and the donkey' is a NP consisting of two smaller NP's `The farmer' and `the donkey' seperated by `and'.
+
+We resolve this by assigning top down, overwriting `word to ref' assignments with fresh variables as we go deeper down the tree.
+This way, for our short sentence, we might end up with the assignment $\{`The': `y', `farmer': `y', `and': `x', `the': `z', `donkey': `z'\}$
+
+I parse the parse tree using parsec\cite{leijen2001parsec}
+
+\begin{lstlisting}
+toMapping :: [Ref] -> PosTree Word -> ([Ref], Map Int Ref)
+toMapping vs (Leaf _ _) = (vs, empty)
+
+toMapping vs tr@(Phrase "NP" subs) = (vs', submaps `union` np)
+  where np = mapInterval 0 (tsize tr) v
+        (v:vs', submaps) = toMapping vs (Phrase "" subs)
+
+toMapping vs (Phrase _ []) = (vs, empty)
+toMapping vs (Phrase _ (sub:subs)) = (vs'', p `union` shifted)
+  where (vs', subm) = toMapping vs (Phrase "" subs)
+        shifted = mapKeys (+tsize sub) subm
+        (vs'', p) = toMapping vs' sub
+\end{lstlisting}
+
+where $tsize :: PosTree Word -> Int$ gives the number of words in the tree
+
 \subsection{Using machine learning}
@@
 
-\section{Perspectives}
+\subsection{Examples}
+Until now we've been very focused on the same one or two simple sentences.
+To provide a better test bench for our approach, it is interesting to follow some sentences through the different levels of our approach.
+
+\begin{enumerate}
+\item
+Some description?
+\begin{description}[style=multiline, leftmargin=10.25em]
+  \item[Natural language] If a farmer who owns a donkey, he beats it 
+  \item[Dependency Graph] ... eh
+  \item[Monadic DPL] $\Bowtie \cdot \triangle \cdot \exists x \cdot farmer(x) \cdot \triangle \cdot owns(x,y) \cdot \triangle \cdot \exists y \cdot donkey(y) \cdot \triangle \cdot \Bowtie \cdot beats(x,y)$
+  \item[DPL] $\exists x \cdot farmer(x) \cdot \exists y \cdot donkey(y) \cdot owns(x,y) \rightarrow beats(x,y)$
+  \item[FOL] $\forall x ( farmer(x) \rightarrow \forall y ( donkey(y) \wedge owns(x,y) \rightarrow beats(x,y)))$
+\end{description}
+
+\item The second item
+\item
+`John wants to marry a Spanish girl'
+
+This one is hard because it is an example of existential non commitment. I bet we'll fail it.
+\end{enumerate}
+
+Extract D Testing 
+We have already seen some screen shots of the working program; however, we provide two stringent tests 
+for our program to ensure it works as intended, along with a test rig to fully analyse the program. In both test 
+programs, I will run through the whole series of options available to the user, and ensure its correctness. 
+However, I will also demonstrate its ability to visualise code, and hopefully provide valuable insights whilst 
+debugging. 
+5.1 Simple Program -BFS and DFS using the Visitor Pattern
+This test program begins by creating an underlying tree structure…
+This kind of debugging is intuitive, and simple to do within this framework. If you have an intuitive 
+understanding of what the underlying model in your program should look like, it is fairly straight forward to 
+spot bugs like this in  small code samples. Assuming a larger program is in use, the user must delve a little 
+deeper into the part of the graph which they suspect the bug to exist in. This is obviously heavily aided by the 
+JDT debugger itself. However, this test still shows the usability of the code in a small program, and shows 
+that the code can cope with the different types of back links and cross links that can occur in a memory 
+graph.
+
+\section{Perspectives/Conclusions}
 2pagish
+
 What I learned
+
+‘a summary of your achievements; a critical appraisal, describing what worked well and what could be 
+improved; a discussion of the lessons you have learnt from the project.’
+In the ‘Conclusions’ you summarise your work and then stand back from the project and assess it as 
+objectively as you can. This shows the examiners that you are capable of evaluating your own work 
+according to the standards of the field. It is acceptable to mention areas that were not very satisfactory
+because this shows that you have learnt from the experience of doing the project.
+There are often three moves starting with a summary of your work, then giving an evaluation, highlighting
+both its achievements and limitations and finally the possible future extensions of the work. The Evaluation 
+move may also be carried out as part of the Summary. The Conclusion is the mirror image of the 
+Introduction, in that it starts with the narrow concerns of the project and widens out to more general 
+statements about further work.
+
+Move 1: Summary 
+Summarises the most important aspects of the project work.
+
+I have successfully developed a novel local type inference algorithm
+
+Move 2: Evaluation
+Indicates what the writer considers to be the most valuable aspects of the project and its limitations. Provides 
+an assessment of how far the aims of the project have been achieved.
+
+my algorithm is exponential: whereas Gagnon's algorithm is 
+polynomial. But experiments of execution time against method length show a typically linear trend whereas 
+Gagnon's show a cubic trend
+
+Move 3: Future Work
+Gives possible extensions of the project. These suggestions may follow from the limitations mentioned in the 
+Evaluation move.
+
+The greatest scope for future work is extending the application of my algorithm from local type inference to 
+global type inference.
+
+Should I make my latex file literate haskell? %bhttp://www.haskell.org/haskellwiki/Literate_programming
+
+\lstinputlisting[language=Haskell]{Main.hs}
 
@@
 Thanks to Samson
+
 This section's content...\cite{visser1999donkey}
+
 the PCFG parser\cite{klein2003accurate}
+
+Érica is still super hot. And she rocks at linguistics.\cite{lander2001initial}
 

### 2013-03-22 Dpl.hs

@@ -1,4 +1,4 @@
-module Dpl (dpl2prop, Type2, s2dpl,
+module Dpl (DPL(..), dpl2prop, Type2, s2dpl,
 	sT, sF, sSwitch, sScope, sPredi, sExist, sComp, sTest0, sTest1) where
 
 import Prop (Ref, Prop(..))
@@ -7,14 +7,14 @@ import Prop (Ref, Prop(..))
 ----------------------------------------
 
 data DPL =
-	    DT | DF
-	  | DPredi Ref [Ref]
-	  | DExist Ref
-	  | DComp DPL DPL
-	  | DOr DPL DPL
-	  | DNot DPL
-	  | DImpl DPL DPL
-	  deriving (Eq, Show)
+    DT | DF
+  | DComp DPL DPL
+  | DOr DPL DPL
+  | DNot DPL
+  | DImpl DPL DPL
+  | DExists Ref
+  | DPredi Ref [Ref]
+  deriving (Eq, Show)
 
 -- TODO: Move exec code here
 
@@ -26,16 +26,21 @@ addDpl p [] = p
 addDpl p (DF:ds) = F
 addDpl p (DT:ds) = addDpl p ds
 addDpl p ((DPredi f ks):ds) = addDpl (And (Predi f ks) p) ds
-addDpl p ((DExist k):ds) = addDpl (Exist k p) ds
-addDpl p ((DComp d1 d2):ds) = addDpl p (d2:d1:ds) -- backwards
+addDpl p ((DExists k):ds) = addDpl (Exist k p) ds
+addDpl p ((DComp d1 d2):ds) = addDpl p (d2:d1:ds) -- backwards: explain
 addDpl p ((DNot d):ds) = addDpl (And (Not (addDpl T [d])) p) ds
 addDpl p ((DImpl d1 d2):ds) = addDpl p ((DNot (DComp d1 (DNot d2))):ds)
-addDpl p ((DOr d1 d2):ds) = addDpl (And (Or (addDpl T [d1]) (addDpl T [d2])) p) ds
+-- This rule is no good at all. Scoping becomes weird
+--addDpl p ((DOr d1 d2):ds) = addDpl (And (Or (addDpl T [d1]) (addDpl T [d2])) p) ds
+
+addDpl p ((DOr d1 d2):ds) = Or (addDpl p (d1:ds)) (addDpl p (d2:ds))
+
 
[…]

### 2013-03-22 Prop.hs

@@ -1,4 +1,6 @@
-module Prop (Ref, Prop(..), simplify) where
+module Prop (Ref, Prop(..), simplify, tostring) where
+
+import Data.List
 
 type Ref = String
 
@@ -18,9 +20,11 @@ simp :: Prop -> Prop
 simp (Not (Exist k p)) = simp (All k (Not p))
 simp (Not (All k p)) = simp (Exist k (Not p))
 simp (Not (And p q)) = simp (Impl p (Not q))
+simp (Not (Impl p q)) = simp (And p (Not q))
 simp (Not (Not p)) = simp p
 simp (Not T) = F
 simp (Not F) = T
+
 
 simp (Impl p F) = simp (Not p)
 simp (Impl p (Impl q r)) = simp (Impl (And p q) r)
@@ -34,6 +38,9 @@ simp (Exist k p) = Exist k (simp p)
 simp (Not p) = Not (simp p)
 simp (Impl p q) = Impl (simp p) (simp q)
 
+-- We still don't have too many rules for Or, as we are quite interested in things like `foo v false`
+simp (Or p q) = Or (simp p) (simp q)
+
 simp p = p
 
 fexp :: (a -> a) -> Int -> (a -> a)
@@ -41,3 +48,13 @@ fexp f e = (iterate (f.) id) !! e
 
 simplify :: Prop -> Prop
 simplify = fexp simp 10
+
+tostring T = "T"
+tostring F = "F"
+tostring (Not p) = "!(" ++ tostring p ++ ")"
+tostring (And p q) = tostring p ++ "&" ++ tostring q
+tostring (Or p q) = "(" ++ tostring p ++ ")|(" ++ tostring q ++ ")"
+tostring (Impl p q) = "(" ++ tostring p ++ ")->(" ++ tostring q ++ ")"
+tostring (Predi f ks) = f ++ "(" ++ concat (intersperse "," ks) ++ ")"
+tostring (Exist k p) = "E" ++ k ++ "(" ++ tostring p ++ ")"
+tostring (All k p) = "A" ++ k ++ "(" ++ tostring p ++ ")"

### 2013-03-22 ToLogic.hs

@@ -2,7 +2,7 @@ module ToLogic (logify) where
 
 import Dpl (dpl2prop, s2dpl, Type2, sT, sF, sSwitch, sScope, sPredi, sExist, sComp, sTest0, sTest1)
 import Prop
-import Stanford (DepTree(..), Word, Sentence)
+import Stanford (DepTree(..), Word, Sentence, PosTree(..))
 
 import Data.Map hiding (lookup, mapMaybe)
 import Data.Maybe
@@ -34,9 +34,85 @@ logify (sentence, refs, deps) = (simplify . dpl2prop . s2dpl) logic
 	where
 		logic = parseVP sentence refs deps
 
+{-
 monte :: PosTree (Int, Word) -> Map Int Ref -> Type2
 monte (Phrase "ROOT" ts) = monte (head ts)
 monte (Phrase "S" [(Phrase "NP" a), (Phrase "VP" b)])
+
+
+parseRoot ws refs (D n ds)
+	|
+	 [("auxpass", p), ("nsubjpass", q) ("tmod", r)])
+
+
+    monteHead (D _ vb [ns@(D "nsubj" _ _), do@(D "dobj" _ _)]) = (monteDobj dobj) (monteNsubj nsubj)
+    monteHead ((Phrase "SBAR" [(Leaf "IN" "if"), (Phrase "S" s1)]):(Leaf "," _):s2) = sSwitch `sComp` monteS s1 `sComp` sSwitch `sComp` monteS s2
+
+parse (P "NP" [("")])
+-}
+
+-- we can use a lookup :: a -> [(a,b)] -> [b]
+-- but then how do we patternmatch to avoid empties? n:ns?
+monteHead (w, es)
+	| (n:subj) <- lookups "nsubj" es
+	, (d:obj) <- lookups "dobj" es
+	= monteNsubj nsubj `sComp`
+
+lookups :: (Eq a) => a -> [(a,b)] -> [b]
+lookups _ []                         = []
+lookups  key ((x,y):xys) | key == x  = y : lookup key xys
+                         | otherwise = lookup key xys
+
+
+monteRoot :: [PosTree Word] -> Type2
+monteRoot [(Phrase "ROOT" xs)] = monteRoot xs
+monteRoot [(Phrase "S" xs)] = monteS xs
+monteRoot p = error (show p)
+
+monteS :: [PosTree Word] -> Type2
[…]

### 2013-03-22 exec.hs

@@ -1,5 +1,6 @@
 
 
+import Dpl (DPL(..))
 import Data.Maybe
 import Control.Monad ((>=>))
 
@@ -7,6 +8,7 @@ type Val = Integer
 type Ref = String
 type Ass = [(Ref, Val)]
 type Rel = Ass -> [Ass]
+type Pre = [[Val]]
 
 -- Rel a = [b in ASS | Ez in a (z R b)]
 -- tanke: den initielle assignment er til de fri variable.
@@ -17,9 +19,25 @@ sigma = ["x","y"]
 domain :: [Val]
 domain = [0,1,2]
 
+predicates :: [(Ref, Pre)]
+predicates = [
+    ("farmer", [[0]]),
+    ("donkey", [[1]]),
+    ("owns", [[0,1]]),
+    ("beats", [[0,1]])
+    ]
+
 everything :: [Ass]
 everything = sequence [[(k,v) | v <- domain] | k <- sigma]
 
+sem :: DPL -> Rel
+sem DT = return
+sem DF = const []
+sem (DComp r s) = sem r >=> sem s
+sem (DNot r) = \a -> if sem r a == [] then [a] else []
+sem (DImpl r s) = \a -> if all (\z -> sem s z /= []) (sem r a) then [a] else []
+sem (DExists k) = \a -> [set k v a | v <- domain]
+sem (DPredi f ks) = \a -> if elem (map (get a) ks) (get predicates f) then [a] else []
 
 true :: Rel
 true = return
@@ -48,10 +66,10 @@ exist :: Ref -> Rel
 exist k a = [set k v a | v <- domain]
 
 predi1 :: [Val] -> Ref -> Rel
-predi1 f k a = if elem (get k a) f then [a] else []
+predi1 f k a = if elem (get a k) f then [a] else []
 
 predi2 :: [(Val,Val)] -> Ref -> Ref -> Rel
-predi2 f k l a = if elem (get k a, get l a) f then [a] else []
+predi2 f k l a = if elem (get a k, get a l) f then [a] else []
 
 -- like or
 union :: Rel -> Rel -> Rel
@@ -62,8 +80,10 @@ forall k r = exist k `impl` r
 --forall k r = rnot (exist k (rnot r))
 
 -- Will fail if ref not in ass
-get :: Ref -> Ass -> Val
-get k = fromJust . (lookup k)
+get :: Eq a => [(a, b)] -> a -> b
[…]

### 2013-03-25 paper.tex
Nearly 7000 words!

@@
 
-\title{Title of my paper}
+\title{Musings in Dynamic Semantics for Donkeys}
 \author{Thomas Dybdahl Ahle}
@@
 
-Érica is super hot
+donkeys sentences
+dynamic logic
+disjunction, errors
+monads
+dependencies
+
+From Donkeys to Predicates: Musings Into Dynamic Semantics and Logification
+Donkification and Logification
+Donkeys into Logic
+From Farmers and Donkeys to Grammars and Predicates
+Dependency ased
+Discourse and Donkeys: A Dependency Based 
+Discourse and Donkeys: A Monoid Based Language Grammar
+Discourse and Donkeys: Monoid Based Language Semantics
+Discourse and Donkeys: Using Monoids for Language Semantics
+Dynamic Discourse and Donkeys: Using Monoids for Language Semantics
 
@@
 
+Let's try it then
+
+How to treat determinants/indeterminants
+  1) ref
+  2) quant
+  3) vp -> sentence function
+
+Ambiguity
+  1) Strong/Weak
+
+How to get a good logical representation
+  1) Looking like language
+  2) Compositionality
+
 \end{abstract}
@@
 
-\section{Introduction}
-5 pagish Donkey example to follow through.
+\section{Introduction (5)}
+% 5 pagish Donkey example to follow through.
+
+% Move 1 Background 
+% Possible Steps 1. Why the area is important
+% 2. Giving background information
+% 3. Reviewing previous research
+% Move 2 Indicating a Problem or Need
+% Move 3 Presenting the Project
+% Possible Steps 1. Purposes, aims, objectives
+% 2. Work carried out
+% 3. Justification or importance of the project
+% 4. Outline of the structure of the report
+
 There are many reasons why it is interesting to represent information in a logical language.
-You can do things such as unification, resolution, and question answering.
-
-Move 1 Background 
-Possible Steps 1. Why the area is important
-2. Giving background information
-3. Reviewing previous research
-Move 2 Indicating a Problem or Need
-Move 3 Presenting the Project
-Possible Steps 1. Purposes, aims, objectives
-2. Work carried out
-3. Justification or importance of the project
-4. Outline of the structure of the report
-
-In this paper we won't do anything about animal violence
+For a computer scientist, the biggest advantage is the computational aspect.
+There is a horde of algorithms for doing querying, question answering, reasoning, unification and resolution.
+All are they at your disposal if only you can obtain a first order logic form.
+
+Getting computers to understand and work with natural language is arguably on of the biggest challenges in information technologies this day.
+Hence if we could move the problem into the space of formal logic, it would be a fantastic accomplishment.
+Unfortunately many people have pointed out, that classical logic doesn't have the power to express most of the sentences people casually express on a daily basis.
+
+To draw up this point, let's first look at a sentence that easily lets itself convert and then a couple that dont:
+\begin{itemize}
+\item `Socrates is a donkey': $donkey(Socrates)$.
+This is a classical logical example, treating Socrates as a constant.
+If we want to do without constants we can write: $\forall x(Socrates(x)\rightarrow donkey(x))$, though this does allow more than one Socrates to exist.
+\item `All his donkeys are treated well': This sentence somewhat willingly let's itself translate into: $\forall y(his(x,y)\wedge donkey(y) \rightarrow treated\_well(y))$, which has a free variable $x$ refering to some person from earlier in the discourse.
+\item `Every farmer beats his donkey': This sentence clearly has two possible meanings: $\exists y(his(x,y)\wedge donkey(y)\wedge \forall z(farmer(z) \rightarrow beats(z,y))$ if there is only one donkey, and $\forall x(farmer(x)\rightarrow \exists y (beats(x,y)\wedge his(x,y)\wedge donkey(y))$ if there is a donkey for each farmer.
+\item `The happy donkey doesn't exist': Here we get into more problems.
+We might write: $\exists x(happy\_donkey(x)\rightarrow doesnt\_exist(x))$, but that asserts some sort of entity representing the non existing donkey.
+Alternatively we could write $\exists x(happy\_donkey(x)\rightarrow \bot)$, but that is another sentence: `There are no happy donkeys'.
+\item `They donkeys are cold': This is an often used example of an `implied action sentence' and thought to have the same meaning as `Please close the stable door'.
+The best we can do here seems to be $please\_close\_the\_stable\_door()$ or perhaps $\exists x(stable\_door(x)\wedge please\_close(x))$.
+\end{itemize}
+
+In this paper we will work with a subset of English that translate reasonably well into logic.
+This unfortunately means that arguably the majority of sentences will be unaccesible to our methods.
+We will take a shot at ambigious sentences like the third one though.
+
+%TODO: Perhaps this should be moved to the aknowledgements or something. It sets a very demoralizing tone during the rest of the paper.
+Even more unfortunate than the problems above is it that we won't be able to do anything about the concerning welfare for donkeys around the world\cite{donkey2013sanctuary}, but must restrict ourselves to simply pointing it out.
+We will follow Geach\cite{geach1962reference} on this, and make `If a farmer owns a donkey, he beats it' our primary object of study.
+%
+\begin{align}
+\exists x (farmer(x) \wedge \exists y (donkey(y)\wedge owns(x,y))) \rightarrow beats(x,y)
+\end{align}
+%
+This is arguably the sentence that first comes to mind for a moderately trained logician.
+Translating an indeterminate `a farmer' into $\exists x(farmer(x))$ seems correct, but clearly the formula has problems with scope.
+%
+\begin{align}
+\exists x (farmer(x) \wedge \exists y (donkey(y)\wedge owns(x,y) \rightarrow beats(x,y))) \label{wrong_donkey_2}
+\end{align}
+%
+Moving the implication into the scope of $x$ and $y$ also doesn't work.
+Sentence \eqref{wrong_donkey_2} evaluates to true as long as a single farmer beats his donkey.
+In fact as long as a farmer exists it evaluates to true, since plugging the farmer into the $donkey$ predicate makes the anticedent false.
+%
+\begin{align}
+\forall x (farmer(x) \rightarrow \forall y ( owns(x,y) \wedge donkey(y) \rightarrow beats(x,y))) \label{correct_donkey}
+\end{align}
+%
+This turns out to be the correct formula. At least as long as you consider the use of `all' in natural language equal to the `all' in logic, which can be vacously true.
+
+So what is going on?
+Most people think of determinates `the donkey' and indeterminates `a donkey' as being a reference to an existing donkey and an `introduction' of some new donkey into the conversation.
+However in a lot of sentences this turns out to not be the case.
+
+How to treat determinants/indeterminants
+  1) ref
+  2) quant
+  3) vp -> sentence function
+
+Another problem with sentence \eqref{correct_donkey} is a lack of compositionality.
+
+How to get a good logical representation
+  1) Looking like language
+  2) Compositionality
+
+Finally let's have a look at the ambiguity we noted in the beginning of this chapter
+
+Ambiguity
+  1) Strong/Weak
+
+If a farmer owns a donkey he beats it
+\begin{equation}
+\forall x . farmer(x) \rightarrow (\forall y . donkey(y) \wedge owns(x,y) \rightarrow beats(x,y))
+\end{equation}
+
+There are a lot of problems with this representation
+\begin{itemize}
+  \item It is not compositional. 
+  dfdf
+  \item It doesn't very well follow the order of the original sentence
+  \item It is just not very easy to get right
+\end{itemize}
+
+More stuff from wikipedia:
+
+Unfortunately, this translation leads to a serious problem of inconsistency.
+Indefinites must sometimes be interpreted as existential quantifiers, and other times as universal quantifiers, without any apparent regularity
+
+Another reason for studying to so called `donkey sentences' is their particular use of referants.
+
+The upshot of the foregoing observations is that, apparently, indefinites are neither quantifiers (dårlig konstruktion) nor referential (no farmer, etc.) terms, and this problem entrains another one, for as long as it unclear what indefinites mean, it will also remain obscure how they can serve as antecedents to pronouns.
+Later on we will see that very similar problems arise in seemingly disparate domains, like the interpretation of tense and presuppositions.
+
+(1)  A definite is used to refer to something that is already familiar at the current stage 
+of the conversation. An indefinite is used to introduce a new referent. 
+This has been  labeled the `familiarity  theory of definiteness.'
+When confronted with (l), the logically minded semanticist will notice immediately 
+that it  presumes  something  very  objectionable:  that  definites and indefinites  are 
+referring expressions.  For  only  if  there  is  a referent at  all  can there be  any question 
+of its familiarity or novelty
+
+In montague it is a function from verb to sentence
+
+If there is a Chompsky universal grammar, we should be able to make a simple logic over it??
 
@@
     \node [block, right of=dr] (monad) {Enhanced DPL Representation};
-    \node [block, right of=monad] (dpl) {Dynamic Propositional Logic};
+    \node [block, right of=monad] (dpl) {Dynamic Predicate Logic};
     \node [block, right of=dpl] (prop) {First Order Logic};
@@
 
-\begin{equation}
-\forall x . farmer(x) \rightarrow (\forall y . donkey(y) \wedge owns(x,y) \rightarrow beats(x,y))
-\end{equation}
-
-This one has problems with scope
-\begin{equation}
-\forall x (farmer(x) \wedge \exists y (donkey(y) \wedge owns(x,y) \rightarrow beats(x,y)))
-\end{equation}
-
-Moving the scope doesn't solve it, since now if there are two donkeys and he only owns one of them, the sentence will be true nomatter if he beats them or not.
-Also if he owns say a donkey and a horse, the left part of the implication would be vacously true and hence the beats part wouldn't matter.
-\begin{equation}
-\forall x \exists y (farmer(x) \wedge donkey(y) \wedge owns(x,y) \rightarrow beats(x,y))
-\end{equation}
-
-If a farmer owns a donkey he beats it
-\begin{equation}
-\forall x . farmer(x) \rightarrow (\forall y . donkey(y) \wedge owns(x,y) \rightarrow beats(x,y))
-\end{equation}
-
-There are a lot of problems with this representation
-\begin{itemize}
-  \item It is not compositional. 
-  dfdf
-  \item It doesn't very well follow the order of the original sentence
-  \item It is just not very easy to get right
-\end{itemize}
-
-More stuff from wikipedia:
-
-Unfortunately, this translation leads to a serious problem of inconsistency.
-Indefinites must sometimes be interpreted as existential quantifiers, and other times as universal quantifiers, without any apparent regularity
-
 More questions:
@@
 [1c]  Jane likes a painting and either it is raining or it isn't.
-
-\section{Formal Stuff}
-10 pagish
-
-This subsection's content...
-
-Background should introduce any information that is necessary for the examiner to understand your project. 
-The information here is generally more specific than the background given in the Introduction.
-Requirements gives the program requirements. These chapters prepare the ground for the Design chapter.
-That is: What do we need to handle to be able to solve this problem
-
-In particular, anaphora resolution is part of DRT, but in dynamic semantic systems anaphora resolution is assumed to be performed by a separate component of the theory
-
-\subsection{DPL}
-Dynamic semantics: http://plato.stanford.edu/entries/dynamic-semantics/
-
-growth of information in time
-Inspired by programming language
-not philosofically, states can be anything. I don't want to get into that
-isomorphic with normal logic (visser siger generalization/projection)
-doen't promise to be a perfect dual of natural language, but it's closer, and has been rewarding
-Dynamic semantics aims to model meaning and interpretation. You can do that without answering broader philosophical questions
-takes the meaning of concepts for given
-dynamic assignment language, discourse representation theory, file change, update semantics, 
-Jan van Eijck, Hans Kamp, Veltman, Groenendijk
-
-compoitionality. A man comes in. He sees a dog. He smiles.
-
-Vi har brug for en logik, der ikke behøver at vide alting fra starten
 
@@
 
-relation between sets of assignments
-$\{\overline{PQR}, \overline{PQ}R, \overline{P}Q\overline{R}, \overline{P}QR, P\overline{QR}, P\overline{Q}R, PQ\overline{R}, PQR\}$
-
-example
-
-real world example
-
-hvad med sudoku. I started er alle assignments mulige. Nae.. Men. hm
-
-something about more information, less models
-the interpretation of false
-
-Something about kleiski combinator?
-
-the thing about quantifiers
-how they would work in a programming language
-$\exists x; A$
-$\exists x; x = 4$ hvis $x=4$ skal virke og reducere mængden af modeller, så må exists gøre dem alle sammen mulige
-
-A man comes in. En fri variabel vi ikke ved noget om.
-A man comes in. It's Thomas..
-
-we take as DPL-meanings binary relations between assignments
-
-later we will use an isomorphism and take them as functions from assignments to sets of assignments.
-
-\begin{equation}
-\phi ::= \bot \mid \top \mid \pi \mid \epsilon \mid \phi \cdot \phi \mid \lnot(\phi)
-\end{equation}
-\begin{equation}
-\alpha [\bot] \beta :\equiv False
-\end{equation}
-\begin{equation}
-\alpha [\top] \beta :\equiv \alpha = \beta
-\end{equation}
-\begin{equation}
-\alpha[P(x_0, ..., xn-1)]\beta :\equiv \alpha = \beta and (\alpha x0, ..., \alpha x n-1 ) \in I(P)
-\end{equation}
-
-\subsection{Connection with first order logic}
-
-normal logik kan også laves om til dpl. Nemt, bare brug not.not til at scope E.
-
-An alternative way of defining the semantics
-
-Shows that we aren't doing anything `odd' or `fancy'
-
-shows sound and complete wrt.
-
-Preconditions
-weakest phi sa <phi> pi <T> is still true
-
-Like hoare stuff
-
-inspired Dynamic assignment language \cite{eijck1992dynamic}
-Jan van Eijck 
-Fer-Jan de Vries
-
-like using a stack
-
-$\alpha \models \langle\psi\rangle\phi$ iff there is an assignment $\beta$ with $\alpha[\psi]\beta$, and $\beta \models \phi$
-
-Definition of true and false $\alpha\models\phi :\leftrightarrow \exists\beta(alpha\phi\beta)$.
-
-%{
-
-%Det her skal med for at vi kan forsvare ligningerne
-
-%α[⊥]β :⇔ 0 ≠ 0.
-%α[⊤]β :⇔ α = β.
-%α[P(x0, …, xn−1)]β :⇔ α = β and ⟨αx0, …, αxn−1 ⟩ ∈ I(P), 
-%where P is a predicate symbol of Σ with arity n.
-%α[∃v]β :⇔ α[v]β, where α[v]β iff αw = βw, for all variables w ≢ v.
-%α[φ · ψ]β :⇔ ∃γ  α[φ]γ[ψ]β.
-%α[~(φ)]β :⇔ α = β and ∀γ  ¬ α[φ]γ.
-%
-
-Det giver ikke mening at skrive $\not\models[\psi]$ da typen af $[\psi]$ er en relation. Eller hvad?
-
-\begin{align}
-\alpha[\psi_1\cdot\psi_2]\beta &:\leftrightarrow \exists\gamma (\alpha[\psi_1]\gamma \wedge \gamma[\psi_2]\beta) \label{sem_and} \\
-\alpha[\psi_1\cup\psi_2]\beta &:\leftrightarrow \alpha[\psi_1]\beta \vee \alpha[\psi_2]\beta \label{sem_or} \\
-\alpha[P(x_1,\dots,x_n)]\beta &:\leftrightarrow \alpha = \beta \wedge P(\alpha_{x_1},\dots,\alpha_{x_n}) \label{sem_pred} \\
-\alpha[\neg\psi]\beta &:\leftrightarrow \alpha = \beta \wedge \alpha\not\models[\psi] \label{sem_neg}\\
-\alpha[\exists x]\beta &:\leftrightarrow \forall\omega (\omega \neq x \rightarrow \alpha_{\omega} = \beta_{\omega}) \label{sem_exists} \\
-\intertext{Using these we can derive the following:}
-\alpha[\bot]\beta &\leftrightarrow \alpha = \beta \wedge \bot \nonumber\\
-                  &\leftrightarrow \bot \label{sem_bot}\\
-\alpha[\top]\beta &\leftrightarrow \alpha = \beta \wedge \top \nonumber\\
-                  &\leftrightarrow \alpha = \beta \label{sem_top}\\
+\section{Formal Stuff (10)}
+
+% This should be around 10 pages
+%Background should introduce any information that is necessary for the examiner to understand your project. 
+%The information here is generally more specific than the background given in the Introduction.
+%Requirements gives the program requirements. These chapters prepare the ground for the Design chapter.
+%That is: What do we need to handle to be able to solve this problem
+
+Since we are going to experiment with converting language into logic, a natural place to start is with classical logic.
+Classical logic, also called Predicate logic (PL), is the logic of conjunction, disjunction and negation extended with universal and existential quantifiers.
+
+In the introduction chapter we saw that PL was rather lacking in terms of compositionality and we predicted that what we really wanted was an `update logic' (a logic with update semantics) of some sort.
+Update semantics are well known and studied in computer language theory, and have in the past couple of decades found their way into a wider range of philosophies.
+
+The idea of all of those is a `growth of information in time'.
+Starting from little, and with each statement adding more.
+Information is here seen in the information theoretic way, where no information is when everything is possible ($\top$) and total information is when nothing is possible ($\bot$).
+
+Logicians and lingusts have come up with many different update logics since the 80s. I will quickly describe the four most cited ones, to give an idea of the landscape we're working in:
+%
+\begin{itemize}
+\item Discourse Representation Theory, 1981 by Hans Kamp\cite{kamp1981theory}, is perhaps the best known approach to using update semantics for understanding language.
+Kamp takes a very philosophical approach, thinking of his logic as a model for how a person builds up a context during a conversation.
+
+For example, say you you listen to `Some farmer owns no donkey.
+He beats it'.
+Discourse Representation Theory (DRT) builds this up from two blocks.
+Each consists of a set of referents and a set of properties using those.
+Underlined referents are referents that need to be `merged' with a properly `introduced' one:
+%
+\begin{align}
+&[_1 x\mid farmer(x), \neg[_2 y\mid donkey(y), owns(x,y)] ] \nonumber\\
+\oplus\ & [_1 \underline{v}, \underline{w}\mid beats(v,w) ] \nonumber
+\end{align}
+%
+The rule for `$\oplus$' is simply to do a disjoint union of the referents and the properties:
+%
+\begin{align}
+&[_1 x, \underline{v}, \underline{w}\mid farmer(x), \neg[_2 y\mid donkey(y), owns(x,y)], beats(v,w) ] \nonumber
+\end{align}
+%
+At this step it is possible to perform anaphora related heuristics. Since $x$ is the first paramter of $farmer$, we assume it should be so for $beats$ as well:
+%
+\begin{align}
+&[_1 x, \underline{v}, w\mid farmer(x), v=x, \neg[_2 y\mid donkey(y), owns(x,y)], beats(v,w) ] \nonumber\\
+&[_1 x, w\mid farmer(x), \neg[_2 y\mid donkey(y), owns(x,y)], beats(x,w) ]\nonumber
+\end{align}
+%
+On the other hand $y$ is not known in the scope of $beats$, so there is no way we can get rid of $w$. And we shouldn't, because in the sentence there is of course no way `it' can refer to `a donkey'.
+
+To deal with our example of `If a farmer owns...', DRT introduces more symbols for weak and strong conditionals.
+
+%In Discourse Representation Theory (DRT) our donkey sentence `If a farmer owns a donkey, he beats it' is represented as
+%\begin{align}
+%&[_1 x\mid Pedro(x), [_2 y, z\mid owns(y,z), donkey(z)] \Rightarrow [_3 v, w\mid beats(v,w)]] \nonumber
+%\intertext{This form is created from the syntax tree, and DRT then provides rules for eliminating variables and thus providing anaphora:}
+%&[_1 x, v: v = x, Pedro(x), [_2 y, z, w: y = x, w = y, owns(y,z), donkey(z)] \Rightarrow [_3 : beats(v,w)] ] \nonumber\\
+%&[_1 x: Pedro(x), [_2 z: donkey(z), owns(x,z)] \Rightarrow [_3 : beats(x,z)] ] \nonumber
+%\end{align}
+%The DRT merge
+
+\item File Change Semantics, 1982 by Irene Heim\cite{heim1983file}, are based around the puzzle of discourse `referents' which may, as we have seen, sometimes act like quantifiers, sometimes referents and sometimes anticedents.
+Heim suggests using so called `file cards' which though developed independently, has a lot in common with DRT.
+
+However, where DRT defines complex rules for merging different `discourse referents', in File Change Semantics (FCS) the part `added onto' a file gets to decide how the merge is made.
+Some how similar to how lambda functions in montague grammers give you a lot of flexibility.
+For example in case of `adding' $p$, an n-ary atomic predicate, to a file $F$:
+%
+\begin{align}
+&Sat(F + p) = \{a \in Sat(F) \mid R(a_{i_1}, \dots, a_{i_n})\} \nonumber
+\end{align}
+%
+Where $Sat$ is the function that takes a file to the set of `individuals' `satisfying' it.
+
+\item Mental Spaces, 1984 by Fauconnier\cite{fauconnier1984espaces}, is not really a logic in that it's not build on a model theoretic interpretation.
+It does however build on the same idea of `information growth' on a context representation.
+Mental Spaces (MS) represent the current context as a graph of predicates and variables:
+
+\tikzstyle{round} = [circle, draw, text width=3em, text centered]
+\begin{tikzpicture}[->, >=stealth', semithick, node distance = 3cm, auto]
+    % Place nodes
+    \node [round] (beats) {beats};
+    \node [round, below left of=beats] (x) {x};
+    \node [round, left of=x] (farmer) {farmer};
+    \node [round, below right of=beats] (z) {z};
+    \node [round, right of=z] (donkey) {donkey};
+    \node [round, below right of=x] (owns) {owns};
+    \node [round, below left of=owns] (y) {y};
+    % Draw edges
+    \path (beats) edge node {agt} (x);
+    \path (beats) edge node {obj} (z);
+    \path (z) edge node {dependent} (x);
+    \path (z) edge node {a} (donkey);
+    \path (x) edge node {every} (farmer);
+    \path (x) edge node {def-as} (y);
+    \path (owns) edge node {agt} (y);
+    \path (owns) edge node {obj} (z);
+\end{tikzpicture}
+
+Notice here in particular the interesting the `dependent' arrow.
+This arrow communicates that there is a one-many relationship between farmer and donkey, and hence that we are talking about a specific donkey for each farmer, not a single poor donkey beaten by everyone.
+
+\item Dynamic Predicate Logic, 1991 by Groenendijk and Stokhof\cite{groenendijk1991dynamic}, is what I will focus on in the remains of this paper.
+In contrast to some of the previous models, the authors of Dynamic Predicate Logic (DPL) don't make any big philosophical claims on connections between their model and the brain.
+Instead they simply create a logic that is isomorphic to PL, while keeping a reasonable ammount of `update semantics' like properties, such as being closer syntactically to natural language.
+
+DPL also doesn't provide any guidiance into resolving anaphora.
+Just like giving meaning to the actual concepts, this must be handled outside of the logic.
+We will see later in this paper that this stance makes sence, since the current state of the art anaphora resolvers are simply based on simple heuristics over noun phrases.
+
+In the end of this chapter we will see how a certain monadic construction by Visser can give us as much flexibility in DPL as we would get with a more involved system such as DRT or FCS.
+\end{itemize}
+
+\subsection{DPL}
+
+Dynamic Predicate Logic has been developed by a wide group of people: Jan van Eijck and Fer-Jan de Vries\cite{eijck1992dynamic}, Hans Kamp, Veltman, Groenendijk and Stokhof.
+It is the linguist version of Dynamic Logic invented by Pratt\cite{pratt1976semantical} which is again build on top of Floyd-Hoare Logic for reasoning about computer programs.
+
+The logic cleverly uses these advances in logical understanding to create an update logic in a pragmatic way, without any external constructs and a minimum of new syntax, while still getting all the benefits of compoitionality etc.
+
+To model a simple sentence `A man comes in. He sees a donkey. He smiles.', we write:
+%
+\begin{align}
+&(\exists x \cdot man(x) \cdot comes\_in(x)) \nonumber\\
+\cdot\ &(\exists y \cdot donkey(y) \cdot sees(x,y)) \nonumber\\
+\cdot\ &(smiles(x)) \nonumber
+\end{align}
+
+(The parenthesis is only for clarity.
+The logic is 100\% associative).
+Notice how we introduce referants casually as we progress in the discourse.
+We even manage to look pretty similar to PL at the same time.
+
+Too understand the semantics of the above, let's start by only considering a dynamic propositional logic with three propositions: $P$, $Q$ and $R$.
+We'll write $p$ to indicate a predicate is true and $\overline{p}$ if it's false.
+We talked about earlier, that the update semantics modeled `growth of information over time', so we'll start with case that every $2^3$ state is possible:
+%
+\begin{align}
+\{\overline{PQR}, \overline{PQ}R, \overline{P}Q\overline{R}, \overline{P}QR, P\overline{QR}, P\overline{Q}R, PQ\overline{R}, PQR\} \label{dpl_all_states}
+\end{align}
+
+Now we'll consider a statement in our predicate logic as a relation between possible states.
+For example $P\overline{Q}R[P\cdot Q]P\overline{Q}R$ is false, while $P\overline{Q}R[P\cup Q]PQR$ is true.
+If we take the image of \eqref{dpl_all_states} under $P\cdot\neg Q$ we get
+%
+\begin{align}
+\{P\overline{QR}, P\overline{Q}R\} \label{dpl_PnQ_states}
+\end{align}
+
+If we then take the image of \eqref{dpl_PnQ_states} under $\overline{P}$ we end up with the empty set of states.
+That is there is no possible assignment satisfying our formula, and we consider it a falsum.
+
+The most interesting question is how we work with quantifiers.
+Which on earth relation could $\exists S$ be?
+We are going to take a hint from computer programming and consider the variable decleration: $S:=\bot$.
+We want to write this in DPL as $\exists S \cdot \neg S$, such that e.g.
+$P\overline{Q}[\exists S \cdot \neg S]P\overline{QS}$ is true.
+Now since $\cdot\neg S$ acts as a filter, removing all assignments that doesn't have $\overline{S}$, it must be that $\exists S$ is a `random reset, introducing $S$ and allowing it every value, such that $P\overline{Q}[\exists S]P\overline{QS}$ and $P\overline{Q}[\exists S]P\overline{Q}S$ are both true.
+The image of \eqref{dpl_all_states} under $\exists S$ has 16 possible states.
+
+Let's have a look at the model theoretic definition of the semantics.
+Visser builds this up in two steps: First a relational algebra and then DPL on top of it.
+But since DPL is bascailly identical to the relational algebra, we are going to try and do it in one step.
+
+Let a signature $\Sigma$ for DPL be a structure $\langle Pred, Ar, Var\rangle$.
+$Pred$ is the set of predicate symbols containing at least $=$, $\top$ and $\bot$.
+$Ar$ is the function from predicate symbols to their arity and in particular we have $Ar(=) \mapsto 2$, $Ar(\top) \mapsto 0$, $Ar(\bot) \mapsto 0$.
+Finally $Var$ is the set of variable symbols we may use.
+We are not going to work with constants in this treatment.
+
+We are further going to define:
+%
+\begin{align}
+Prop_\Sigma &= \{P(x_1,\dots,x_n) \mid P \in Pred,\ Ar(P)=n,\ x_1,\dots,x_n\in Var\}\\
+Reset_\Sigma &= \{\exists x \mid x\in Var\}\\
+Atom_\Sigma &= Var \cup Prop_\Sigma \cup Reset_\Sigma
+\end{align}
+
+A model $\mathcal{M}$ of a signature $\Sigma$ is a tuple $\langle\mathcal{D},\mathcal{I}\rangle$ where $\mathcal{D}$ is a domain which we will always set as $\{\top,\bot\}$, and $\mathcal{I}$ is the interpretation function $Prop_\Sigma\to\mathcal{D}$.
+
+The set of formulas we will say to be the smallest set $Form_\Sigma$ satisfying:
+\begin{align}
+&Atom_\Sigma \subseteq Form_\Sigma\\
+\text{If } \psi\in Form_\Sigma \text{ then } &\neg(\psi)\in Form_\Sigma\\
+\text{If } \psi_1,\psi_2\in Form_\Sigma \text{ then } &\psi_1\cdot\psi_2\in Form_\Sigma\\
+\text{If } \psi_1,\psi_2\in Form_\Sigma \text{ then } &\psi_1\cup\psi_2\in Form_\Sigma
+\end{align}
+
+We are also going to say that an assignment can satisfy a formula, defined as:
+\begin{align}
+\alpha\models\psi :\equiv \exists\beta(\alpha[\psi]\beta)
+\end{align}
+
+Let $[\cdot]$ be the interpretation function from syntax to relation. Then we define the semantics of our logic pointwise for all assignments $\alpha$ and $\beta$:
+\begin{align}
+\alpha[\psi_1\cdot\psi_2]\beta :\equiv&\ \exists\gamma (\alpha[\psi_1]\gamma \wedge \gamma[\psi_2]\beta) \label{sem_and} \\
+\alpha[\psi_1\cup\psi_2]\beta :\equiv&\ \alpha[\psi_1]\beta \vee \alpha[\psi_2]\beta \label{sem_union} \\
+\alpha[P(x_1,\dots,x_n)]\beta :\equiv&\ \alpha = \beta \wedge \mathcal{I}(P(\alpha_{x_1},\dots,\alpha_{x_n})) \label{sem_pred} \\
+\alpha[\neg\psi]\beta :\equiv&\ \alpha = \beta \wedge \alpha\not\models[\psi] \label{sem_neg}\\
+\alpha[\exists x]\beta :\equiv&\ \forall\omega (\omega \neq x \rightarrow \alpha_{\omega} = \beta_{\omega}) \label{sem_exists} \\
+\intertext{Using those basic axioms we can derive for our built in propositions:}
+\alpha[\bot]\beta \leftrightarrow&\ \alpha = \beta \wedge \bot \nonumber\\
+                  \leftrightarrow&\ \bot \label{sem_bot}\\
+\alpha[\top]\beta \leftrightarrow&\ \alpha = \beta \wedge \top \nonumber\\
+                  \leftrightarrow&\ \alpha = \beta \label{sem_top}\\
 \intertext{Taking $\psi_1\rightarrow\psi_2$ to mean $\neg(\psi_1\wedge\neg\psi_2)$ we also get the following neat equation, originally by Groenendijk and Stokhof\cite{groenendijk1991dynamic}:}
 \alpha[\psi_1\rightarrow\psi_2]\beta
- &:\leftrightarrow \alpha[\neg(\psi_1\wedge\neg\psi_2)]\beta \nonumber\\
- &\leftrightarrow \alpha = \beta \wedge \alpha\not\models[\psi_1\wedge\neg\psi_2] \nonumber\\
- &\leftrightarrow \alpha = \beta \wedge \neg\exists\gamma(\alpha[\psi_1\wedge\neg\psi_2]\gamma) \nonumber\\
- &\leftrightarrow \alpha = \beta \wedge \neg\exists\gamma(\exists\delta(\alpha[\psi_1]\delta\wedge\delta[\neg\psi_2]\gamma)) \nonumber\\
- &\leftrightarrow \alpha = \beta \wedge \neg\exists\gamma(\exists\delta(\alpha[\psi_1]\delta\wedge\delta=\gamma\wedge\delta\not\models[\psi_2])) \nonumber\\
- &\leftrightarrow \alpha = \beta \wedge \forall\gamma(\forall\delta(\delta=\gamma\rightarrow(\alpha[\psi_1]\delta\rightarrow\delta\models[\psi_2]))) \nonumber\\
- &\leftrightarrow \alpha = \beta \wedge \forall\gamma(\alpha[\psi_1]\gamma\rightarrow\gamma\models[\psi_2]) \label{sem_impl}
-\end{align}
-
-Hence we can read implication as a the filter, that if you can get `through' $\psi_1$ then you must also be able to get through $\psi_2$.
-Also notice that $\gamma$ doesn't have to equal $\alpha$.
-Hence variables introduced in $\psi_1$ can bind variables in $\psi_2$.
+ :\equiv&\ \alpha[\neg(\psi_1\wedge\neg\psi_2)]\beta \nonumber\\
+ \leftrightarrow&\ \alpha = \beta \wedge \alpha\not\models[\psi_1\wedge\neg\psi_2] \nonumber\\
+ \leftrightarrow&\ \alpha = \beta \wedge \neg\exists\gamma(\alpha[\psi_1\wedge\neg\psi_2]\gamma) \nonumber\\
+ \leftrightarrow&\ \alpha = \beta \wedge \neg\exists\gamma(\exists\delta(\alpha[\psi_1]\delta\wedge\delta[\neg\psi_2]\gamma)) \nonumber\\
+ \leftrightarrow&\ \alpha = \beta \wedge \neg\exists\gamma(\exists\delta(\alpha[\psi_1]\delta\wedge\delta=\gamma\wedge\delta\not\models[\psi_2])) \nonumber\\
+ \leftrightarrow&\ \alpha = \beta \wedge \forall\gamma(\forall\delta(\delta=\gamma\rightarrow(\alpha[\psi_1]\delta\rightarrow\delta\models[\psi_2]))) \nonumber\\
+ \leftrightarrow&\ \alpha = \beta \wedge \forall\gamma(\alpha[\psi_1]\gamma\rightarrow\gamma\models[\psi_2]) \label{sem_impl}
+\end{align}
+
+In my implementation section, I describe how the above can be efficiently implemented in Haskell using an isomorphism between relations $\mathcal{P}(A\times B)$ and functions $A\to\mathcal{P}(B)$.
+I also show how the kleiski combinator takes the role of $\cdot$.
+
+However let's first look at some of the implications of the above definitions.
+For example, how can we intuitively understand the DPL notion of $\rightarrow$?
+Well, first notice that many of the eqations start with $\alpha=\beta\wedge\dots$.
+These equations we can think of intuitively as having a closed scope.
+We saw earlier how DRT nicely determined that the `it' in `A farmer doesn't own a donkey.
+He beats it' couldn't refer to `a donkey' because the negation didn't let it's variable out.
+In DPL we get the same effect.
+Because an assignment must be the same before and after a negated clause, all new variables introduced inside the clause have disappeared.
+
+The DPL $\psi_1\rightarrow\psi_2$ in particular can be seen as the filter, that if you can get `through' $\psi_1$ then you must also be able to get through $\psi_2$.
+But notice that $\gamma$ doesn't have to equal $\alpha$.
+Hence variables introduced in $\psi_1$ can well bind variables in $\psi_2$.
 This will show to be extreamly important in the treatment of donkey sentences later.
@@
 This is going to introduce a lot of weird situations later.
-An alternative definition of `or': $\neg(\neg\psi_1\cdot\neg\psi_2)$ will also be considered.
-
-For \eqref{sem_pred} $P$ is a predicate symbol of $\Sigma$ with arity n.
-
-static negation. DOesn't allow new things. LIke a filter
-
-some people have a closed/static or: $\neg(\neg\psi_1\wedge\neg\psi_2)$ \cite{groenendijk1991dynamic}
-
-%Notice that \eqref{sem_bot} and \eqref{sem_top} are special cases of \eqref{sem_pred}.
-
-We are going to introduce formuals that allow us to recursively convert any DRA formula into PL.
-
-To smoothen the following proofs, we will introduce a special syntax `$\langle\psi\rangle\phi$' where $\psi$ is a DRA formula, $\phi$ is a PL formula and the whole thing is a PL formula.
-For some assignment $\alpha$, we define `$\alpha\models\langle\psi\rangle\phi$' to mean $\exists\beta(\alpha[\psi]\beta \wedge \beta\models\phi)$.
-
-\begin{align}
-\alpha\models\langle\exists x\rangle\phi
- & \leftrightarrow \exists\beta (\alpha[\exists x]\beta \wedge \beta\models\phi) \nonumber\\
- & \leftrightarrow \exists\beta (\forall\omega (\omega \neq x \rightarrow \alpha_{\omega} = \beta_{\omega}) \wedge \beta\models\phi) \nonumber\\
- & \leftrightarrow \exists\beta (\exists\nu (\forall\omega (\alpha[x:=\nu]_{\omega} = \beta_{\omega})) \wedge \beta\models\phi) \nonumber\\
- & \leftrightarrow \exists\beta (\exists\nu (\alpha[x:=\nu] = \beta) \wedge \beta\models\phi) \nonumber\\
-% & \leftrightarrow \exists\nu (\exists\beta (\alpha[x:=\nu] = \beta \wedge \beta\models\phi)) \nonumber\\
- & \leftrightarrow \exists\nu (\alpha[x:=\nu]\models\phi) \nonumber\\
- & \leftrightarrow \alpha\models\exists x (\phi) \label{conv_exists}\\\nonumber\\
-%
-\alpha\models\langle\psi_1\cdot\psi_2\rangle\phi
- & \leftrightarrow \exists\beta (\alpha[\psi_1\cdot\psi_2]\beta \wedge \beta\models\phi) \nonumber\\
- & \leftrightarrow \exists\beta(\exists\gamma (\alpha[\psi_1]\gamma \wedge \gamma[\psi_2]\beta) \wedge \beta\models\phi) \nonumber\\
- & \leftrightarrow \exists\gamma (\alpha[\psi_1]\gamma \wedge \exists\beta(\gamma[\psi_2]\beta \wedge \beta\models\phi)) \nonumber\\
- & \leftrightarrow \exists\gamma (\alpha[\psi_1]\gamma \wedge \gamma\models\langle\psi_2\rangle\phi) \nonumber\\
- & \leftrightarrow \alpha\models\langle\psi_1\rangle\langle\psi_2\rangle\phi \label{conv_and}\\\nonumber\\
-%
-\alpha\models\langle\psi_1 \cup \psi_2\rangle\phi
- & \leftrightarrow \exists\beta (\alpha[\psi_1 \cup \psi_2]\beta \wedge \beta\models\phi) \nonumber\\
- & \leftrightarrow \exists\beta ((\alpha[\psi_1]\beta \vee \alpha[\psi_2]\beta) \wedge \beta\models\phi) \nonumber\\
- & \leftrightarrow \exists\beta (\alpha[\psi_1]\beta \wedge \beta\models\phi) \vee \exists\beta (\alpha[\psi_2]\beta \wedge \beta\models\phi) \nonumber\\
- & \leftrightarrow (\alpha\models\langle\psi_1\rangle\phi) \vee (\alpha\models\langle\psi_2\rangle\phi) \nonumber\\
- & \leftrightarrow \alpha\models\langle\psi_1\rangle\phi \vee \langle\psi_2\rangle\phi \label{conv_or}\\\nonumber\\
-%
-\alpha\models\langle P(x_1,\dots,x_n)\rangle\phi
- & \leftrightarrow \exists\beta (\alpha[P(x_1,\dots,x_n)]\beta \wedge \beta\models\phi) \nonumber\\
- & \leftrightarrow \exists\beta (\alpha = \beta \wedge P(\alpha_{x_1},\dots,\alpha_{x_n}) \wedge \beta\models\phi) \nonumber\\
- & \leftrightarrow P(\alpha_{x_1},\dots,\alpha_{x_n}) \wedge \alpha\models\phi \nonumber\\
- & \leftrightarrow \alpha\models P(x_1,\dots,x_n) \wedge \phi \label{conv_pred}\\\nonumber\\
-%
-\alpha\models\langle\neg\psi\rangle\phi
- & \leftrightarrow \exists\beta (\alpha[\neg\psi]\beta \wedge \beta\models\phi) \nonumber\\
- & \leftrightarrow \exists\beta (\alpha=\beta \wedge \neg\exists\gamma(\alpha[\psi]\gamma) \wedge \beta\models\phi) \nonumber\\
- & \leftrightarrow \neg\exists\gamma(\alpha[\psi]\gamma) \wedge \alpha\models\phi \nonumber\\
- & \leftrightarrow \neg\exists\gamma(\alpha[\psi]\gamma \wedge \gamma\models\top) \wedge \alpha\models\phi \nonumber\\
- & \leftrightarrow \alpha\not\models\langle\psi\rangle\top \wedge \alpha\models\phi \nonumber\\
- & \leftrightarrow \alpha\models \neg\langle\psi\rangle\top \wedge \phi \label{conv_neg}
-\end{align}
-
-Notice that $\alpha\models\langle\bot\rangle\phi$ and $\alpha\models\langle\top\rangle\phi$ follow trivially from \eqref{conv_pred}.
-
-\begin{haskell}
-addDpl :: Prop -> [DPL] -> Prop
-addDpl p [] = p
-addDpl p (DF:ds) = F
-addDpl p (DT:ds) = addDpl p ds
-addDpl p ((DPredi f ks):ds) = addDpl (And (Predi f ks) p) ds
-addDpl p ((DExists k):ds) = addDpl (Exist k p) ds
-addDpl p ((DComp d1 d2):ds) = addDpl p (d2:d1:ds) -- backwards
-addDpl p ((DNot d):ds) = addDpl (And (Not (addDpl T [d])) p) ds
-addDpl p ((DImpl d1 d2):ds) = addDpl p ((DNot (DComp d1 (DNot d2))):ds)
-addDpl p ((DOr d1 d2):ds) = addDpl (And (Or (addDpl T [d1]) (addDpl T [d2])) p) ds
-\end{haskell}
+In the next section we will look closer at these problems, and also consider the alternative definition of `or': $\neg(\neg\psi_1\cdot\neg\psi_2)$.
 
@@
 
+Having defined seamingly very sound semantics for DPL above, we may start to notice certain oddities about our choice of using the relation union for `or'.
+An obvious alternative would have been the closed/static $\neg(\neg\psi_1\wedge\neg\psi_2)$ or $\neg\psi_1\rightarrow\psi_2$ considered by Groenendijk\cite{groenendijk1991dynamic}.
+
+Actually there is an obvious definition of `or' that we have totally ignored in the previous chapter:
+\begin{align}
+\alpha[\psi_1\vee\psi_2]\beta
+ :\equiv&\ \alpha[\neg(\neg\psi_1\wedge\neg\psi_2)]\beta \nonumber\\
+ \leftrightarrow&\ \alpha[\neg\psi_1\rightarrow\psi_2)]\beta \nonumber\\
+ \leftrightarrow&\ \alpha = \beta \wedge \forall\gamma(\alpha[\neg\psi_1]\gamma\rightarrow\gamma\models[\psi_2]) \nonumber\\
+ \leftrightarrow&\ \alpha = \beta \wedge \forall\gamma((\alpha=\gamma\wedge\gamma\not\models\psi_1)\rightarrow\gamma\models[\psi_2]) \nonumber\\
+ \leftrightarrow&\ \alpha = \beta \wedge \forall\gamma(\alpha=\gamma\rightarrow(\gamma\models\psi_1\vee\gamma\models\psi_2)) \nonumber\\
+ \leftrightarrow&\ \alpha = \beta \wedge (\alpha\models\psi_1\vee\alpha\models\psi_2) \label{sem_or}
+\end{align}
+
+So does this fix our problem of binding between clauses? No. In fact this definition doesn't even allow binding from a clause and out.
+
 a farmer doesn't own a horse or a donkey
-
-Ex.farmer(x) . ~(Ey.horse(y).owns(x,y) v Ey.donkey(y).owns(x,y))
-Ex.farmer(x) . ~Ey.horse(y).owns(x,y) . ~Ey.donkey(y).owns(x,y)
+\begin{align}
+\exists x\cdot farmer(x) \cdot \neg(\exists y\cdot horse(y)\cdot owns(x,y) v \exists y\cdot donkey(y)\cdot owns(x,y))\nonumber\\
+\exists x\cdot farmer(x) \cdot \neg\exists y\cdot horse(y)\cdot owns(x,y) \cdot \neg \exists y\cdot donkey(y)\cdot owns(x,y)\nonumber
+\end{align}
 
 a farmer beats a horse or a donkey
-
-Ex.farmer(x).(Ey.horse(y) v Ez.donkey(z)).beats(x,y) = {Nothing, Ex.farmer(x).EyJust .horse(y).beats(x,y)}
-
-Ex.farmer(x).~((Ey.horse(y) v Ez.donkey(z)).beats(x,y)) =
-Ex.farmer(x).(~Ey.horse(y).~Ez.donkey(z) v ~beats(x,y)) = {Nothing, Ex.farmer(x).~Ey.horse(y).~Ez.donkey(z)}
-
-Ex.farmer(x).~(tall(x) v blue(x)) = Ex.famer(x).~tall(x).~blue(x)
+\begin{align}
+&Ex.farmer(x).(Ey.horse(y) v Ez.donkey(z)).beats(x,y) = \\
+&\{Nothing, Ex.farmer(x).EyJust .horse(y).beats(x,y)\}\\
+\\
+&Ex.farmer(x).~((Ey.horse(y) v Ez.donkey(z)).beats(x,y)) = \\
+&Ex.farmer(x).(~Ey.horse(y).~Ez.donkey(z) v ~beats(x,y)) = \\
+&\{Nothing, Ex.farmer(x).&~Ey.horse(y).~Ez.donkey(z)\}\\
+\\
+&Ex.farmer(x).~(tall(x) v blue(x)) = Ex.famer(x).~tall(x).~blue(x)
+\end{align}
 
 There is a farmer who maybe has a donkey. He beats it.
-Ex.farmer(x).(Ey.donkey(y) v false).beats(x,y)
+$Ex.farmer(x).(Ey.donkey(y) v false).beats(x,y)$
 With union semantics for v this would be equal to
-Ex.farmer(x).Ey.donkey(y).beats(x,y)
+$Ex.farmer(x).Ey.donkey(y).beats(x,y)$
 which is stupid.
 Intead with sets of relations, this would be equal to
-Ex.farmer(x).Ey.donkey(y).beats(x,y) v Ex.farmer(x).false.beats(x,y)
+$Ex.farmer(x).Ey.donkey(y).beats(x,y) v Ex.farmer(x).false.beats(x,y)$
 the right side being an error? no
@@
 If a farmer is not poor he owns a donkey. He beats it.
-Ex.farmer(x).~poor(x) -> Ey.donkey(y).owns(x,y).beats(x,y)
+$Ex.farmer(x).~poor(x) -> Ey.donkey(y).owns(x,y).beats(x,y)$
 
 A farmer is either poor or he owns a donkey. He beats it.
-Ex.farmer(x).(poor(x) v Ey.donkey(y).owns(x,y)).beats(x,y) =
-
-(Ex v false) . f(x)
-(exist x `union` false) [] = [[("x",0)],[("x",1)],[("x",2)]
-[Nothing, Just [[("x",0)],[("x",1)],[("x",2)]]]
-
-(Ex v false) = [Just [[("x",0)],[("x",1)],[("x",2)], Just []]
-not (Ex v false) =
-
-not (Ex v f(x))
-= ~Ex.~f(x)
+$Ex.farmer(x).(poor(x) v Ey.donkey(y).owns(x,y)).beats(x,y) =$
+
+%(Ex v false) . f(x)
+%(exist x `union` false) [] = [[("x",0)],[("x",1)],[("x",2)]
+%[Nothing, Just [[("x",0)],[("x",1)],[("x",2)]]]
+
+%(Ex v false) = [Just [[("x",0)],[("x",1)],[("x",2)], Just []]
+%not (Ex v false) =
+
+%not (Ex v f(x))
+%= ~Ex.~f(x)
 
 A farmer owns a horse or a donkey
-Ex.farmer(x).~(~(Ey.horse(y)).~(Ey.donkey(y))).owns(x,y)   this surely doesn't work
+$Ex.farmer(x).~(~(Ey.horse(y)).~(Ey.donkey(y))).owns(x,y)$   this surely doesn't work
 
@@
 
+hvad hvis vi puttede vissers disjoint union stuff into the monad?
+
+\subsection{Connection with first order logic}
+
+An interesting property of DPL is that it is isomorphic with classical logic.
+Any PL formula corresponds to one in DPL and vice versa.
+This gives an alternative way to define and understand the semantics of DPL, and it relaxes us that nothing `weird' is going on.
+We might also be a bit disappointed that after all this, we don't possess any extra power, but let's forget about that for now, and see how the conversion works.
+
+Let $\phi$ be a formula in negation normal form, or something. Hmm, come back here once the disjunction chapter is done.
+\begin{align}
+&[\exists x(\phi)] \equiv \neg(\neg(\exists x \cdot [\phi]))\\
+&[\forall x(\phi)] \equiv \exists x \rightarrow [\phi]\\
+&[\phi_1\wedge\phi_2] \equiv [\phi_1]\cdot[\phi_2]\\
+&[\phi_1\vee\phi_2] \equiv \dots\\
+&[\neg(\exists x)] \equiv \dots\\
+&[\neg(\phi_1\wedge\phi_2)] \equiv \neg[\phi_1]\cdot\neg[\phi_2]\\
+&[\neg(\phi_1\vee\phi_2)] \equiv \dots
+\end{align}
+
+Similarly we may create PL formulas with the same meaning as any DPL formula.
+But perhaps we should be a bit conserned on how to interpret the meaning of a PL formula and a DPL formula being equivalent.
+After all DPL formulas are relations where PL formulas a propositions.
+In the above it was all easy, but let's think again about what we are doing.
+
+Remember that we said an assignment could model a DPL relation.
+If we consider the truth value of a DPL relation to be whether there exists any relation that models it $\exists\alpha(\alpha\models\psi)$, we can use a trick.
+The trick is to think of the weakest precondion such an assignment must satisfy, and model that in PL.
+In fact the precondition is not just for our assignment, but for the entire model.
+For example to model $\exists x\cdot farmer(x)$ the starting assignment might well be empty ($\bot$), but our model must contain an entity for which $farmer$ is true.
+
+To smoothen the following derivations, we will introduce a special syntax `$\langle\psi\rangle\phi$' where $\psi$ is a DRA formula, $\phi$ is a PL formula and the whole thing is a PL formula.
+For some assignment $\alpha$, we define `$\alpha\models\langle\psi\rangle\phi$' to mean $\exists\beta(\alpha[\psi]\beta \wedge \beta\models\phi)$.
+%
+\begin{align}
+\alpha\models\langle\exists x\rangle\phi
+ & \leftrightarrow \alpha\models\exists x (\phi) \label{conv_exists}\\
+\alpha\models\langle\psi_1\cdot\psi_2\rangle\phi
+ & \leftrightarrow \alpha\models\langle\psi_1\rangle\langle\psi_2\rangle\phi \label{conv_and}\\
+\alpha\models\langle\psi_1 \cup \psi_2\rangle\phi
+ & \leftrightarrow \alpha\models\langle\psi_1\rangle\phi \vee \langle\psi_2\rangle\phi \label{conv_or}\\
+\alpha\models\langle P(x_1,\dots,x_n)\rangle\phi
+ & \leftrightarrow \alpha\models P(x_1,\dots,x_n) \wedge \phi \label{conv_pred}\\
+\alpha\models\langle\neg\psi\rangle\phi
+ & \leftrightarrow \alpha\models \neg\langle\psi\rangle\top \wedge \phi \label{conv_neg}
+\end{align}
+
+Notice that $\alpha\models\langle\bot\rangle\phi\leftrightarrow a\models\bot$ and $\alpha\models\langle\top\rangle\phi\leftrightarrow a\models\phi$ follow trivially from \eqref{conv_pred}.
+For my full derivation of these equations, see the appendix.
+
+In the implementation we will use the above equations to recursively convert our DPL formulas to PL. The $\langle\rangle$ notation easily lends itself to calculation by a stak manipulation function.
+
+The rules will look like:
+\begin{haskell}
+addDpl :: Prop -> [DPL] -> Prop
+addDpl p ((DNot d):ds) = addDpl (And (Not (addDpl T [d])) p) ds
+\end{haskell}
+
 \subsubsection{Weird disjunctive subjects}
@@
 %- ∃y(book(y) ∧ ∀x(student(x) ⇒ read(y)(x)))
+
+Every farmer owns a donkey. It is pink. Second sentence decides the reading.
 
@@
 
-\section{Implementation}
-8pagish
-This subsubsection's content...
+\section{Implementation (8)}
 
@@
 
-\section{Perspectives/Conclusions}
-2pagish
+\section{Perspectives/Conclusions (2)}
 
@@
 
-\section*{Acknowledgements}
+\section{Acknowledgements}
 Thanks to Samson
@@
 
+\newpage
+\section{Appendix}
+\subsection{Derivation of DPL preconditions}
+\begin{align}
+\alpha\models\langle\exists x\rangle\phi
+ & \leftrightarrow \exists\beta (\alpha[\exists x]\beta \wedge \beta\models\phi) \nonumber\\
+ & \leftrightarrow \exists\beta (\forall\omega (\omega \neq x \rightarrow \alpha_{\omega} = \beta_{\omega}) \wedge \beta\models\phi) \nonumber\\
+ & \leftrightarrow \exists\beta (\exists\nu (\forall\omega (\alpha[x:=\nu]_{\omega} = \beta_{\omega})) \wedge \beta\models\phi) \nonumber\\
+ & \leftrightarrow \exists\beta (\exists\nu (\alpha[x:=\nu] = \beta) \wedge \beta\models\phi) \nonumber\\
+ & \leftrightarrow \exists\nu (\exists\beta (\alpha[x:=\nu] = \beta \wedge \beta\models\phi)) \nonumber\\
+ & \leftrightarrow \exists\nu (\alpha[x:=\nu]\models\phi) \nonumber\\
+ & \leftrightarrow \alpha\models\exists x (\phi) \\
+%
+\intertext{Here we take $\alpha[x:=\nu]$ to mean the $\alpha$ with $\alpha_x$ assigned to the value of $\nu$.}
+%
+\alpha\models\langle\psi_1\cdot\psi_2\rangle\phi
+ & \leftrightarrow \exists\beta (\alpha[\psi_1\cdot\psi_2]\beta \wedge \beta\models\phi) \nonumber\\
+ & \leftrightarrow \exists\beta(\exists\gamma (\alpha[\psi_1]\gamma \wedge \gamma[\psi_2]\beta) \wedge \beta\models\phi) \nonumber\\
+ & \leftrightarrow \exists\gamma (\alpha[\psi_1]\gamma \wedge \exists\beta(\gamma[\psi_2]\beta \wedge \beta\models\phi)) \nonumber\\
+ & \leftrightarrow \exists\gamma (\alpha[\psi_1]\gamma \wedge \gamma\models\langle\psi_2\rangle\phi) \nonumber\\
+ & \leftrightarrow \alpha\models\langle\psi_1\rangle\langle\psi_2\rangle\phi
+%
+\intertext{c}
+%
+\alpha\models\langle\psi_1 \cup \psi_2\rangle\phi
+ & \leftrightarrow \exists\beta (\alpha[\psi_1 \cup \psi_2]\beta \wedge \beta\models\phi) \nonumber\\
+ & \leftrightarrow \exists\beta ((\alpha[\psi_1]\beta \vee \alpha[\psi_2]\beta) \wedge \beta\models\phi) \nonumber\\
+ & \leftrightarrow \exists\beta (\alpha[\psi_1]\beta \wedge \beta\models\phi) \vee \exists\beta (\alpha[\psi_2]\beta \wedge \beta\models\phi) \nonumber\\
+ & \leftrightarrow (\alpha\models\langle\psi_1\rangle\phi) \vee (\alpha\models\langle\psi_2\rangle\phi) \nonumber\\
+ & \leftrightarrow \alpha\models\langle\psi_1\rangle\phi \vee \langle\psi_2\rangle\phi
+%
+\intertext{d}
+%
+\alpha\models\langle P(x_1,\dots,x_n)\rangle\phi
+ & \leftrightarrow \exists\beta (\alpha[P(x_1,\dots,x_n)]\beta \wedge \beta\models\phi) \nonumber\\
+ & \leftrightarrow \exists\beta (\alpha = \beta \wedge P(\alpha_{x_1},\dots,\alpha_{x_n}) \wedge \beta\models\phi) \nonumber\\
+ & \leftrightarrow P(\alpha_{x_1},\dots,\alpha_{x_n}) \wedge \alpha\models\phi \nonumber\\
+ & \leftrightarrow \alpha\models P(x_1,\dots,x_n) \wedge \phi
+%
+\intertext{f}
+%
+\alpha\models\langle\neg\psi\rangle\phi
+ & \leftrightarrow \exists\beta (\alpha[\neg\psi]\beta \wedge \beta\models\phi) \nonumber\\
+ & \leftrightarrow \exists\beta (\alpha=\beta \wedge \neg\exists\gamma(\alpha[\psi]\gamma) \wedge \beta\models\phi) \nonumber\\
+ & \leftrightarrow \neg\exists\gamma(\alpha[\psi]\gamma) \wedge \alpha\models\phi \nonumber\\
+ & \leftrightarrow \neg\exists\gamma(\alpha[\psi]\gamma \wedge \gamma\models\top) \wedge \alpha\models\phi \nonumber\\
+ & \leftrightarrow \alpha\not\models\langle\psi\rangle\top \wedge \alpha\models\phi \nonumber\\
+ & \leftrightarrow \alpha\models \neg\langle\psi\rangle\top \wedge \phi
+%
+\intertext{e}\nonumber
+\end{align}
+
 \bibliographystyle{plain}

### 2013-04-25 Main.hs
(Some what) Working montague interpretor!

@@ -1,7 +1,7 @@
 module Main (main) where
 
 import Stanford (run, runDry)
-import Prop (Prop)
+import Prop (Prop, tostring)
 import ToLogic (logify)
 
 import Data.List
@@ -36,7 +36,7 @@ main = do
 			"A man comes in. He sees a dog. He smiles." -- very sequential
 			]
 		putStrLn "Test2"
-		sequence (map print props)
+		sequence (map (print.tostring) props)
 		return ()
 
 {-

### 2013-04-25 Stanford.hs

@@ -1,23 +1,27 @@
-module Stanford (DepTree(..), Word, Sentence, run, runDry, runOnFile, PosTree(..), toMapping, postreeS) where
+module Stanford (Store, Word, Index, run, runDry, runOnFile, IndexTree(..), postrees) where
 
 import Prop (Ref)
 
 import Text.ParserCombinators.Parsec
 import Data.Map hiding (map, (\\))
 
-import Data.List hiding (union, insert)
+import Data.List hiding (union, insert, lookup)
 import Text.XML.Light
 import Control.Monad
 import Data.Maybe
 import System.Process
 
+type Index = (Int, Int)
 type Word = String
 type Sentence = [Word]
 
 type PosTag = String
-data PosTree a = Phrase PosTag [PosTree a] | Leaf PosTag a deriving (Show)
+data PosTree = Phrase PosTag [PosTree] | Leaf PosTag Word deriving (Show)
+data IndexTree = P PosTag ([IndexTree],Index) | L PosTag (Word,Index) deriving (Show)
 
 data DepTree = Dep Int [(Ref, DepTree)] deriving (Show, Eq)
+
+type Store = Index -> Ref
 
 ------------------
 
@@ -31,13 +35,16 @@ lemmas doc = map strContent (findElements (unqual "lemma") doc)
 
 ------------------
 
-postree :: Element -> PosTree Word
-postree doc = postreeS dat
+postrees :: Element -> [PosTree]
+postrees doc = map rootfilter roots
 	where
-		dat = (strContent . fromJust) (findElement (unqual "parse") doc)
+		parses = findElements (unqual "parse") doc
+		roots = map (postreeRoot . strContent) parses
+		rootfilter (Phrase "ROOT" [(Phrase "S" s)]) = (Phrase "S" s)
+		rootfilter root = error ("Cant parse strange sentence " ++ show root)
 
-postreeS :: String -> PosTree Word
-postreeS dat = tree
+postreeRoot :: String -> PosTree
+postreeRoot dat = tree
 	where
 		Right tree = parse posEither "(unknown)" dat
 
@@ -60,25 +67,20 @@ posLeaf =
 		word <- many (noneOf " ()")
 		return (Leaf tag word)
 
-toMapping :: [Ref] -> PosTree Word -> ([Ref], Map Int Ref)
-toMapping vs (Leaf _ _) = (vs, empty)
[…]

### 2013-04-25 ToLogic.hs

@@ -2,127 +2,67 @@ module ToLogic (logify) where
 
 import Dpl (dpl2prop, s2dpl, Type2, sT, sF, sSwitch, sScope, sPredi, sExist, sComp, sTest0, sTest1)
 import Prop
-import Stanford (DepTree(..), Word, Sentence, PosTree(..))
+import Stanford (Word, Store, Index, IndexTree(..))
 
 import Data.Map hiding (lookup, mapMaybe)
 import Data.Maybe
 
-parseVP :: Sentence -> Map Int Ref -> DepTree -> Type2
-parseVP ws refs (Dep n ds) = foldl sComp sT parts `sComp` predi
-	where
-		parts = mapMaybe (\(name, dep) -> (case name of
-			 "nsubj" -> Just (parseNP ws refs dep)
-			 "dobj"  -> Just (parseNP ws refs dep)
-			 "advcl" -> Just (parseAdvcl ws refs dep)
-			 otherwise -> Nothing)) ds
-		Just (Dep sub _) = lookup "nsubj" ds
-		Just (Dep obj _) = lookup "dobj" ds
-		predi = sPredi (ws !! n) [refs ! sub, refs ! obj]
+logify :: (Store, [IndexTree]) -> Prop
+logify (refs, tree) = (simplify . dpl2prop . s2dpl) logic
+	where logic = intRoot tree refs
 
-parseAdvcl :: Sentence -> Map Int Ref -> DepTree -> Type2
-parseAdvcl ws refs dep = sSwitch `sComp` (parseVP ws refs dep) `sComp` sSwitch
+type SyntaxB = ([IndexTree], Index)
+type SyntaxL = (Word, Index)
+type Predicate = Ref -> Type2
+type Predicate2 = Ref -> Ref -> Type2
 
--- Check if we are the primary mention, if true then add an exists and a test
-parseNP :: Sentence -> Map Int Ref -> DepTree -> Type2
-parseNP ws refs (Dep n ds) = case isJust (lookup "det" ds) of
-	 True -> sExist var `sComp` sPredi (ws !! n) [var]
-	 False -> sT
-	where var = refs ! n
+intRoot :: [IndexTree] -> Store -> Type2
+intRoot ((P "S" s):ss) r = intS s r `sComp` intRoot ss r
+intRoot [] r = sT
 
[…]

### 2013-04-25 Stanford.hs
Simplified indextree

@@ -17,7 +17,7 @@ type Sentence = [Word]
 
 type PosTag = String
 data PosTree = Phrase PosTag [PosTree] | Leaf PosTag Word deriving (Show)
-data IndexTree = P PosTag ([IndexTree],Index) | L PosTag (Word,Index) deriving (Show)
+data IndexTree = P PosTag [IndexTree] | L PosTag (Word,Index) deriving (Show)
 
 data DepTree = Dep Int [(Ref, DepTree)] deriving (Show, Eq)
 
@@ -76,10 +76,10 @@ toIndexTrees trees = map toIndexTree (zip trees [0..])
 toIndexTree :: (PosTree, Int) -> IndexTree
 toIndexTree (tree, sentence) = snd (toIndexTree_ 0 tree)
 	where
-		toIndexTree_ i (Phrase tag (t:ts)) = (k, P tag (it:its,(10,0)))
+		toIndexTree_ i (Phrase tag (t:ts)) = (k, P tag (it:its))
 			where (j, it) = toIndexTree_ i t
-			      (k, P _ (its,_)) = toIndexTree_ j (Phrase tag ts)
-		toIndexTree_ i (Phrase tag []) = (i, P tag ([],(10,0)))
+			      (k, P _ its) = toIndexTree_ j (Phrase tag ts)
+		toIndexTree_ i (Phrase tag []) = (i, P tag [])
 		toIndexTree_ i (Leaf tag word) = (i+1, L tag (word,(sentence,i)))
 
 ------------------

### 2013-04-25 ToLogic.hs

@@ -11,7 +11,7 @@ logify :: (Store, [IndexTree]) -> Prop
 logify (refs, tree) = (simplify . dpl2prop . s2dpl) logic
 	where logic = intRoot tree refs
 
-type SyntaxB = ([IndexTree], Index)
+type SyntaxB = [IndexTree]
 type SyntaxL = (Word, Index)
 type Predicate = Ref -> Type2
 type Predicate2 = Ref -> Ref -> Type2
@@ -21,23 +21,23 @@ intRoot ((P "S" s):ss) r = intS s r `sComp` intRoot ss r
 intRoot [] r = sT
 
 intS :: SyntaxB -> Store -> Type2
-intS ([P "NP" np, P "VP" vp], _) r = intNP np r (intVP vp r)
-intS (((P "SBAR" ([(L "IN" ("If",_)), (P "S" s1)],_)) : (L "," _) : s2), _) r =
-		sSwitch `sComp` intS s1 r `sComp` sSwitch `sComp` intS (s2,(0,0)) r
-intS ([P "S" s1, L "CC" ("and",_), P "S" s2], _) r = intS s1 r `sComp` intS s2 r
+intS [P "NP" np, P "VP" vp] r = intNP np r (intVP vp r)
+intS (P "SBAR" ([(L "IN" ("If",_)), (P "S" s1)]) : (L "," _) : s2) r =
+		sSwitch `sComp` intS s1 r `sComp` sSwitch `sComp` intS s2 r
+intS [P "S" s1, L "CC" ("and",_), P "S" s2] r = intS s1 r `sComp` intS s2 r
 intS other r = error ("Unsuported by intS: " ++ show other)
 
 intNP :: SyntaxB -> Store -> Predicate -> Type2
 -- intNP ([L "NNP" (nnp,_)], i) r vb = sPredi nnp [v] `sComp`
-intNP ([L "PRP" (prp,i)], _) r = \vb -> vb (r i)
-intNP ([L "DT" dt, L "NN" nn], _) r = intDT dt r (intNN nn r)
+intNP [L "PRP" (prp,i)] r = \vb -> vb (r i)
+intNP [L "DT" dt, L "NN" nn] r = intDT dt r (intNN nn r)
 intNP other r = error ("Unsuported by intNP: " ++ show other)
 
 intVP :: SyntaxB -> Store -> Predicate
[…]

### 2013-04-26 Stanford.hs
Nicer way to do it

@@ -1,4 +1,4 @@
-module Stanford (Store, Word, Index, run, runDry, runOnFile, IndexTree(..), postrees) where
+module Stanford (Store, Word, Index, run, runDry, runOnFile, PosTree(..), postrees) where
 
 import Prop (Ref)
 
@@ -16,8 +16,7 @@ type Word = String
 type Sentence = [Word]
 
 type PosTag = String
-data PosTree = Phrase PosTag [PosTree] | Leaf PosTag Word deriving (Show)
-data IndexTree = P PosTag [IndexTree] | L PosTag (Word,Index) deriving (Show)
+data PosTree a = P PosTag [PosTree a] | L PosTag a deriving (Show)
 
 data DepTree = Dep Int [(Ref, DepTree)] deriving (Show, Eq)
 
@@ -30,20 +29,23 @@ type Store = Index -> Ref
 -- Use "lemma" instead of "word" to get standardized forms
 -- (did -> do, n't -> not)
 
-lemmas :: Element -> Sentence
-lemmas doc = map strContent (findElements (unqual "lemma") doc)
+lemmas :: Element -> [Sentence]
+lemmas doc = map lemma (findElements (unqual "sentence") sentences)
+	where
+		Just sentences = findElement (unqual "sentences") doc
+		lemma doc' = map strContent (findElements (unqual "lemma") doc')
 
 ------------------
 
-postrees :: Element -> [PosTree]
+postrees :: Element -> [PosTree Word]
 postrees doc = map rootfilter roots
 	where
 		parses = findElements (unqual "parse") doc
 		roots = map (postreeRoot . strContent) parses
-		rootfilter (Phrase "ROOT" [(Phrase "S" s)]) = (Phrase "S" s)
+		rootfilter (P "ROOT" [(P "S" s)]) = (P "S" s)
 		rootfilter root = error ("Cant parse strange sentence " ++ show root)
 
-postreeRoot :: String -> PosTree
+postreeRoot :: String -> PosTree Word
 postreeRoot dat = tree
 	where
 		Right tree = parse posEither "(unknown)" dat
@@ -59,28 +61,32 @@ posTree =
 		tag <- many (noneOf " ")
 		char ' ' -- maybe not needed?
 		subs <- sepBy1 posEither (char ' ')
-		return (Phrase tag subs)
+		return (P tag subs)
 posLeaf =
 	do
 		tag <- many (noneOf " ")
[…]

### 2013-04-26 ToLogic.hs

@@ -2,27 +2,27 @@ module ToLogic (logify) where
 
 import Dpl (dpl2prop, s2dpl, Type2, sT, sF, sSwitch, sScope, sPredi, sExist, sComp, sTest0, sTest1)
 import Prop
-import Stanford (Word, Store, Index, IndexTree(..))
+import Stanford (Word, Store, Index, PosTree(..))
 
 import Data.Map hiding (lookup, mapMaybe)
 import Data.Maybe
 
-logify :: (Store, [IndexTree]) -> Prop
-logify (refs, tree) = (simplify . dpl2prop . s2dpl) logic
-	where logic = intRoot tree refs
-
-type SyntaxB = [IndexTree]
-type SyntaxL = (Word, Index)
+type IWord = (Word, Index)
+type SyntaxB = [PosTree IWord]
 type Predicate = Ref -> Type2
 type Predicate2 = Ref -> Ref -> Type2
 
-intRoot :: [IndexTree] -> Store -> Type2
+logify :: (Store, [PosTree IWord]) -> Prop
+logify (refs, tree) = (simplify . dpl2prop . s2dpl) logic
+	where logic = intRoot tree refs
+
+intRoot :: SyntaxB -> Store -> Type2
 intRoot ((P "S" s):ss) r = intS s r `sComp` intRoot ss r
 intRoot [] r = sT
 
 intS :: SyntaxB -> Store -> Type2
 intS [P "NP" np, P "VP" vp] r = intNP np r (intVP vp r)
-intS (P "SBAR" ([(L "IN" ("If",_)), (P "S" s1)]) : (L "," _) : s2) r =
+intS (P "SBAR" ([(L "IN" ("if",_)), (P "S" s1)]) : (L "," _) : s2) r =
 		sSwitch `sComp` intS s1 r `sComp` sSwitch `sComp` intS s2 r
 intS [P "S" s1, L "CC" ("and",_), P "S" s2] r = intS s1 r `sComp` intS s2 r
 intS other r = error ("Unsuported by intS: " ++ show other)
@@ -34,7 +34,7 @@ intNP [L "DT" dt, L "NN" nn] r = intDT dt r (intNN nn r)
 intNP other r = error ("Unsuported by intNP: " ++ show other)
 
 intVP :: SyntaxB -> Store -> Predicate
-intVP [L ('V':'B':_) ("is",_), P "NP" np] r = \v -> sPredi (stringify np) [v]
[…]

### 2013-04-28 paper.tex
Before transfer to Visser syntax

@@
 
-\title{Musings in Dynamic Semantics for Donkeys}
+%Musings in Dynamic Semantics for Donkeys
+%From Donkeys to Predicates: Musings Into Dynamic Semantics and Logification
+%Donkification and Logification
+%Donkeys into Logic
+%From Farmers and Donkeys to Grammars and Predicates
+%Dependency ased
+%Discourse and Donkeys: A Dependency Based 
+%Discourse and Donkeys: A Monoid Based Language Grammar
+%Discourse and Donkeys: Monoid Based Language Semantics
+%Discourse and Donkeys: Using Monoids for Language Semantics
+%Dynamic Discourse and Donkeys: Using Monoids for Language Semantics in Practice
+
+\title{Dynamic Discourse and Donkeys: Using Monoids for Language Semantics in Practice}
 \author{Thomas Dybdahl Ahle}
@@
 \begin{abstract}
-Your abstract goes here...
-
-donkeys sentences
-dynamic logic
-disjunction, errors
-monads
-dependencies
-
-From Donkeys to Predicates: Musings Into Dynamic Semantics and Logification
-Donkification and Logification
-Donkeys into Logic
-From Farmers and Donkeys to Grammars and Predicates
-Dependency ased
-Discourse and Donkeys: A Dependency Based 
-Discourse and Donkeys: A Monoid Based Language Grammar
-Discourse and Donkeys: Monoid Based Language Semantics
-Discourse and Donkeys: Using Monoids for Language Semantics
-Dynamic Discourse and Donkeys: Using Monoids for Language Semantics
-
-‘a brief description of what you did; no more than a page.’
-The abstract provides a summary of the project report. It may contain some or all of the following moves, 
-often, but not always, in the following order. Only use one or two sentences for each move. Keep the 
-language and sentence structure simple. The reader should be able to read the abstract and obtain the 
-necessary information quickly and efficiently.
-Move 1:  Background to the project
-Move 2: Purpose of the project
-Move 3: Problem tackled
-Move 4: Work carried out
-Move 5: Results 
-Move 6: Conclusions or implications
-Move 7: Achievements of the project
-
-Let's try it then
-
-How to treat determinants/indeterminants
-  1) ref
-  2) quant
-  3) vp -> sentence function
-
-Ambiguity
-  1) Strong/Weak
-
-How to get a good logical representation
-  1) Looking like language
-  2) Compositionality
+
+%‘a brief description of what you did; no more than a page.’
+%The abstract provides a summary of the project report.
+It may contain some or all of the following moves,  often, but not always, in the following order.
+Only use one or two sentences for each move.
+Keep the  language and sentence structure simple.
+The reader should be able to read the abstract and obtain the  necessary information quickly and efficiently.
+%Move 1:  Background to the project
+%Move 2: Purpose of the project
+%Move 3: Problem tackled
+%Move 4: Work carried out
+%Move 5: Results 
+%Move 6: Conclusions or implications
+%Move 7: Achievements of the project
+
+A donkey sentence is a sentence with a pronoun that is bound in semantics, but not in syntax.
+This make them difficult to capture in the scope of ordinary logic.
+In this paper we explore some of the ways that have been proposed to tackle this problem, in particular Albert Visser's e-monoid extension of Dynamic Predicate Logic, which allows us to write formulas deceivingly similarly looking to natural language.
+
+We explain those concepts in depth, deriving or proving every result stated.
+We then start looking into implementing the logic on a computer, experimenting with different ways of defining the semantics.
+We further explore state of the art tools for dependency parsing and coreference resolution, the workings of which we also explain in depth.
+
+Finally we put it all together in the Haskell programming language.
+We test out our machinery on a range of different sentences, showing all the converting steps from natural language to propositional logic.
+The problem is hard and the result is far from perfect, but it shows some interesting and less explored possibilities in computational linguistics.
 
@@
 
-To draw up this point, let's first look at a sentence that easily lets itself convert and then a couple that dont:
+To draw up this point, let's first look at a sentence that easily lets itself convert and then a couple that don't:
 \begin{itemize}
@@
 If we want to do without constants we can write: $\forall x(Socrates(x)\rightarrow donkey(x))$, though this does allow more than one Socrates to exist.
-\item `All his donkeys are treated well': This sentence somewhat willingly let's itself translate into: $\forall y(his(x,y)\wedge donkey(y) \rightarrow treated\_well(y))$, which has a free variable $x$ refering to some person from earlier in the discourse.
+\item `All his donkeys are treated well': This sentence somewhat willingly let's itself translate into: $\forall y(his(x,y)\wedge donkey(y) \rightarrow treated\_well(y))$, which has a free variable $x$ referring to some person from earlier in the discourse.
 \item `Every farmer beats his donkey': This sentence clearly has two possible meanings: $\exists y(his(x,y)\wedge donkey(y)\wedge \forall z(farmer(z) \rightarrow beats(z,y))$ if there is only one donkey, and $\forall x(farmer(x)\rightarrow \exists y (beats(x,y)\wedge his(x,y)\wedge donkey(y))$ if there is a donkey for each farmer.
 \item `The happy donkey doesn't exist': Here we get into more problems.
-We might write: $\exists x(happy\_donkey(x)\rightarrow doesnt\_exist(x))$, but that asserts some sort of entity representing the non existing donkey.
+We might write: $\exists x(happy\_donkey(x)\rightarrow doesn't\_exist(x))$, but that asserts some sort of entity representing the non existing donkey.
 Alternatively we could write $\exists x(happy\_donkey(x)\rightarrow \bot)$, but that is another sentence: `There are no happy donkeys'.
@@
 The best we can do here seems to be $please\_close\_the\_stable\_door()$ or perhaps $\exists x(stable\_door(x)\wedge please\_close(x))$.
+\item `Jane likes a painting': This is easy to translate to logic, but are we comfortable that it has the same meaning as the sentence `It is not the case that Jane doesn't like a painting' and even `Jane likes a painting and either it is raining or it isn't'?
+It is hard to imagine a language with strong logical foundations, where those three sentences wouldn't be equal.
 \end{itemize}
@@
 In this paper we will work with a subset of English that translate reasonably well into logic.
-This unfortunately means that arguably the majority of sentences will be unaccesible to our methods.
-We will take a shot at ambigious sentences like the third one though.
+This unfortunately means that arguably the majority of sentences will be inaccessible to our methods.
+We will take a shot at ambiguous sentences like the third one though.
 
@@
 \begin{align}
-\exists x (farmer(x) \wedge \exists y (donkey(y)\wedge owns(x,y))) \rightarrow beats(x,y)
-\end{align}
-%
-This is arguably the sentence that first comes to mind for a moderately trained logician.
+\exists x (farmer(x) \wedge \exists y (donkey(y)\wedge owns(x,y)) \rightarrow beats(x,y))
+\end{align}
+%
+This is arguably the formula that first comes to mind for a moderately trained logician, or simply what you expect by looking at the structure of the syntax tree.
 Translating an indeterminate `a farmer' into $\exists x(farmer(x))$ seems correct, but clearly the formula has problems with scope.
@@
 %
-Moving the implication into the scope of $x$ and $y$ also doesn't work.
+Moving the implication into the scope of $x$ also doesn't work.
 Sentence \eqref{wrong_donkey_2} evaluates to true as long as a single farmer beats his donkey.
-In fact as long as a farmer exists it evaluates to true, since plugging the farmer into the $donkey$ predicate makes the anticedent false.
+In fact as long as a farmer exists it evaluates to true, since plugging the farmer into the $donkey$ predicate makes the antecedent false.
 %
@@
 %
-This turns out to be the correct formula. At least as long as you consider the use of `all' in natural language equal to the `all' in logic, which can be vacously true.
+This turns out to be the correct formula. At least as long as you consider the use of `all' in natural language equal to the `all' in logic, which can be vacuously true.
 
@@
 However in a lot of sentences this turns out to not be the case.
-
-How to treat determinants/indeterminants
-  1) ref
-  2) quant
-  3) vp -> sentence function
+There doesn't seam to any simple way to translate them into logic.
+Even less does it seem intuitive how to use them as antecedents to pronouns.
+
+The common way to deal with this problem is to make a big jump in power.
+Using a Montague Grammar we can define any noun phrase (NP) as a function from a verb phrase (VP) to a sentence.
+This way `if a farmer' might be a function $VP \to S \to S$, but even with this power, a working function is not easy to construct\cite{montague1973proper}.
 
 Another problem with sentence \eqref{correct_donkey} is a lack of compositionality.
-
-How to get a good logical representation
-  1) Looking like language
-  2) Compositionality
-
-Finally let's have a look at the ambiguity we noted in the beginning of this chapter
-
-Ambiguity
-  1) Strong/Weak
-
-If a farmer owns a donkey he beats it
-\begin{equation}
-\forall x . farmer(x) \rightarrow (\forall y . donkey(y) \wedge owns(x,y) \rightarrow beats(x,y))
-\end{equation}
-
-There are a lot of problems with this representation
-\begin{itemize}
-  \item It is not compositional. 
-  dfdf
-  \item It doesn't very well follow the order of the original sentence
-  \item It is just not very easy to get right
-\end{itemize}
-
-More stuff from wikipedia:
-
-Unfortunately, this translation leads to a serious problem of inconsistency.
-Indefinites must sometimes be interpreted as existential quantifiers, and other times as universal quantifiers, without any apparent regularity
-
-Another reason for studying to so called `donkey sentences' is their particular use of referants.
-
-The upshot of the foregoing observations is that, apparently, indefinites are neither quantifiers (dårlig konstruktion) nor referential (no farmer, etc.) terms, and this problem entrains another one, for as long as it unclear what indefinites mean, it will also remain obscure how they can serve as antecedents to pronouns.
-Later on we will see that very similar problems arise in seemingly disparate domains, like the interpretation of tense and presuppositions.
-
-(1)  A definite is used to refer to something that is already familiar at the current stage 
-of the conversation. An indefinite is used to introduce a new referent. 
-This has been  labeled the `familiarity  theory of definiteness.'
-When confronted with (l), the logically minded semanticist will notice immediately 
-that it  presumes  something  very  objectionable:  that  definites and indefinites  are 
-referring expressions.  For  only  if  there  is  a referent at  all  can there be  any question 
-of its familiarity or novelty
-
-In montague it is a function from verb to sentence
-
-If there is a Chompsky universal grammar, we should be able to make a simple logic over it??
-
-% Define block styles
+This property which is vital if we want to build up a meaning of a conversation piece by piece.
+Say for example that our primary sentence was followed by the sentence `He also feeds it': $feed(x,y)$.
+How do we combine the first and the second sentence to give the meaning $\forall x(farmer(x)\rightarrow \forall y(owns(x,y)\wedge donkey(y)\rightarrow beat(x,y)\wedge feed(x,y)))$?
+
+The point is that there is simply not a way to do this in PL.
+All variables in a formula are always closed off in their scope, and there are no logical operators that can `push' a new predicate into a closed scope.
+Our previous case might be a bit unfair, since we were trying to push something into a complicated implication, but even in case of the simple sentences `There is a farmer.
+He owns a donkey.' we wouldn't be able to build a meaning from the pieces.
+
+One could argue that we can simply obtain compositionality using Montague functions like we discussed earlier (and will discuss more deeply in a later chapter), but what would be really nice is to have a simple logic that simply has these things built in.
+In this paper we are going to look into not only one, but a couple of such logics.
+In these logics we will be able to write things such as:
+%
+\begin{align}
+(\exists x \cdot farmer(x)) \cdot (\exists y \cdot donkey \cdot owns(x,y))
+\end{align}
+%
+Here the variables survive throughout the formula, which allows us to concatenate it with other formulas. The parenthesis are associative, so we don't even need them.
+Later in the paper we will also look at a logic allowing us to write:
+%
+\begin{align}
+(\bowtie \cdot \exists x \cdot Farmer(x) \cdot \exists y \cdot donkey(y) \cdot owns(x,y) \cdot \bowtie beats(x,y)) \cdot feeds(x,y)
+\end{align}
+%
+Which allows us to be compositional even in the case of implications.
+
+It is interesting to consider whether such logics can give us something on the way of translating speech to logic.
+In this paper we are going to look into exactly that, and we will combine it with the best available tools for dependency analysis and coreference resolution.
+This is the game plan:
+\vspace{1em}
 \tikzstyle{decision} = [diamond, draw, fill=blue!20,
@@
 \tikzstyle{line} = [draw, -latex']
-
 \begin{tikzpicture}[node distance = 3cm, auto]
@@
 
-% It is not possible
-It is clear that not all sentences can be unambigiously translated from natural language to logic. A sentence like 	
-\begin{quotation}
-A farmer beats a donkey
-\end{quotation}
-Might be translated as 
-\begin{equation}
-\exists x ( farmer(x) \wedge \exists y ( donkey(y) \wedge beats(x,y)))
-\end{equation}
-Indicating that ``a farmer'' refers to some specific individual. Or we can might see it as a general rule, and translate it as
-\begin{equation}
-\forall x ( farmer(x) \rightarrow \exists y ( donkey(y) \wedge beats(x,y)))
-\end{equation}
-Indicating that every farmer has some donkey he beats. Or we could even read it as
-\begin{equation}
-\exists y ( donkey(y) \wedge \forall x ( farmer(x) \rightarrow beats(x,y)))
-\end{equation}
-Indicating that the poor donkey is common between all farmers.
-% This seams wrong to me
-
-Additionally, the indefinite article `a' is normally understood as an existential quantifier
-
-Week and strong readings, check this out \cite{kanazawa1994weak}
-
-In this paper we will not try to solve the above problem.
-(Though it could be interesting to generate multiple possibilities and perhaps use context to filter out the correct one).
-Instead we will try to show how some old method for doing this conversion can be enhanced with some new tools
-
-More questions:
-
-Should those three have the same meaning?
-[1a]  Jane likes a painting.
-[1b]  It is not the case that Jane doesn't like a painting.
-[1c]  Jane likes a painting and either it is raining or it isn't.
-
-direct translation not possible
-`He saw the man with the telescope'
-and even if we understand the syntax
-`The students all attended one class'
-has at least two meanings.
-interpretation of discourse is influenced and guided by the common ground that exists between speaker and hearer
+Going from step 2 to step 3 we are going to try and eliminate some of the ambiguity we saw earlier by outputting multiple possible sentences.
+In particular we will try and make a strong and a weak reading of every implication.
+However before we get into the details of how everything is implemented in code, we should look more into the nifty details of the theory applied.
 
@@
 
-Logicians and lingusts have come up with many different update logics since the 80s. I will quickly describe the four most cited ones, to give an idea of the landscape we're working in:
+Logicians and linguists have come up with many different update logics since the 80s. I will quickly describe the four most cited ones, to give an idea of the landscape we're working in:
 %
@@
 %
-At this step it is possible to perform anaphora related heuristics. Since $x$ is the first paramter of $farmer$, we assume it should be so for $beats$ as well:
+At this step it is possible to perform anaphora related heuristics. Since $x$ is the first parameter of $farmer$, we assume it should be so for $beats$ as well:
 %
@@
 
-\item File Change Semantics, 1982 by Irene Heim\cite{heim1983file}, are based around the puzzle of discourse `referents' which may, as we have seen, sometimes act like quantifiers, sometimes referents and sometimes anticedents.
+\item File Change Semantics, 1982 by Irene Heim\cite{heim1983file}, are based around the puzzle of discourse `referents' which may, as we have seen, sometimes act like quantifiers, sometimes referents and sometimes antecedents.
 Heim suggests using so called `file cards' which though developed independently, has a lot in common with DRT.
@@
 However, where DRT defines complex rules for merging different `discourse referents', in File Change Semantics (FCS) the part `added onto' a file gets to decide how the merge is made.
-Some how similar to how lambda functions in montague grammers give you a lot of flexibility.
+Some how similar to how lambda functions in Montague grammars give you a lot of flexibility.
 For example in case of `adding' $p$, an n-ary atomic predicate, to a file $F$:
@@
     \node [round, right of=z] (donkey) {donkey};
-    \node [round, below right of=x] (owns) {owns};
-    \node [round, below left of=owns] (y) {y};
+    \node [round, below of=x] (y) {y};
+    \node [round, below of=z] (owns) {owns};
     % Draw edges
@@
 In contrast to some of the previous models, the authors of Dynamic Predicate Logic (DPL) don't make any big philosophical claims on connections between their model and the brain.
-Instead they simply create a logic that is isomorphic to PL, while keeping a reasonable ammount of `update semantics' like properties, such as being closer syntactically to natural language.
-
-DPL also doesn't provide any guidiance into resolving anaphora.
+Instead they simply create a logic that is isomorphic to PL, while keeping a reasonable amount of `update semantics' like properties, such as being closer syntactically to natural language.
+
+DPL also doesn't provide any guidance into resolving anaphora.
 Just like giving meaning to the actual concepts, this must be handled outside of the logic.
-We will see later in this paper that this stance makes sence, since the current state of the art anaphora resolvers are simply based on simple heuristics over noun phrases.
+We will see later in this paper that this stance makes sense, since the current state of the art anaphora resolvers are simply based on simple heuristics over noun phrases.
 
@@
 
-The logic cleverly uses these advances in logical understanding to create an update logic in a pragmatic way, without any external constructs and a minimum of new syntax, while still getting all the benefits of compoitionality etc.
+The logic cleverly uses these advances in logical understanding to create an update logic in a pragmatic way, without any external constructs and a minimum of new syntax, while still getting all the benefits of compositionality etc.
 
@@
 The logic is 100\% associative).
-Notice how we introduce referants casually as we progress in the discourse.
+Notice how we introduce referents casually as we progress in the discourse.
 We even manage to look pretty similar to PL at the same time.
@@
 Which on earth relation could $\exists S$ be?
-We are going to take a hint from computer programming and consider the variable decleration: $S:=\bot$.
+We are going to take a hint from computer programming and consider the variable declaration: $S:=\bot$.
 We want to write this in DPL as $\exists S \cdot \neg S$, such that e.g.
@@
 Visser builds this up in two steps: First a relational algebra and then DPL on top of it.
-But since DPL is bascailly identical to the relational algebra, we are going to try and do it in one step.
+But since DPL is basically identical to the relational algebra, we are going to try and do it in one step.
 
@@
 For example, how can we intuitively understand the DPL notion of $\rightarrow$?
-Well, first notice that many of the eqations start with $\alpha=\beta\wedge\dots$.
+Well, first notice that many of the equations start with $\alpha=\beta\wedge\dots$.
 These equations we can think of intuitively as having a closed scope.
@@
 
+Of course this `static negation' might not always be the right thing.
+Consider the sentences `It is not that case that a farmer doesn't own a donkey.
+He beats it'.
+This should be a good sentence, but we can't handle it.
+In the next chapter we will look at how this might be handled.
+
 The DPL $\psi_1\rightarrow\psi_2$ in particular can be seen as the filter, that if you can get `through' $\psi_1$ then you must also be able to get through $\psi_2$.
@@
 Hence variables introduced in $\psi_1$ can well bind variables in $\psi_2$.
-This will show to be extreamly important in the treatment of donkey sentences later.
+This will show to be extremely important in the treatment of donkey sentences later.
 
@@
 
-Having defined seamingly very sound semantics for DPL above, we may start to notice certain oddities about our choice of using the relation union for `or'.
-An obvious alternative would have been the closed/static $\neg(\neg\psi_1\wedge\neg\psi_2)$ or $\neg\psi_1\rightarrow\psi_2$ considered by Groenendijk\cite{groenendijk1991dynamic}.
-
-Actually there is an obvious definition of `or' that we have totally ignored in the previous chapter:
+A major problem with our use of `relation union' as disjunction is the way it allows seemingly sound sentences to combine into syntactically invalid ones.
+Consider the sentence `Either a farmer owns a donkey or he doesn't'.
+We would write that as
+\begin{align}
+\exists x\cdot farmer(x)\cdot(\exists y\cdot owns(x,y)\cdot donkey(y)\vee \bot) \label{owns_donkey_or_not}
+\end{align}
+Now concatenating this sentence with `he beats it': $beat(x,y)$ would give us a problem.
+In fact this problem would be more than a syntactic problem, because before we know the model and the existence of the donkey, we don't know if the sentence is valid or not.
+
+Visser\cite{visser1999donkey} suggests that we should build our logic on sets of relations read disjunctively instead of just relations.
+This `possible world' scenario would transform \eqref{owns_donkey_or_not}$\cdot beats(x,y)$ into $\{\exists x\cdot farmer(x)\cdot\exists y\cdot donkey(y)\cdot owns(x,y)\cdot beats(x,y), \exists x\cdot farmer(x)\cdot\bot\cdot beats(x,y)\} = \{\exists x\cdot farmer(x)\cdot\exists y\cdot donkey(y)\cdot owns(x,y)\cdot beats(x,y), \bot\}$.
+This clearly is always well defined.
+
+Visser's idea doesn't fully save us though. If we now consider `a farmer either owns a donkey or owns a horse':
+\begin{align}
+\exists x\cdot farmer(x)\cdot(&\exists y\cdot owns(x,y)\cdot donkey(y)\nonumber\\
+                              &\vee \exists z\cdot owns(x,z)\cdot horse(z))
+\end{align}
+And combines that with `He beats the donkey': $beat(x,y)$, we get, in Visser's model, $\{\exists x\cdot farmer(x)\cdot\exists y\cdot owns(x,y)\cdot donkey(y)\cdot beats(x,y), \exists x\cdot farmer(x)\cdot\exists z\cdot owns(x,z)\cdot horse(z)\cdot beats(x,y)\}$.
+The second one is clearly not valid (though we wouldn't get a problem if there is no horse or farmer) which we might catch `at compile time' and replace it with something like `$\bot$' or `error'.
+This sort of error handling is discussed and implemented in the implementation section, but is not terribly interesting from a formal point of view.
+
+Much more interesting are the following two questions: How could we have translated \label{owns_donkey_or_horse} so it would actually work?
+After all it seemed like a reasonable sentence until variable naming got us.
+And does the way variable binding works with our definition actually make sense?
+
+Clearly just renaming $z$ to $y$ in the horse term would have solved the first problem.
+However to work in a compositional way, that would either require great luck, or renaming, and we don't have a renaming operator.
+
+A trick we could consider is to declare an `object' variable outside of the disjunction:
+\begin{align}
+\exists s\cdot farmer(s)\cdot\exists o\cdot(
+      &\exists y\cdot owns(s,y)\cdot donkey(y)\cdot o=y \nonumber\\
+      &\vee \exists z\cdot owns(s,z)\cdot horse(z)\cdot o=z)\cdot beats(s,o)
+\end{align}
+%
+This even inspires on idea that we might be able to build everything up around subjects, objects, indirect objects etc.
+This is explored in the implementation, but it does suffer from many problems.
+E.g.
+we might well have a sentence very similar to the above: `A farmer owns a donkey and owns a horse.
+He beat them'.
+Here we suddenly have two objects in the same sentence.
+
+Sentences like the above are easy to make in English because the absence of noun genders make it easy to use the same pronoun to refer to a lot of things at once.
+By choosing the right actors however, it is always possible to run into problems.
+
+The problem is perhaps the non-linearity in the way variables pass through the disjunction. Somehow we lose track of their history when the `results' are merged together.
+
+This leads on to the other question we mentioned: Is the binding flow through our disjunction operator really any good?
+We are is inspired by the sentence `Either a farmer or his scapegoat must go to court.
+In any case he will be picked up by sunset'.
+In terms of DPL formulas, this would be modelled something like $(\psi_1\cup\psi_2)\cdot\psi_3$.
+Clearly we have a problem here: Variables bound in $\psi_1$ are not accessible in $\psi_2$.
+
+This sentence suggests that we might want to define a disjunction operator with a linear flow of variables: $\psi_1\to\psi_2\to\psi_3$ instead of $\psi_1\to\psi_3\text{ and }\psi_2\to\psi_3$.
+However there is a problem: We currently define `false' as the absence of possible assignments.
+Hence if $\psi_1$ is false, there is no way we can define disjunction to push it's variables into $\psi_2$.
+If any variables were left undefined, $\psi_1$ wouldn't be false.
+
+There is a cure to this problem.
+When we extend our logic with more power in the monad chapter, we get a second chance.
+However Let's first have a look at an obvious alternative definition for disjunction considered by Groenendijk\cite{groenendijk1991dynamic}.
+We'll call it `static disjunction':
 \begin{align}
@@
  :\equiv&\ \alpha[\neg(\neg\psi_1\wedge\neg\psi_2)]\beta \nonumber\\
- \leftrightarrow&\ \alpha[\neg\psi_1\rightarrow\psi_2)]\beta \nonumber\\
+ \leftrightarrow&\ \alpha[\neg\psi_1\rightarrow\psi_2]\beta \nonumber\\
  \leftrightarrow&\ \alpha = \beta \wedge \forall\gamma(\alpha[\neg\psi_1]\gamma\rightarrow\gamma\models[\psi_2]) \nonumber\\
@@
 
-So does this fix our problem of binding between clauses? No. In fact this definition doesn't even allow binding from a clause and out.
-
-a farmer doesn't own a horse or a donkey
-\begin{align}
-\exists x\cdot farmer(x) \cdot \neg(\exists y\cdot horse(y)\cdot owns(x,y) v \exists y\cdot donkey(y)\cdot owns(x,y))\nonumber\\
-\exists x\cdot farmer(x) \cdot \neg\exists y\cdot horse(y)\cdot owns(x,y) \cdot \neg \exists y\cdot donkey(y)\cdot owns(x,y)\nonumber
-\end{align}
-
-a farmer beats a horse or a donkey
-\begin{align}
-&Ex.farmer(x).(Ey.horse(y) v Ez.donkey(z)).beats(x,y) = \\
-&\{Nothing, Ex.farmer(x).EyJust .horse(y).beats(x,y)\}\\
-\\
-&Ex.farmer(x).~((Ey.horse(y) v Ez.donkey(z)).beats(x,y)) = \\
-&Ex.farmer(x).(~Ey.horse(y).~Ez.donkey(z) v ~beats(x,y)) = \\
-&\{Nothing, Ex.farmer(x).&~Ey.horse(y).~Ez.donkey(z)\}\\
-\\
-&Ex.farmer(x).~(tall(x) v blue(x)) = Ex.famer(x).~tall(x).~blue(x)
-\end{align}
-
-There is a farmer who maybe has a donkey. He beats it.
-$Ex.farmer(x).(Ey.donkey(y) v false).beats(x,y)$
-With union semantics for v this would be equal to
-$Ex.farmer(x).Ey.donkey(y).beats(x,y)$
-which is stupid.
-Intead with sets of relations, this would be equal to
-$Ex.farmer(x).Ey.donkey(y).beats(x,y) v Ex.farmer(x).false.beats(x,y)$
-the right side being an error? no
-
-If a farmer is not poor he owns a donkey. He beats it.
-$Ex.farmer(x).~poor(x) -> Ey.donkey(y).owns(x,y).beats(x,y)$
-
-A farmer is either poor or he owns a donkey. He beats it.
-$Ex.farmer(x).(poor(x) v Ey.donkey(y).owns(x,y)).beats(x,y) =$
-
-%(Ex v false) . f(x)
-%(exist x `union` false) [] = [[("x",0)],[("x",1)],[("x",2)]
-%[Nothing, Just [[("x",0)],[("x",1)],[("x",2)]]]
-
-%(Ex v false) = [Just [[("x",0)],[("x",1)],[("x",2)], Just []]
-%not (Ex v false) =
-
-%not (Ex v f(x))
-%= ~Ex.~f(x)
-
-A farmer owns a horse or a donkey
-$Ex.farmer(x).~(~(Ey.horse(y)).~(Ey.donkey(y))).owns(x,y)$   this surely doesn't work
-
-nothing survives out of a not (or impl) it's normal. Don't worry about it
-
-Let's just say that if some of 
-maybe not (bla v bla v bla v ERROR) should be ERROR
-it could also be something more involved, like ERROR(bla v bla v bla v ERROR), but let's not do that
-fmap
-
-We can also try to define semantics using first order logic. This totally doesn't work and convert reasonable statements into mess. In particular it is hard to have or not being closed scopewise.
-
-The difference between those two:
-
-$addDpl p ((DOr d1 d2):ds) = addDpl (And (Or (addDpl T [d1]) (addDpl T [d2])) p) ds$
-
-$addDpl p ((DOr d1 d2):ds) = Or (addDpl p (d1:ds)) (addDpl p (d2:ds))$
-
-Fra Dynamic Montague Grammar:
-This means that the extended dynamic semantics is also able to deal with a donkey
-disjunction such as (49), and in such a way that it is equivalent with the donkey
-conditional (50). It can further be noted that this notion of dynamic disjunction is
-like dynamic implication in this respect that sentences which follow it, are in fact
-conjoined with its second argument:
-Fact 24 [either p or q] ; r = either p or [q ; r]
-This is needed for a treatment of examples such as:
-(54) Either there is no bathroom here, or it is in a funny place. In any case, it is
-not on the ground floor.
-
-Den ene giver mulighed for at binde fra a til b. Den anden giver mulighed for at begge dele binder res
-
-hvad hvis vi puttede vissers disjoint union stuff into the monad?
+The idea to this follows natural from De Morgan's laws of classical logic.
+Clearly the semantics are different from the ones we defined for $\cup$.
+However not in a good way.
+This static disjunction doesn't allow any variables to escape from inside it's reach.
+
+So does this fix our problem of binding between clauses?
+No.
+In fact this definition doesn't even allow binding from a clause and out.
+The villain is $\neg\psi$.
+From the first mention we defined it to be static so it wouldn't allow unsafe, arbitrary assignments as long as they didn't satisfy $\psi$.
+We wanted our logic to be reasonably monotonic, and we got what we deserved.
+$\neg$ not only closed down this version of disjunction, but our implication definition as well.
+
+We can imagine defining $\neg$ dynamically: $\neg\psi_1\cdot\psi_2 = \neg(\psi_1\cdot\psi_2)$.
+However not in our current model, and it won't solve any of our problems with disjunction.
+It does however allow this sentence: `It is not the case that a farmer doesn't own a donkey.
+He beats it'.
+While it made sense not to allow variables to escape from a single $\neg$, we want them to escape from a double.
 
@@
 Similarly we may create PL formulas with the same meaning as any DPL formula.
-But perhaps we should be a bit conserned on how to interpret the meaning of a PL formula and a DPL formula being equivalent.
+But perhaps we should be a bit concerned on how to interpret the meaning of a PL formula and a DPL formula being equivalent.
 After all DPL formulas are relations where PL formulas a propositions.
@@
 If we consider the truth value of a DPL relation to be whether there exists any relation that models it $\exists\alpha(\alpha\models\psi)$, we can use a trick.
-The trick is to think of the weakest precondion such an assignment must satisfy, and model that in PL.
+The trick is to think of the weakest precondition such an assignment must satisfy, and model that in PL.
 In fact the precondition is not just for our assignment, but for the entire model.
@@
  & \leftrightarrow \alpha\models\langle\psi_1\rangle\phi \vee \langle\psi_2\rangle\phi \label{conv_or}\\
+ \alpha\models\langle\psi_1 \vee \psi_2\rangle\phi
 \alpha\models\langle P(x_1,\dots,x_n)\rangle\phi
@@
 
-Notice that $\alpha\models\langle\bot\rangle\phi\leftrightarrow a\models\bot$ and $\alpha\models\langle\top\rangle\phi\leftrightarrow a\models\phi$ follow trivially from \eqref{conv_pred}.
-For my full derivation of these equations, see the appendix.
-
-In the implementation we will use the above equations to recursively convert our DPL formulas to PL. The $\langle\rangle$ notation easily lends itself to calculation by a stak manipulation function.
+See the appendix for my full derivations of every equation.
+Also notice that $\alpha\models\langle\bot\rangle\phi\leftrightarrow a\models\bot$ and $\alpha\models\langle\top\rangle\phi\leftrightarrow a\models\phi$ follow trivially from \eqref{conv_pred}.
+Equations for `convenience' operators such as $\rightarrow$ and $\vee$ can be derived too:
+\begin{align}
+\langle\psi_1\vee\psi_2\rangle\phi
+& \leftrightarrow \langle\neg(\neg\psi_1\cdot\neg\psi_2)\rangle\phi \nonumber\\
+& \leftrightarrow \neg\langle\neg\psi_1\cdot\neg\psi_2\rangle\top \wedge\phi \nonumber\\
+& \leftrightarrow \neg\langle\neg\psi_1\rangle\langle\neg\psi_2\rangle\top \wedge\phi \nonumber\\
+& \leftrightarrow \neg\langle\neg\psi_1\rangle(\neg\langle\psi_2\rangle\top) \wedge\phi \nonumber\\
+& \leftrightarrow \neg(\neg\langle\psi_1\rangle\top \wedge \neg\langle\psi_2\rangle\top) \wedge\phi \nonumber\\
+& \leftrightarrow (\langle\psi_1\rangle\top \vee \langle\psi_2\rangle\top) \wedge\phi \label{conv_vee}
+\end{align}
+
+In the implementation we will use the above equations to recursively convert our DPL formulas to PL. The $\langle\rangle$ notation easily lends itself to calculation by a stack manipulation function.
 
@@
 
-\subsubsection{Weird disjunctive subjects}
-
-problem wouldn't work in Spanish
-A farmer or their donkey was arrested this morning.
-
-\begin{equation}
-Ex(farmer(x)) ^ Ey(donkey(y)) ^ (a-t-m(x) v a-t-m(y))
-ExEy(farmer(x) ^ donkey(y) ^ owns(x,y) ^ (a-t-m(x) v a-t-m(y)))
-Ex(farmer(x)) v Ey(donkey(y))
-((Ex . farmer(x) . Es.s=x) v (Ey . donkey(y) . owns(x,y) . Es.s=y)) . a-t-m(s)
-((Ex . farmer(x) . (Es.s=x v (Ey . donkey(y) . owns(x,y) . Es.s=y)) v (Ey . donkey(y) . owns(x,y) . Es.s=y)) . a-t-m(s)
-\end{equation}
-
-If a farmer is Spanish or a Japanese she beats her donkey
-It is a donkey or a horse
-You beat it or I do
-Either a farmer or his donkey is happy
-
-either = xor?
-
-Vissers dynamic bracketing
-
-\subsection{Errors}
-
-Or should this be in implementation? Hard for me to say
-
 \subsection{Monadic variants}
@@
 
-To resque us from this problem Albert Visser suggests constructing a special e-monoid over the dynamic relational algebra.
-The idea is that we are going to simultaniously work on two `streams' of dpl, and switch easily between them with a special element $\Bowtie$.
+To rescue us from this problem Albert Visser suggests constructing a special e-monoid over the dynamic relational algebra.
+The idea is that we are going to simultaneously work on two `streams' of dpl, and switch easily between them with a special element $\Bowtie$.
 
@@
 Notice how we keep the negative stream in the first position of the tuple, and the positive stream at the second position.
-The third position idicates to what stream we are currently writing.
+The third position indicates to what stream we are currently writing.
 For example, if we are writing to the negative stream, and we apply an element that itself asks us to continue at its negative stream, we go back to the positive stream.
@@
 Notice $\Bowtie$ and $\triangle$ are commutative, since they work entirely on their own indicator.
-It's an interesting question to consider what sentences might require three or more levels of scoping, but those seam to be so rare that it doesn't now make much sense to clutter our algebra with an extra infinity of streams.
-
-We define the monad as follows:
+It's an interesting question to consider what sentences might require three or more levels of scoping.
+We could define $\triangle$ to go up in scope and $\bigtriangledown$ to go down.
+Instead of ${0,1}$ we would work on $\mathbb{Z}$.
+
+For now we define the monad as follows:
 %
@@
 %
-This is quite interesting for certain centences containing `surprises', like
+This is quite interesting for certain sentences containing `surprises', like
 %
@@
 \subsection{Montague semantic}
-Montague grammars are an approach to combinatorical semantics based on types and lambda calculus. It is interesting because it 
+
+$\exists a(farmer(a) \wedge \exists g(donkey(g) \wedge \forall o(o=a\vee o=g \rightarrow \exists v(friend(v) \wedge visit(o,v)))))$
+
+Montague (1970b): `I reject the contention that an important
+theoretical difference exists between formal and natural
+languages.'
+
+To us, the revolutionary idea in Montague's PTQ paper (and
+earlier papers) is the claim that natural language is not impossibly
+incoherent, as his teacher Tarski had led us to believe, but that
+large portions of its semantics can be treated by combining known
+tools from logic, tools like functions of finite type, the lambda-calculus,
+generalized quantifiers, tense and modal logic, and all the rest.
+
+Montague grammars are an approach to combinatorial semantics based on types and lambda calculus. It is interesting because it 
 Types
@@
 
-montague can make any logic compositional
+`...
+we are convinced that the capacities of [Montague Grammar] have not been exploited to the limit, that sometimes an analysis is carried out in a rival framework simply because it is more fashionable' - Groendijk \& Stokhof in `Dynamic Montague Grammar'
+
+Montague can make any logic compositional
 
@@
   Transitive verb (owns, beats): Entity -> Entity -> Bool
-  Attributive adjective (good, inteligent, former): (Entity -> Bool) -> (Entity -> Bool)
+  Attributive adjective (good, intelligent, former): (Entity -> Bool) -> (Entity -> Bool)
   Noun phrase: (Entity -> Bool) -> Bool
@@
 
-Let's return multiple intepretations
+Let's return multiple interpretations
+
+% It is not possible
+It is clear that not all sentences can be unambiguously translated from natural language to logic. A sentence like  
+\begin{quotation}
+A farmer beats a donkey
+\end{quotation}
+Might be translated as 
+\begin{equation}
+\exists x ( farmer(x) \wedge \exists y ( donkey(y) \wedge beats(x,y)))
+\end{equation}
+Indicating that ``a farmer'' refers to some specific individual. Or we can might see it as a general rule, and translate it as
+\begin{equation}
+\forall x ( farmer(x) \rightarrow \exists y ( donkey(y) \wedge beats(x,y)))
+\end{equation}
+Indicating that every farmer has some donkey he beats. Or we could even read it as
+\begin{equation}
+\exists y ( donkey(y) \wedge \forall x ( farmer(x) \rightarrow beats(x,y)))
+\end{equation}
+Indicating that the poor donkey is common between all farmers.
+% This seams wrong to me
+
+Additionally, the indefinite article `a' is normally understood as an existential quantifier
+
+Week and strong readings, check this out \cite{kanazawa1994weak}
 
@@
 
-\subsection{Corefferences}
-Corefferences (or anaphora) is essential in order to create advanced discourse, since it is what allows us maintain a topic beond the first mention.
-Without corefferences we couldn't even express simple logical statements such as transitivity: `a man's brother's brother is his brother' etc.
-
-Identifying corefferences (or resolving anaphora) is hence an important part of converting natural language into logic.
+\subsection{Coreferences}
+Coreferences (or anaphora) is essential in order to create advanced discourse, since it is what allows us maintain a topic beyond the first mention.
+Without coreferences we couldn't even express simple logical statements such as transitivity: `a man's brother's brother is his brother' etc.
+
+Identifying coreferences (or resolving anaphora) is hence an important part of converting natural language into logic.
 The issue is not touched upon in Visser's paper, which casually sidesteps the problem by assigning variables to all mentioned entities.
@@
 For the subtask of resolving simple pronouns a solution could be to work in terms of subjects, objects, indirect objects and so on.
-However if not earlier, this certainly fails for pronouns seperated from their antecedent by sentence bounderies.
-Also we want to resolve other kinds of corefferences, such as `John, the farmer, owns a donkey', where `the farmer' and `John' reffers to the same entity.
+However if not earlier, this certainly fails for pronouns separated from their antecedent by sentence boundaries.
+Also we want to resolve other kinds of coreferences, such as `John, the farmer, owns a donkey', where `the farmer' and `John' refers to the same entity.
 
@@
 In the sentence `John and his wife had a farm.
-He took care of the donkeys' we could substitute `He' with `She' and to change the refferant of the pronoun.
+He took care of the donkeys' we could substitute `He' with `She' and to change the referent of the pronoun.
 In this case we need semantic knowledge of genders to proceed.
 
-In my project I have taken advantage of the corefference tagger built into Stanford CoreNLP\cite{lee2013deterministic}\cite{lee2011stanford}\cite{raghunathan2010multi}.
-In the following section I will summerize the workings of this machine, made to work across large texts as well as simple sentences.
-
-Correference tagging is one part of computational linguistics where machine learning algorithms still haven't been able to outperform large itterative applications of linguistic heuristics.
-Of course the procedure assumes (most likely) machine learned postagging and named entity recognization, but the main algorithm is transparent and deterministic.
+In my project I have taken advantage of the coreference tagger built into Stanford CoreNLP\cite{lee2013deterministic}\cite{lee2011stanford}\cite{raghunathan2010multi}.
+In the following section I will summarize the workings of this machine, made to work across large texts as well as simple sentences.
+
+Coreference tagging is one part of computational linguistics where machine learning algorithms still haven't been able to outperform large iterative applications of linguistic heuristics.
+Of course the procedure assumes (most likely) machine learned postagging and named entity recognition, but the main algorithm is transparent and deterministic.
 
@@
 \item[Mention detection]
-The first step is to identify all noun phrases (NP) and personal (PRP) and possesive (PRP\$) pronouns. This information is easily pulled from the postree. Notice we also identify nested mentions:
+The first step is to identify all noun phrases (NP) and personal (PRP) and possessive (PRP\$) pronouns. This information is easily pulled from the postree. Notice we also identify nested mentions:
 
@@
 
-All mentions are initially assigned to seperate entities (entity ids).
+All mentions are initially assigned to separate entities (entity ids).
 The entities are also attached information we can easily pull from the postree, such as `a girl: {number: singular, gender: female}' which are used for the following rules.
@@
 
-We also note restrictions for future use, so we don't risk corefferencing an `I' with a `he' etc.
+We also note restrictions for future use, so we don't risk coreferencing an `I' with a `he' etc.
 \item[String Match]
-The second step is to merge entities reffered to by exactly the same string. In our case we have two mentions of `John':
+The second step is to merge entities referred to by exactly the same string. In our case we have two mentions of `John':
 
@@
 
-Like the `String Match' sieve, we are not doing anything fancy here, and we will incorrectly corefference the two mentions of `John' in: `[John, who loves donkeys] and [John, who hates donkeys] were brothers.'
+Like the `String Match' sieve, we are not doing anything fancy here, and we will incorrectly coreference the two mentions of `John' in: `[John, who loves donkeys] and [John, who hates donkeys] were brothers.'
 \item[Precise Constructs]
-The `Precise Constructs' sieve uses a lot of common patterns that link metions. In our case two `X is Y' patterns are found:
+The `Precise Constructs' sieve uses a lot of common patterns that link mentions. In our case two `X is Y' patterns are found:
 
@@
 \item[Proper Head Noun Match] could be called `Strict Head Match D'.
-It is the weakest form of the sieve, as it has only the three contraints: two mentions of the same entity cannot have one included in the other, they cannot contain words reffering to conflicting geographical locations, and they cannot have different numbers.
+It is the weakest form of the sieve, as it has only the three constraints: two mentions of the same entity cannot have one included in the other, they cannot contain words referring to conflicting geographical locations, and they cannot have different numbers.
 e.g.
@@
 
-As always there are a few contraints put on what can be merged. One rule is that the distance between pronun and antecedent must be maximum 3 sentences.
+As always there are a few constraints put on what can be merged. One rule is that the distance between pronoun and antecedent must be maximum 3 sentences.
 \end{description}
@@
 
-The Stanford procedure above is currently the strongest competitor in various correference competitions.
+The Stanford procedure above is currently the strongest competitor in various coreference competitions.
 
@@
 \section{Implementation (8)}
+
+Incremental dynamics\quote{van2001incremental}
+
+DPL Semantics - Functional Version
+DPL Semantics - Relational Version
+A state is an element of $D^V$ (function from variables to domain)
+In functional semantics a state is a function $D^V -> P(D^V)$
+In relational semantics it is a relation $P(D^V x D^V)$
 
@@
 
-And since we are going to deal with some large computer generated formulas, I've implemented a simple simplification function, which simlply applies basic identities for $\top$ and $\bot$, and a comnbination of De Morgan's laws and the implication identity to eliminate negations.
-
-One interesting part is that since the simplifyer has no look ahead, it will sometimes not find all applicable simplifications in one pass. For example
+And since we are going to deal with some large computer generated formulas, I've implemented a simple simplification function, which simply applies basic identities for $\top$ and $\bot$, and a combination of De Morgan's laws and the implication identity to eliminate negations.
+
+One interesting part is that since the simplifier has no look ahead, it will sometimes not find all applicable simplifications in one pass. For example
 %
@@
 
-\subsection{My implementaiton of DPL semantics}
+\subsection{My implementation of DPL semantics}
 
@@
 In addition to being essential for composition, free variables are also very useful for information extraction type sentences.
-We can immagine writing a curious researcher writing into a search engine `farmers who like donkeys' which translates into $farmer(x) \cdot likes(x,y) \cdot \triangle \cdot \exists y \cdot donkey(y)$.
+We can imagine writing a curious researcher writing into a search engine `farmers who like donkeys' which translates into $farmer(x) \cdot likes(x,y) \cdot \triangle \cdot \exists y \cdot donkey(y)$.
 
@@
 
-The coreference parser does however not find entities that are only mentioned once. Hence to make the later processing simpler, we assign those entitie to free variable now as well.
-
-The task is slightly complicated by the fact that NP's can be nested, so `The farmer and the donkey' is a NP consisting of two smaller NP's `The farmer' and `the donkey' seperated by `and'.
+The coreference parser does however not find entities that are only mentioned once. Hence to make the later processing simpler, we assign those entity to free variable now as well.
+
+The task is slightly complicated by the fact that NP's can be nested, so `The farmer and the donkey' is a NP consisting of two smaller NP's `The farmer' and `the donkey' separated by `and'.
 
@@
 
+\subsubsection{Errors}
+
+What do we do when they fail?
+Also something from the disjunction chapter.
+
+What if my states always contain the entire domain as their domain. Just add a $\bot$ to the entity set, then everything will be mapped to that as default.
+
 \subsection{Using machine learning}
@@
 
-My main problem here is that such translation requires large ammounts of dualingo data. It was out of scope for my project to try create such.
-
-\subsection{Handcoding}
-This is not as bad as it sounds. It is similar to montague and what the stanfordparser does for dependencies.
+My main problem here is that such translation requires large amounts of bilingual data. It was out of scope for my project to try create such.
+
+\subsection{Hand coding}
+This is not as bad as it sounds. It is similar to Montague and what the Stanford Parser does for dependencies.
 
@@
 We have already seen some screen shots of the working program; however, we provide two stringent tests 
-for our program to ensure it works as intended, along with a test rig to fully analyse the program. In both test 
+for our program to ensure it works as intended, along with a test rig to fully analyze the program. In both test 
 programs, I will run through the whole series of options available to the user, and ensure its correctness. 
-However, I will also demonstrate its ability to visualise code, and hopefully provide valuable insights whilst 
+However, I will also demonstrate its ability to visualize code, and hopefully provide valuable insights whilst 
 debugging. 
@@
 
+In the papers it is always assumed that the postree and things are `trivial' or `already sorted', but how do the pieces really stack up when put together?
+
 What I learned
@@
 improved; a discussion of the lessons you have learnt from the project.’
-In the ‘Conclusions’ you summarise your work and then stand back from the project and assess it as 
+In the ‘Conclusions’ you summarize your work and then stand back from the project and assess it as 
 objectively as you can. This shows the examiners that you are capable of evaluating your own work 
@@
 Move 1: Summary 
-Summarises the most important aspects of the project work.
+Summarizes the most important aspects of the project work.
 
@@
 
+Things learned from courses
+
 Should I make my latex file literate haskell? %bhttp://www.haskell.org/haskellwiki/Literate_programming
@@
 
-\section{Acknowledgements}
-Thanks to Samson
-
-This section's content...\cite{visser1999donkey}
+\section{Acknowledgments}
+I would like to thank Samson for guiding me through
 
@@
 
-Érica is still super hot. And she rocks at linguistics.\cite{lander2001initial}
-
-\subsection{Checklist}
-Things learned from courses
+Érica for linguistic discussions
 

### 2013-04-28 Main.hs

@@ -18,29 +18,45 @@ main :: IO ()
 main = do
 		putStrLn "Test1"
 		props <- super [
+			"Every farmer beats some donkey.",
 			"If a farmer owns a donkey, he beats it",
 			"If a farmer owns it and he beats it, it is a donkey", -- fails because 'donkey' becomes head somehow. (It even has a nsubj, huh?)
 			"If a farmer owns it and beats it, it is a donkey", -- fails because co-ref doesn't handle single occourences
-			"She is a mother of five", -- mother becomes the root, because is is not a 'real' verb
+			"A man comes in. He sees a dog. He smiles.", -- very sequential
 			"A farmer. A Donkey. He owns it. He beats it", -- fails because we assume all sentences have a VP root
-			"If you like me, you will hate my cousin",
-			"Do you still beat your wife?",
+			"All donkeys and a horse sing a song.",
+			"A farmer starves and beats a donkey.",
+			"A farmer starves a horse and beats a donkey.",
+			"If a man walks, he talks.",
+			"She is a mother of five", -- mother becomes the root, because is is not a 'real' verb
+			"No man walks in the park. He yodels.",
+			"A farmer and a donkey visit a friend.",
+			--"A farmer and a donkey live together.", -- I cannot do adverbs
+			--"If Mary is swimming or dancing, then Sue is.", -- I cannot do proper nouns
+			-- "If John beat a horse or a donkey, his wife helped it",
+			"No man walks in the park. He yodels and beats a donkey.",
+			"A farmer owns a donkey. It is brown and he beats it every day.",
+			"A farmer and entrepreneur beat a donkey.", -- I'm not sure how to handle 'his'
 			-- "The man could not make his donkey walk.", fails because SP things walk is a noun
-			"The man couldn't make his donkey see.",
+			"Some farmer couldn't make his donkey see.",
 			"If there's a farmer and a donkey, the farmer beats the donkey.",
 			"If there's a farmer who owns a donkey, the farmer beats the donkey.",
+
+			"If you like me, you will hate my cousin",
+			"Do you still beat your wife?",
 			"James won a Nobel prize", -- commitment
-			"The King of France is bold", -- unique, what if there's no king?
+			"The King of France is bald", -- unique, what if there's no king?
 			"John regrets that his dog died.", -- what is the object?
-			"When did you stop beating your wife",
-			"A man comes in. He sees a dog. He smiles." -- very sequential
[…]

### 2013-04-28 Prop.hs

@@ -49,12 +49,33 @@ fexp f e = (iterate (f.) id) !! e
 simplify :: Prop -> Prop
 simplify = fexp simp 10
 
-tostring T = "T"
-tostring F = "F"
-tostring (Not p) = "!(" ++ tostring p ++ ")"
-tostring (And p q) = tostring p ++ "&" ++ tostring q
-tostring (Or p q) = "(" ++ tostring p ++ ")|(" ++ tostring q ++ ")"
-tostring (Impl p q) = "(" ++ tostring p ++ ")->(" ++ tostring q ++ ")"
-tostring (Predi f ks) = f ++ "(" ++ concat (intersperse "," ks) ++ ")"
-tostring (Exist k p) = "E" ++ k ++ "(" ++ tostring p ++ ")"
-tostring (All k p) = "A" ++ k ++ "(" ++ tostring p ++ ")"
+-- Printing
+
+wrap :: Int -> (String, Int) -> String
+wrap n (p, k) | k > n     = "(" ++ p ++ ")"
+              | otherwise = p
+
+pretty :: Prop -> (String, Int)
+pretty T = ("⊤", 0)
+pretty F = ("⊥", 0)
+pretty (Not p) = ("¬" ++ wrap 0 (pretty p), 0)
+pretty (And p q) = (wrap 1 (pretty p) ++ "∧" ++ wrap 1 (pretty q), 1)
+pretty (Or p q) = (wrap 2 (pretty p) ++ "∨" ++ wrap 2 (pretty q), 2)
+pretty (Impl p q) = (wrap 3 (pretty p) ++ "→" ++ wrap 3 (pretty q), 3)
+pretty (Predi f ks) = (f ++ "(" ++ concat (intersperse "," ks) ++ ")", 0)
+pretty (Exist k p) = ("∃" ++ k ++ wrap 0 (pretty p), 0)
+pretty (All k p) = ("∀" ++ k ++ wrap 0 (pretty p), 0)
+
+tex :: Prop -> (String, Int)
+tex T = ("\\top", 0)
+tex F = ("\\bot", 0)
+tex (Not p) = ("\\neg " ++ wrap 0 (tex p), 0)
+tex (And p q) = (wrap 1 (tex p) ++ " \\wedge " ++ wrap 1 (tex q), 1)
+tex (Or p q) = (wrap 2 (tex p) ++ " \\ve " ++ wrap 2 (tex q), 2)
[…]

### 2013-04-28 Stanford.hs

@@ -3,7 +3,7 @@ module Stanford (Store, Word, Index, run, runDry, runOnFile, PosTree(..), postre
 import Prop (Ref)
 
 import Text.ParserCombinators.Parsec
-import Data.Map hiding (map, (\\))
+import Data.Map hiding (map, (\\), mapMaybe, null)
 
 import Data.List hiding (union, insert, lookup)
 import Text.XML.Light
@@ -24,11 +24,7 @@ type Store = Index -> Ref
 
 ------------------
 
--- Run these functions on each <sentence>
-
--- Use "lemma" instead of "word" to get standardized forms
--- (did -> do, n't -> not)
-
+-- lemmas are standardized forms of words. E.g. did -> do, n't -> not
 lemmas :: Element -> [Sentence]
 lemmas doc = map lemma (findElements (unqual "sentence") sentences)
 	where
@@ -42,7 +38,7 @@ postrees doc = map rootfilter roots
 	where
 		parses = findElements (unqual "parse") doc
 		roots = map (postreeRoot . strContent) parses
-		rootfilter (P "ROOT" [(P "S" s)]) = (P "S" s)
+		rootfilter (P "ROOT" [n]) = n
 		rootfilter root = error ("Cant parse strange sentence " ++ show root)
 
 postreeRoot :: String -> PosTree Word
@@ -69,16 +65,56 @@ posLeaf =
 		word <- many (noneOf " ()")
 		return (L tag word)
 
--- Why use an index tree if we can move the referants stright in there?
--- Because a lot of things don't have references...
 
 cleanTrees :: [PosTree Word] -> [Sentence] -> [PosTree (Word,Index)]
-cleanTrees trees ls = [clean t l s | (t, l, s) <- zip3 trees ls [0..]]
-	where clean t l s = substitute t (zip l (zip (repeat s) [0..]))
+cleanTrees trees ls = map cleanElements $ map cleanTags $ improveTrees trees ls
 
--- also simplify VB's
+improveTrees :: [PosTree Word] -> [Sentence] -> [PosTree (Word,Index)]
+improveTrees trees ls = [improve t l s | (t, l, s) <- zip3 trees ls [0..]]
+	where improve t l s = substitute t (zip l (zip (repeat s) [0..]))
 
--- Replaces every word in f
+-- simplifies certain tags with tedious information
[…]

### 2013-04-28 ToLogic.hs

@@ -4,7 +4,8 @@ import Dpl (dpl2prop, s2dpl, Type2, sT, sF, sSwitch, sScope, sPredi, sExist, sCo
 import Prop
 import Stanford (Word, Store, Index, PosTree(..))
 
-import Data.Map hiding (lookup, mapMaybe)
+import Data.List
+import Data.Map hiding (lookup, mapMaybe, map)
 import Data.Maybe
 
 type IWord = (Word, Index)
@@ -18,27 +19,53 @@ logify (refs, tree) = (simplify . dpl2prop . s2dpl) logic
 
 intRoot :: SyntaxB -> Store -> Type2
 intRoot ((P "S" s):ss) r = intS s r `sComp` intRoot ss r
+intRoot ((P "NP" np):ss) r = intNP np r (\x -> sT) `sComp` intRoot ss r
 intRoot [] r = sT
 
 intS :: SyntaxB -> Store -> Type2
+--intS [P "NP" [L "EX" _], P "VP" [L "VB" _, P "NP" np]] r = intNP np r (intVP vp r)
 intS [P "NP" np, P "VP" vp] r = intNP np r (intVP vp r)
 intS (P "SBAR" ([(L "IN" ("if",_)), (P "S" s1)]) : (L "," _) : s2) r =
 		sSwitch `sComp` intS s1 r `sComp` sSwitch `sComp` intS s2 r
-intS [P "S" s1, L "CC" ("and",_), P "S" s2] r = intS s1 r `sComp` intS s2 r
+--intS [P "CC" [P "S" s1, P "S" s2]] r = intS s1 r `sComp` intS s2 r
+intS [P "S" s1, L "CC" _, P "S" s2] r = intS s1 r `sComp` intS s2 r
 intS other r = error ("Unsuported by intS: " ++ show other)
 
 intNP :: SyntaxB -> Store -> Predicate -> Type2
 -- intNP ([L "NNP" (nnp,_)], i) r vb = sPredi nnp [v] `sComp`
-intNP [L "PRP" (prp,i)] r = \vb -> vb (r i)
-intNP [L "DT" dt, L "NN" nn] r = intDT dt r (intNN nn r)
-intNP other r = error ("Unsuported by intNP: " ++ show other)
+--intNP ([L "NN" (nnp,_)], i) r vb = sPredi nnp [v] `sComp`
+intNP [L "PRP" (prp,i)] r vb = vb (r i)
+intNP [L "DT" dt, L "NN" nn] r vb = intDT dt r (intNN nn r) vb
[…]

### 2013-04-28 Visser.hs

@@ -1,0 +1,59 @@
+import Prop (Ref, Prop(..))
+
+-- Syntax
+
+data Visser =
+    VT | VF
+  | VC Visser Visser
+  | VD Visser Visser
+  | VE Ref
+  | VP Ref [Ref]
+  | VSw
+  | VSc
+  | VQ0 Visser
+  | VQ1 Visser
+  deriving (Eq, Show)
+
+-- Interpretation
+{-
+type IVisser = (DPL, DPL, DPL, DPL, Integer, Integer)
+
+intVisser :: Visser -> IVisser
+intVisser VT = (DT, DT, DT, DT, 1, 0)
+intVisser VF = (DT, DT, DT, DF, 1, 0)
+intVisser VSw = (DT, DT, DT, DT, -1, 0)
+intVisser VSc = (DT, DT, DT, DT, 1, 1)
+intVisser (VP p rs) = (DT, DT, DT, DPredi p rs, 1, 0)
+intVisser (VE r) = (DT, DT, DT, DExists r, 1, 0)
+intVisser (VC vis1 vis2) = compVisser (intVisser vis1) (intVisser vis2)
+intVisser (VQ0 vis) = (qm1, DT, qp1, DImpl qm0 qp0, 1, 0)
+    where (qm1, qm0, qp1, qp0, a, i) = intVisser vis
+intVisser (VQ1 vis) = (DT, DT, DT, DImpl (DComp qm1 qm0) (DComp qp1 qp0), 1, 0)
+    where (qm1, qm0, qp1, qp0, a, i) = intVisser vis
+
+compVisser :: IVisser -> IVisser -> IVisser
+compVisser (qm1, qm0, qp1, qp0, 1, 0) (rm1, rm0, rp1, rp0, b, j)  = (DComp qm1 rm1, DComp qm0 rm0, DComp qp1 rp1, DComp qp0 rp0, b, j)
+compVisser (qm1, qm0, qp1, qp0, 1, 1) (rm1, rm0, rp1, rp0, b, j)  = (DComp qm1 rm0, DComp qm0 rm1, DComp qp1 rp0, DComp qp0 rp1, b, 1-j)
+compVisser (qm1, qm0, qp1, qp0, -1, 0) (rm1, rm0, rp1, rp0, b, j) = (DComp qm1 rp1, DComp qm0 rp0, DComp qp1 rm1, DComp qp0 rm0, -b, j)
[…]

### 2013-04-28 paper.tex
Works very well now, and it's clean too!

@@
 
+% Samson suggests using the word 'Computational'
+
 %Musings in Dynamic Semantics for Donkeys
@@
 \subsection{Montague semantic}
+
+\begin{align}
+&\exists a \cdot farmer(a) \rightarrow \exists g \cdot donkey(g) \cdot beat(a,g) \\
+&\exists a \cdot farmer(a) \cdot (\exists b \cdot donkey(b) \cdot own(a,b)) \rightarrow beat(a,b) \\
+&\exists a \cdot farmer(a) \cdot own(a,b) \cdot beat(a,b) \rightarrow a\_donkey(b) \\
+&\exists c \cdot farmer(c) \cdot (own(c,a) \cdot beat(c,a)) \rightarrow a\_donkey(a) \\
+&\exists a \cdot man(a) \cdot come\_in(a) \cdot (\exists i \cdot dog(i) \cdot see(a,i) \cdot smile(a)) \\
+&\exists a \cdot farmer(a) \cdot (\exists b \cdot donkey(b) \cdot (own(a,b) \cdot beat(a,b))) \\
+&\exists a \cdot donkey(a) \rightarrow \exists g \cdot horse(g) \cdot (\exists o \cdot eq\_on\_of(o,a,g) \rightarrow \exists v \cdot song(v) \cdot sing(o,v)) \\
+&\exists a \cdot farmer(a) \cdot (\exists p \cdot donkey(p) \cdot (starve(a,p) \cdot beat(a,p))) \\
+&\exists a \cdot farmer(a) \cdot (\exists g \cdot horse(g) \cdot starve(a,g) \cdot (\exists ac \cdot donkey(ac) \cdot beat(a,ac))) \\
+&\exists a \cdot man(a) \cdot walk(a) \rightarrow talk(a) \\
+&a\_mother\_of\_five(a) \\
+&\neg (\exists a \cdot man(a) \cdot walk\_in\_the\_park(a)) \cdot yodel(a) \\
+&\exists a \cdot farmer(a) \cdot (\exists g \cdot donkey(g) \cdot (\exists o \cdot eq\_on\_of(o,a,g) \rightarrow \exists v \cdot friend(v) \cdot visit(o,v))) \\
+&\neg (\exists a \cdot man(a) \cdot walk\_in\_the\_park(a)) \cdot (\exists r \cdot donkey(r) \cdot (yodel(a,r) \cdot beat(a,r))) \\
+&\exists a \cdot farmer(a) \cdot (\exists b \cdot donkey(b) \cdot own(a,b)) \cdot (brown(b) \cdot beat\_every\_day(a,b)) \\
+&\exists a \cdot (farmer(a) \cdot entrepreneur(a)) \cdot (\exists p \cdot donkey(p) \cdot beat(a,p))\end{align}
 

### 2013-04-28 Dpl.hs

@@ -1,117 +1,125 @@
-module Dpl (DPL(..), dpl2prop, Type2, s2dpl,
-	sT, sF, sSwitch, sScope, sPredi, sExist, sComp, sTest0, sTest1) where
+module Dpl (Dpl(..), tex, pretty, simplify, Assignment, intDpl, intDpl2, intDpl3) where
 
-import Prop (Ref, Prop(..))
+import Fol (Ref, Sigma, Fol(..), Entity, Domain, Predicate)
 
--- Type 1: Basic dynamic predicate logic
-----------------------------------------
+import Data.List (intersperse)
+import Utils (wrap, replace)
+import Control.Monad ((>=>))
 
-data DPL =
+-----------------------------------------------------------
+-- Syntax
+
+data Dpl =
     DT | DF
-  | DComp DPL DPL
-  | DOr DPL DPL
-  | DNot DPL
-  | DImpl DPL DPL
-  | DExists Ref
-  | DPredi Ref [Ref]
+  | DC Dpl Dpl
+  | DD Dpl Dpl
+  | DN Dpl
+  | DI Dpl Dpl
+  | DE Ref
+  | DP Ref [Ref]
   deriving (Eq, Show)
 
--- TODO: Move exec code here
 
-dpl2prop :: DPL -> Prop
-dpl2prop dpl = addDpl T [dpl]
 
-addDpl :: Prop -> [DPL] -> Prop
+
+tex :: Dpl -> String
+tex = fst . tex' where
+  tex' DT = ("\\top", 0)
+  tex' DF = ("\\bot", 0)
+  tex' (DC p q) = (wrap 1 (tex' p) ++ " \\cdot " ++ wrap 1 (tex' q), 1)
+  tex' (DD p q) = (wrap 2 (tex' p) ++ " \\vee " ++ wrap 2 (tex' q), 2)
+  tex' (DI p q) = (wrap 3 (tex' p) ++ " \\rightarrow " ++ wrap 3 (tex' q), 3)
+  tex' (DN p) = ("\\neg " ++ wrap 0 (tex' p), 0)
+  tex' (DE k) = ("\\exists " ++ k, 0)
+  tex' (DP f ks) = (replace "-" "\\_" f ++ "(" ++ concat (intersperse "," ks) ++ ")", 0)
+
+pretty :: Dpl -> String
+pretty = fst . pretty' where
+  pretty' DT = ("⊤", 0)
+  pretty' DF = ("⊥", 0)
[…]

### 2013-04-28 Dpl_old.hs

@@ -1,0 +1,117 @@
+module Dpl (DPL(..), dpl2prop, Type2, s2dpl,
+	sT, sF, sSwitch, sScope, sPredi, sExist, sComp, sTest0, sTest1) where
+
+import Prop (Ref, Prop(..))
+
+-- Type 1: Basic dynamic predicate logic
+----------------------------------------
+
+data DPL =
+    DT | DF
+  | DComp DPL DPL
+  | DOr DPL DPL
+  | DNot DPL
+  | DImpl DPL DPL
+  | DExists Ref
+  | DPredi Ref [Ref]
+  deriving (Eq, Show)
+
+-- TODO: Move exec code here
+
+dpl2prop :: DPL -> Prop
+dpl2prop dpl = addDpl T [dpl]
+
+addDpl :: Prop -> [DPL] -> Prop
+addDpl p [] = p
+addDpl p (DF:ds) = F
+addDpl p (DT:ds) = addDpl p ds
+addDpl p ((DPredi f ks):ds) = addDpl (And (Predi f ks) p) ds
+addDpl p ((DExists k):ds) = addDpl (Exist k p) ds
+addDpl p ((DComp d1 d2):ds) = addDpl p (d2:d1:ds) -- backwards: explain
+addDpl p ((DNot d):ds) = addDpl (And (Not (addDpl T [d])) p) ds
+addDpl p ((DImpl d1 d2):ds) = addDpl p ((DNot (DComp d1 (DNot d2))):ds)
+-- This rule is no good at all. Scoping becomes weird
+--addDpl p ((DOr d1 d2):ds) = addDpl (And (Or (addDpl T [d1]) (addDpl T [d2])) p) ds
+
+addDpl p ((DOr d1 d2):ds) = Or (addDpl p (d1:ds)) (addDpl p (d2:ds))
+
+
+-- If there's a farmer and a donkey, the farmer beats the donkey
+r1 = DImpl (DComp (DComp (DComp (DExists "x") (DPredi "farmer" ["x"])) (DExists "y")) (DPredi "donkey" ["y"])) (DPredi "beats" ["x", "y"])
+
+r7 = DComp (DOr (DExists "x") DF) (DPredi "farmer" ["x"])
[…]

### 2013-04-28 Fol.hs

@@ -1,0 +1,99 @@
+module Fol (Ref, Sigma, Fol(..), simplify, Entity, Domain, Predicate, pretty, tex) where
+
+import Utils (replace, wrap)
+import Data.List (intersperse)
+
+-----------------------------------------------------------
+-- Syntax
+
+type Ref = String
+type Sigma = [Ref]
+
+data Fol =
+	  T
+	| F
+	| Not   Fol
+  | And   Fol Fol
+  | Or    Fol Fol
+  | Predi Ref [Ref]
+  | Exists Ref Fol
+  | All   Ref Fol
+  | Impl  Fol Fol
+  deriving (Eq, Show)
+
+simp :: Fol -> Fol
+simp (Not (Exists k p)) = simp (All k (Not p))
+simp (Not (All k p)) = simp (Exists k (Not p))
+simp (Not (And p q)) = simp (Impl p (Not q))
+simp (Not (Impl p q)) = simp (And p (Not q))
+simp (Not (Not p)) = simp p
+simp (Not T) = F
+simp (Not F) = T
+
+simp (Impl p F) = simp (Not p)
+simp (Impl p (Impl q r)) = simp (Impl (And p q) r)
+
+simp (And p T) = simp p
+simp (And p F) = F
+
+-- Otherwise just let the call go through
+simp (And p q) = And (simp p) (simp q)
+simp (All k p) = All k (simp p)
+simp (Exists k p) = Exists k (simp p)
+simp (Not p) = Not (simp p)
+simp (Impl p q) = Impl (simp p) (simp q)
+-- We still don't have too many rules for Or, as we are quite interested in things like `foo v false`
+simp (Or p q) = Or (simp p) (simp q)
+simp p = p
+
+simplify :: Fol -> Fol
+simplify p = iterate simp p !! 10
+
+pretty :: Fol -> String
+pretty = fst . pretty' where
+  pretty' T = ("⊤", 0)
+  pretty' F = ("⊥", 0)
+  pretty' (Not p) = ("¬" ++ wrap 0 (pretty' p), 0)
+  pretty' (And p q) = (wrap 1 (pretty' p) ++ "∧" ++ wrap 1 (pretty' q), 1)
+  pretty' (Or p q) = (wrap 2 (pretty' p) ++ "∨" ++ wrap 2 (pretty' q), 2)
[…]

### 2013-04-28 Main.hs

@@ -1,8 +1,10 @@
 module Main (main) where
 
 import Stanford (run, runDry)
-import Prop (Prop, tostring)
-import ToLogic (logify)
+import ToLogic (montague)
+import qualified Visser as V
+import qualified Dpl as D
+import qualified Fol as F
 
 import Data.List
 import Text.XML.Light
@@ -10,47 +12,64 @@ import Control.Monad
 import Data.Maybe
 import System.Process
 
-super :: [String] -> IO [Prop]
---super sentences = run sentences >>= (return . map logify)
-super sentences = runDry sentences >>= (return . map logify)
+super :: [String] -> IO [V.Visser]
+super sentences = run sentences >>= (return . map montague)
+--super sentences = runDry sentences >>= (return . map montague)
+
+sentences = [
+	"Every farmer beats some donkey.",
+	"If a farmer owns a donkey, he beats it",
+	"If a farmer owns it and he beats it, it is a donkey",
+	"If a farmer owns it and beats it, it is a donkey",
+	"A man comes in. He sees a donkey. He smiles.", -- very sequential
+	"A farmer. A Donkey. He owns it. He beats it",
+	"A farmer. A Donkey. If he owns it, he beats it.",
+	"All donkeys and a horse sing a song.",
+	"A farmer starves and beats a donkey.",
+	"A farmer starves a horse and beats a donkey.",
+	"If a man walks, he talks.",
+	"She is a mother of five", -- mother becomes the root, because is is not a 'real' verb
+	"No man walks in the park. He yodels.", -- coref checker screws this up, but dpl fixes it
+	"A farmer and a donkey visit a friend.",
+	"No man walks in the park. He yodels and beats a donkey.",
+	"A farmer owns a donkey. It is brown and he beats it every day.",
+	"Some farmer and entrepreneur beats a donkey.", -- I'm not sure how to handle 'his'
+	"A man and a woman owned a donkey.",
+	"A man owned a horse and a donkey.",
+	"A man and a woman owned a horse and a donkey.",
+	"A man and a woman owned a horse and a donkey. The donkey did not walk. The man beat it.",
+	"A man and a woman owned a horse and a donkey. If the donkey did not walk, the man beat it.",
+	"A man and a woman owned a horse and a donkey. If the donkey did not walk and the man was a farmer, he beat it.",
+
+	"A farmer and his wife own a horse and a donkey.", -- posessives...
[…]

### 2013-04-28 Stanford.hs

@@ -1,6 +1,6 @@
-module Stanford (Store, Word, Index, run, runDry, runOnFile, PosTree(..), postrees) where
+module Stanford (Store, IWord, run, runDry, runOnFile, PosTree(..), postrees) where
 
-import Prop (Ref)
+import Fol (Ref)
 
 import Text.ParserCombinators.Parsec
 import Data.Map hiding (map, (\\), mapMaybe, null)
@@ -13,6 +13,7 @@ import System.Process
 
 type Index = (Int, Int)
 type Word = String
+type IWord = (Word, Index)
 type Sentence = [Word]
 
 type PosTag = String
@@ -66,10 +67,10 @@ posLeaf =
 		return (L tag word)
 
 
-cleanTrees :: [PosTree Word] -> [Sentence] -> [PosTree (Word,Index)]
+cleanTrees :: [PosTree Word] -> [Sentence] -> [PosTree IWord]
 cleanTrees trees ls = map cleanElements $ map cleanTags $ improveTrees trees ls
 
-improveTrees :: [PosTree Word] -> [Sentence] -> [PosTree (Word,Index)]
+improveTrees :: [PosTree Word] -> [Sentence] -> [PosTree IWord]
 improveTrees trees ls = [improve t l s | (t, l, s) <- zip3 trees ls [0..]]
 	where improve t l s = substitute t (zip l (zip (repeat s) [0..]))
 
@@ -185,14 +186,14 @@ alphabet = [c:"" | c <- "abcdefghijklmnopqrstuvwxyz"]
 variables :: [Ref]
 variables = alphabet ++ [b++a | b <- variables, a <- alphabet]
 
-runDry :: [String] -> IO [(Store, [PosTree (Word,Index)])]
+runDry :: [String] -> IO [(Store, [PosTree IWord])]
 runDry sentences =
 	do
 		sequence [runOnFile (name ++ ".xml") | name <- names]
 	where
 		names = ["input" ++ show i | i <- [1..length sentences]]
 
-run :: [String] -> IO [(Store, [PosTree (Word,Index)])]
+run :: [String] -> IO [(Store, [PosTree IWord])]
 run sentences =
 	do
 		sequence (zipWith writeFileLn names sentences)
@@ -205,7 +206,7 @@ run sentences =
 		names = ["input" ++ show i | i <- [1..length sentences]]
 		writeFileLn path s = writeFile path (s ++ "\n")
 
[…]

### 2013-04-28 ToLogic.hs

@@ -1,50 +1,72 @@
-module ToLogic (logify) where
+module ToLogic (montague) where
 
-import Dpl (dpl2prop, s2dpl, Type2, sT, sF, sSwitch, sScope, sPredi, sExist, sComp, sTest0, sTest1)
-import Prop
-import Stanford (Word, Store, Index, PosTree(..))
+import Fol (Ref)
+import Visser (Visser(..))
+import Stanford (IWord, Store, PosTree(..))
 
 import Data.List
 import Data.Map hiding (lookup, mapMaybe, map)
 import Data.Maybe
 
-type IWord = (Word, Index)
-type SyntaxB = [PosTree IWord]
-type Predicate = Ref -> Type2
-type Predicate2 = Ref -> Ref -> Type2
+type Predicate = Ref -> Visser
+type Predicate2 = Ref -> Ref -> Visser
 
-logify :: (Store, [PosTree IWord]) -> Prop
-logify (refs, tree) = (simplify . dpl2prop . s2dpl) logic
-	where logic = intRoot tree refs
+montague :: (Store, [PosTree IWord]) -> Visser
+montague (refs, tree) = intRoot tree refs
 
-intRoot :: SyntaxB -> Store -> Type2
-intRoot ((P "S" s):ss) r = intS s r `sComp` intRoot ss r
-intRoot ((P "NP" np):ss) r = intNP np r (\x -> sT) `sComp` intRoot ss r
-intRoot [] r = sT
+intRoot :: [PosTree IWord] -> Store -> Visser
+intRoot ((P "S" s):ss) r = VQ0 (intS s r) `VC` intRoot ss r
+intRoot ((P "NP" np):ss) r = intNP np r (\x -> VT) `VC` intRoot ss r
+intRoot [] r = VT
 
-intS :: SyntaxB -> Store -> Type2
---intS [P "NP" [L "EX" _], P "VP" [L "VB" _, P "NP" np]] r = intNP np r (intVP vp r)
+intS :: [PosTree IWord] -> Store -> Visser
 intS [P "NP" np, P "VP" vp] r = intNP np r (intVP vp r)
 intS (P "SBAR" ([(L "IN" ("if",_)), (P "S" s1)]) : (L "," _) : s2) r =
-		sSwitch `sComp` intS s1 r `sComp` sSwitch `sComp` intS s2 r
[…]

### 2013-04-28 Utils.hs

@@ -1,0 +1,15 @@
+module Utils (wrap, replace) where
+
+import Data.List (isPrefixOf)
+
+-- This is a standard subtring replacement function
+replace :: String -> String -> String -> String
+replace pat repl target
+  | target == ""            = ""
+  | pat `isPrefixOf` target = repl ++ replace pat repl (drop (length pat) target)
+  | otherwise               = head target : replace pat repl (tail target)
+
+-- This is used for pretty printing
+wrap :: Int -> (String, Int) -> String
+wrap n (p, k) | k > n     = "(" ++ p ++ ")"
+              | otherwise = p

### 2013-04-28 Visser.hs

@@ -1,4 +1,10 @@
-import Prop (Ref, Prop(..))
+module Visser (Visser(..), simplify, tex, pretty, intVisser1, intVisser2, intVisser3) where
+
+import Fol (Ref, Fol(..))
+import Dpl (Dpl(..))
+
+import Data.List (intersperse)
+import Utils (replace)
 
 -- Syntax
 
@@ -14,46 +20,93 @@ data Visser =
   | VQ1 Visser
   deriving (Eq, Show)
 
--- Interpretation
-{-
-type IVisser = (DPL, DPL, DPL, DPL, Integer, Integer)
-
-intVisser :: Visser -> IVisser
-intVisser VT = (DT, DT, DT, DT, 1, 0)
-intVisser VF = (DT, DT, DT, DF, 1, 0)
-intVisser VSw = (DT, DT, DT, DT, -1, 0)
-intVisser VSc = (DT, DT, DT, DT, 1, 1)
-intVisser (VP p rs) = (DT, DT, DT, DPredi p rs, 1, 0)
-intVisser (VE r) = (DT, DT, DT, DExists r, 1, 0)
-intVisser (VC vis1 vis2) = compVisser (intVisser vis1) (intVisser vis2)
-intVisser (VQ0 vis) = (qm1, DT, qp1, DImpl qm0 qp0, 1, 0)
-    where (qm1, qm0, qp1, qp0, a, i) = intVisser vis
-intVisser (VQ1 vis) = (DT, DT, DT, DImpl (DComp qm1 qm0) (DComp qp1 qp0), 1, 0)
-    where (qm1, qm0, qp1, qp0, a, i) = intVisser vis
-
-compVisser :: IVisser -> IVisser -> IVisser
-compVisser (qm1, qm0, qp1, qp0, 1, 0) (rm1, rm0, rp1, rp0, b, j)  = (DComp qm1 rm1, DComp qm0 rm0, DComp qp1 rp1, DComp qp0 rp0, b, j)
-compVisser (qm1, qm0, qp1, qp0, 1, 1) (rm1, rm0, rp1, rp0, b, j)  = (DComp qm1 rm0, DComp qm0 rm1, DComp qp1 rp0, DComp qp0 rp1, b, 1-j)
[…]

### 2013-04-28 Dpl.hs
Variable cleaning

@@ -28,7 +28,7 @@ tex = fst . tex' where
   tex' DF = ("\\bot", 0)
   tex' (DC p q) = (wrap 1 (tex' p) ++ " \\cdot " ++ wrap 1 (tex' q), 1)
   tex' (DD p q) = (wrap 2 (tex' p) ++ " \\vee " ++ wrap 2 (tex' q), 2)
-  tex' (DI p q) = (wrap 3 (tex' p) ++ " \\rightarrow " ++ wrap 3 (tex' q), 3)
+  tex' (DI p q) = (wrap 2 (tex' p) ++ " \\rightarrow " ++ wrap 2 (tex' q), 3)
   tex' (DN p) = ("\\neg " ++ wrap 0 (tex' p), 0)
   tex' (DE k) = ("\\exists " ++ k, 0)
   tex' (DP f ks) = (replace "-" "\\_" f ++ "(" ++ concat (intersperse "," ks) ++ ")", 0)
@@ -39,7 +39,7 @@ pretty = fst . pretty' where
   pretty' DF = ("⊥", 0)
   pretty' (DC p q) = (wrap 1 (pretty' p) ++ "⋅" ++ wrap 1 (pretty' q), 1)
   pretty' (DD p q) = (wrap 2 (pretty' p) ++ "∨" ++ wrap 2 (pretty' q), 2)
-  pretty' (DI p q) = (wrap 3 (pretty' p) ++ "→" ++ wrap 3 (pretty' q), 3)
+  pretty' (DI p q) = (wrap 2 (pretty' p) ++ "→" ++ wrap 2 (pretty' q), 3)
   pretty' (DN p) = ("¬" ++ wrap 0 (pretty' p), 0)
   pretty' (DE k) = ("∃" ++ k, 0)
   pretty' (DP f ks) = (f ++ "(" ++ concat (intersperse "," ks) ++ ")", 0)

### 2013-04-28 Fol.hs

@@ -56,7 +56,7 @@ pretty = fst . pretty' where
   pretty' (Not p) = ("¬" ++ wrap 0 (pretty' p), 0)
   pretty' (And p q) = (wrap 1 (pretty' p) ++ "∧" ++ wrap 1 (pretty' q), 1)
   pretty' (Or p q) = (wrap 2 (pretty' p) ++ "∨" ++ wrap 2 (pretty' q), 2)
-  pretty' (Impl p q) = (wrap 3 (pretty' p) ++ "→" ++ wrap 3 (pretty' q), 3)
+  pretty' (Impl p q) = (wrap 2 (pretty' p) ++ "→" ++ wrap 2 (pretty' q), 3)
   pretty' (Predi f ks) = (f ++ "(" ++ concat (intersperse "," ks) ++ ")", 0)
   pretty' (Exists k p) = ("∃" ++ k ++ wrap 0 (pretty' p), 0)
   pretty' (All k p) = ("∀" ++ k ++ wrap 0 (pretty' p), 0)
@@ -68,7 +68,7 @@ tex = fst . tex' where
   tex' (Not p) = ("\\neg " ++ wrap 0 (tex' p), 0)
   tex' (And p q) = (wrap 1 (tex' p) ++ " \\wedge " ++ wrap 1 (tex' q), 1)
   tex' (Or p q) = (wrap 2 (tex' p) ++ " \\ve " ++ wrap 2 (tex' q), 2)
-  tex' (Impl p q) = (wrap 3 (tex' p) ++ " \\rightarrow " ++ wrap 3 (tex' q), 3)
+  tex' (Impl p q) = (wrap 2 (tex' p) ++ " \\rightarrow " ++ wrap 2 (tex' q), 3)
   tex' (Predi f ks) = (replace "-" "\\_" f ++ "(" ++ concat (intersperse "," ks) ++ ")", 0)
   tex' (Exists k p) = ("\\exists " ++ k ++ wrap 0 (tex' p), 0)
   tex' (All k p) = ("\\forall " ++ k ++ wrap 0 (tex' p), 0)

### 2013-04-28 Main.hs

@@ -13,8 +13,8 @@ import Data.Maybe
 import System.Process
 
 super :: [String] -> IO [V.Visser]
-super sentences = run sentences >>= (return . map montague)
---super sentences = runDry sentences >>= (return . map montague)
+--super sentences = run sentences >>= (return . map montague)
+super sentences = runDry sentences >>= (return . map montague)
 
 sentences = [
 	"Every farmer beats some donkey.",

### 2013-04-28 Stanford.hs

@@ -1,4 +1,4 @@
-module Stanford (Store, IWord, run, runDry, runOnFile, PosTree(..), postrees) where
+module Stanford (Store, IWord, run, runDry, runOnFile, PosTree(..), postrees, variables) where
 
 import Fol (Ref)
 

### 2013-04-28 ToLogic.hs

@@ -1,7 +1,7 @@
 module ToLogic (montague) where
 
 import Fol (Ref)
-import Visser (Visser(..))
+import Visser (Visser(..), cleanVariables)
 import Stanford (IWord, Store, PosTree(..))
 
 import Data.List
@@ -12,7 +12,7 @@ type Predicate = Ref -> Visser
 type Predicate2 = Ref -> Ref -> Visser
 
 montague :: (Store, [PosTree IWord]) -> Visser
-montague (refs, tree) = intRoot tree refs
+montague (refs, tree) = cleanVariables $ intRoot tree refs
 
 intRoot :: [PosTree IWord] -> Store -> Visser
 intRoot ((P "S" s):ss) r = VQ0 (intS s r) `VC` intRoot ss r

### 2013-04-28 Visser.hs

@@ -1,9 +1,10 @@
-module Visser (Visser(..), simplify, tex, pretty, intVisser1, intVisser2, intVisser3) where
+module Visser (Visser(..), simplify, tex, pretty, cleanVariables, intVisser1, intVisser2, intVisser3) where
 
 import Fol (Ref, Fol(..))
 import Dpl (Dpl(..))
+import Stanford (variables)
 
-import Data.List (intersperse)
+import Data.List (intersperse, nub)
 import Utils (replace)
 
 -- Syntax
@@ -59,6 +60,26 @@ simp p = p
 
 simplify :: Visser -> Visser
 simplify p = iterate simp p !! 2
+
+cleanVariables :: Visser -> Visser
+cleanVariables p = foldl rename p (zip used im ++ zip im variables)
+  where
+    used = nub (listem p)
+    im = map (++"0") used
+    listem (VE k) = [k]
+    listem (VP f ks) = ks
+    listem (VC p q) = listem p ++ listem q
+    listem (VD p q) = listem p ++ listem q
+    listem (VQ0 p) = listem p
+    listem (VQ1 p) = listem p
+    listem p = []
+    rename (VE k) (x,y) = if k == x then VE y else VE k
+    rename (VP f ks) (x,y) = VP f [if k == x then y else k | k <- ks]
+    rename (VC p q) (x,y) = VC (rename p (x,y)) (rename q (x,y))
+    rename (VD p q) (x,y) = VD (rename p (x,y)) (rename q (x,y))
+    rename (VQ0 p) (x,y) = VQ0 (rename p (x,y))
+    rename (VQ1 p) (x,y) = VQ1 (rename p (x,y))
+    rename p (x,y) = p
 
 -- Interpretation
 

### 2013-04-28 Dpl_old.hs
Removed old files

@@ -1,117 +1,0 @@
-module Dpl (DPL(..), dpl2prop, Type2, s2dpl,
-	sT, sF, sSwitch, sScope, sPredi, sExist, sComp, sTest0, sTest1) where
-
-import Prop (Ref, Prop(..))
-
--- Type 1: Basic dynamic predicate logic
-----------------------------------------
-
-data DPL =
-    DT | DF
-  | DComp DPL DPL
-  | DOr DPL DPL
-  | DNot DPL
-  | DImpl DPL DPL
-  | DExists Ref
-  | DPredi Ref [Ref]
-  deriving (Eq, Show)
-
--- TODO: Move exec code here
-
-dpl2prop :: DPL -> Prop
-dpl2prop dpl = addDpl T [dpl]
-
-addDpl :: Prop -> [DPL] -> Prop
-addDpl p [] = p
-addDpl p (DF:ds) = F
-addDpl p (DT:ds) = addDpl p ds
-addDpl p ((DPredi f ks):ds) = addDpl (And (Predi f ks) p) ds
-addDpl p ((DExists k):ds) = addDpl (Exist k p) ds
-addDpl p ((DComp d1 d2):ds) = addDpl p (d2:d1:ds) -- backwards: explain
-addDpl p ((DNot d):ds) = addDpl (And (Not (addDpl T [d])) p) ds
-addDpl p ((DImpl d1 d2):ds) = addDpl p ((DNot (DComp d1 (DNot d2))):ds)
--- This rule is no good at all. Scoping becomes weird
---addDpl p ((DOr d1 d2):ds) = addDpl (And (Or (addDpl T [d1]) (addDpl T [d2])) p) ds
-
-addDpl p ((DOr d1 d2):ds) = Or (addDpl p (d1:ds)) (addDpl p (d2:ds))
-
-
--- If there's a farmer and a donkey, the farmer beats the donkey
-r1 = DImpl (DComp (DComp (DComp (DExists "x") (DPredi "farmer" ["x"])) (DExists "y")) (DPredi "donkey" ["y"])) (DPredi "beats" ["x", "y"])
-
-r7 = DComp (DOr (DExists "x") DF) (DPredi "farmer" ["x"])
-r8 = DComp (DOr (DComp (DExists "x") (DPredi "farmer" ["x"])) (DComp (DExists "x") (DPredi "donkey" ["x"]))) (DPredi "farmer" ["x"])
[…]

### 2013-04-28 Main.hs

@@ -13,8 +13,8 @@ import Data.Maybe
 import System.Process
 
 super :: [String] -> IO [V.Visser]
---super sentences = run sentences >>= (return . map montague)
-super sentences = runDry sentences >>= (return . map montague)
+super sentences = run sentences >>= (return . map montague)
+--super sentences = runDry sentences >>= (return . map montague)
 
 sentences = [
 	"Every farmer beats some donkey.",

### 2013-04-28 Prop.hs

@@ -1,81 +1,0 @@
-module Prop (Ref, Prop(..), simplify, tostring) where
-
-import Data.List
-
-type Ref = String
-
-data Prop =
-	  T
-	| F
-	| Not   Prop
-    | And   Prop Prop
-    | Or    Prop Prop
-    | Predi Ref [Ref]
-    | Exist Ref Prop
-    | All Ref Prop
-    | Impl Prop Prop
-    deriving (Eq, Show)
-
-simp :: Prop -> Prop
-simp (Not (Exist k p)) = simp (All k (Not p))
-simp (Not (All k p)) = simp (Exist k (Not p))
-simp (Not (And p q)) = simp (Impl p (Not q))
-simp (Not (Impl p q)) = simp (And p (Not q))
-simp (Not (Not p)) = simp p
-simp (Not T) = F
-simp (Not F) = T
-
-
-simp (Impl p F) = simp (Not p)
-simp (Impl p (Impl q r)) = simp (Impl (And p q) r)
-
-simp (And p T) = simp p
-simp (And p F) = F
-
-simp (And p q) = And (simp p) (simp q)
-simp (All k p) = All k (simp p)
-simp (Exist k p) = Exist k (simp p)
-simp (Not p) = Not (simp p)
-simp (Impl p q) = Impl (simp p) (simp q)
-
--- We still don't have too many rules for Or, as we are quite interested in things like `foo v false`
-simp (Or p q) = Or (simp p) (simp q)
-
-simp p = p
-
-fexp :: (a -> a) -> Int -> (a -> a)
-fexp f e = (iterate (f.) id) !! e
-
-simplify :: Prop -> Prop
-simplify = fexp simp 10
-
--- Printing
-
-wrap :: Int -> (String, Int) -> String
-wrap n (p, k) | k > n     = "(" ++ p ++ ")"
-              | otherwise = p
-
-pretty :: Prop -> (String, Int)
-pretty T = ("⊤", 0)
-pretty F = ("⊥", 0)
-pretty (Not p) = ("¬" ++ wrap 0 (pretty p), 0)
-pretty (And p q) = (wrap 1 (pretty p) ++ "∧" ++ wrap 1 (pretty q), 1)
-pretty (Or p q) = (wrap 2 (pretty p) ++ "∨" ++ wrap 2 (pretty q), 2)
[…]

### 2013-04-28 donkey.hs

@@ -1,43 +1,0 @@
--- http://stackoverflow.com/questions/3228856/predicate-logic-in-haskell
-
-import Data.List
-
-data Prop =
-	  Not   Prop
-    | And   Prop Prop
-    | Or    Prop Prop
-    | Impl  Prop Prop
-    | Equiv Prop Prop
-    | Equals Obj Obj
-    | ForAll (Obj -> Prop)
-    | Exists (Obj -> Prop)
-    deriving (Eq, Ord)
-
-data Dra =
-	  DraComp Dra Dra
-	| DraId
-	| DraBottom
-	| DraImpl Dra Dra
-	| Set
-
-class Pol p where
-	plus :: 
-
-data Pol = 1 | -1
-type Dlp = (Prop, Prop, Pol)
-
-true :: Dlp
-true = (true, true, 1)
-
-false :: Dlp
-false = (true, false, 1)
-
-switch :: Dlp
-switch = (true, true, -1)
-
-mul :: Dlp -> Dlp -> Dlp
-mul (qm, qp, 1) (rm, rp, b) = (pmul qm rm, pmul qp rp, b)
-mul (qm, qp, -1) (rm, rp, b) = (pmul qm rp, pmul qp rm, -b)
-
-test :: Dlp -> Prop
-test (qm, qp, a) = Impl qm qp

### 2013-04-28 dra.hs

@@ -1,66 +1,0 @@
-import Data.List
-
-type Var = Integer
-type Rels = Var -> Var -> Bool
-type M = [Var] -> Rels
-
--- monaden skal have constanter, værdier, variable, funktioner...
-
-create :: [(Var,Var)] -> Rels
-create pairs x y = elem (x,y) pairs
-
-result :: Rels -> M
-result r vs = r
-
-new = result . create
-
--- ($>) :: M -> (Rels -> M) -> M
--- (r $> f) vs = f (filter r (zip vs' vs')) where vs' = r vs
-
-draComp :: M -> M -> M
-draComp r s vs x y = any (\z -> r vs x z && s vs z y) vs
-
-draId :: M
-draId vs x y = x == y
-
-draBot :: M
-draBot vs x y = False
-
-impl a b = not a || b
-
-draImpl :: M -> M -> M
-draImpl r s vs x y = x == y && all (\z -> r vs x z `impl` any (\u -> s vs z u) vs) vs
-
-draNot :: M -> M
-draNot r = r `draImpl` draBot
-
-cross :: [a] -> [b] -> [(a, b)]
-cross [] ys = []
-cross (x:xs) ys = [(x, y) | y <- ys] ++ cross xs ys
-
-draDiag :: [Var] -> M
-draDiag xs vs x y = x == y && elem x xs
-
-draShow :: M -> [Var] -> [(Var,Var)]
-draShow r vs = filter (uncurry (r vs)) (cross vs vs)
-
-draDom :: M -> [Var] -> [Var]
-draDom r vs = map fst (draShow r vs)
-
-test :: [(Var,Var)] -> [Var] -> Bool
-test pairs vs =
-    draShow (draNot r) vs == draShow (draDiag (vs \\ draDom r vs)) vs
-    && draShow ((draNot.draNot) r) vs == draShow (draId `draImpl` r) vs
-    && draShow (draId `draImpl` r) vs == draShow (draDiag (draDom r vs)) vs
-    where r = (result . create) pairs
-
-testAll :: [Var] -> Bool
-testAll vs = and [test pairs vs | pairs <- subsequences (cross vs vs)]
-
--- Der må være forskel på værdier og variable
--- exists skal tage en variabel, og assigne den til alle forskellige værdier
--- indtil det, den bliver kædet med, virker.
--- Så faktisk skal der være noget effekt at et element, nemlig at det kan ændre
--- assigmenten på variablerne
--- Kræver måske renaming...
-exists 

### 2013-04-28 xml.hs

@@ -1,113 +1,0 @@
-import Text.XML.Light
-import Control.Monad
-import System.IO.Unsafe
-import Data.List
-import Data.Maybe
---main = interact parse
--- stanford-corenlp-full-2012-11-12/input.txt.xml
-parse xml = show doc where Just doc = parseXMLDoc xml
--- readFile "stanford-corenlp-full-2012-11-12/input.txt.xml" >>= 
-
-type Sentence = [String]
-
-data DepTree = Dep Int [(String, DepTree)] deriving Show
-data Coref = Reprs Int Int Int | Mention Int Int Int
-
-deps :: Element -> String -> [DepTree]
-deps doc name = map (buildTree deps) roots
-	where
-		deps = listDeps doc name
-		lefts = nub [gov | (typ, gov, dep) <- deps]
-		rights = nub [dep | (typ, gov, dep) <- deps]
-		roots = lefts \\ rights
-
-buildTree :: [(String, Int, Int)] -> Int -> DepTree
-buildTree deps root = Dep root [(name, buildTree deps dep) | (name, gov, dep) <- deps, gov == root]
-
-listDeps :: Element -> String -> [(String, Int, Int)]
-listDeps doc name = map parseDep deps
-	where
-		Just dep = findElement (unqual name) doc
-		deps = findElements (unqual "dep") dep
-
-parseDep :: Element -> (String, Int, Int)
-parseDep e = (typ, read gov, read dep)
-	where
-		Just typ = findAttr (unqual "type") e
-		Just gov = findChild (unqual "governor") e >>= (findAttr (unqual "idx"))
-		Just dep = findChild (unqual "dependent") e >>= (findAttr (unqual "idx"))
-
- 
-
---super :: Sentence -> Element -> Scop
-
--- let s = (unsafePerformIO . readFile) "stanford-corenlp-full-2012-11-12/input.txt.xml"
--- let Just doc = parseXMLDoc s
-
--- listDeps doc "basic-dependencies"
--- [("nn",2,1),("nsubj",4,2),("cop",4,3),("prep",4,5),("pobj",5,6)]
-{-
-Stanford University is located in California
-Dep 4 [
-	("nsubj", Dep 2 [
-				("nn", Dep 1 [])]),
-	("cop", Dep 3 []),
-	("prep", Dep 5 [
-		("pobj", Dep 6 [])])]
--}
-
[…]

### 2013-04-28 src/Stanford.hs
cleaned up directory structure

@@ -56,7 +56,7 @@ posEither =
 posTree =
 	do
 		tag <- many (noneOf " ")
-		char ' ' -- maybe not needed?
+		char ' '
 		subs <- sepBy1 posEither (char ' ')
 		return (P tag subs)
 posLeaf =
@@ -176,34 +176,35 @@ allRefs doc vs = searcher
 	      searcher i = findWithDefault (vs' !! diagonal i) i comap
 	      diagonal (j,k) = (j+k)*(j+k+1)`div`2 + j
 
-
--- TODO: To help us avoid running out of variables, we might want to
---       flatten the map once in a while
-
 --------------------
 
 alphabet = [c:"" | c <- "abcdefghijklmnopqrstuvwxyz"]
+
 variables :: [Ref]
 variables = alphabet ++ [b++a | b <- variables, a <- alphabet]
+
+filenames = ["/tmp/donkey_input" ++ show i | i <- [0..]]
 
 runDry :: [String] -> IO [(Store, [PosTree IWord])]
 runDry sentences =
 	do
 		sequence [runOnFile (name ++ ".xml") | name <- names]
 	where
-		names = ["input" ++ show i | i <- [1..length sentences]]
+		names = take (length sentences) filenames
 
 run :: [String] -> IO [(Store, [PosTree IWord])]
 run sentences =
 	do
 		sequence (zipWith writeFileLn names sentences)
-		writeFileLn "inputlist" (intercalate "\n" names)
-		(_, _, _, h) <- createProcess (shell "stanford-corenlp-full-2012-11-12/corenlp.sh -filelist inputlist")
+		writeFileLn listname (intercalate "\n" names)
+		(_, _, _, h) <- createProcess (shell (stanfordpath++" -outputDirectory /tmp -filelist "++listname))
 		exitCode <- waitForProcess h
 		putStrLn ("Exit code: " ++ show exitCode)
 		sequence [runOnFile (name ++ ".xml") | name <- names]
 	where
-		names = ["input" ++ show i | i <- [1..length sentences]]
+		stanfordpath = "../stanford-corenlp-full-2012-11-12/corenlp.sh"
+		listname = "/tmp/donkey_inputlist"
+		names = take (length sentences) filenames
 		writeFileLn path s = writeFile path (s ++ "\n")
 
 runOnFile :: FilePath -> IO (Store, [PosTree IWord])

### 2013-04-28 src/Dpl.hs
Cleaned up files.

@@ -56,7 +56,9 @@ simp (DC DF p) = DF
 -- simp (DD p DF) = simp p
 -- simp (DD DF p) = simp p
 simp (DI p DF) = DN p
--- simp (DI DT p) = simp p -- this is incorrect, as they lhs creates scope
+-- scope related
+simp (DI DT (DI DT p)) = DT `DI` simp p
+simp (DN (DN p)) = DT `DI` simp p
 -- Fall throughs
 simp (DC p q) = simp p `DC` simp q
 simp (DD p q) = simp p `DD` simp q

### 2013-04-28 src/Main.hs

@@ -13,50 +13,91 @@ import Data.Maybe
 import System.Process
 
 super :: [String] -> IO [V.Visser]
-super sentences = run sentences >>= (return . map montague)
---super sentences = runDry sentences >>= (return . map montague)
+--super sentences = run sentences >>= (return . map montague)
+super sentences = runDry sentences >>= (return . map montague)
 
 sentences = [
 	"Every farmer beats some donkey.",
 	"If a farmer owns a donkey, he beats it",
 	"If a farmer owns it and he beats it, it is a donkey",
-	"If a farmer owns it and beats it, it is a donkey",
-	"A man comes in. He sees a donkey. He smiles.", -- very sequential
-	"A farmer. A Donkey. He owns it. He beats it",
-	"A farmer. A Donkey. If he owns it, he beats it.",
-	"All donkeys and a horse sing a song.",
-	"A farmer starves and beats a donkey.",
-	"A farmer starves a horse and beats a donkey.",
-	"If a man walks, he talks.",
-	"She is a mother of five", -- mother becomes the root, because is is not a 'real' verb
-	"No man walks in the park. He yodels.", -- coref checker screws this up, but dpl fixes it
-	"A farmer and a donkey visit a friend.",
-	"No man walks in the park. He yodels and beats a donkey.",
-	"A farmer owns a donkey. It is brown and he beats it every day.",
-	"Some farmer and entrepreneur beats a donkey.", -- I'm not sure how to handle 'his'
-	"A man and a woman owned a donkey.",
-	"A man owned a horse and a donkey.",
-	"A man and a woman owned a horse and a donkey.",
-	"A man and a woman owned a horse and a donkey. The donkey did not walk. The man beat it.",
-	"A man and a woman owned a horse and a donkey. If the donkey did not walk, the man beat it.",
-	"A man and a woman owned a horse and a donkey. If the donkey did not walk and the man was a farmer, he beat it.",
+	"If a farmer owns a donkey, he beat it. If he rewards it, he doesn't own it.",
 
-	"A farmer and his wife own a horse and a donkey.", -- posessives...
+	-- these two become a bit weird due to the scoping hack for 'a'
+	"If a farmer owns a donkey, he owns a horse. If he doesn't own it, he owns a cow.",
+	"Every dog sees a cat. It chases it.",
+
+	-- sequentials
+	"A man comes in. He sees a donkey. He smiles.", -- very sequential
[…]

### 2013-04-29 run.sh
Added simple run script Fixed problem with variables introduced in the 'then' clause of ifs

@@ -1,0 +1,4 @@
+#!/bin/bash
+cd "$(dirname $0)/src"
+runghc Main.hs
+

### 2013-04-29 src/Main.hs

@@ -19,12 +19,12 @@ super sentences = runDry sentences >>= (return . map montague)
 sentences = [
 	"Every farmer beats some donkey.",
 	"If a farmer owns a donkey, he beats it",
-	"If a farmer owns it and he beats it, it is a donkey",
 	"If a farmer owns a donkey, he beat it. If he rewards it, he doesn't own it.",
+	"A dog barks. If it is beaten, it whines.", -- a donkey brays, but Stanford doesn't like it
 
 	-- these two become a bit weird due to the scoping hack for 'a'
 	"If a farmer owns a donkey, he owns a horse. If he doesn't own it, he owns a cow.",
-	"Every dog sees a cat. It chases it.",
+	"If a woman is American, she loves a soldier. If she is Dutch, she loves a bike-rider.",
 
 	-- sequentials
 	"A man comes in. He sees a donkey. He smiles.", -- very sequential
@@ -34,16 +34,16 @@ sentences = [
 	"A farmer starves a horse and beats a donkey.", -- double vp
 	"All donkeys and a horse sing a song.", -- double np
 	"Some farmer and entrepreneur beats a donkey.", -- double description
+	"A man and a woman owned a horse and a donkey. If the donkey did not walk and the man was a farmer, he beat it.", -- cross product
 
 	-- free variables
+	"If a farmer owns it and he beats it, it is a donkey",
 	"She is a mother of five", -- mother becomes the root, because is is not a 'real' verb
 	"If he owns it, he beats it.", -- double free variable
 	"No man walks in the park. He yodels.", -- coref checker screws this up, but dpl fixes it
 
-	"A farmer owns a donkey. It is brown and he beats it every day.", -- temporal
-	"If a woman is American, she loves a soldier. If she is Dutch, she loves a bike-rider.",
-	"A man and a woman owned a horse and a donkey. If the donkey did not walk and the man was a farmer, he beat it.", -- cross product
-	"A dog barks. If it is beaten, it whines.",
+	-- stringifying
+	"A farmer owns a donkey. It is brown and he beats it every day.",
 
 ------------- I Can do until this line
 
@@ -57,7 +57,7 @@ sentences = [
 	"If Mary is swimming or dancing, then Sue is.",
 	"If John beat a horse or a donkey, his wife helped it",
 	"He was quite angry. John. He was MAD.",
-	"The King of France is bald", -- commitment
+	"The King of France is bald",
 
[…]

### 2013-04-29 src/ToLogic.hs

@@ -15,14 +15,14 @@ montague :: (Store, [PosTree IWord]) -> Visser
 montague (refs, tree) = cleanVariables $ intRoot tree refs
 
 intRoot :: [PosTree IWord] -> Store -> Visser
-intRoot ((P "S" s):ss) r = VQ0 (intS s r) `VC` intRoot ss r
+intRoot ((P "S" s):ss) r = intS s r `VC` intRoot ss r
 intRoot ((P "NP" np):ss) r = intNP np r (\x -> VT) `VC` intRoot ss r
 intRoot [] r = VT
 
 intS :: [PosTree IWord] -> Store -> Visser
 intS [P "NP" np, P "VP" vp] r = intNP np r (intVP vp r)
 intS (P "SBAR" ([(L "IN" ("if",_)), (P "S" s1)]) : (L "," _) : s2) r =
-		VSw `VC` intS s1 r `VC` VSw `VC` intS s2 r
+		VQ0 (VSw `VC` intS s1 r `VC` VSw `VC` VQ1 (intS s2 r))
 intS [P "S" s1, L "CC" ("and",_), P "S" s2] r = intS s1 r `VC` intS s2 r
 intS other r = error ("Unsuported by intS: " ++ show other)
 
@@ -31,15 +31,13 @@ intNP [L "PRP" (prp,i)] r vb = vb (r i)
 intNP [L "DT" dt, L "NN" nn] r vb = intDT dt r (intNN nn r) vb
 intNP [L "DT" dt, L "NN" nn1, L "CC" ("and",_), L "NN" nn2] r vb = intDT dt r (\x -> intNN nn1 r x `VC` intNN nn2 r x) vb
 intNP [P "NP" np1, L "CC" ("and",_), P "NP" np2] r vb = intNP np1 r (\x -> intNP np2 r (inner x))
-	where
-		inner = \x y -> VQ0 (VSw `VC` VE (x++y) `VC` VP "eq-on-of" [(x++y),x,y] `VC` VSw `VC` vb (x++y))
-		--z = x++y -- just some free variable
+	where inner x y = VQ0 (VSw `VC` VE (x++y) `VC` VP "eq-on-of" [(x++y),x,y] `VC` VSw `VC` vb (x++y)) -- hopefully x+y is free
 intNP other r vb = error ("Unsuported by intNP: " ++ show other)
 
 intVP :: [PosTree IWord] -> Store -> Predicate
[…]

### 2013-04-30 paper.tex
Wrote about montague grammars

--- 2013-04-28 paper.tex
@@
 \documentclass[12pt]{article}
+\usepackage{booktabs}
 
@@
 \subsection{Montague semantic}
-
-$\exists a(farmer(a) \wedge \exists g(donkey(g) \wedge \forall o(o=a\vee o=g \rightarrow \exists v(friend(v) \wedge visit(o,v)))))$
-
-Montague (1970b): `I reject the contention that an important
-theoretical difference exists between formal and natural
-languages.'
-
-To us, the revolutionary idea in Montague's PTQ paper (and
-earlier papers) is the claim that natural language is not impossibly
-incoherent, as his teacher Tarski had led us to believe, but that
-large portions of its semantics can be treated by combining known
-tools from logic, tools like functions of finite type, the lambda-calculus,
-generalized quantifiers, tense and modal logic, and all the rest.
-
-Montague grammars are an approach to combinatorial semantics based on types and lambda calculus. It is interesting because it 
-Types
-Quantification
-
-`...
-we are convinced that the capacities of [Montague Grammar] have not been exploited to the limit, that sometimes an analysis is carried out in a rival framework simply because it is more fashionable' - Groendijk \& Stokhof in `Dynamic Montague Grammar'
-
-Montague can make any logic compositional
-
-Types of:
-  One-place predicate constant (farmer, sleeps, walks): Entity -> Bool
-  Transitive verb (owns, beats): Entity -> Entity -> Bool
-  Attributive adjective (good, intelligent, former): (Entity -> Bool) -> (Entity -> Bool)
-  Noun phrase: (Entity -> Bool) -> Bool
-  Proper noun: Entity
-  Quantifiers/Determiners: a, the, every, no: (Entity -> Bool) -> (Entity -> Bool) -> Bool
-  Sentence: Bool
+Montague semantics is a theory of natural language that believes its semantics can be described thoroughly in terms of standard mathematical models.
+Logician Richard Montague (1930-1971) famously wrote
+\begin{quotation}
+There is in my opinion no important theoretical difference between natural languages and the artificial languages of logicians; indeed I consider it possible to comprehend the syntax and semantics of both kinds of languages with a single natural and mathematically precise theory.
+\cite{montague1970universal}
+\end{quotation}
+
+He believed that just as Chomsky had formulated the syntax of language in a mathematical framework, so too should it be possible to define the meaning of every sentence, based on the meaning of its parts and a set of compositional rules.
+
+While the above remains a hot topic in the philosophy of language, we are going to take a more practical stand.
+In our quest to automate the translation of language into Visser's monadic DPL, it is interesting that Montague was able to define such a large fragment of English using familiar tools of compositionality and lambda calculus.
+
+The basic idea is very simple.
+For each postag we define a type.
+E.g.
+$S$ has the type $Bool$ and $VP$ has the type $Entity \rightarrow Bool$, the intuition being that the VP returns true if the entity `does what the VP described' and false otherwise.
+An NP is given type type $VP \rightarrow Bool$ which allows it to create a scope for a variable and pass it to the VP, or whatever it sees fitting for that NP.
+
+The point is that we always give a specific phrase type the same logical type. This allows our composition rules to know exactly what type each component has. Here are some more examples:
+\begin{itemize}
+\item Proper noun (John, Oxford): Entity
+\item One-place predicate constant (farmer, sleeps, walks): $Entity \rightarrow Bool$
+\item Transitive verb (owns, beats): $Entity \rightarrow Entity \rightarrow Bool$
+\item Attributive adjective (good, intelligent, former): $(Entity \rightarrow Bool) \rightarrow (Entity \rightarrow Bool)$
+\item Quantifiers/Determiners (a, the, every, no): $(Entity \rightarrow Bool) \rightarrow (Entity \rightarrow Bool) \rightarrow Bool$
+\end{itemize}
+
+To translate a sentence like `Every farmer beats a donkey' we first create a parse tree [$_\text{S}$ [$_\text{NP}$ [$_\text{DT}$ Every] [$_\text{NN}$ farmer]] [$_\text{VP}$ [$_\text{VBZ}$ beats] [$_\text{NP}$ [$_\text{DT}$ a] [$_\text{NN}$ donkey]]]].
+
+Next we create objects for the basic elements (a computer would do this recursively, but it is easier for us to do it level by level):
+\begin{equation*}\begin{aligned}
+&\text{Every, $r_1$}&=& \lambda xy(\Bowtie\cdot\triangle\cdot\exists r_1\cdot (xr_1)\cdot\triangle\cdot\Bowtie\cdot (yr_1)) & :: (E\rightarrow B)\rightarrow(E\rightarrow B)\rightarrow B\\
+&\text{farmer}&=& \lambda x(farmer(x)) & :: E\rightarrow B\\
+&\text{beats}&=& \lambda xy(beats(xy)) & :: E\rightarrow E\rightarrow B\\
+&\text{a, $r_2$}&=& \lambda xy(\triangle\cdot\exists r_2\cdot (xr_2)\cdot\triangle\cdot (yr_2)) & :: (E\rightarrow B)\rightarrow(E\rightarrow B)\rightarrow B\\
+&\text{donkey}&=& \lambda x(donkey(x)) & :: E\rightarrow B\\
+\midrule
+&\text{Every farmer}&=& \lambda y(\Bowtie\cdot\triangle\cdot\exists r_1\cdot farmer(r_1)\cdot\triangle\cdot\Bowtie\cdot (yr_1)) & :: (E\rightarrow B)\rightarrow B\\
+&\text{a donkey}&=& \lambda y(\triangle\cdot\exists r_2\cdot donkey(r_2)\cdot\triangle\cdot (yr_2)) & :: (E\rightarrow B)\rightarrow B\\
+\midrule
+&\text{beats a donkey}&=& \lambda z(\triangle\cdot\exists r_2\cdot donkey(r)\cdot\triangle\cdot beats(z,r_2)) & :: E\rightarrow B\\
+\end{aligned}\end{equation*}
+
+Until we can then finally compose the top NP and VP to get the sentence:
+\begin{equation*}\begin{aligned}
+&\Bowtie\cdot\triangle\cdot\exists r_1\cdot farmer(r_1)\cdot\triangle\cdot\Bowtie\cdot \triangle\cdot\exists r_2\cdot donkey(r_2)\cdot\triangle\cdot beats(r_1,r_2)\\
+=\text{ }&\Bowtie\cdot\triangle\cdot\exists r_1\cdot farmer(r_1)\cdot\triangle\cdot\Bowtie\cdot beats(r_1,r_2)\cdot \triangle\cdot\exists r_2\cdot donkey(r_2)\cdot\triangle
+\end{aligned}\end{equation*}
+
+You might have noticed we didn't discuss how the variables $r_1$ and $r_2$ were assigned.
+This is of course of utter importance in relation to anaphora resolution.
+However in the section \ref{sec:coreferences} we will discuss a way out of this problem.
+
+Later in the implementation section we will look at more issues regarding to using Montague Grammars in practise. First we need to look at a very important source of ambiguity in sentences.
 
@@
 
-Let's return multiple interpretations
-
-% It is not possible
-It is clear that not all sentences can be unambiguously translated from natural language to logic. A sentence like  
+In the process of defining semantics for language, ambiguity in sentenses is of course a serious problem. A sentence like  
 \begin{quotation}
@@
 Every farmer owns a donkey. It is pink. Second sentence decides the reading.
+
+\subsection{Coreferences}\label{sec:coreferences}
+Coreferences (or anaphora) is essential in order to create advanced discourse, since it is what allows us maintain a topic beyond the first mention.
+Without coreferences we couldn't even express simple logical statements such as transitivity: `a man's brother's brother is his brother' etc.
+
+Identifying coreferences (or resolving anaphora) is hence an important part of converting natural language into logic.
+The issue is not touched upon in Visser's paper, which casually sidesteps the problem by assigning variables to all mentioned entities.
+
+For the subtask of resolving simple pronouns a solution could be to work in terms of subjects, objects, indirect objects and so on.
+However if not earlier, this certainly fails for pronouns separated from their antecedent by sentence boundaries.
+Also we want to resolve other kinds of coreferences, such as `John, the farmer, owns a donkey', where `the farmer' and `John' refers to the same entity.
+
+It quickly becomes clear that the problem is not only syntactical.
+In the sentence `John and his wife had a farm.
+He took care of the donkeys' we could substitute `He' with `She' and to change the referent of the pronoun.
+In this case we need semantic knowledge of genders to proceed.
+
+In my project I have taken advantage of the coreference tagger built into Stanford CoreNLP\cite{lee2013deterministic}\cite{lee2011stanford}\cite{raghunathan2010multi}.
+In the following section I will summarize the workings of this machine, made to work across large texts as well as simple sentences.
+
+Coreference tagging is one part of computational linguistics where machine learning algorithms still haven't been able to outperform large iterative applications of linguistic heuristics.
+Of course the procedure assumes (most likely) machine learned postagging and named entity recognition, but the main algorithm is transparent and deterministic.
+
+The idea is to iteratively go over the text with different `sieves'. The below example explains each step with an example I have adopted from \cite{lee2013deterministic}:
+
+\begin{quotation}
+John is a farmer. He owned a stubborn donkey.
+A girl was looking at the donkey.
+``It was my favorite,'' John said to her.
+\end{quotation}
+
+\begin{description}
+\item[Mention detection]
+The first step is to identify all noun phrases (NP) and personal (PRP) and possessive (PRP\$) pronouns. This information is easily pulled from the postree. Notice we also identify nested mentions:
+
+[John]$_1^1$ is [a farmer]$_2^2$. [He]$_3^3$ owned [a stubborn donkey]$_4^4$.\newline
+[A girl]$_5^5$ was looking at [the donkey]$_6^6$.\newline
+``[It]$_7^7$ was [[my]$_9^9$ favorite]$_8^8$,'' [John]$_{10}^{10}$ said to [her]$_{11}^{11}$.
+
+All mentions are initially assigned to separate entities (entity ids).
+The entities are also attached information we can easily pull from the postree, such as `a girl: {number: singular, gender: female}' which are used for the following rules.
+\item[Speaker Sieve]
+The purpose of this first sieve is to link self referring pronouns to speakers.
+Speakers are identified simply by their connection to verbs such as `say' and proximity to the quotations.
+In our case the pronoun `my' gets linked to `John':
+
+[John]$_1^1$ is [a farmer]$_2^2$. [He]$_3^3$ owned [a stubborn donkey]$_4^4$.\newline
+[A girl]$_5^5$ was looking at [the donkey]$_6^6$.\newline
+``[It]$_7^7$ was [\textbf{[my]$_9^1$} favorite]$_8^8$,'' \textbf{[John]$_{10}^{9}$} said to [her]$_{11}^{11}$.
+
+In conversational text we would store information about the speakers in our entities. Thus we could e.g. match mentions of `you' to the speaker of the previous quote we saw.
+
+We also note restrictions for future use, so we don't risk coreferencing an `I' with a `he' etc.
+\item[String Match]
+The second step is to merge entities referred to by exactly the same string. In our case we have two mentions of `John':
+
+\textbf{[John]$_1^1$} is [a farmer]$_2^2$. [He]$_3^3$ owned [a stubborn donkey]$_4^4$.\newline
+[A girl]$_5^5$ was looking at [the donkey]$_6^6$.\newline
+``[It]$_7^7$ was [[my]$_9^1$ favorite]$_8^8$,'' \textbf{[John]$_{10}^{1}$} said to [her]$_{11}^{11}$.
+\item[Relaxed String Match]
+This third sieve doesn't get activated by our example.
+It will try to remove relative clauses and other text following the head word of a mention, and do an exact match of the result.
+For example [John] and [John, whose donkey loves him] will correctly be merged.
+
+Like the `String Match' sieve, we are not doing anything fancy here, and we will incorrectly coreference the two mentions of `John' in: `[John, who loves donkeys] and [John, who hates donkeys] were brothers.'
+\item[Precise Constructs]
+The `Precise Constructs' sieve uses a lot of common patterns that link mentions. In our case two `X is Y' patterns are found:
+
+\textbf{[John]$_1^1$} is \textbf{[a farmer]$_2^1$}. [He]$_3^3$ owned [a stubborn donkey]$_4^4$.\newline
+[A girl]$_5^5$ was looking at [the donkey]$_6^6$.\newline
+``\textbf{[It]$_7^7$} was \textbf{[}[my]$_9^1$ \textbf{favorite]$_8^7$},'' [John]$_{10}^1$ said to [her]$_{11}^{11}$.
+
+Other patterns include acronyms and appositives.
+For example [Canadian Donkey \& Mule Association] and [CDMA] are linked, because the second is tagged as an acronym and it matches the upper case letters of the first one.
+An appositive is usually a pattern `X, Y, ...' where Y is another description of X.
+\item[Strict Head Match A, B, C] are simply rules strip mentions down to their head word and try to find exact matches.
+This is similar to `Relaxed String Match', but more radical.
+In our case we correctly get a link between `a stubborn donkey' and `the donkey':
+
+[John]$_1^1$ is [a farmer]$_2^1$. [He]$_3^3$ owned \textbf{[a stubborn donkey]$_4^4$}.\newline
+[A girl]$_5^5$ was looking at \textbf{[the donkey]$_6^4$}.\newline
+``[It]$_7^7$ was [[my]$_9^1$ favorite]$_8^7$,'' [John]$_{10}^1$ said to [her]$_{11}^{11}$.
+
+Linking mentions based on their head word is dangerous.
+e.g.
+`University of Oxford' and `University of Cambridge' both have `University' as their head word, but are (obviously?) different entities.
+To combat this a lot of strict rules must be observed.
+These are however gradually weakened in sieve B and C.
+\item[Proper Head Noun Match] could be called `Strict Head Match D'.
+It is the weakest form of the sieve, as it has only the three constraints: two mentions of the same entity cannot have one included in the other, they cannot contain words referring to conflicting geographical locations, and they cannot have different numbers.
+e.g.
+[donkeys] and [at least 5 donkeys].
+\item[Pronoun Match]
+This last sieve is perhaps the most important one.
+It is done in the end so that we may have as much data in our entities as possible for matching.
+Using guesses on gender, number and animacy we can make our final merges:
+
+\textbf{[John]$_1^1$} is [a farmer]$_2^1$. \textbf{[He]$_3^1$} owned [a stubborn donkey]$_4^4$.\newline
+\textbf{[A girl]$_5^5$} was looking at \textbf{[the donkey]$_6^4$}.\newline
+``\textbf{[It]$_7^4$} was [[my]$_9^1$ favorite]$_8^4$,'' [John]$_{10}^1$ said to \textbf{[her]$_{11}^{5}$}.
+
+As always there are a few constraints put on what can be merged. One rule is that the distance between pronoun and antecedent must be maximum 3 sentences.
+\end{description}
+
+After the final step different post processing options are usually performed. For example singleton mentions are often discarded. That would mean we got rid of `a musician' and `my favorite'.
+
+The reason the rules are applied in the order above is to give the most certain rules the highest priority.
+
+The Stanford procedure above is currently the strongest competitor in various coreference competitions.
 
@@
 
-\subsection{Coreferences}
-Coreferences (or anaphora) is essential in order to create advanced discourse, since it is what allows us maintain a topic beyond the first mention.
-Without coreferences we couldn't even express simple logical statements such as transitivity: `a man's brother's brother is his brother' etc.
-
-Identifying coreferences (or resolving anaphora) is hence an important part of converting natural language into logic.
-The issue is not touched upon in Visser's paper, which casually sidesteps the problem by assigning variables to all mentioned entities.
-
-For the subtask of resolving simple pronouns a solution could be to work in terms of subjects, objects, indirect objects and so on.
-However if not earlier, this certainly fails for pronouns separated from their antecedent by sentence boundaries.
-Also we want to resolve other kinds of coreferences, such as `John, the farmer, owns a donkey', where `the farmer' and `John' refers to the same entity.
-
-It quickly becomes clear that the problem is not only syntactical.
-In the sentence `John and his wife had a farm.
-He took care of the donkeys' we could substitute `He' with `She' and to change the referent of the pronoun.
-In this case we need semantic knowledge of genders to proceed.
-
-In my project I have taken advantage of the coreference tagger built into Stanford CoreNLP\cite{lee2013deterministic}\cite{lee2011stanford}\cite{raghunathan2010multi}.
-In the following section I will summarize the workings of this machine, made to work across large texts as well as simple sentences.
-
-Coreference tagging is one part of computational linguistics where machine learning algorithms still haven't been able to outperform large iterative applications of linguistic heuristics.
-Of course the procedure assumes (most likely) machine learned postagging and named entity recognition, but the main algorithm is transparent and deterministic.
-
-The idea is to iteratively go over the text with different `sieves'. The below example explains each step with an example I have adopted from \cite{lee2013deterministic}:
-
+\subsubsection{Semantics via dependencies}
+It is an interesting idea to consider that we might be able to define our semantics over a dependency graph, in a similar way to the Montague approach described earlier.
+The dependency graph adds a lot of flexibility where the phrase structure tree used in the Montague approach is limited to the direct syntax of the tree.
+A good example is our favourite sentence `if a farmer owns a donkey, he beats it', where there isn't a very strong syntactical relationship between `farmer' and `he'.
+It's also been suggested that it is mostly due to luck that a lot of English fits relatively well to the Montague approach, where as in some languages the structure of the phrase tree might not say much about the actual semantics.
+
+I've worked hard on trying to define a Montague like set of equations for Enlish, but over the dependency graph.
+This has however failed, mostly due to one simple reason: In a Montague grammar we expect every type of phrase to have a specific logical type.
+E.g.
+all NPs are `VP -> Bool' and VPs are `Entity -> Bool'.
+This makes it fit very nice in a programming context, where we can define a function `parseNP :: PhraseTree -> (VP -> Bool)'.With the dependency approach we don't have this option.
+A `subj' arrow can mean many different things.
+Sometimes it points at a noun, sometimes at a proper noun, sometimes something else.
+
+Another problem worth considering is the fact that all information in the graph is drawn syntactically from the phrase tree.
+That is, the graph doesn't contain any information not in the tree, but the other direction is entirely possible.
+While it is hard to say exactly what's missing, it is worth noting.
+
+\section{Implementation (8)}
+
+Montague grammars have been used for many years in the quest of machine translation.
+The idea was that to use logic as a `universal language' to which we could translate any sentence, and ascend back into whatever language we'd like.
+Hence you would think that the web was flooded with code and long machine readable lists of rules, but this is however not the case.
+
+Perhaps the lack of usable material is due to the fact that most computer scientists abandoned the formal approach two decades ago for instead to try their luck with machine learning and vector space models.
+These models are however starting to show their weaknesses, specially in very logical sentences, and people are starting to wonder if there are still things to be learned from Montague's approach.
 \begin{quotation}
-John is a farmer. He owned a stubborn donkey.
-A girl was looking at the donkey.
-``It was my favorite,'' John said to her.
+`...
+we are convinced that the capacities of [Montague Grammar] have not been exploited to the limit, that sometimes an analysis is carried out in a rival framework simply because it is more fashionable' - Groendijk \& Stokhof in `Dynamic Montague Grammar'\cite{groenendijk1990dynamic}
 \end{quotation}
 
-\begin{description}
-\item[Mention detection]
-The first step is to identify all noun phrases (NP) and personal (PRP) and possessive (PRP\$) pronouns. This information is easily pulled from the postree. Notice we also identify nested mentions:
-
-[John]$_1^1$ is [a farmer]$_2^2$. [He]$_3^3$ owned [a stubborn donkey]$_4^4$.\newline
-[A girl]$_5^5$ was looking at [the donkey]$_6^6$.\newline
-``[It]$_7^7$ was [[my]$_9^9$ favorite]$_8^8$,'' [John]$_{10}^{10}$ said to [her]$_{11}^{11}$.
-
-All mentions are initially assigned to separate entities (entity ids).
-The entities are also attached information we can easily pull from the postree, such as `a girl: {number: singular, gender: female}' which are used for the following rules.
-\item[Speaker Sieve]
-The purpose of this first sieve is to link self referring pronouns to speakers.
-Speakers are identified simply by their connection to verbs such as `say' and proximity to the quotations.
-In our case the pronoun `my' gets linked to `John':
-
-[John]$_1^1$ is [a farmer]$_2^2$. [He]$_3^3$ owned [a stubborn donkey]$_4^4$.\newline
-[A girl]$_5^5$ was looking at [the donkey]$_6^6$.\newline
-``[It]$_7^7$ was [\textbf{[my]$_9^1$} favorite]$_8^8$,'' \textbf{[John]$_{10}^{9}$} said to [her]$_{11}^{11}$.
-
-In conversational text we would store information about the speakers in our entities. Thus we could e.g. match mentions of `you' to the speaker of the previous quote we saw.
-
-We also note restrictions for future use, so we don't risk coreferencing an `I' with a `he' etc.
-\item[String Match]
-The second step is to merge entities referred to by exactly the same string. In our case we have two mentions of `John':
-
-\textbf{[John]$_1^1$} is [a farmer]$_2^2$. [He]$_3^3$ owned [a stubborn donkey]$_4^4$.\newline
-[A girl]$_5^5$ was looking at [the donkey]$_6^6$.\newline
-``[It]$_7^7$ was [[my]$_9^1$ favorite]$_8^8$,'' \textbf{[John]$_{10}^{1}$} said to [her]$_{11}^{11}$.
-\item[Relaxed String Match]
-This third sieve doesn't get activated by our example.
-It will try to remove relative clauses and other text following the head word of a mention, and do an exact match of the result.
-For example [John] and [John, whose donkey loves him] will correctly be merged.
-
-Like the `String Match' sieve, we are not doing anything fancy here, and we will incorrectly coreference the two mentions of `John' in: `[John, who loves donkeys] and [John, who hates donkeys] were brothers.'
-\item[Precise Constructs]
-The `Precise Constructs' sieve uses a lot of common patterns that link mentions. In our case two `X is Y' patterns are found:
-
-\textbf{[John]$_1^1$} is \textbf{[a farmer]$_2^1$}. [He]$_3^3$ owned [a stubborn donkey]$_4^4$.\newline
-[A girl]$_5^5$ was looking at [the donkey]$_6^6$.\newline
-``\textbf{[It]$_7^7$} was \textbf{[}[my]$_9^1$ \textbf{favorite]$_8^7$},'' [John]$_{10}^1$ said to [her]$_{11}^{11}$.
-
-Other patterns include acronyms and appositives.
-For example [Canadian Donkey \& Mule Association] and [CDMA] are linked, because the second is tagged as an acronym and it matches the upper case letters of the first one.
-An appositive is usually a pattern `X, Y, ...' where Y is another description of X.
-\item[Strict Head Match A, B, C] are simply rules strip mentions down to their head word and try to find exact matches.
-This is similar to `Relaxed String Match', but more radical.
-In our case we correctly get a link between `a stubborn donkey' and `the donkey':
-
-[John]$_1^1$ is [a farmer]$_2^1$. [He]$_3^3$ owned \textbf{[a stubborn donkey]$_4^4$}.\newline
-[A girl]$_5^5$ was looking at \textbf{[the donkey]$_6^4$}.\newline
-``[It]$_7^7$ was [[my]$_9^1$ favorite]$_8^7$,'' [John]$_{10}^1$ said to [her]$_{11}^{11}$.
-
-Linking mentions based on their head word is dangerous.
-e.g.
-`University of Oxford' and `University of Cambridge' both have `University' as their head word, but are (obviously?) different entities.
-To combat this a lot of strict rules must be observed.
-These are however gradually weakened in sieve B and C.
-\item[Proper Head Noun Match] could be called `Strict Head Match D'.
-It is the weakest form of the sieve, as it has only the three constraints: two mentions of the same entity cannot have one included in the other, they cannot contain words referring to conflicting geographical locations, and they cannot have different numbers.
-e.g.
-[donkeys] and [at least 5 donkeys].
-\item[Pronoun Match]
-This last sieve is perhaps the most important one.
-It is done in the end so that we may have as much data in our entities as possible for matching.
-Using guesses on gender, number and animacy we can make our final merges:
-
-\textbf{[John]$_1^1$} is [a farmer]$_2^1$. \textbf{[He]$_3^1$} owned [a stubborn donkey]$_4^4$.\newline
-\textbf{[A girl]$_5^5$} was looking at \textbf{[the donkey]$_6^4$}.\newline
-``\textbf{[It]$_7^4$} was [[my]$_9^1$ favorite]$_8^4$,'' [John]$_{10}^1$ said to \textbf{[her]$_{11}^{5}$}.
-
-As always there are a few constraints put on what can be merged. One rule is that the distance between pronoun and antecedent must be maximum 3 sentences.
-\end{description}
-
-After the final step different post processing options are usually performed. For example singleton mentions are often discarded. That would mean we got rid of `a musician' and `my favorite'.
-
-The reason the rules are applied in the order above is to give the most certain rules the highest priority.
-
-The Stanford procedure above is currently the strongest competitor in various coreference competitions.
-
-\subsection{Semantics via dependencies}
-This subsection's content...
-
-\section{Implementation (8)}
+Nonetheless the lack of machine tested composition rules means that nearly all of the rules I shall present are of my own creation.
+Because these rules are tested in practice on a large set of English sentences, they hopefully avoid some mistakes and oversights that I have found in purely theoretical papers.
+It seams like the correct approach to start with a small working fragment, and slowly add more phrase rules and sentences.
+The opposite of this approach is what is often observed when logicians and philosophers go astray in very abstract corners of English, without having covered the basic ground.
 
@@
 
+The free variable thing is a problem in the sense that we have decided the meaning of a sentence is a truth value. Maybe we should change it to assignment -> Truth...
+
 Something that's very interesting is when to use free variables and when to use bound variables.
@@
 
+The sentences are all simplified through a number of syntactical identities defined for each language. The purpose of this is only to improve readability.
+
 \begin{enumerate}
 \item
-Some description?
-\begin{description}[style=multiline, leftmargin=10.25em]
-  \item[Natural language] If a farmer who owns a donkey, he beats it 
-  \item[Dependency Graph] ... eh
-  \item[Monadic DPL] $\Bowtie \cdot \triangle \cdot \exists x \cdot farmer(x) \cdot \triangle \cdot owns(x,y) \cdot \triangle \cdot \exists y \cdot donkey(y) \cdot \triangle \cdot \Bowtie \cdot beats(x,y)$
-  \item[DPL] $\exists x \cdot farmer(x) \cdot \exists y \cdot donkey(y) \cdot owns(x,y) \rightarrow beats(x,y)$
-  \item[FOL] $\forall x ( farmer(x) \rightarrow \forall y ( donkey(y) \wedge owns(x,y) \rightarrow beats(x,y)))$
-\end{description}
-
-\item The second item
+Let's start up with a very sequential example to showcase the strength of DPL. Notice how the sentences are easily concattenated with no need for scope handling.
+\begin{align*}
+&\text{\textbf{English}}&&\text{A man comes in. He sees a donkey. He smiles.} \\
+&\text{\textbf{Visser}}&&\triangle \cdot \exists a \cdot man(a) \cdot \triangle \cdot come\_in(a) \cdot \triangle \cdot \exists b \cdot donkey(b) \cdot \triangle \cdot see(a,b) \cdot smile(a) \\
+&\text{\textbf{DPL}}&&\exists a \cdot man(a) \cdot \exists b \cdot donkey(b) \cdot come\_in(a) \cdot see(a,b) \cdot smile(a) \\
+&\text{\textbf{FOL}}&&\exists a(man(a) \wedge \exists b(donkey(b) \wedge come\_in(a) \wedge see(a,b) \wedge smile(a))) \\
+\end{align*}
 \item
-`John wants to marry a Spanish girl'
-
-This one is hard because it is an example of existential non commitment. I bet we'll fail it.
+Now let's try a real donkey sentence. Notice how in this case the program chooses the weak reading of `some donkey'.
+\begin{align*}
+&\text{\textbf{English}}&&\text{Every farmer beats some donkey.} \\
+&\text{\textbf{Visser}}&&\Bowtie \cdot \triangle \cdot \exists a \cdot farmer(a) \cdot \triangle \cdot \Bowtie \cdot \triangle \cdot \exists b \cdot donkey(b) \cdot \triangle \cdot beat(a,b) \\
+&\text{\textbf{DPL}}&&\exists a \cdot farmer(a) \rightarrow \exists b \cdot donkey(b) \cdot beat(a,b) \\
+&\text{\textbf{FOL}}&&\forall a(farmer(a) \rightarrow \exists b(donkey(b) \wedge beat(a,b))) \\
+\end{align*}
+\item
+We can also handle cases where `a farmer' is introduced inside the anticedent of an `if'.
+\begin{align*}
+&\text{\textbf{English}}&&\text{If a farmer owns a donkey, he beats it} \\
+&\text{\textbf{Visser}}&&\Bowtie \cdot \triangle \cdot \exists a \cdot farmer(a) \cdot \exists b \cdot donkey(b) \cdot \triangle \cdot own(a,b) \cdot \Bowtie \cdot beat(a,b) \\
+&\text{\textbf{DPL}}&&\exists a \cdot farmer(a) \cdot \exists b \cdot donkey(b) \cdot own(a,b) \rightarrow beat(a,b) \\
+&\text{\textbf{FOL}}&&\forall a(farmer(a) \rightarrow \forall b(donkey(b) \wedge own(a,b) \rightarrow beat(a,b))) \\
+\end{align*}
+\item
+This is an example from Visser \cite{visser1999donkey} containing sentences with no verbs.
+Notice how the Visser logic has to introduce $?_0$ to prevent the antecedent from becoming part of the consequent.
+\begin{align*}
+&\text{\textbf{English}}&&\text{A farmer. A Donkey. If he owns it, it is beaten.} \\
+&\text{\textbf{Visser}}&&\triangle \cdot \exists a \cdot farmer(a) \cdot \exists b \cdot donkey(b) \cdot \triangle \cdot ?_0(\Bowtie \cdot own(a,b) \cdot \Bowtie \cdot beat(b)) \\
+&\text{\textbf{DPL}}&&\exists a \cdot farmer(a) \cdot \exists b \cdot donkey(b) \cdot (own(a,b) \rightarrow beat(b)) \\
+&\text{\textbf{FOL}}&&\exists a(farmer(a) \wedge \exists b(donkey(b) \wedge (own(a,b) \rightarrow beat(b)))) \\
+\end{align*}
+$?_0$ however only scopes off the consequent `channel' of the monoid. Hence this more complicated example also works as expected.
+\begin{align*}
+&\text{\textbf{English}}&&\text{If a farmer owns a donkey, he beats it.}\\
+&&&\text{If he rewards it, he doesn't own it.} \\
+&\text{\textbf{Visser}}&&?_0(\Bowtie \cdot \triangle \cdot \exists a \cdot farmer(a) \cdot \exists b \cdot donkey(b) \cdot \triangle \cdot own(a,b) \cdot \Bowtie \cdot beat(a,b))\\&&& \cdot ?_0(\Bowtie \cdot reward(a,b) \cdot \Bowtie \cdot ?_0(\Bowtie \cdot own(a,b) \cdot \Bowtie \cdot \bot)) \\
+&\text{\textbf{DPL}}&&\exists a \cdot farmer(a) \cdot \exists b \cdot donkey(b) \rightarrow\\&&& (own(a,b) \rightarrow beat(a,b))\cdot (reward(a,b) \rightarrow \neg own(a,b)) \\
+&\text{\textbf{FOL}}&&\forall a(farmer(a) \rightarrow \forall b(donkey(b) \rightarrow\\&&& (own(a,b) \rightarrow beat(a,b))\wedge (reward(a,b) \rightarrow \neg own(a,b)))) \\
+\end{align*}
+\item
+A big consequence of/decision in the grammar created, is that no new entities can be introduced in the `then part' of an `if'.
+The reason for this is that such objects are not guarenteed to exist on the ground of the sentence.
+We enforce this rule using the $?_1$ function.
+\begin{align*}
+&\text{\textbf{English}}&&\text{If a farmer owns a donkey, he owns a horse.}\\&&&\text{If he doesn't own it, he owns a cow.} \\
+&\text{\textbf{Visser}}&&?_0(\Bowtie \cdot \triangle \cdot \exists a \cdot farmer(a) \cdot \exists b \cdot donkey(b) \cdot \triangle \cdot own(a,b) \cdot \Bowtie\\&&&\cdot ?_1(\triangle \cdot \exists c \cdot horse(c) \cdot \triangle \cdot own(a,c))) \cdot\\&&& ?_0(\Bowtie\cdot ?_0(\Bowtie \cdot own(a,b) \cdot \Bowtie \cdot \bot) \cdot \Bowtie \cdot ?_1(\triangle \cdot \exists d \cdot cow(d) \cdot \triangle \cdot own(a,d))) \\
+&\text{\textbf{DPL}}&&\exists a \cdot farmer(a) \cdot \exists b \cdot donkey(b) \rightarrow (own(a,b) \rightarrow \exists c \cdot horse(c) \cdot own(a,c))\\&&& \cdot (\neg own(a,b) \rightarrow \exists d \cdot cow(d) \cdot own(a,d)) \\
+&\text{\textbf{FOL}}&&\forall a(farmer(a) \rightarrow \forall b(donkey(b) \rightarrow (own(a,b) \rightarrow \exists c(horse(c) \wedge own(a,c)))\\&&&\wedge (\neg own(a,b) \rightarrow \exists d(cow(d) \wedge own(a,d))))) \\
+\end{align*}
+\item
+Because of the doubtful semantic meaning of disjunctive sentences discussed, I have instead chosen to focus on conjunctive problems.
+And there are a lot of those too.
+The first one is a case of a conjunctive VP.
+Notice that we treat `starves' as a transitive verb, even though the sentence is clearly ambigious. In this case the phrase parser makes the decision for us.
+\begin{align*}
+&\text{\textbf{English}}&&\text{A farmer starves a horse and beats a donkey.} \\
+&\text{\textbf{Visser}}&&\triangle \cdot \exists a \cdot farmer(a) \cdot  \exists b \cdot horse(b) \cdot \triangle \cdot starve(a,b)\\&&&\cdot \triangle \cdot \exists c \cdot donkey(c) \cdot \triangle \cdot beat(a,c) \\
+&\text{\textbf{DPL}}&&\exists a \cdot farmer(a) \cdot \exists b \cdot horse(b) \cdot \exists c \cdot donkey(c) \cdot starve(a,b) \cdot beat(a,c) \\
+&\text{\textbf{FOL}}&&\exists a(farmer(a) \wedge \exists b(horse(b) \wedge \exists c(donkey(c) \wedge starve(a,b) \wedge beat(a,c)))) \\
+\end{align*}
+\item
+In Montague's PTQ\cite{montague1973proper} he doesn't give rules for conjunction of NPs, but only sentences and VP's.
+In \cite{rooth1982conjunction} and \cite{partee1983generalized} Mats Rooth and Barbara Partee discuss different issues arrising, but never reach any usable construction.
+Personally I've found the below to be very useful for conjunctive NPs that act as a union.
+\begin{align*}
+&\text{\textbf{English}}&&\text{All donkeys and a horse sing a song.} \\
+&\text{\textbf{Visser}}&&\Bowtie \cdot \triangle \cdot \exists a \cdot donkey(a) \cdot \triangle \cdot \Bowtie \cdot \triangle \cdot \exists b \cdot horse(b) \cdot \triangle\\&&& \cdot ?_0(\Bowtie \cdot \exists c \cdot eq\_on\_of(c,a,b) \cdot \Bowtie \cdot \triangle \cdot \exists d \cdot song(d) \cdot \triangle \cdot sing(c,d)) \\
+&\text{\textbf{DPL}}&&\exists a \cdot donkey(a) \rightarrow \exists b \cdot horse(b) \cdot \exists d \cdot song(d)\\&&& \cdot (\exists c \cdot eq\_on\_of(c,a,b) \rightarrow sing(c,d)) \\
+&\text{\textbf{FOL}}&&\forall a(donkey(a) \rightarrow \exists b(horse(b) \wedge \exists d(song(d)\\&&& \wedge \forall c(eq\_on\_of(c,a,b) \rightarrow sing(c,d))))) \\
+\end{align*}
+\item
+This is a more simple example of a conjunctive NN.
+\begin{align*}
+&\text{\textbf{English}}&&\text{Some farmer and entrepreneur owns a donkey.} \\
+&\text{\textbf{Visser}}&&\triangle \cdot \exists a \cdot farmer(a) \cdot entrepreneur(a) \cdot  \exists b \cdot donkey(b) \cdot \triangle \cdot own(a,b) \\
+&\text{\textbf{DPL}}&&\exists a \cdot farmer(a) \cdot entrepreneur(a) \cdot \exists b \cdot donkey(b) \cdot own(a,b) \\
+&\text{\textbf{FOL}}&&\exists a(farmer(a) \wedge entrepreneur(a) \wedge \exists b(donkey(b) \wedge own(a,b))) \\
+\end{align*}
+\item
+The great thing about the construct introduced for conjunctive NPs is that it also works well when we need a cross product of subjects and objects, as this example shows.
+\begin{align*}
+&\text{\textbf{English}}&&\text{A man and a woman owned a horse and a donkey.}\\&&&\text{If the donkey did not walk, and the man was a farmer, he beat it.} \\
+&\text{\textbf{Visser}}&&\triangle \cdot \exists a \cdot man(a) \cdot  \exists b \cdot woman(b) \cdot \triangle \cdot ?_0(\Bowtie \cdot \exists c \cdot eq\_on\_of(c,a,b) \cdot \Bowtie\\&&&\cdot \triangle \cdot \exists d \cdot horse(d) \cdot  \exists e \cdot donkey(e) \cdot \triangle \cdot ?_0(\Bowtie \cdot \exists f \cdot eq\_on\_of(f,d,e) \cdot \Bowtie\\&&&\cdot own(c,f))) \cdot ?_0(\Bowtie \cdot ?_0(\Bowtie \cdot walk(e) \cdot \Bowtie \cdot \bot) \cdot a\_farmer(a) \cdot \Bowtie \cdot ?_1(beat(a,e))) \\
+&\text{\textbf{DPL}}&&\exists a \cdot man(a) \cdot \exists b \cdot woman(b) \cdot \exists d \cdot horse(d) \cdot \exists e \cdot donkey(e)\\&&&\cdot (\exists c \cdot eq\_on\_of(c,a,b) \cdot \exists f \cdot eq\_on\_of(f,d,e) \rightarrow own(c,f))\\&&&\cdot (\neg walk(e) \cdot a\_farmer(a) \rightarrow beat(a,e)) \\
+&\text{\textbf{FOL}}&&\exists a(man(a) \wedge \exists b(woman(b) \wedge \exists d(horse(d) \wedge \exists e(donkey(e)\\&&&\wedge \forall c(eq\_on\_of(c,a,b) \rightarrow \forall f(eq\_on\_of(f,d,e) \rightarrow own(c,f)))\\&&&\wedge (\neg walk(e) \wedge a\_farmer(a) \rightarrow beat(a,e)))))) \\
+\end{align*}
+\item
+Finally it is interesting to see how the program handles pronouns.
+I've chosen to make `free pronouns' not be an error, but a feature.
+Free pronouns are treated as free variables, which allow you to create sentences that might be use e.g.
+to filter a database.
+\begin{align*}
+&\text{\textbf{English}}&&\text{She is a mother of five}\hspace{15em} \\
+&\text{\textbf{Visser}}&&a\_mother\_of\_five(a) \\
+&\text{\textbf{DPL}}&&a\_mother\_of\_five(a) \\
+&\text{\textbf{FOL}}&&a\_mother\_of\_five(a) \\
+\end{align*}
+\item
+The following sentence is very similar to the original donkey sentences, but it features a free variable.
+\begin{align*}
+&\text{\textbf{English}}&&\text{If a farmer owns it and he beats it, it is a donkey} \\
+&\text{\textbf{Visser}}&&\Bowtie \cdot \triangle \cdot \exists a \cdot farmer(a) \cdot \triangle \cdot own(a,b) \cdot beat(a,b) \cdot \Bowtie \cdot a\_donkey(b) \\
+&\text{\textbf{DPL}}&&\exists a \cdot farmer(a) \rightarrow (own(a,b) \cdot beat(a,b) \rightarrow a\_donkey(b)) \\
+&\text{\textbf{FOL}}&&\forall a(farmer(a) \wedge own(a,b) \wedge beat(a,b) \rightarrow a\_donkey(b)) \\
+\end{align*}
+\item
+A final example that really shows how DPL scope and free variables all play along together is this the case of a `non existent' antecedent.
+The stanford coref tagger mistakingly tags the two men in this sentence as the same, but because negation creates a closed scope, the final sentence still comes out right.
+\begin{align*}
+&\text{\textbf{English}}&&\text{No man walks in the park. He yodels.} \\
+&\text{\textbf{Visser}}&&?_0(\Bowtie \cdot \exists a \cdot man(a) \cdot walk\_in\_the\_park(a) \cdot \Bowtie \cdot \bot) \cdot yodel(a) \\
+&\text{\textbf{DPL}}&&\neg (\exists a \cdot man(a) \cdot walk\_in\_the\_park(a)) \cdot yodel(a) \\
+&\text{\textbf{FOL}}&&\forall a(man(a) \rightarrow \neg walk\_in\_the\_park(a)) \wedge yodel(a) \\
+\end{align*}
+
 \end{enumerate}
@@
 
-\lstinputlisting[language=Haskell]{Main.hs}
+%$\lstinputlisting[language=Haskell]{src/Main.hs}
 

### 2013-04-30 src/Main.hs

@@ -19,31 +19,30 @@ super sentences = runDry sentences >>= (return . map montague)
 sentences = [
 	"Every farmer beats some donkey.",
 	"If a farmer owns a donkey, he beats it",
-	"If a farmer owns a donkey, he beat it. If he rewards it, he doesn't own it.",
-	"A dog barks. If it is beaten, it whines.", -- a donkey brays, but Stanford doesn't like it
-
-	-- these two become a bit weird due to the scoping hack for 'a'
-	"If a farmer owns a donkey, he owns a horse. If he doesn't own it, he owns a cow.",
-	"If a woman is American, she loves a soldier. If she is Dutch, she loves a bike-rider.",
 
 	-- sequentials
 	"A man comes in. He sees a donkey. He smiles.", -- very sequential
 	"A farmer. A Donkey. If he owns it, he beats it.", -- sequential
 
+	-- ifs
+	"A dog barks. If it is beaten, it whines.", -- a donkey brays, but Stanford doesn't like it
+	"If a farmer owns a donkey, he beats it. If he rewards it, he doesn't own it.",
+	"If a farmer owns a donkey, he owns a horse. If he doesn't own it, he owns a cow.",
+	"If a woman is American, she loves a soldier. If she is Dutch, she loves a bike-rider.",
+
+	-- stringifying
+	"A farmer owns a donkey. It is brown and he beats it every day.",
+
 	-- doubles
 	"A farmer starves a horse and beats a donkey.", -- double vp
 	"All donkeys and a horse sing a song.", -- double np
 	"Some farmer and entrepreneur beats a donkey.", -- double description
-	"A man and a woman owned a horse and a donkey. If the donkey did not walk and the man was a farmer, he beat it.", -- cross product
+	"A man and a woman owned a horse and a donkey. If the donkey did not walk, and the man was a farmer, he beat it.", -- cross product
 
 	-- free variables
-	"If a farmer owns it and he beats it, it is a donkey",
 	"She is a mother of five", -- mother becomes the root, because is is not a 'real' verb
-	"If he owns it, he beats it.", -- double free variable
+	"If a farmer owns it and he beats it, it is a donkey",
 	"No man walks in the park. He yodels.", -- coref checker screws this up, but dpl fixes it
-
-	-- stringifying
-	"A farmer owns a donkey. It is brown and he beats it every day.",
 
 ------------- I Can do until this line
 

### 2013-04-30 src/ToLogic.hs

@@ -24,6 +24,7 @@ intS [P "NP" np, P "VP" vp] r = intNP np r (intVP vp r)
 intS (P "SBAR" ([(L "IN" ("if",_)), (P "S" s1)]) : (L "," _) : s2) r =
 		VQ0 (VSw `VC` intS s1 r `VC` VSw `VC` VQ1 (intS s2 r))
 intS [P "S" s1, L "CC" ("and",_), P "S" s2] r = intS s1 r `VC` intS s2 r
+intS [P "S" s1, L "," _, L "CC" ("and",_), P "S" s2] r = intS s1 r `VC` intS s2 r
 intS other r = error ("Unsuported by intS: " ++ show other)
 
 intNP :: [PosTree IWord] -> Store -> Predicate -> Visser

### 2013-04-30 src/Visser.hs

@@ -81,6 +81,8 @@ cleanVariables p = foldl rename p (zip used im ++ zip im variables)
     rename (VQ1 p) (x,y) = VQ1 (rename p (x,y))
     rename p (x,y) = p
 
+
+
 -- Interpretation
 
 type IVisser1 = (Dpl, Dpl, Integer)

### 2013-05-06 paper.tex
More work and a license

@@
 
+\usepackage{stmaryrd}
 \usepackage[utf8]{inputenc}
@@
 %Dynamic Discourse and Donkeys: Using Monoids for Language Semantics in Practice
-
-\title{Dynamic Discourse and Donkeys: Using Monoids for Language Semantics in Practice}
+%Dynamic Semantics Computationally with the Donkey Monoid
+%Computational Donkey Monoids: Dynamic Semantics in Practice
+%Computation with Donkey Monoids: Dynamic Semantics in Practice
+%Computational Semantics with Donkey Monoids
+%Montague Grammar for Donkey Monoids
+%From Donkeys to Monoids: A Computational Dynamic Grammar
+%From Donkeys to Monoids: A Dynamic Semantics Framework
+%From Donkeys to Monoids: A Computational Approach to Dynamic Semantics
+%From Donkeys to Monoids: A Computational Experiment with Dynamic Semantics
+%From Donkeys to Monoids: A Computational Framework for Dynamic Semantics
+
+\title{From Donkeys to Monoids: A Computational Framework for Dynamic Semantics}
 \author{Thomas Dybdahl Ahle}
-\date{March 9, 2013}
+\date{May 30, 2013}
 
@@
 
-A donkey sentence is a sentence with a pronoun that is bound in semantics, but not in syntax.
-This make them difficult to capture in the scope of ordinary logic.
-In this paper we explore some of the ways that have been proposed to tackle this problem, in particular Albert Visser's e-monoid extension of Dynamic Predicate Logic, which allows us to write formulas deceivingly similarly looking to natural language.
-
-We explain those concepts in depth, deriving or proving every result stated.
-We then start looking into implementing the logic on a computer, experimenting with different ways of defining the semantics.
-We further explore state of the art tools for dependency parsing and coreference resolution, the workings of which we also explain in depth.
-
-Finally we put it all together in the Haskell programming language.
-We test out our machinery on a range of different sentences, showing all the converting steps from natural language to propositional logic.
-The problem is hard and the result is far from perfect, but it shows some interesting and less explored possibilities in computational linguistics.
+Donkey Monoids are a set of variations of Dynamic Predicate Logic (DPL) introduced by Albert Visser.
+The rationale behind them is a stride towards a logic that can more easily approximate natural language in aspects such as polarity and scope.
+
+Visser didn't test his ideas computationally and neither did most people involved with DPL and logification of natural language in general.
+In this paper I supply such an experiment, combining technologies introduced over the last four decades, creating a fully functioning `English to logic' framework.
+
+The transformation rules of the supported English fragment are defined as a Montague Grammar, the semantics of which is discussed and defined in the first part of the paper.
+I apply the framework to a large collection of sentences, showing the conversion process step by step.
+
+Finally I discuss ways to support a larger fragment of English, how we can combat and interprete the inherent ambiguity, and how the flexibility of Donkey Monoids compare with the classical Montague lambda structures.
 
 \end{abstract}
+
+\smallskip\smallskip
+\noindent \small{\textbf{Keywords:} Montague Grammar, Dynamic Predicate Logic, Natural Language, Meaning, Donkey, Monoid, Computational, Haskell}
 
@@
 There are many reasons why it is interesting to represent information in a logical language.
-For a computer scientist, the biggest advantage is the computational aspect.
+For a computer scientist, the biggest advantage is that we can compute with logic.
 There is a horde of algorithms for doing querying, question answering, reasoning, unification and resolution.
-All are they at your disposal if only you can obtain a first order logic form.
-
-Getting computers to understand and work with natural language is arguably on of the biggest challenges in information technologies this day.
-Hence if we could move the problem into the space of formal logic, it would be a fantastic accomplishment.
-Unfortunately many people have pointed out, that classical logic doesn't have the power to express most of the sentences people casually express on a daily basis.
-
-To draw up this point, let's first look at a sentence that easily lets itself convert and then a couple that don't:
-\begin{itemize}
+They key to this powerhouse is obtaining a first order logic form of your information.
+
+\begin{description}
+\item[Knowledge bases] In the field of Knowledge Representation and Reasoning, they long ago discovered that formal logic provides a good language for buliding ontologies.
+Typical users are crime fighters or life scientists that need to draw conclusions based on complex systems\cite{malone2010modeling} where most of the known facts are only described in natural language.
+Today those people use mainly have to manually build large models in gui tools, but an automated approach is much wanted.
+
+\item[Semantic search] From the early days of database systems logic was identified\cite{fagin1995reasoning} as a good language for querying.
+Today technologies like Facebook Graph Search allows users to execute searches like `Restaurants in London my friends have been to'.
+The missing piece is a reliable translation system.
+
+\item[Robots] Since the beginning of Artificial Intelligence, logic based approaches have done well.\cite{russell2010artificial} Robots need to store relationships they learn about the world and react to it.
+For modern `robots' like apples Siri to be able to communicate with people, a translation layer between natural language and logic is a plausible path.
+\end{description}
+
+HVAD MED AT SKRIVE LIDT MERE OM HVAD VI VIL HER??
+
+It would seem getting computers to understand and work with natural language is on of the biggest challenges in information technologies this day.
+Moving the problems into the space of formal logic is a classical and tempting path.
+Unfortunately, as many people have pointed out, classical logic simply doesn't have the power to express most of the sentences people casually express every day.
+
+We won't let us self get too down over this, but it is worth it to outline where the typical problems arrise, and to get a feeling of what a formal logic translation of natural language actually looks like.
+We'll start out with a sentence that easily lets itself translate and then a couple that don't:
+\begin{enumerate}
 \item `Socrates is a donkey': $donkey(Socrates)$.
 This is a classical logical example, treating Socrates as a constant.
-If we want to do without constants we can write: $\forall x(Socrates(x)\rightarrow donkey(x))$, though this does allow more than one Socrates to exist.
+If we want to do without constants we can write: $\forall x(Socrates(x)\rightarrow donkey(x))$.
+Notice this allows for more than one Socrates to exist.
 \item `All his donkeys are treated well': This sentence somewhat willingly let's itself translate into: $\forall y(his(x,y)\wedge donkey(y) \rightarrow treated\_well(y))$, which has a free variable $x$ referring to some person from earlier in the discourse.
@@
 It is hard to imagine a language with strong logical foundations, where those three sentences wouldn't be equal.
-\end{itemize}
+\end{enumerate}
 
@@
 %
-This is arguably the formula that first comes to mind for a moderately trained logician, or simply what you expect by looking at the structure of the syntax tree.
+This is arguably the formula that first comes to mind for a moderately trained logician, or simply what you expect by looking at the structure of the constituent tree.
 Translating an indeterminate `a farmer' into $\exists x(farmer(x))$ seems correct, but clearly the formula has problems with scope.
@@
 
+TODO: Requirements?
+
 \section{Formal Stuff (10)}
@@
 
+We won't actually use `or' in the implementation.
+Example of the problems we still have in the logic.
+Can't tell what variables are available where.
+
 A major problem with our use of `relation union' as disjunction is the way it allows seemingly sound sentences to combine into syntactically invalid ones.
@@
 \subsection{Monadic variants}
-This might be what I call DRA underneath
+
+Let's call them `Visser variants' or `VDPL'. Hm, I like that.
+
+What should we call those? Visser calls them DPL variants, e.g. DPL$_{\Bowtie\triangle}$
+
 In the previous subsection we saw that DPL allowed us to rewrite the donkey sentence as
@@
 
-To translate a sentence like `Every farmer beats a donkey' we first create a parse tree [$_\text{S}$ [$_\text{NP}$ [$_\text{DT}$ Every] [$_\text{NN}$ farmer]] [$_\text{VP}$ [$_\text{VBZ}$ beats] [$_\text{NP}$ [$_\text{DT}$ a] [$_\text{NN}$ donkey]]]].
+To translate a sentence like `Every farmer beats a donkey' we first create a constituent tree [$_\text{S}$ [$_\text{NP}$ [$_\text{DT}$ Every] [$_\text{NN}$ farmer]] [$_\text{VP}$ [$_\text{VBZ}$ beats] [$_\text{NP}$ [$_\text{DT}$ a] [$_\text{NN}$ donkey]]]].
 
@@
 \begin{equation*}\begin{aligned}
-&\text{Every, $r_1$}&=& \lambda xy(\Bowtie\cdot\triangle\cdot\exists r_1\cdot (xr_1)\cdot\triangle\cdot\Bowtie\cdot (yr_1)) & :: (E\rightarrow B)\rightarrow(E\rightarrow B)\rightarrow B\\
-&\text{farmer}&=& \lambda x(farmer(x)) & :: E\rightarrow B\\
-&\text{beats}&=& \lambda xy(beats(xy)) & :: E\rightarrow E\rightarrow B\\
-&\text{a, $r_2$}&=& \lambda xy(\triangle\cdot\exists r_2\cdot (xr_2)\cdot\triangle\cdot (yr_2)) & :: (E\rightarrow B)\rightarrow(E\rightarrow B)\rightarrow B\\
-&\text{donkey}&=& \lambda x(donkey(x)) & :: E\rightarrow B\\
+&\text{Every, $r_1$}=&& \lambda xy(\Bowtie\cdot\triangle\cdot\exists r_1\cdot (xr_1)\cdot\triangle\cdot\Bowtie\cdot (yr_1)) &&:: (E\rightarrow B)\rightarrow(E\rightarrow B)\rightarrow B\\
+&\text{farmer}=&& \lambda x(farmer(x)) &&:: E\rightarrow B\\
+&\text{beats}=&& \lambda xy(beats(xy)) &&:: E\rightarrow E\rightarrow B\\
+&\text{a, $r_2$}=&& \lambda xy(\triangle\cdot\exists r_2\cdot (xr_2)\cdot\triangle\cdot (yr_2)) &&:: (E\rightarrow B)\rightarrow(E\rightarrow B)\rightarrow B\\
+&\text{donkey}=&& \lambda x(donkey(x)) &&:: E\rightarrow B\\
 \midrule
-&\text{Every farmer}&=& \lambda y(\Bowtie\cdot\triangle\cdot\exists r_1\cdot farmer(r_1)\cdot\triangle\cdot\Bowtie\cdot (yr_1)) & :: (E\rightarrow B)\rightarrow B\\
-&\text{a donkey}&=& \lambda y(\triangle\cdot\exists r_2\cdot donkey(r_2)\cdot\triangle\cdot (yr_2)) & :: (E\rightarrow B)\rightarrow B\\
+&\text{Every farmer}=&& \lambda y(\Bowtie\cdot\triangle\cdot\exists r_1\cdot farmer(r_1)\cdot\triangle\cdot\Bowtie\cdot (yr_1)) &&:: (E\rightarrow B)\rightarrow B\\
+&\text{a donkey}=&& \lambda y(\triangle\cdot\exists r_2\cdot donkey(r_2)\cdot\triangle\cdot (yr_2)) &&:: (E\rightarrow B)\rightarrow B\\
 \midrule
-&\text{beats a donkey}&=& \lambda z(\triangle\cdot\exists r_2\cdot donkey(r)\cdot\triangle\cdot beats(z,r_2)) & :: E\rightarrow B\\
+&\text{beats a donkey}=&& \lambda z(\triangle\cdot\exists r_2\cdot donkey(r)\cdot\triangle\cdot beats(z,r_2)) &&:: E\rightarrow B\\
 \end{aligned}\end{equation*}
@@
 \item[Mention detection]
-The first step is to identify all noun phrases (NP) and personal (PRP) and possessive (PRP\$) pronouns. This information is easily pulled from the postree. Notice we also identify nested mentions:
+The first step is to identify all noun phrases (NP) and personal (PRP) and possessive (PRP\$) pronouns.
+This information is easily pulled from the constituent tree.
+Notice we also identify nested mentions:
 
@@
 All mentions are initially assigned to separate entities (entity ids).
-The entities are also attached information we can easily pull from the postree, such as `a girl: {number: singular, gender: female}' which are used for the following rules.
+The entities are also attached information we can easily pull from the constituent tree, such as `a girl: {number: singular, gender: female}' which are used for the following rules.
 \item[Speaker Sieve]
@@
 
-\subsection{Dependencies}
-
-Extracted from parse trees using tregex\cite{de2006generating}
+\subsection{Dependency grammars}
+
+This is an important thing to mention.
+
+Extracted from constituent trees using tregex\cite{de2006generating}
 
@@
 
-\subsubsection{Semantics via dependencies}
+%\subsubsection{Semantics via dependencies}
+
+%This might need updating based on: http://en.wikipedia.org/wiki/Dependency_grammar and http://en.wikipedia.org/wiki/Phrase_structure_grammar
+
 It is an interesting idea to consider that we might be able to define our semantics over a dependency graph, in a similar way to the Montague approach described earlier.
-The dependency graph adds a lot of flexibility where the phrase structure tree used in the Montague approach is limited to the direct syntax of the tree.
+The dependency graph adds a lot of flexibility where the constituent tree used in the Montague approach is limited to the direct syntax of the tree.
 A good example is our favourite sentence `if a farmer owns a donkey, he beats it', where there isn't a very strong syntactical relationship between `farmer' and `he'.
-It's also been suggested that it is mostly due to luck that a lot of English fits relatively well to the Montague approach, where as in some languages the structure of the phrase tree might not say much about the actual semantics.
+It's also been suggested that it is mostly due to luck that a lot of English fits relatively well to the Montague approach, where as in some languages the structure of the constituent tree might not say much about the actual semantics.
 
@@
 all NPs are `VP -> Bool' and VPs are `Entity -> Bool'.
-This makes it fit very nice in a programming context, where we can define a function `parseNP :: PhraseTree -> (VP -> Bool)'.With the dependency approach we don't have this option.
+This makes it fit very nice in a programming context, where we can define a function `parseNP :: ConstituentTree -> (VP -> Bool)'.With the dependency approach we don't have this option.
 A `subj' arrow can mean many different things.
@@
 
-Another problem worth considering is the fact that all information in the graph is drawn syntactically from the phrase tree.
+Another problem worth considering is the fact that all information in the graph is drawn syntactically from the constituent tree.
 That is, the graph doesn't contain any information not in the tree, but the other direction is entirely possible.
@@
 \section{Implementation (8)}
+
+alpha/beta renaming: We don't need it because we use globally unique variable names
 
@@
 
-I parse the parse tree using parsec\cite{leijen2001parsec}
+I parse the constituent tree using parsec\cite{leijen2001parsec}
 
 \begin{lstlisting}
-toMapping :: [Ref] -> PosTree Word -> ([Ref], Map Int Ref)
+toMapping :: [Ref] -> ConstituentTree Word -> ([Ref], Map Int Ref)
 toMapping vs (Leaf _ _) = (vs, empty)
@@
 
-where $tsize :: PosTree Word -> Int$ gives the number of words in the tree
+where $tsize :: ConstituentTree Word -> Int$ gives the number of words in the tree
 
@@
 
-\subsection{Using machine learning}
+\subsection{The grammar}
+This is not as bad as it sounds. It is similar to Montague and what the Stanford Parser does for dependencies.
+
+Let $\phi$, $\psi$, $\omega$ be lists of phrase nodes, let $\alpha$, $\beta$, $\gamma$ leaf nodes.
+Then we can define a set of interpretation functions $\llbracket\cdot\rrbracket_\text{type}$ from phrase nodes to logic:
+
+\begin{equation*}\begin{aligned}
+\ \llbracket (_\text{S}\phi)\psi\rrbracket_\text{ROOT} &:= \llbracket \phi\rrbracket_\text{S}\cdot\llbracket \psi\rrbracket_\text{ROOT}\\
+\llbracket (_\text{NP}\phi)\psi\rrbracket_\text{ROOT} &:= \llbracket \phi\rrbracket_\text{NP}(\lambda x.\top)\cdot\llbracket \psi\rrbracket_\text{ROOT}\\
+\midrule\ 
+%
+\llbracket (_\text{NP}\phi) (_\text{VP}\psi)\rrbracket_\text{S} &:= \llbracket \phi\rrbracket_\text{NP} \llbracket \psi\rrbracket_\text{VP}\\
+\llbracket (_\text{S}\phi) (_\text{CC}\text{and}) (_\text{S}\psi)\rrbracket_\text{S} &:= \llbracket \phi\rrbracket_\text{S} \cdot \llbracket \psi\rrbracket_\text{S}\\
+\llbracket (_\text{SBAR}(_\text{IN}\text{if})(_\text{S}\phi))(_\text{,},)\psi\rrbracket_\text{S} &:=\ ?_0(\Bowtie\cdot \llbracket \phi\rrbracket_\text{S}\cdot\Bowtie\ \cdot\ ?_1(\llbracket \psi\rrbracket_\text{S}))\\
+\midrule\ 
+%
+\llbracket (_\text{PRP}e)\rrbracket_\text{NP} &:= \lambda v . ve\\
+&\text{where $e$ is a predetermined variable name}\\
+\llbracket (_\text{DT}\alpha) (_\text{NN}\beta)\rrbracket_\text{NP} &:= \lambda v . \llbracket \alpha\rrbracket_\text{DT} \llbracket \beta\rrbracket_\text{NN} v\\
+\llbracket (_\text{DT}\alpha) (_\text{NN}\beta) (_\text{CC}\text{and}) (_\text{NN}\gamma)\rrbracket_\text{NP} &:= \lambda v .
+\llbracket \alpha\rrbracket_\text{DT} (\lambda y.
+\llbracket \beta\rrbracket_\text{NN}y \cdot \llbracket \gamma\rrbracket_\text{NN}y) v\\
+\llbracket (_\text{NP}\phi) (_\text{CC}\text{and}) (_\text{NP}\psi)\rrbracket_\text{NP} &:= \lambda v.\llbracket \phi\rrbracket_\text{NP}(\lambda x.\llbracket \psi\rrbracket_\text{NP}(\lambda y.\forall z.z=x \vee z=y \rightarrow vz))\\
+& \text{where $z$ is a globally free variable}\\
+\midrule\ 
+%
+\llbracket \text{a}, e\rrbracket_\text{DT} := \llbracket \text{some}, e\rrbracket_\text{DT} &:= \lambda vw.\triangle\cdot\exists e\cdot ve\cdot\triangle\cdot we\\
+\llbracket \text{all}, e\rrbracket_\text{DT} := \llbracket \text{every}, e\rrbracket_\text{DT} &:= \lambda vw.?_0(\Bowtie\cdot\triangle\cdot\exists e\cdot ve\cdot\triangle\cdot\Bowtie\cdot we)\\
+\llbracket \text{no}, e\rrbracket_\text{DT} &:= \lambda vw.?_0(\Bowtie\cdot\exists e\cdot ve\cdot we\cdot\Bowtie\cdot\bot)\\
+\llbracket \text{the}, e\rrbracket_\text{DT} &:= \lambda vw.we\\
+\midrule\ 
+%
+\llbracket (_\text{VB}\alpha)\rrbracket_\text{VP} &:= \llbracket \alpha\rrbracket_\text{VB}\\
+\llbracket (_\text{VB}\alpha)(_\text{NP}\phi)\rrbracket_\text{VP} &:= \lambda x.\llbracket \phi\rrbracket_\text{NP} (\llbracket \alpha\rrbracket_\text{TVB} x)\\
+\llbracket (_\text{VP}\phi)(_\text{CC}\text{and})(_\text{VP}\psi)\rrbracket_\text{VP} &:= \lambda x.\llbracket \phi\rrbracket_\text{VP}x\cdot \llbracket \psi\rrbracket_\text{VP}x\\
+\llbracket (_\text{VB}\alpha)(_\text{CC}\text{and})(_\text{VB}\beta)(_\text{NP}\phi)\rrbracket_\text{VP} &:= \lambda x.\llbracket \phi\rrbracket_\text{NP}(\lambda y.\llbracket \alpha\rrbracket_\text{TVB}xy\cdot \llbracket \beta\rrbracket_\text{TVB}xy)\\
+\llbracket (_\text{VB}\text{do})(_\text{RB}\text{not})(_\text{VP}\phi)\rrbracket_\text{VP} &:= \lambda x.?_0(\Bowtie\cdot\llbracket \phi\rrbracket_\text{VP}x\cdot\Bowtie\cdot\bot)
+\end{aligned}\end{equation*}
+
+\subsubsection{Using machine learning}
 The $\bowtie$ and $\triangle$ notation makes logic look close enough to natural language that it starts looking tempting to use machine learning methods such as those used in translation to translate.
@@
 
-\subsection{Hand coding}
-This is not as bad as it sounds. It is similar to Montague and what the Stanford Parser does for dependencies.
-
-\subsection{Examples}
-Until now we've been very focused on the same one or two simple sentences.
-To provide a better test bench for our approach, it is interesting to follow some sentences through the different levels of our approach.
+\subsection{Testing}
+In the previous sections we have tried to be as formal and careful as possible in our derivations and proofs, but as Knuth famously said:
+
+\begin{quotation}
+Beware of bugs in the above code; I have only proved it correct, not tried it.
+\end{quotation}
+
+In the actual implementation of my program I have used the \emph{QuickCheck} and \emph{HUnit} test environments to write unit tests for algebraic properties and actual inputs to my program.
+To enable QuickTest of the different properties on constituent trees, a random generator of sentences was implemented.
+
+What remains to test is the validity of the Montague grammar introduced in the previous section.
+As reasonable as it seems we want to reduce the likelihood of it containing conflicting definitions.
+In a transformation grammar over natural language we are always running the risk of have been too narrow sighted in its definition.
+The classic example being `($_\text{DT}$a)' interpreted as $\exists$, but then actually being an $\forall$ if used in an `if' phrase.
+
+In DPL we don't have that particular problem, but instead we can run into problems with the way the scopes of composed elements affect each other.
+In Visser logic it is even worse, as $\Bowtie$s and $\triangle$s introduced much later in the discourse, can get transposed far around the final statement.
+
+Hence, as reasonable as the introduced grammar seems, we want to reduce the likelihood of conflicts as much as possible.
+Until now we have been very focused on the same one or two simple sentences, or whatever sentences were relevant to the features we were discussing, but now we are going to introduce a much larger set of complicated test cases.
+Ideally we want test cases for testing every combination of grammar rules immaginable.
+
+In the test of this program, we used a branch of 100 sentences.
+Some copied from the papers referenced, others invented based on features considered or the daily life of farmers and donkeys as immagined by the author.
+The sentences were converted to Visser, DPL and FOL, and the meanings were checked manually to make sense.
+
+In the following section we will run through the most interesting sentences used, and hopefully make it a bit clearer how the different steps of translation work.
+
+\subsubsection{Examples}
+
+In this section we translate 12 sentences from English to the Visser logic to DPL and finally to FOL as described in the earlier sections.
 
@@
 &\text{\textbf{English}}&&\text{A man and a woman owned a horse and a donkey.}\\&&&\text{If the donkey did not walk, and the man was a farmer, he beat it.} \\
-&\text{\textbf{Visser}}&&\triangle \cdot \exists a \cdot man(a) \cdot  \exists b \cdot woman(b) \cdot \triangle \cdot ?_0(\Bowtie \cdot \exists c \cdot eq\_on\_of(c,a,b) \cdot \Bowtie\\&&&\cdot \triangle \cdot \exists d \cdot horse(d) \cdot  \exists e \cdot donkey(e) \cdot \triangle \cdot ?_0(\Bowtie \cdot \exists f \cdot eq\_on\_of(f,d,e) \cdot \Bowtie\\&&&\cdot own(c,f))) \cdot ?_0(\Bowtie \cdot ?_0(\Bowtie \cdot walk(e) \cdot \Bowtie \cdot \bot) \cdot a\_farmer(a) \cdot \Bowtie \cdot ?_1(beat(a,e))) \\
+&\text{\textbf{Visser}}&&\triangle \cdot \exists a \cdot man(a) \cdot  \exists b \cdot woman(b) \cdot \triangle \cdot ?_0(\Bowtie \cdot \exists c \cdot eq\_on\_of(c,a,b) \cdot \Bowtie\\&&&\cdot \triangle \cdot \exists d \cdot horse(d) \cdot  \exists e \cdot donkey(e) \cdot \triangle \cdot ?_0(\Bowtie \cdot \exists f \cdot eq\_on\_of(f,d,e) \cdot \Bowtie\\&&&\cdot own(c,f))) \cdot ?_0(\Bowtie \cdot ?_0(\Bowtie \cdot walk(e) \cdot \Bowtie \cdot \bot) \cdot a\_farmer(a) \cdot \Bowtie \cdot beat(a,e)) \\
 &\text{\textbf{DPL}}&&\exists a \cdot man(a) \cdot \exists b \cdot woman(b) \cdot \exists d \cdot horse(d) \cdot \exists e \cdot donkey(e)\\&&&\cdot (\exists c \cdot eq\_on\_of(c,a,b) \cdot \exists f \cdot eq\_on\_of(f,d,e) \rightarrow own(c,f))\\&&&\cdot (\neg walk(e) \cdot a\_farmer(a) \rightarrow beat(a,e)) \\
@@
 
-Extract D Testing 
-We have already seen some screen shots of the working program; however, we provide two stringent tests 
-for our program to ensure it works as intended, along with a test rig to fully analyze the program. In both test 
-programs, I will run through the whole series of options available to the user, and ensure its correctness. 
-However, I will also demonstrate its ability to visualize code, and hopefully provide valuable insights whilst 
-debugging. 
-5.1 Simple Program -BFS and DFS using the Visitor Pattern
-This test program begins by creating an underlying tree structure…
-This kind of debugging is intuitive, and simple to do within this framework. If you have an intuitive 
-understanding of what the underlying model in your program should look like, it is fairly straight forward to 
-spot bugs like this in  small code samples. Assuming a larger program is in use, the user must delve a little 
-deeper into the part of the graph which they suspect the bug to exist in. This is obviously heavily aided by the 
-JDT debugger itself. However, this test still shows the usability of the code in a small program, and shows 
-that the code can cope with the different types of back links and cross links that can occur in a memory 
-graph.
-
 \section{Perspectives/Conclusions (2)}
 
-In the papers it is always assumed that the postree and things are `trivial' or `already sorted', but how do the pieces really stack up when put together?
-
-What I learned
+Move 1: Summary 
+Summarizes the most important aspects of the project work.
+
+I have successfully developed a novel local type inference algorithm
 
@@
 improved; a discussion of the lessons you have learnt from the project.’
-In the ‘Conclusions’ you summarize your work and then stand back from the project and assess it as 
-objectively as you can. This shows the examiners that you are capable of evaluating your own work 
-according to the standards of the field. It is acceptable to mention areas that were not very satisfactory
-because this shows that you have learnt from the experience of doing the project.
-There are often three moves starting with a summary of your work, then giving an evaluation, highlighting
-both its achievements and limitations and finally the possible future extensions of the work. The Evaluation 
+
+In the papers it is always assumed that the constituent tree and things are `trivial' or `already sorted', but how do the pieces really stack up when put together?
+
+\begin{itemize}
+\item Studied computational linguistics and learned about state of the art technologies
+\item Found out what was going on with dynamic logics and 
+similar discourse algebras
+\item Ran into many issues some of which have been discussed in great deal before (like strong vs. weak readings) and some that haven't had as much attention (like semantics of `and' and `or').
+\item Combined the technologies and theory in a way not tried before, which yielded very good results
+\end{itemize}
+
+DPL made it possible to specify a lot of rules that would otherwise have been terrible complicated. Most notably `if'.
+
+Monoids were an interesting alternative to lambda composition. It could have been interesting to see how far they could go.
+
+It would be interesting to return more than one result in the case of ambiguity. Weak/strong readings and things like transitive verbs. `A farmer starves and beats a donkey'.
+
+A lot more rules could relatively easily be added to the grammar.
+For some things like adverbs we would need a richer language, like $furiously(beats(x,y))$, but there's certainly still a lot of low hanging fruits.
+
+Level 1, easy:
+* More uses of `no' and `nobody'
+* possessives pronouns, 
+* `existential there'
+* more kinds of SBARs like started by `then', `whom' etc.
+
+Level 2, more work:
+* adverbs
+* modals verbs are a big task
+
+Level 3, requires new understanding of language:
+* Existential non-commitment (The golden mountain does not exist)
+* The ancient greeks thought the morning star was the evening star
+* Understanding words never used before
+
+Level X, not working due to the framework:
+* Every dog sees a cat. It chases it.
+* Buffalo buffalo Buffalo buffalo buffalo buffalo Buffalo buffalo
+
+Move 2: Evaluation
+Indicates what the writer considers to be the most valuable aspects of the project and its limitations. Provides 
+an assessment of how far the aims of the project have been achieved.
+
+In the ‘Conclusions’ you summarize your work and then stand back from the project and assess it as  objectively as you can.
+This shows the examiners that you are capable of evaluating your own work according to the standards of the field.
+It is acceptable to mention areas that were not very satisfactory because this shows that you have learnt from the experience of doing the project.
+
+Everything was perfect!
+It feel
+
+my algorithm is exponential: whereas Gagnon's algorithm is 
+polynomial. But experiments of execution time against method length show a typically linear trend whereas 
+Gagnon's show a cubic trend
+
+Move 3: Future Work
+Gives possible extensions of the project. These suggestions may follow from the limitations mentioned in the 
+Evaluation move.
 move may also be carried out as part of the Summary. The Conclusion is the mirror image of the 
@@
 
-Move 1: Summary 
-Summarizes the most important aspects of the project work.
-
-I have successfully developed a novel local type inference algorithm
-
-Move 2: Evaluation
-Indicates what the writer considers to be the most valuable aspects of the project and its limitations. Provides 
-an assessment of how far the aims of the project have been achieved.
-
-my algorithm is exponential: whereas Gagnon's algorithm is 
-polynomial. But experiments of execution time against method length show a typically linear trend whereas 
-Gagnon's show a cubic trend
-
-Move 3: Future Work
-Gives possible extensions of the project. These suggestions may follow from the limitations mentioned in the 
-Evaluation move.
-
-The greatest scope for future work is extending the application of my algorithm from local type inference to 
-global type inference.
-
-Things learned from courses
-
-Should I make my latex file literate haskell? %bhttp://www.haskell.org/haskellwiki/Literate_programming
-
-%$\lstinputlisting[language=Haskell]{src/Main.hs}
+The greatest scope for future work is extending the application of my algorithm from local type inference to global type inference.
 
 \section{Acknowledgments}
-I would like to thank Samson for guiding me through
-
-the PCFG parser\cite{klein2003accurate}
-
-Érica for linguistic discussions
-
+I would like to thank Samson Abrahamski for introducing me to the world of Natural Language parsing and semantics, for helpful and stimulating discussins and for keeping my on the right path
+
+Thanks to Phil Blunsom and Edward Grefenstette for exciting and enlightening lectures into the classical and modern approaches.
+
+Thanks to the Natural Language Processing Group at Stanford University for making their software and results easily available for use in my project\cite{klein2003accurate}.
+
+Thanks to Érica Baena Kato for keeping me awake with interesting linguistic perspectives.
 \newpage

### 2013-05-06 src/Main.hs

@@ -17,6 +17,10 @@ super :: [String] -> IO [V.Visser]
 super sentences = runDry sentences >>= (return . map montague)
 
 sentences = [
+	"If a man eats a sausage, he is happy.",
+	"Every person has a pet.",
+	"Every man supports every woman.",
+	"Every man supports no woman.",
 	"Every farmer beats some donkey.",
 	"If a farmer owns a donkey, he beats it",
 
@@ -91,6 +95,7 @@ sentences = [
 
 	-- adverbs
 	"A farmer and a donkey live together.",
+	"Colorless green ideas sleep furiously",
 	"Do you still beat your wife?",
 
 	-- Existential non-commitment
@@ -99,7 +104,8 @@ sentences = [
 	"The golden mountain does not exist",
 
 	-- Missing features in coref tagger
-	"Every dog sees a cat. It chases it."
+	"Every dog sees a cat. It chases it.",
+	"Buffalo buffalo Buffalo buffalo buffalo buffalo Buffalo buffalo"
 	]
 
 main :: IO ()

### 2013-05-06 src/ToLogic.hs

@@ -21,7 +21,7 @@ intRoot [] r = VT
 
 intS :: [PosTree IWord] -> Store -> Visser
 intS [P "NP" np, P "VP" vp] r = intNP np r (intVP vp r)
-intS (P "SBAR" ([(L "IN" ("if",_)), (P "S" s1)]) : (L "," _) : s2) r =
+intS ((P "SBAR" [(L "IN" ("if",_)), (P "S" s1)]) : (L "," _) : s2) r =
 		VQ0 (VSw `VC` intS s1 r `VC` VSw `VC` VQ1 (intS s2 r))
 intS [P "S" s1, L "CC" ("and",_), P "S" s2] r = intS s1 r `VC` intS s2 r
 intS [P "S" s1, L "," _, L "CC" ("and",_), P "S" s2] r = intS s1 r `VC` intS s2 r
@@ -60,7 +60,7 @@ intDT :: IWord -> Store -> Predicate -> Predicate -> Visser
 intDT ("some", i) r = \p q -> VSc `VC` VE m `VC` p m `VC` VSc `VC` q m where m = r i
 intDT ("a", i) r = intDT ("some", i) r
 intDT ("the", i) r = \p q -> q m where m = r i
-intDT ("every", i) r = \p q -> VSw `VC` VSc `VC` VE m `VC` p m `VC` VSc `VC` VSw `VC` q m where m = r i
+intDT ("every", i) r = \p q -> VQ0 (VSw `VC` VSc `VC` VE m `VC` p m `VC` VSc `VC` VSw `VC` q m) where m = r i
 intDT ("all", i) r = intDT ("every", i) r
 intDT ("no", i) r = \p q -> VQ0 (VSw `VC` VE m `VC` p m `VC` q m `VC` VSw `VC` VF) where m = r i
 intDT other r = error ("Unsuported by intDT: " ++ show other)

### 2013-05-09 paper.tex
Fixed introduction

@@
 %From Donkeys to Monoids: A Computational Framework for Dynamic Semantics
-
-\title{From Donkeys to Monoids: A Computational Framework for Dynamic Semantics}
+%Donkeys to Monoids: A Computational Framework for Dynamic Semantics
+
+\title{Donkeys to Monoids: A Computational Framework for Dynamic Semantics}
 \author{Thomas Dybdahl Ahle}
@@
 The rationale behind them is a stride towards a logic that can more easily approximate natural language in aspects such as polarity and scope.
-
-Visser didn't test his ideas computationally and neither did most people involved with DPL and logification of natural language in general.
+Indeed their name derives from the famous `Donkey Sentence' by Geach\cite{geach1962reference} which exhibit these issues so well.
+
+Visser didn't test his ideas computationally and neither did most people involved with DPL.
+In fact most of the classical literature on logification of natural language has this `problem'.
 In this paper I supply such an experiment, combining technologies introduced over the last four decades, creating a fully functioning `English to logic' framework.
@@
 I apply the framework to a large collection of sentences, showing the conversion process step by step.
-
-Finally I discuss ways to support a larger fragment of English, how we can combat and interprete the inherent ambiguity, and how the flexibility of Donkey Monoids compare with the classical Montague lambda structures.
+Finally I discuss ways to enlarge the supported fragment of English, how we can combat and interprete the inherent ambiguity we meet, and how the flexibility of Donkey Monoids relate to the classical Montague lambda structures.
 
@@
 
+% No donkeys were hurt in the making of this paper
+
 There are many reasons why it is interesting to represent information in a logical language.
@@
 
-HVAD MED AT SKRIVE LIDT MERE OM HVAD VI VIL HER??
-
-It would seem getting computers to understand and work with natural language is on of the biggest challenges in information technologies this day.
+It would seem that getting computers to understand and work with natural language is on of the biggest challenges in information technologies this day.
 Moving the problems into the space of formal logic is a classical and tempting path.
-Unfortunately, as many people have pointed out, classical logic simply doesn't have the power to express most of the sentences people casually express every day.
-
-We won't let us self get too down over this, but it is worth it to outline where the typical problems arrise, and to get a feeling of what a formal logic translation of natural language actually looks like.
-We'll start out with a sentence that easily lets itself translate and then a couple that don't:
+It does however remain open whether our formal logics actually have the power to express most of the sentences people casually voice on a daily basis.
+In this paper we will assume it does for enough practical cases to be useful.
+
+The reason logics like DPL and Visser Monoids exist is a famous sentence noted by Peter Thomas Geach in 1962\cite{geach1962reference}: `Every farmer who owns a donkey beats it'.
+As we will see, this sentence illustrates that even if we restrict ourselves to factual sentences, formal first order logic still runs into a lot of problems when representing many natural sentences.
+In fact this sentence broke early automatic approaches to logification, and led to the construction of Discourse Representation Theory (DRT)\cite{kamp1993discourse}.
+
+Before I go into more details on my automatic logification scheme, let's have a look at some of the issues we run into if we try approach the task manually.
+This should help us get a feeling of what a formal logic translation of natural language actually looks like:
 \begin{enumerate}
+
 \item `Socrates is a donkey': $donkey(Socrates)$.
@@
 If we want to do without constants we can write: $\forall x(Socrates(x)\rightarrow donkey(x))$.
-Notice this allows for more than one Socrates to exist.
-\item `All his donkeys are treated well': This sentence somewhat willingly let's itself translate into: $\forall y(his(x,y)\wedge donkey(y) \rightarrow treated\_well(y))$, which has a free variable $x$ referring to some person from earlier in the discourse.
-\item `Every farmer beats his donkey': This sentence clearly has two possible meanings: $\exists y(his(x,y)\wedge donkey(y)\wedge \forall z(farmer(z) \rightarrow beats(z,y))$ if there is only one donkey, and $\forall x(farmer(x)\rightarrow \exists y (beats(x,y)\wedge his(x,y)\wedge donkey(y))$ if there is a donkey for each farmer.
+See how this allows for more than one Socrates to exist.
+
+\item `All his donkeys are treated well': This sentence somewhat willingly let's itself translate into: $\forall y(his(x,y)\wedge donkey(y) \rightarrow treated\_well(y))$, which has a free variable $x$ referring to some (male?) entity in the context of the discourse.
+
+\item `Every farmer beats his donkey': This sentence clearly has two possible meanings: $\forall x(farmer(x)\rightarrow \exists y (beats(x,y)\wedge his(x,y)\wedge donkey(y))$ if `his' refers to the farmers or $\exists y(his(x,y)\wedge donkey(y)\wedge \forall z(farmer(z) \rightarrow beats(z,y))$ if it refers to somebody else.
+There is no way we can know without more context information.
+
 \item `The happy donkey doesn't exist': Here we get into more problems.
@@
 Alternatively we could write $\exists x(happy\_donkey(x)\rightarrow \bot)$, but that is another sentence: `There are no happy donkeys'.
-\item `They donkeys are cold': This is an often used example of an `implied action sentence' and thought to have the same meaning as `Please close the stable door'.
-The best we can do here seems to be $please\_close\_the\_stable\_door()$ or perhaps $\exists x(stable\_door(x)\wedge please\_close(x))$.
-\item `Jane likes a painting': This is easy to translate to logic, but are we comfortable that it has the same meaning as the sentence `It is not the case that Jane doesn't like a painting' and even `Jane likes a painting and either it is raining or it isn't'?
+In any case it makes it hard to say what the logical meaning of `the' really is.
+
+\item `The donkeys are being stubborn!': This is an often used example of an `implied action' sentence and thought to have the same meaning as `Please go out and beat them!'.
+The best we can do here seems to be $please\_beat\_the\_doneys()$ or perhaps $\forall x(donkey(x)\rightarrow please\_beat(x))$ or even $please(\forall x(donkey(x)\rightarrow beat(x)))$.
+
+\item `Jane beats a donkey': This is easy to translate to $\exists x(donkey(x)\wedge beat(Jane,x))$, but are we comfortable that this is also our interpretation of the sentence `It is not the case that Jane doesn't beat a donkey' and even `Jane beats a donkey and either it is raining or it isn't'?
 It is hard to imagine a language with strong logical foundations, where those three sentences wouldn't be equal.
@@
 
-In this paper we will work with a subset of English that translate reasonably well into logic.
-This unfortunately means that arguably the majority of sentences will be inaccessible to our methods.
-We will take a shot at ambiguous sentences like the third one though.
-
-%TODO: Perhaps this should be moved to the aknowledgements or something. It sets a very demoralizing tone during the rest of the paper.
-Even more unfortunate than the problems above is it that we won't be able to do anything about the concerning welfare for donkeys around the world\cite{donkey2013sanctuary}, but must restrict ourselves to simply pointing it out.
-We will follow Geach\cite{geach1962reference} on this, and make `If a farmer owns a donkey, he beats it' our primary object of study.
-%
-\begin{align}
-\exists x (farmer(x) \wedge \exists y (donkey(y)\wedge owns(x,y)) \rightarrow beats(x,y))
-\end{align}
-%
-This is arguably the formula that first comes to mind for a moderately trained logician, or simply what you expect by looking at the structure of the constituent tree.
-Translating an indeterminate `a farmer' into $\exists x(farmer(x))$ seems correct, but clearly the formula has problems with scope.
+These examples should hopefully have shown that some uses of English, especially example 4 and 5, are well beyond what we can hope to automatically logify.
+However you might think the sentences of `factual nature' like 1 and 2 shouldn't be too hard?
+Unfortunately, as sentences get longer and more complex we will see that they are, and this is where Donkey Monads, Dynamic Predicate Logic (DPL) and Discourse Representation Theory (DRT) have a role to play.
+Let's take a close look at that `Donkey Sentence': `If a farmer owns a donkey, he beats it'.
+
+From the examples before, and perhaps from common intuition, we might find it natural to think of the sentence as two subsentences combined by `if': `if (a farmer owns a donkey) (he beats it)'.
+Translating the parentheses first: `A farmer owns a donkey' logifies to $\exists x (farmer(x) \wedge \exists y (donkey(y)\wedge owns(x,y))$, and `he beats it' translates to $beats(x,y)$.
+Hence the whole formula should probably be:
+%
+\begin{align}
+\exists x (farmer(x) \wedge \exists y (donkey(y)\wedge owns(x,y))) \rightarrow beats(x,y)
+\end{align}
+%
+Which is totally wrong. The $beats(x,y)$ clause ends up refering to the variables outside of their scope, and even if we try to push it inside the parentheses:
 %
@@
 %
-Moving the implication into the scope of $x$ also doesn't work.
-Sentence \eqref{wrong_donkey_2} evaluates to true as long as a single farmer beats his donkey.
-In fact as long as a farmer exists it evaluates to true, since plugging the farmer into the $donkey$ predicate makes the antecedent false.
-%
-\begin{align}
-\forall x (farmer(x) \rightarrow \forall y ( owns(x,y) \wedge donkey(y) \rightarrow beats(x,y))) \label{correct_donkey}
-\end{align}
-%
-This turns out to be the correct formula. At least as long as you consider the use of `all' in natural language equal to the `all' in logic, which can be vacuously true.
-
-So what is going on?
-Most people think of determinates `the donkey' and indeterminates `a donkey' as being a reference to an existing donkey and an `introduction' of some new donkey into the conversation.
-However in a lot of sentences this turns out to not be the case.
-There doesn't seam to any simple way to translate them into logic.
-Even less does it seem intuitive how to use them as antecedents to pronouns.
-
-The common way to deal with this problem is to make a big jump in power.
-Using a Montague Grammar we can define any noun phrase (NP) as a function from a verb phrase (VP) to a sentence.
-This way `if a farmer' might be a function $VP \to S \to S$, but even with this power, a working function is not easy to construct\cite{montague1973proper}.
-
-Another problem with sentence \eqref{correct_donkey} is a lack of compositionality.
-This property which is vital if we want to build up a meaning of a conversation piece by piece.
-Say for example that our primary sentence was followed by the sentence `He also feeds it': $feed(x,y)$.
-How do we combine the first and the second sentence to give the meaning $\forall x(farmer(x)\rightarrow \forall y(owns(x,y)\wedge donkey(y)\rightarrow beat(x,y)\wedge feed(x,y)))$?
-
-The point is that there is simply not a way to do this in PL.
-All variables in a formula are always closed off in their scope, and there are no logical operators that can `push' a new predicate into a closed scope.
-Our previous case might be a bit unfair, since we were trying to push something into a complicated implication, but even in case of the simple sentences `There is a farmer.
-He owns a donkey.' we wouldn't be able to build a meaning from the pieces.
-
-One could argue that we can simply obtain compositionality using Montague functions like we discussed earlier (and will discuss more deeply in a later chapter), but what would be really nice is to have a simple logic that simply has these things built in.
-In this paper we are going to look into not only one, but a couple of such logics.
-In these logics we will be able to write things such as:
-%
-\begin{align}
-(\exists x \cdot farmer(x)) \cdot (\exists y \cdot donkey \cdot owns(x,y))
-\end{align}
-%
-Here the variables survive throughout the formula, which allows us to concatenate it with other formulas. The parenthesis are associative, so we don't even need them.
-Later in the paper we will also look at a logic allowing us to write:
-%
-\begin{align}
-(\bowtie \cdot \exists x \cdot Farmer(x) \cdot \exists y \cdot donkey(y) \cdot owns(x,y) \cdot \bowtie beats(x,y)) \cdot feeds(x,y)
-\end{align}
-%
-Which allows us to be compositional even in the case of implications.
-
-It is interesting to consider whether such logics can give us something on the way of translating speech to logic.
-In this paper we are going to look into exactly that, and we will combine it with the best available tools for dependency analysis and coreference resolution.
-This is the game plan:
-\vspace{1em}
-\tikzstyle{decision} = [diamond, draw, fill=blue!20,
-    text width=4.5em, text badly centered, node distance=2.5cm, inner sep=0pt]
-\tikzstyle{block} = [rectangle, draw, fill=blue!20,
-    text width=5em, text centered, rounded corners, minimum height=4em]
-\tikzstyle{line} = [draw, -latex']
-\begin{tikzpicture}[node distance = 3cm, auto]
-    % Place nodes
-    \node [block] (nl) {Natural Language};
-    \node [block, right of=nl] (dr) {Dependency Representation};
-    \node [block, right of=dr] (monad) {Enhanced DPL Representation};
-    \node [block, right of=monad] (dpl) {Dynamic Predicate Logic};
-    \node [block, right of=dpl] (prop) {First Order Logic};
-    % Draw edges
-    \path [line] (nl) -- (dr);
-    \path [line] (dr) -- (monad);
-    \path [line] (monad) -- (dpl);
-    \path [line] [dotted] (dpl) -- (prop);
+It doesn't work either.
+Sentence \eqref{wrong_donkey_2}, while syntactically correct, evaluates to true as long as a single farmer beats his donkey.
+In fact \textbf{as long as any farmer exists} \eqref{wrong_donkey_2} evaluates to true.
+This is because $\exists y$ also spans over the farmer entity (and he's not a donkey) which makes the antecedent false.
+Clearly this is not the formula we are after.
+
+What turns out to be the correct logification is this:
+%
+\begin{align}
+\forall x (farmer(x) \rightarrow \forall y (owns(x,y) \wedge donkey(y) \rightarrow beats(x,y))) \label{correct_donkey}
+\end{align}
+%
+A far cry from our original `guess'.
+Moreover the above version even have the oddity that the interpretation of `all' is `the logical one': The formula \eqref{correct_donkey} is, in addition to the intended situations, also considered true if no farmers exist in the domain, or none of them own donkeys.
+
+Now the above example may seem a bit contrived.
+Why did I have to try to combine the meaning from the obviously wrong pieces?
+Wasn't some other strategy more obvious?
+Before we go into details of such logification strategies, let's first see how easy it could have been, if we instead of first order logic had used Visser's Donkey Monoids:
+%
+\begin{align}
+\Bowtie \cdot \exists x \cdot farmer(x) \cdot owns(x,y) \cdot\triangle\cdot\exists y \cdot donkey(y) \cdot\triangle\cdot\Bowtie\cdot beats(x,y)
+\end{align}
+%
+If we read $\exists$ as `a' this is basically identical to the English sentence.
+The donkey monoid even supports placing the transitive verb between the two referants.
+Clearly some magic is going on, that we will have to investigate further.
+
+Neat as the Donkey Monoids are, interested as we are in automation, we should be very concerned about those magical $\Bowtie$s and $\triangle$s that have popped up in our formula.
+What do they mean?
+How do they work?
+And how do we know when to use them?
+Visser himself, while answering question 1 and 2 in detail, doesn't give many hints on question 3.
+Only that they are probably controlled by adverbs and similar `helping words'.
+
+So if we don't know how to use them, how can we know if the Donkey Monoids are good logics for natural language?
+In the later chapters I will discuss their formal meaning in more detail and also what semantic meaning I intend to give them.
+In the final chapters we will see just how much flexibility this approach allows us.
+
+We need a framework to encode our rules into, and luckily the field of computational semantics offer several solutions.
+Basically there are two well described ways to go: statistical vector space methods, which is very much the most popular alternative right now, and logical methods.
+As you might have guessed I will lean in the dirrection of the later.
+The most well described methods are Montague Grammars and Discourse Representation Theory (invented after the donkey sentence).
+
+In the early sixties the American logician, Richard Montague, was a teaching a class in which the students had to do logification of simple sentences.
+He noticed that there were many patterns in the way they did it, and conjectured that it would be possible to formalize the entire process.
+So he set out to write a series of papers that became very influential in the philosophy of language, linguistics and computer science.
+
+The Montague Grammar, as Montague's approach came to be know, works amazingly well, it's simplicity taken into account.
+It is not hard to see how one might convince oneself that it could be used to describe all of language.
+The basic idea is to define logical formulas recursively over the so called constituent tree of a sentence.
+The tree is based on the idea of context free grammars for natural language developed mainly by Noam Chomsky\cite{chomsky2002syntactic}.
+Figure \ref{fig:const_tree} is an example of what it might look like.
+
+\tikzstyle{level 1}=[level distance=1cm,sibling distance=3.5cm]
+\tikzstyle{level 2}=[level distance=1cm,sibling distance=2.25cm]
+\tikzstyle{level 3}=[level distance=1cm,sibling distance=1.1cm]
+\begin{figure}
+\centering
+\begin{tikzpicture}[grow=down, sloped]
+\node {S}
+    child {
+        node {NP}
+        child {
+            node {Det}
+            edge from parent
+            child {
+                node {the}
+                edge from parent
+            }
+        }
+        child {
+            node {N}
+            child {
+                node {farmer}
+                edge from parent
+            }
+            edge from parent
+        }
+        edge from parent
+    }
+    child {
+        node {VP}
+        child {
+            node {V}
+            edge from parent
+            child {
+                node {beats}
+                edge from parent
+            }
+        }
+        child {
+            node {NP}
+            edge from parent
+            child {
+                node {Det}
+                edge from parent
+                child {
+                    node {a}
+                    edge from parent
+                }
+            }
+            child {
+                node {JJ}
+                edge from parent
+                child {
+                    node {lazy}
+                    edge from parent
+                }
+            }
+            child {
+                node {N}
+                edge from parent
+                child {
+                    node {donkey}
+                    edge from parent
+                }
+            }
+        }
+        edge from parent
+    };
 \end{tikzpicture}
-
-Going from step 2 to step 3 we are going to try and eliminate some of the ambiguity we saw earlier by outputting multiple possible sentences.
-In particular we will try and make a strong and a weak reading of every implication.
-However before we get into the details of how everything is implemented in code, we should look more into the nifty details of the theory applied.
-
-TODO: Requirements?
+\caption{A Constituent Tree} \label{fig:const_tree}
+\end{figure}
+
+We will see later how exactly the Montague Grammar obtains a full sentence formula from its reqursive process, but in any case for a computer scientist the introduction of trees is always a promising step.
+
+The main thing to notice now is that for Montague's approach to work, it is crucial that we can always get the meaning of a composed phrase from a composition of the meaning of its subphrases.
+This compositional property is exactly what we saw earlier that first order logic doesn't posses, and Montague goes through many hoops to simmulate it.
+
+I do however avoid many of those problems, since my goal here is not to test the validity of Montague's proposed rules, but to experiment with how Donkey Monoids and to an extent DPL behaves when put in the framework described.
+To draw a full circle I am however then going to translate the Donkey Monoids into first order logic so we can easily check if our formulas have the meanings we expect.
+
+This is the chain I am going to build:
+
+\begin{description}
+\item[1 Natural Language] We start with a file of simple English natural language sentences. Each line is processed on its own, but may contain more than one sentence.
+
+\item[2 POS tagger] This is the first of three Stanford NLP programs my Haskell script is going to pass the text to.
+The POS tagger tokenizes the sentences and statstically tries to figures out part of speach for each word.
+This is our first source of error.
+
+\item[3 Constituent Grammar] It might be that the reason Montague and other early pioneers didn't test their programs mechanically was their lack of access to a good parser.
+The Stanford Parser is the best on the market, but it still an error rate of around 5\% which is one in every two sentences.
+
+\item[4 Correference Resolution] A lot have been said about how anaphora might be resolved using formal Montague Grammar like approaches.
+DRT is especially focused on this topic.
+None the less the current champeon of anaphora competitions\cite{lee2011stanford}\cite{raghunathan2010multi} is the heuristical Stanford Deterministic Coreference Resolution System, the workings of which I will explain in detail.
+
+\item[5 Donkey Monoid] For us the step from 4 to 5 is the most interesting, as it is where we go from syntax to semantics.
+If we manage to push our original text through to here, not much can go wrong.
+Only issue is if the logical meaning we create is not the one intended by the speaker.
+
+\item[6 Dynamic Predicate Logic] Since the Donkey Monoids are variants of DPL, they are easy to convert into it. It is much more surprising that DPL lets itself convert into first order logic.
+
+\item[7 First Order Logic] We end up translating our dynamic logics into this old friend.
+This is mainly done to test the conversion algorithm, and because the first order logic is what we are used to read.
+The results can be surprising.
+\end{description}
+
+However before we can go into the many technical details, we need to develop a very good understanding of the logics and ideas we are working with.
 
@@
 
+Visser: The constructions studied in this paper can be viewed as adding control elements to the language.
+A control element is a symbol representing an instruction concerning storage of bits of meaning.
+Thus, control elements serve to steer semantic processing.
+
 \begin{itemize}
@@
 Thanks to Érica Baena Kato for keeping me awake with interesting linguistic perspectives.
+
+Thanks to The Donkey Sanctuary\cite{donkey2013sanctuary} for actually taking donkeys welfare seriously and rescuing neglected donkeys in the UK and all over the world.
+
 \newpage

### 2013-05-09 paper.tex
Annotated appendix

@@
 
-Since we are going to experiment with converting language into logic, a natural place to start is with classical logic.
-Classical logic, also called Predicate logic (PL), is the logic of conjunction, disjunction and negation extended with universal and existential quantifiers.
-
-In the introduction chapter we saw that PL was rather lacking in terms of compositionality and we predicted that what we really wanted was an `update logic' (a logic with update semantics) of some sort.
-Update semantics are well known and studied in computer language theory, and have in the past couple of decades found their way into a wider range of philosophies.
-
-The idea of all of those is a `growth of information in time'.
-Starting from little, and with each statement adding more.
+In these chapters I'm going to describe the logical theories by Visser and earlier dynamic semantists.
+I will introduce some addtions of my own and prove interesting properties about them.
+Finally, but most importantly, I will go into depth with the tools and ideas we need to write a strong logification framework.
+
+In the previous chapter I showed how First Order Logic lack certain compositional properties.
+A possible solution to this problem is to use a so called `update logic' (a logic with update semantics).
+Update semantics are logics inspired on the semantics of programming languages where a variable, once introduced, will stick around for a period not know at the time of it's introduction.
+
+The idea of those logics is a `growth of information in time'.
+Starting from no information, and with each statement adding more.
 Information is here seen in the information theoretic way, where no information is when everything is possible ($\top$) and total information is when nothing is possible ($\bot$).
@@
 
-\subsection{The problem with disjunction}
+\subsubsection{The problem with disjunction}
 
@@
 
-\subsection{Connection with first order logic}
+\subsubsection{Connection with first order logic}
 
@@
 
-\subsection{Monadic variants}
-
-Let's call them `Visser variants' or `VDPL'. Hm, I like that.
-
-What should we call those? Visser calls them DPL variants, e.g. DPL$_{\Bowtie\triangle}$
+\subsection{Donkey Monads*}
 
@@
 
-\subsection{Montague semantic}
+\subsection{Montague Semantics}
 Montague semantics is a theory of natural language that believes its semantics can be described thoroughly in terms of standard mathematical models.
@@
 
-\subsection{Weak and strong readings}
+\subsubsection{Weak and strong readings*}
 
@@
 
-\subsection{Dependency grammars}
+\subsection{Dependency grammars*}
 
@@
 
-\tikzstyle{level 1}=[level distance=2.5cm, sibling distance=3.5cm]
+\tikzstyle{level 1}=[level distance=2cm, sibling distance=3.5cm]
+\tikzstyle{level 2}=[level distance=2.5cm, sibling distance=3.5cm]
+\tikzstyle{level 3}=[level distance=2cm, sibling distance=3.5cm]
 \begin{figure}
@@
 \subsection{Derivation of DPL preconditions}
-\begin{align}
+Remember that we have defined, for assignments $\alpha$ and $\beta$, DPL formula $\psi$ and FOL formula $\phi$: $\alpha\models\langle\psi\rangle\phi$ to mean $\exists\beta(\alpha\llbracket\psi\rrbracket\beta\wedge\beta\models\phi)$.
+$\alpha\models\phi$ is just the usual $\models$ from FOL.
+
+Since the proof is over the syntax of DPL, we use $\llbracket\cdot\rrbracket$ as the interpretation function so that $\alpha\llbracket\psi\rrbracket\beta$ is the FOL statement `$(\alpha,\beta)\in\Psi$', where $\Psi$ is the relation defined by $\psi$.
+\begin{align}
+%
+\intertext{Let's start out with the most tricky derivation. How does the resetting, unbounded $\exists$ of DPL translate into the defining, bounded $\eixsts$ of FOL?}
+%
 \alpha\models\langle\exists x\rangle\phi
- & \leftrightarrow \exists\beta (\alpha[\exists x]\beta \wedge \beta\models\phi) \nonumber\\
+ & \leftrightarrow \exists\beta (\alpha\llbracket\exists x\rrbracket\beta \wedge \beta\models\phi) \nonumber\\
  & \leftrightarrow \exists\beta (\forall\omega (\omega \neq x \rightarrow \alpha_{\omega} = \beta_{\omega}) \wedge \beta\models\phi) \nonumber\\
@@
 %
-\intertext{Here we take $\alpha[x:=\nu]$ to mean the $\alpha$ with $\alpha_x$ assigned to the value of $\nu$.}
+\intertext{Where we have taken $\alpha[x:=\nu]$ to mean the $\alpha$ with $\alpha_x$ assigned to the value of $\nu$.
+The derivation for $\cdot$ goes more smoothly:}
 %
 \alpha\models\langle\psi_1\cdot\psi_2\rangle\phi
- & \leftrightarrow \exists\beta (\alpha[\psi_1\cdot\psi_2]\beta \wedge \beta\models\phi) \nonumber\\
- & \leftrightarrow \exists\beta(\exists\gamma (\alpha[\psi_1]\gamma \wedge \gamma[\psi_2]\beta) \wedge \beta\models\phi) \nonumber\\
- & \leftrightarrow \exists\gamma (\alpha[\psi_1]\gamma \wedge \exists\beta(\gamma[\psi_2]\beta \wedge \beta\models\phi)) \nonumber\\
- & \leftrightarrow \exists\gamma (\alpha[\psi_1]\gamma \wedge \gamma\models\langle\psi_2\rangle\phi) \nonumber\\
+ & \leftrightarrow \exists\beta (\alpha\llbracket\psi_1\cdot\psi_2\rrbracket\beta \wedge \beta\models\phi) \nonumber\\
+ & \leftrightarrow \exists\beta(\exists\gamma (\alpha\llbracket\psi_1\rrbracket\gamma \wedge \gamma\llbracket\psi_2\rrbracket\beta) \wedge \beta\models\phi) \nonumber\\
+ & \leftrightarrow \exists\gamma (\alpha\llbracket\psi_1\rrbracket\gamma \wedge \exists\beta(\gamma\llbracket\psi_2\rrbracket\beta \wedge \beta\models\phi)) \nonumber\\
+ & \leftrightarrow \exists\gamma (\alpha\llbracket\psi_1\rrbracket\gamma \wedge \gamma\models\langle\psi_2\rangle\phi) \nonumber\\
  & \leftrightarrow \alpha\models\langle\psi_1\rangle\langle\psi_2\rangle\phi
 %
-\intertext{c}
+\intertext{$\cup$ interestingly doesn't translate to $\langle\psi_1\vee\psi_2\rangle\phi$ as we might have expected:}
 %
 \alpha\models\langle\psi_1 \cup \psi_2\rangle\phi
- & \leftrightarrow \exists\beta (\alpha[\psi_1 \cup \psi_2]\beta \wedge \beta\models\phi) \nonumber\\
- & \leftrightarrow \exists\beta ((\alpha[\psi_1]\beta \vee \alpha[\psi_2]\beta) \wedge \beta\models\phi) \nonumber\\
- & \leftrightarrow \exists\beta (\alpha[\psi_1]\beta \wedge \beta\models\phi) \vee \exists\beta (\alpha[\psi_2]\beta \wedge \beta\models\phi) \nonumber\\
+ & \leftrightarrow \exists\beta (\alpha\llbracket\psi_1 \cup \psi_2\rrbracket\beta \wedge \beta\models\phi) \nonumber\\
+ & \leftrightarrow \exists\beta ((\alpha\llbracket\psi_1\rrbracket\beta \vee \alpha\llbracket\psi_2\rrbracket\beta) \wedge \beta\models\phi) \nonumber\\
+ & \leftrightarrow \exists\beta (\alpha\llbracket\psi_1\rrbracket\beta \wedge \beta\models\phi) \vee \exists\beta (\alpha\llbracket\psi_2\rrbracket\beta \wedge \beta\models\phi) \nonumber\\
  & \leftrightarrow (\alpha\models\langle\psi_1\rangle\phi) \vee (\alpha\models\langle\psi_2\rangle\phi) \nonumber\\
@@
 %
-\intertext{d}
+\intertext{For predicates it's mostly a matter of juggling variables, values and assigments:}
 %
 \alpha\models\langle P(x_1,\dots,x_n)\rangle\phi
- & \leftrightarrow \exists\beta (\alpha[P(x_1,\dots,x_n)]\beta \wedge \beta\models\phi) \nonumber\\
+ & \leftrightarrow \exists\beta (\alpha\llbracket P(x_1,\dots,x_n)\rrbracket\beta \wedge \beta\models\phi) \nonumber\\
  & \leftrightarrow \exists\beta (\alpha = \beta \wedge P(\alpha_{x_1},\dots,\alpha_{x_n}) \wedge \beta\models\phi) \nonumber\\
@@
 %
-\intertext{f}
+\intertext{$\neg$ is interesting because it has to create a scope:}
 %
 \alpha\models\langle\neg\psi\rangle\phi
- & \leftrightarrow \exists\beta (\alpha[\neg\psi]\beta \wedge \beta\models\phi) \nonumber\\
- & \leftrightarrow \exists\beta (\alpha=\beta \wedge \neg\exists\gamma(\alpha[\psi]\gamma) \wedge \beta\models\phi) \nonumber\\
- & \leftrightarrow \neg\exists\gamma(\alpha[\psi]\gamma) \wedge \alpha\models\phi \nonumber\\
- & \leftrightarrow \neg\exists\gamma(\alpha[\psi]\gamma \wedge \gamma\models\top) \wedge \alpha\models\phi \nonumber\\
+ & \leftrightarrow \exists\beta (\alpha\llbracket\neg\psi\rrbracket\beta \wedge \beta\models\phi) \nonumber\\
+ & \leftrightarrow \exists\beta (\alpha=\beta \wedge \neg\exists\gamma(\alpha\llbracket\psi\rrbracket\gamma) \wedge \beta\models\phi) \nonumber\\
+ & \leftrightarrow \neg\exists\gamma(\alpha\llbracket\psi\rrbracket\gamma) \wedge \alpha\models\phi \nonumber\\
+ & \leftrightarrow \neg\exists\gamma(\alpha\llbracket\psi\rrbracket\gamma \wedge \gamma\models\top) \wedge \alpha\models\phi \nonumber\\
  & \leftrightarrow \alpha\not\models\langle\psi\rangle\top \wedge \alpha\models\phi \nonumber\\
@@
 %
-\intertext{e}\nonumber
+\intertext{Because $\psi$ is this way `evaluated on $\top$ only', we know that no variables introduced in $\psi$ can affect $\phi$.}\nonumber
 \end{align}

### 2013-05-09 paper.tex
Wrote implementation stuff

@@
         child {
-            node {Det}
+            node {DT}
             edge from parent
@@
             child {
-                node {Det}
+                node {DT}
                 edge from parent
@@
 
-alpha/beta renaming: We don't need it because we use globally unique variable names
-
-Montague grammars have been used for many years in the quest of machine translation.
-The idea was that to use logic as a `universal language' to which we could translate any sentence, and ascend back into whatever language we'd like.
-Hence you would think that the web was flooded with code and long machine readable lists of rules, but this is however not the case.
-
-Perhaps the lack of usable material is due to the fact that most computer scientists abandoned the formal approach two decades ago for instead to try their luck with machine learning and vector space models.
-These models are however starting to show their weaknesses, specially in very logical sentences, and people are starting to wonder if there are still things to be learned from Montague's approach.
+The requirements for the testing framework are very simple: Support as big a chunk of English as possible, while keeping to the intended meaning.
+
+Since this program is going to be used for research by computational lingusts, easy access to modify the code is more useful than big user interfaces and parameterization.
+Our interface is going to be simple text based, with an input file of text you want translated and a command you run on said file to start producing logifications.
+
+\subsection{Interfacing with Stanford}
+
+The Stanford Parser tools provide a rich Java library for interfacing directly with the system.
+They have classes for things like constituent trees, dependencies, correferences and more.
+Obviously I couldn't use any of that.
+Instead I ran a simple shell script they include, which outputs the relevant information in XML.
+
+Haskell has a nice XML library in Text.XML.Light.
+However the constituent found by the parser were not actually in xml, but in a simple plain text format.
+To parse this I wrote my own parser using the great monadic Parsec library.
+
+The next step for my code was to tag the relevant tree leafs with correct variable names, so it could go into the Montague code.
+As stated before the Stanford toolchain provided me with correference tagging, but unfortunately not in a great format for my needs.
+If a sentence said `A man and a dog, they had a shower', it would assign the same `mention' to `a man' and `a dog', but nothing to `a shower'.
+Luckily Haskell is quite good at such tree manipulation.
+One thing I made sure was that I always use distinct variables so I never have to perform $\alpha$-conversion.
+
+One major issue with the Stanford tools is the fact that they are very slow.
+It uses less than a second per input text, but actually starting the program takes more than two minutes.
+This is due to the suite loading large compressed files of language statistics into RAM.
+This issue is the reason my program works with batches of inputs.
+
+\subsection{Syntax and Semantics}
+
+An interesting issue is the fact that we deal with 4 types of logic.
+Each having a syntactic and semantic part.
+That means we have 8 kinds of truth ($\top$), 8 kinds of implication, 8 kinds of everything.
+And this is with only two Donkey Monoids implemented.
+
+The mentioned types are the following:
+
+\begin{haskell}
+data Fol = T | F | ,,,
+type IFol = Interpretation -> Bool
+
+data Dpl = DT | DF | ,,,
+type IDpl = Interpretation -> Assignment -> Assignment -> Bool
+type IDpl2 = Interpretation -> Assignment -> [Assignment]
+
+data Visser1 = VT | VF | ,,,
+type IVisser1 = (Dpl, Dpl, Integer)
+data Visser2 = VT | VF | ,,,
+type IVisser2 = (Dpl, Dpl, Dpl, Dpl, Integer, Integer)
+\end{haskell}
+
+Notice that DPL has two possible interpretations\footnote{This is more of curiosity due to the fact that a DPL formula is a relation between two assignments, and that an assignment can either be thought of as a set of pairs: $P(D^V \times D^V)$ or a function from type $X$ to a set of type $X$: $D^V \rightarrow P(D^V)$.
+The first one is nice for proofs, but the second one has a nice and short programmatic definition.}.
+None of these are actually used in the final program, but they serve well to test this middle layer of the program.
+
+For each syntactic type I've defined a number of functions:
+\begin{description}
+\item[`simplify'] which takes a formula of the given syntactic type and repeatedly applies a number of identities thought to simplify the logic\footnote{This is extra important because the conversion of Donkey Monoids into DPL and DPL into FOL creates a lot of `dummy' $\top$s.}.
+\item[`tex' and `pretty'] which prints the formula in either \TeX or Unicode format.
+It's in a way main part of the program's user interface.
+Hence I have taken extra care to make sure it has visually pleasing bracketing.
+\item[`int'] which converts the formula to its semantic counterpart.
+There are also interpretation functions that interpret a logic in the logic `below' it, hence creating one directional conversion functions.
+\end{description}
+
+In terms of the semantic types we have to be careful.
+Types such as `Interpretation $\rightarrow$ Bool' forgets to allow for basic `errors' such as references to unbound variables.
+For FOL we can check this statically and not allow the syntax, but for a DPL sentence $(\exists x\cdot P(x) \cup \exists y)\cdot Q(x)$ the existance of $x$ in $Q(x)$ depends on the value of $P(x)$.
+
+One option in case of an error is to simply return $\bot$ and say the formula evaluated to false, but perhaps we want to distinguish the cases.
+I have experimented with a lot of types, such as `Interpretation $\rightarrow$ Maybe Bool', which can return the value `Nothing' in case of an error.
+It doesn't make the formulas any nicer though.
+If we decide to even implement Visser's `many world' semantics as discussed earlier, we get absolutely terrible formulas that don't honour the simplicity of the logic in any way.
+
+In the current version of my program, I have decided to go with the simplest possible type, and simply report errors through Haskell's error system.
+
+\subsection{The grammar}
+
+\newsavebox{\LstBox}
+\begin{lrbox}{\LstBox}
+\begin{haskell}
+data PosTree a = P PosTag [PosTree a] | L PosTag a
+type IWord = (Word, Index)
+\end{haskell}
+\end{lrbox}
+
+Once I have sculpted the input text into a nice tree shape\footnote{I use the type `PosTree IWord' where:\\
+\usebox{\LstBox}}, I parse it to my `montague' function. As I described in the early chapters, this applies a set of rules recursively to create a formal interpretation.
+
+Now, since Montague grammars have been around for a long time, and used in commercial machine translating systems from the 70ies and 80ies, you would think the web was would be flooded with code and long machine readable lists of rules I could use to bootstrap my system.
+This doesn't seem to be the case.
+Perhaps this is due to people abandoning this approach for vector models before the internet had really picked up the speed of today.
+
+Some people are however wondering if there might still be things to learn from Montague's approach, like Groendijk \& Stokhof in Dynamic Montague Grammar\cite{groenendijk1990dynamic}:
 \begin{quotation}
 `...
-we are convinced that the capacities of [Montague Grammar] have not been exploited to the limit, that sometimes an analysis is carried out in a rival framework simply because it is more fashionable' - Groendijk \& Stokhof in `Dynamic Montague Grammar'\cite{groenendijk1990dynamic}
+we are convinced that the capacities of [Montague Grammar] have not been exploited to the limit, that sometimes an analysis is carried out in a rival framework simply because it is more fashionable'
 \end{quotation}
 
-Nonetheless the lack of machine tested composition rules means that nearly all of the rules I shall present are of my own creation.
-Because these rules are tested in practice on a large set of English sentences, they hopefully avoid some mistakes and oversights that I have found in purely theoretical papers.
-It seams like the correct approach to start with a small working fragment, and slowly add more phrase rules and sentences.
+Nonetheless, while that particular paper does have some interesting lists, it is worrying that not even Montague's original papers supply more than a few examples.
+
+This means that nearly all of the rules I shall present are of my own creation.
+Because these rules are tested in practice on a large set of English sentences, they will hopefully avoid some simple mistakes and oversights that are easy to make in purely theoretical settings.
+It seams like a good approach to start from a small working fragment, and slowly add more phrase rules and sentences.
 The opposite of this approach is what is often observed when logicians and philosophers go astray in very abstract corners of English, without having covered the basic ground.
 
-Incremental dynamics\quote{van2001incremental}
-
-DPL Semantics - Functional Version
-DPL Semantics - Relational Version
-A state is an element of $D^V$ (function from variables to domain)
-In functional semantics a state is a function $D^V -> P(D^V)$
-In relational semantics it is a relation $P(D^V x D^V)$
-
-There is not a big first order logic component in my project, but in the following it will often be useful to refer to the following structure
-
-\begin{haskell}
-data Prop =
-    T | F
-  | Not   Prop
-  | And   Prop Prop
-  | Or    Prop Prop
-  | Predi Ref [Ref]
-  | Exist Ref Prop
-  | All Ref Prop
-  | Impl Prop Prop
-\end{haskell}
-
-And since we are going to deal with some large computer generated formulas, I've implemented a simple simplification function, which simply applies basic identities for $\top$ and $\bot$, and a combination of De Morgan's laws and the implication identity to eliminate negations.
-
-One interesting part is that since the simplifier has no look ahead, it will sometimes not find all applicable simplifications in one pass. For example
-%
-\begin{haskell}
-simplify (Not (Impl p F)) = Not (simplify (Impl p F))
-                          = Not (Not simplify p)
-                          = Not (Not q)
-\end{haskell}
-%
-Hence we need to run it a couple of times to get a good result. We can run it until the output doesn't change, or even simpler, just use a function like
-%
-\begin{haskell}
-fexp :: (a -> a) -> Int -> (a -> a)
-fexp f e = (iterate (f.) id) !! e
-\end{haskell}
-
-\subsection{My implementation of DPL semantics}
-
-Syntactically we will be using the following structure
-
-\begin{lstlisting}
-data DPL =
-    DT | DF
-  | DComp DPL DPL
-  | DOr DPL DPL
-  | DNot DPL
-  | DImpl DPL DPL
-  | DExists Ref
-  | DPredi Ref [Ref]
-\end{lstlisting}
-
-Below are some direct implementation of the semantics for DPL. We use the types
-\begin{lstlisting}
-type Assignment = [(Ref, Val)]
-type Relation = Assignment -> [Assignment]
-\end{lstlisting}
-This differs from the indirect set of types are used in the earlier formalisms:
-\begin{lstlisting}
-type Assignment = [(Ref, Val)]
-type Relation = Assignment -> Assignment -> Bool
-\end{lstlisting}
-
-intuitive
-
-uses well known isomorphism
-
-more direct
-
-The later is very difficult to implement, since the semantics require quantification over all assignments. Like for composition.
-
-\begin{lstlisting}
-:: Relation
-true = id
-false = const []
-
-predi1 :: [Val] -> Ref -> Rel
-predi1 f k = filter (\as -> elem (get k as) f)
-
-predi2 :: [(Val,Val)] -> Ref -> Ref -> Rel
-predi2 f k l = filter (\as -> elem (get k as, get l as) f)
-
-:: Relation -> Relation
-neg r = filter (\as -> null (r [as]))
-
-:: Relation -> Relation -> Relation
-comp = flip (.)
-impl s r = rnot (s `comp` (rnot r))
-\end{lstlisting}
-Then comes the most interesting one. To ``randomly reset'' a variable, we have to make sure that we have all assignments in the cross product of new values for the variable and what was there before.
-
-Bum bum. Maybe we should just use a set, then we could overwrite variable assignments
-\begin{lstlisting}
-exist :: Ref -> [Val] -> Rel
-exist k vs ass = [a : as | a <- news, as <- olds]
-	where olds = map (filter ((/=k) . fst)) ass
-	      news = [(k,v) | v <- vs]
-\end{lstlisting}
-
-\subsection{Handling variables}
-One thing I haven't found much \cite{visser1999donkey}
-Stanford: \cite{lee2013deterministic}
-\cite{lee2011stanford}
-\cite{raghunathan2010multi}
-Coreferences
-Using the parse tree and NPs
-
-The free variable thing is a problem in the sense that we have decided the meaning of a sentence is a truth value. Maybe we should change it to assignment -> Truth...
-
-Something that's very interesting is when to use free variables and when to use bound variables.
-We can think of free variables as actors known by the context and bound variables as actors introduced in the sentence.
-So if we had `The farmer had a friend.
-His name was Anders', we could write the first part as $\exists y \cdot friends(x,y)$ and the second part as $name(y, `Anders')$.
-This is important to keep compositionality, but it does of course raise the interesting question of how to make sure that the variable for `Anders' in the first and the second sentence is the same.
-
-In addition to being essential for composition, free variables are also very useful for information extraction type sentences.
-We can imagine writing a curious researcher writing into a search engine `farmers who like donkeys' which translates into $farmer(x) \cdot likes(x,y) \cdot \triangle \cdot \exists y \cdot donkey(y)$.
-
-So how do we know when to include an $\exists$ and when not to?
-
-The coreference parser does however not find entities that are only mentioned once. Hence to make the later processing simpler, we assign those entity to free variable now as well.
-
-The task is slightly complicated by the fact that NP's can be nested, so `The farmer and the donkey' is a NP consisting of two smaller NP's `The farmer' and `the donkey' separated by `and'.
-
-We resolve this by assigning top down, overwriting `word to ref' assignments with fresh variables as we go deeper down the tree.
-This way, for our short sentence, we might end up with the assignment $\{`The': `y', `farmer': `y', `and': `x', `the': `z', `donkey': `z'\}$
-
-I parse the constituent tree using parsec\cite{leijen2001parsec}
-
-\begin{lstlisting}
-toMapping :: [Ref] -> ConstituentTree Word -> ([Ref], Map Int Ref)
-toMapping vs (Leaf _ _) = (vs, empty)
-
-toMapping vs tr@(Phrase "NP" subs) = (vs', submaps `union` np)
-  where np = mapInterval 0 (tsize tr) v
-        (v:vs', submaps) = toMapping vs (Phrase "" subs)
-
-toMapping vs (Phrase _ []) = (vs, empty)
-toMapping vs (Phrase _ (sub:subs)) = (vs'', p `union` shifted)
-  where (vs', subm) = toMapping vs (Phrase "" subs)
-        shifted = mapKeys (+tsize sub) subm
-        (vs'', p) = toMapping vs' sub
-\end{lstlisting}
-
-where $tsize :: ConstituentTree Word -> Int$ gives the number of words in the tree
-
-\subsubsection{Errors}
-
-What do we do when they fail?
-Also something from the disjunction chapter.
-
-What if my states always contain the entire domain as their domain. Just add a $\bot$ to the entity set, then everything will be mapped to that as default.
-
-\subsection{The grammar}
-This is not as bad as it sounds. It is similar to Montague and what the Stanford Parser does for dependencies.
+The following describes the set of rules I use in my program:
 
@@
 
-\subsubsection{Using machine learning}
-The $\bowtie$ and $\triangle$ notation makes logic look close enough to natural language that it starts looking tempting to use machine learning methods such as those used in translation to translate.
-
-My main problem here is that such translation requires large amounts of bilingual data. It was out of scope for my project to try create such.
+The mentioned `montague' function basically passes the tree to $\llbracket\cdot\rrbracket_\text{ROOT}$ and does a bit of cleaning of the result.
+
+\subsection{Machine Learning}
+
+Machine Learning is one gaping hole we haven't mentioned in this paper.
+Obviously something as dynamic as natural language we can never just define a set of rules like we did above.
+Programming a computer to learn a set of rules like that is unfortunately not something much research has gone into, and so no large banks of input/output pairs exist to learn from.
+
+Another idea for logification using machine learning is to entirely skip the Montague step.
+The $\Bowtie$ and $\triangle$ monoid makes logic look close enough to natural language, that it starts to become tempting utilizing machine learning methods from translation.
+Unfortunately this idea runs into the same barrier as the above: that for superviced learning, we first need a large base of already translated sentences.
 
@@
 \begin{quotation}
-Beware of bugs in the above code; I have only proved it correct, not tried it.
+`Beware of bugs in the above code; I have only proved it correct, not tried it.'
 \end{quotation}
@@
 In the actual implementation of my program I have used the \emph{QuickCheck} and \emph{HUnit} test environments to write unit tests for algebraic properties and actual inputs to my program.
-To enable QuickTest of the different properties on constituent trees, a random generator of sentences was implemented.
+To enable QuickTest to test the different preparation algorithms on constituent trees, I implemented a random generator\footnote{In conclusion QuickTest wasn't as useful as I had hoped, since defining my functions as properties to be tested, required far more code than the functions themselves.}.
 
@@
 In DPL we don't have that particular problem, but instead we can run into problems with the way the scopes of composed elements affect each other.
-In Visser logic it is even worse, as $\Bowtie$s and $\triangle$s introduced much later in the discourse, can get transposed far around the final statement.
+In Visser logic it is even worse, as $\Bowtie$s and $\triangle$s can teleport elements introduced much later in the discourse far around the final statement.
 
@@
 
+This concludes the implementation part of the paper.
+
 \section{Perspectives/Conclusions (2)}
+
+Issue: Stanford is very slow. Loads huge libraries of English statistics
+Output: unicode or \TeX
+Input: I chose to just put everything in the haskell file, so I could add comments
 
@@
 %
-\intertext{Let's start out with the most tricky derivation. How does the resetting, unbounded $\exists$ of DPL translate into the defining, bounded $\eixsts$ of FOL?}
+\intertext{Let's start out with the most tricky derivation. How does the resetting, unbounded $\exists$ of DPL translate into the defining, bounded $\exists$ of FOL?}
 %
@@
  & \leftrightarrow \exists\beta (\alpha\llbracket\exists x\rrbracket\beta \wedge \beta\models\phi) \nonumber\\
+ & \text{This is the definition from the DPL chapter:}\nonumber\\
  & \leftrightarrow \exists\beta (\forall\omega (\omega \neq x \rightarrow \alpha_{\omega} = \beta_{\omega}) \wedge \beta\models\phi) \nonumber\\
+ & \text{The trick is this switch to assignments:}\nonumber\\
  & \leftrightarrow \exists\beta (\exists\nu (\forall\omega (\alpha[x:=\nu]_{\omega} = \beta_{\omega})) \wedge \beta\models\phi) \nonumber\\
@@
  & \leftrightarrow \exists\nu (\exists\beta (\alpha[x:=\nu] = \beta \wedge \beta\models\phi)) \nonumber\\
+ & \text{Here we apply the $\models$ definition:}\nonumber\\
  & \leftrightarrow \exists\nu (\alpha[x:=\nu]\models\phi) \nonumber\\
@@
 
+\subsection{Source Code*}
+
+Should I include this?
+
 \bibliographystyle{plain}

### 2013-05-09 paper.tex
Conclusions

@@
 
-\section{Perspectives/Conclusions (2)}
-
-Issue: Stanford is very slow. Loads huge libraries of English statistics
-Output: unicode or \TeX
-Input: I chose to just put everything in the haskell file, so I could add comments
-
-Move 1: Summary 
-Summarizes the most important aspects of the project work.
-
-I have successfully developed a novel local type inference algorithm
-
-‘a summary of your achievements; a critical appraisal, describing what worked well and what could be 
-improved; a discussion of the lessons you have learnt from the project.’
-
-In the papers it is always assumed that the constituent tree and things are `trivial' or `already sorted', but how do the pieces really stack up when put together?
-
-Visser: The constructions studied in this paper can be viewed as adding control elements to the language.
-A control element is a symbol representing an instruction concerning storage of bits of meaning.
-Thus, control elements serve to steer semantic processing.
+\section{Conclusions and Perspectives (2)}
+
+I have successfully developed a testing framework that lets researchers define logics and interpretation functions and test them on large inputs of interesting text.
+I have discussed many possible directions I could have taken, I have proved every result involved, and I have done extensive testing.
+
+My program can correctly logify entire paragraphs of text containing advanced anaphora, conjunctive nouns, verbs and adjectives, implications and scoping.
+It handles sentences with free variables as well as entirely factual sentences.
+
+A few examples of sentences that my framework already supports, but which fail due to short commings in the Stanford toolkit are `Every dog sees a cat.
+It chases it.' where the anaphora doesn't get resolved, and `Buffalo buffalo Buffalo buffalo buffalo buffalo Buffalo buffalo' that manages to slightly confuse the parser.
+
+My tests of Visser's Donkey Monoids show that they can be used for automatic logification, but that they lose some of their charm.
+Mainly it seems that the $\Bowtie$ and $\triangle$ switches are a bit too strong, and hence require extensive use of the scoping function $?(\cdot)$.
+At least the rules I came up with essentially uses Donkey Monoids as if they were DPL.
+
+My tests also show that the Montague grammar can be made very simple when implemented using dynamic logics.
+I couldn't however entire eliminate the need for lambda calculus, but perhaps a more refined Donkey Monoid will be up for the challenge.
+
+\subsection{Future work}
+
+I had a lot of issues with the Stanford Parser not being able to parse sentences correctly.
+In every paper I have read, this part always `just works', but I didn't have any options, but to discard the sentence.
+It would be interesting to make a fault tolerant logification system.
+Perhaps some knowledge from the semantics could be passed back to the syntax parser for better results.
+
+It would also be interesting to return more than one result in the case of ambiguity.
+We have seen how weak and strong readings of the actors in the sentences can give very different results.
+Also conjunction of transitive verbs like `A farmer starves and beats a donkey' are inherently ambigious.
+Perhaps if we returned many possible logifications for each sentence, we could take the interferance to get the real meaning.
+
+In terms of expanding the scope of our translations, the fruits I have identified are hanging at different heights:
+
+At the lowest level there are a lot of trivial additions.
+Support in more cases for more negatives like `no' and `nobody' hasn't been explored well.
+Possessive pronouns is something I've struggeled with, but if we accept that `He took his jacket' translates to $\exists y\cdot owns(x,y)\cdot jacket(y)\cdot took(x,y)$ it shouldn't be hard.
+Also there are things like `existential there' and more kinds of SBARs like started by `then', `whom' etc.
+that we can add, but for the expense of additional complicated rules.
+
+Higher up are things like adverbs.
+This will require a more complicated treatment of predicates to hopefully allow something like $furiously(beat(x,y))$.
+How exactly these things can be logified is still an open question in linguistics.
+Another fruit at this level is model verbs.
+Sentences like `It rains.
+It might rain' and `It might rain.
+It rains' should be treated differently.
+For this we could look into the field of modal logics.
+
+Even higher up are lots of classical linguistic problems.
+Sentences such as `The golden mountain does not exist', `The ancient greeks thought the morning star was the evening star' have had entire papers written on them, without any definive answers.
+Probably we'll have to consider the use case to know how a well defined computational logic should work.
+Another thing at the top level is novel words and sentence structures.
+These will clearly require some sort of machine learning.
+
+Finally I have mentioned that we might be able to expand the Donkey Monoids with more functionality as we sees fit during Montague rule making.
+The possibility for pushing control elements into the semantic of the logic is really the beauty of Visser's approach and unfortunately a road I haven't studied much in this paper.
+It is not hard to immagine $(Dpl,Dpl,Dpl,Dpl,Integer,Integer)$ exchanged for something like $(Dpl,...,PosTag)$ and coresponding composition rules.
+
+\subsection{Personal Report*}
+
+What the...
 
@@
 \end{itemize}
-
-DPL made it possible to specify a lot of rules that would otherwise have been terrible complicated. Most notably `if'.
-
-Monoids were an interesting alternative to lambda composition. It could have been interesting to see how far they could go.
-
-It would be interesting to return more than one result in the case of ambiguity. Weak/strong readings and things like transitive verbs. `A farmer starves and beats a donkey'.
-
-A lot more rules could relatively easily be added to the grammar.
-For some things like adverbs we would need a richer language, like $furiously(beats(x,y))$, but there's certainly still a lot of low hanging fruits.
-
-Level 1, easy:
-* More uses of `no' and `nobody'
-* possessives pronouns, 
-* `existential there'
-* more kinds of SBARs like started by `then', `whom' etc.
-
-Level 2, more work:
-* adverbs
-* modals verbs are a big task
-
-Level 3, requires new understanding of language:
-* Existential non-commitment (The golden mountain does not exist)
-* The ancient greeks thought the morning star was the evening star
-* Understanding words never used before
-
-Level X, not working due to the framework:
-* Every dog sees a cat. It chases it.
-* Buffalo buffalo Buffalo buffalo buffalo buffalo Buffalo buffalo
-
-Move 2: Evaluation
-Indicates what the writer considers to be the most valuable aspects of the project and its limitations. Provides 
-an assessment of how far the aims of the project have been achieved.
-
-In the ‘Conclusions’ you summarize your work and then stand back from the project and assess it as  objectively as you can.
-This shows the examiners that you are capable of evaluating your own work according to the standards of the field.
-It is acceptable to mention areas that were not very satisfactory because this shows that you have learnt from the experience of doing the project.
-
-Everything was perfect!
-It feel
-
-my algorithm is exponential: whereas Gagnon's algorithm is 
-polynomial. But experiments of execution time against method length show a typically linear trend whereas 
-Gagnon's show a cubic trend
-
-Move 3: Future Work
-Gives possible extensions of the project. These suggestions may follow from the limitations mentioned in the 
-Evaluation move.
-move may also be carried out as part of the Summary. The Conclusion is the mirror image of the 
-Introduction, in that it starts with the narrow concerns of the project and widens out to more general 
-statements about further work.
-
-The greatest scope for future work is extending the application of my algorithm from local type inference to global type inference.
 

### 2013-05-09 paper.tex
done

@@
 
-%In Discourse Representation Theory (DRT) our donkey sentence `If a farmer owns a donkey, he beats it' is represented as
-%\begin{align}
-%&[_1 x\mid Pedro(x), [_2 y, z\mid owns(y,z), donkey(z)] \Rightarrow [_3 v, w\mid beats(v,w)]] \nonumber
-%\intertext{This form is created from the syntax tree, and DRT then provides rules for eliminating variables and thus providing anaphora:}
-%&[_1 x, v: v = x, Pedro(x), [_2 y, z, w: y = x, w = y, owns(y,z), donkey(z)] \Rightarrow [_3 : beats(v,w)] ] \nonumber\\
-%&[_1 x: Pedro(x), [_2 z: donkey(z), owns(x,z)] \Rightarrow [_3 : beats(x,z)] ] \nonumber
-%\end{align}
-%The DRT merge
-
 \item File Change Semantics, 1982 by Irene Heim\cite{heim1983file}, are based around the puzzle of discourse `referents' which may, as we have seen, sometimes act like quantifiers, sometimes referents and sometimes antecedents.
@@
 
-\subsection{Donkey Monads*}
-
-In the previous subsection we saw that DPL allowed us to rewrite the donkey sentence as
-
-\begin{equation}
-\exists x \cdot farmer(x) \cdot \exists y \cdot donkey(y) \cdot owns(x,y) \rightarrow beats(x,y)
-\end{equation}
-
-($\cdot$ binds stronger than $\rightarrow$)
-
-But think about the slight modification to the sentence:
+\subsection{Donkey Monoids}
+
+Imagine translating this sentence to DPL:
 %
@@
 %
-We would have to write that as
+We would have to write something along the lines of:
 %
@@
 
-Should we have one more example here?
+Another example from Visser is:
+\begin{quotation}
+A man keeps his word. He is honest
+\end{quotation}
+which roughly translates to:
+\begin{equation}
+\Bowtie\cdot\exists x\cdot man(x)\cdot\Bowtie\cdot keepword(x)\cdot honest(x)
+\end{equation}
+If you are interested in a lot more examples, see Visser's paper\cite{visser1999donkey}.
 
@@
 
-More examples on the utilities of the above monad can be found in Vissers paper.
-
 An even more interesting element than $\Bowtie$ is $\triangle$. $\triangle$ is a scope modifier, and it allows us to write:
@@
 %
-That is pretty close to the order of the natural language!
+That is very close to the order of the natural language!
 
@@
 
-\subsubsection{Weak and strong readings*}
-
-In the process of defining semantics for language, ambiguity in sentenses is of course a serious problem. A sentence like  
+\subsubsection{Weak and strong readings}
+
+`Weak' and `strong' are quite accepted terms in linguistics for describing different ways to interprete quantifiers.
+In line with information theory, we can say that weak readings are about weak information and strong readings about strong information.
+
+Makoto Kanazawa has written an entire paper on issues with weak and strong readings of donkey sentences\cite{kanazawa1994weak}.
+For the donkey sentence, we may have at least the following readings of the second `a':
+\begin{enumerate}
+\item Weak reading: A farmer who owns a donkey beats a donkey he owns. 
+\item Strong reading: A farmer who owns a donkey beats every donkey he owns 
+\item Unique reading: A farmer who owns a donkey beats the donkey he owns. 
+\end{enumerate}
+
+We have equally many reaings for the first `A', creating nine possible interpretations.
+This is because `a' is a very ambigious quantifier.
+Other quantifiers such as `every' and `all' are more clearly defined.
+`Some' is another example of an ambigious quantifier.
+
+Often we can use the context of a sentence to determine which meaning is the correct one. For example:
 \begin{quotation}
-A farmer beats a donkey
+Every farmer owns a donkey. It is pink.
 \end{quotation}
-Might be translated as 
+Here it is clear that all the farmers share one donkey. Hence perhaps the best solution to the problem is to generate all possible sentences and let the context decide which ones survive.
+
+\subsection{Dependency Graphs}
+
+\tikzstyle{level 1}=[level distance=2cm, sibling distance=3.5cm]
+\tikzstyle{level 2}=[level distance=2.5cm, sibling distance=3.5cm]
+\tikzstyle{level 3}=[level distance=2cm, sibling distance=3.5cm]
+\begin{figure}
+\centering
+\begin{tikzpicture}[grow=down, sloped]
+\node {root}
+    child {
+        node {owns}
+        child {
+            node {If}
+            edge from parent
+            node[above] {mark}
+        }
+        child {
+            node {farmer}
+            child {
+                node {a}
+                edge from parent
+                node[above] {det}
+            }
+            edge from parent
+            node[above] {nsubj}
+        }
+        child {
+            node {beats}
+            child {
+                node {he}
+                edge from parent
+                node[above] {nsubj}
+            }
+            child {
+                node {it}
+                edge from parent
+                node[above] {dobj}
+            }
+            edge from parent
+            node[above] {advcl}
+        }
+        child {
+            node {donkey}
+            child {
+                node {a}
+                edge from parent
+                node[above] {det}
+            }
+            edge from parent
+            node[above] {dobj}
+        }
+        edge from parent
+        node[above] {root}
+    };
+\end{tikzpicture}
+\caption{Dependency graph for `If a farmer owns a donkey, he beats it'} \label{fig:dependency_graph}
+\end{figure}
+
+Dependency graphs are an alternative way of parsing sentences.
+Where constituent trees keep the original word order, dependency graphs are allowed to move thing around in any way they like\footnote{Until now we've been lucky that things next to each other often were related, but in some languages other than English, words can jump anywhere they like, and our approach might fail.}.
+Also dependency graphs don't have to be trees.
+
+The flexibility of dependency graphs allow them to do a lot of interesting things.
+In our favourite example of the donkey and the farmer we had the problem that there wasn't a very strong syntactical relationship between `farmer' and `he'.
+In figure \ref{fig:dependency_graph} we see how the same sentence would look if parsed with a dependency parser.
+
+The first thing we notice is that dependency graphs focus on verbs.
+The verbs become the heads of the sentences, and everything else becomes the ` subject', `the object' or `the something else' of the verb.
+In our example `he beats it' becomes an `adverbial clause modifier' of the main verb.
+This means it modifies the verb, here in the sence of making the main verb a condition.
+
+Dependency graphs naturally seems to offer more semantic information than constituent trees, and hence that they would be a good intermediate step towards our goal of logification.
+However the first disappointment arrise as we start to look at how they are created.
+
+The Stanford Parser makes dependency graphs using a special kind of regular expressions for trees called `tregex'\cite{de2006generating}. It defines tregex rules such as:
+%
 \begin{equation}
-\exists x ( farmer(x) \wedge \exists y ( donkey(y) \wedge beats(x,y)))
+NP < NN\ \$\ VP
 \end{equation}
-Indicating that ``a farmer'' refers to some specific individual. Or we can might see it as a general rule, and translate it as
+Which means an NP that is the parent of an NN and sister of a VP.
+
+Another more complicated example is:
 \begin{equation}
-\forall x ( farmer(x) \rightarrow \exists y ( donkey(y) \wedge beats(x,y)))
+NP < (NN < dog)\ \$\ (VP <\!\!<\# (barks > VBZ))
 \end{equation}
-Indicating that every farmer has some donkey he beats. Or we could even read it as
-\begin{equation}
-\exists y ( donkey(y) \wedge \forall x ( farmer(x) \rightarrow beats(x,y)))
-\end{equation}
-Indicating that the poor donkey is common between all farmers.
-% This seams wrong to me
-
-Additionally, the indefinite article `a' is normally understood as an existential quantifier
-
-Week and strong readings, check this out \cite{kanazawa1994weak}
-
-drt
-
-The upshot of the foregoing observations is that, apparently, indefinites are neither quantifiers nor referential terms, and this problem entrains another one, for as long as it unclear what indefinites mean, it will also remain obscure how they can serve as antecedents to pronouns
-
-Is this the same as the problem of multiple `scope bearing operators?'
-Sentences with multiple scope bearing operators - e.g.,
-quantified noun phrases or negations - are often ambiguous.
-
-%$• “Every student reads a book”
-%- ∀x(student(x) ⇒ ∃y(book(y) ∧ read(y)(x)))
-%- ∃y(book(y) ∧ ∀x(student(x) ⇒ read(y)(x)))
-
-Every farmer owns a donkey. It is pink. Second sentence decides the reading.
+Which means an NP over an NN over `dog' and with a sister VP headed by `barks' which is a verb in present singular.
+
+Obviously this means we can't get any information from the graph that we couldn't also get from the tree.
+In fact it probably contains less.
+On the flip side however, the regular expressions are clearly more powerful than simple recursions as we define them in the Montague grammars.
+
+I've worked hard on trying to define a Montague like set of equations for Enlish, but over the dependency graph.
+This has however failed, mostly due to one simple reason: In a Montague grammar we expect every type of phrase to have a specific logical type.
+E.g.
+all NPs are `VP $\rightarrow$ Bool' and VPs are `Entity $\rightarrow$ Bool'.
+This makes it fit very nice in a programming context, where we can define a function `parseNP :: ConstituentTree $\rightarrow$ (VP $\rightarrow$ Bool)'.
+With the dependency approach we don't have this option.
+A `subj' arrow can mean many different things.
+Sometimes it points at a noun, sometimes at a proper noun, sometimes something else.
+
+In conclusion, Montague like grammars for dependency graphs could be very interesting to make.
+It would however require some welldefined rules for what we can expect to find under each type of arrow.
+This was a blind road.
 
@@
 The Stanford procedure above is currently the strongest competitor in various coreference competitions.
-
-\subsection{Dependency grammars*}
-
-This is an important thing to mention.
-
-Extracted from constituent trees using tregex\cite{de2006generating}
-
-\tikzstyle{level 1}=[level distance=2cm, sibling distance=3.5cm]
-\tikzstyle{level 2}=[level distance=2.5cm, sibling distance=3.5cm]
-\tikzstyle{level 3}=[level distance=2cm, sibling distance=3.5cm]
-\begin{figure}
-\centering
-\begin{tikzpicture}[grow=down, sloped]
-\node {root}
-    child {
-        node {owns}
-        child {
-            node {If}
-            edge from parent
-            node[above] {mark}
-        }
-        child {
-            node {farmer}
-            child {
-                node {a}
-                edge from parent
-                node[above] {det}
-            }
-            edge from parent
-            node[above] {nsubj}
-        }
-        child {
-            node {beats}
-            child {
-                node {he}
-                edge from parent
-                node[above] {nsubj}
-            }
-            child {
-                node {it}
-                edge from parent
-                node[above] {dobj}
-            }
-            edge from parent
-            node[above] {advcl}
-        }
-        child {
-            node {donkey}
-            child {
-                node {a}
-                edge from parent
-                node[above] {det}
-            }
-            edge from parent
-            node[above] {dobj}
-        }
-        edge from parent
-        node[above] {root}
-    };
-\end{tikzpicture}
-\caption{M1} \label{fig:M1}
-\end{figure}
-
-\begin{equation}
-NP < NN \$ VP
-\end{equation}
-
-An NP over an NN and w/ sister VP
-
-\begin{equation}
-NP < (NN < dog) \$ (VP <<\# (barks > VBZ))
-\end{equation}
-
-An NP both over an NN over `dog' and with
-a sister VP headed by `barks' under VBZ
-
-%\subsubsection{Semantics via dependencies}
-
-%This might need updating based on: http://en.wikipedia.org/wiki/Dependency_grammar and http://en.wikipedia.org/wiki/Phrase_structure_grammar
-
-It is an interesting idea to consider that we might be able to define our semantics over a dependency graph, in a similar way to the Montague approach described earlier.
-The dependency graph adds a lot of flexibility where the constituent tree used in the Montague approach is limited to the direct syntax of the tree.
-A good example is our favourite sentence `if a farmer owns a donkey, he beats it', where there isn't a very strong syntactical relationship between `farmer' and `he'.
-It's also been suggested that it is mostly due to luck that a lot of English fits relatively well to the Montague approach, where as in some languages the structure of the constituent tree might not say much about the actual semantics.
-
-I've worked hard on trying to define a Montague like set of equations for Enlish, but over the dependency graph.
-This has however failed, mostly due to one simple reason: In a Montague grammar we expect every type of phrase to have a specific logical type.
-E.g.
-all NPs are `VP -> Bool' and VPs are `Entity -> Bool'.
-This makes it fit very nice in a programming context, where we can define a function `parseNP :: ConstituentTree -> (VP -> Bool)'.With the dependency approach we don't have this option.
-A `subj' arrow can mean many different things.
-Sometimes it points at a noun, sometimes at a proper noun, sometimes something else.
-
-Another problem worth considering is the fact that all information in the graph is drawn syntactically from the constituent tree.
-That is, the graph doesn't contain any information not in the tree, but the other direction is entirely possible.
-While it is hard to say exactly what's missing, it is worth noting.
 
@@
 \section{Acknowledgments}
-I would like to thank Samson Abrahamski for introducing me to the world of Natural Language parsing and semantics, for helpful and stimulating discussins and for keeping my on the right path
-
-Thanks to Phil Blunsom and Edward Grefenstette for exciting and enlightening lectures into the classical and modern approaches.
+I would like to thank Samson Abrahamski for introducing me to the world of Natural Language parsing and semantics and for helpful and stimulating discussions.
+
+Also thanks to Phil Blunsom and Edward Grefenstette for exciting and enlightening lectures into the classical and modern approaches.
 
@@
 
-Thanks to Érica Baena Kato for keeping me awake with interesting linguistic perspectives.
-
-Thanks to The Donkey Sanctuary\cite{donkey2013sanctuary} for actually taking donkeys welfare seriously and rescuing neglected donkeys in the UK and all over the world.
+Thanks to Érica Baena Kato for providing interesting linguistic perspectives on many issues.
+
+Finnaly I wish to thanks to The Donkey Sanctuary\cite{donkey2013sanctuary} for taking donkey's welfare seriously and rescuing neglected donkeys in the UK and all over the world.
 

### 2013-05-19 paper.tex
Made Main.hs read from a file Added more appendices

@@
 \usepackage{enumitem}
+\usepackage{courier}
 
@@
 
-\lstloadlanguages{Haskell}
-\lstnewenvironment{haskell}
-    {\lstset{}%
-      \csname lst@SetFirstLabel\endcsname}
-    {\csname lst@SaveFirstLabel\endcsname}
-    \lstset{
-      basicstyle=\small\ttfamily,
+\lstdefinestyle{haskellstyle}{%
+  basicstyle=\scriptsize\ttfamily,
+  language=Haskell,
   escapeinside={/+}{+/},
   flexiblecolumns=false,
-      xleftmargin=0.05\textwidth,
+  xleftmargin=-0.05\textwidth,
   basewidth={0.5em,0.45em},
+  breaklines=true,
   literate={+}{{$+$}}1
@@
            {/=}{{$\neq$}}2
-               {.}{{$\circ$}}2
+%           { . }{{$\circ$}}2
            {<<}{{$\ll$}}2
@@
            {`elem`}{{$\in$}}1
-    }
+           {⋈}{{$\Bowtie$}}1
+           {⋅}{{$\cdot$}}1
+           {△}{{$\triangle$}}1
+           {∀}{{$\forall$}}1
+           {∃}{{$\exists$}}1
+           {⊥}{{$\bot$}}1
+           {⊤}{{$\top$}}1
+           {→}{{$\Rightarrow$}}1
+           {∨}{{$\vee$}}1
+           {¬}{{$\neg$}}1
+           {∧}{{$\wedge$}}1
+}
+
+\lstdefinestyle{bashstyle}{%
+  basicstyle=\scriptsize\ttfamily,
+  escapeinside={/+}{+/},
+  flexiblecolumns=false,
+  xleftmargin=-0.05\textwidth,
+  breaklines=true,
+  basewidth={0.6em,0.45em},
+  literate = {⋈}{{$\Bowtie$}}1
+           {⋅}{{$\cdot$}}1
+           {△}{{$\triangle$}}1
+           {∀}{{$\forall$}}1
+           {∃}{{$\exists$}}1
+           {⊥}{{$\bot$}}1
+           {⊤}{{$\top$}}1
+           {→}{{$\Rightarrow$}}1
+           {∨}{{$\vee$}}1
+           {¬}{{$\neg$}}1
+           {∧}{{$\wedge$}}1
+}
+
+\lstnewenvironment{haskell}{\lstset{style=haskellstyle}}{}
 
@@
 
-Donkey Monoids are a set of variations of Dynamic Predicate Logic (DPL) introduced by Albert Visser.
-The rationale behind them is a stride towards a logic that can more easily approximate natural language in aspects such as polarity and scope.
+In this paper I build a fully automatic system for translation of natural language into formal logic.
+I combine a variety of tools and theories to target standard issues such as anaphora resolution, sentence polarity etc.
+In particular I base my framework on dynamic logics in form of Dynamic Predicate Logic and a set of enhancements by Albert Visser\cite{visser1999donkey}.
+
+`Donkey Monoids', as I will call these monoidal DPL variants, are part of a movement in linguistics towards logics that can more naturally approximate human language in aspects such as polarity and scope.
 Indeed their name derives from the famous `Donkey Sentence' by Geach\cite{geach1962reference} which exhibit these issues so well.
@@
 Visser didn't test his ideas computationally and neither did most people involved with DPL.
-In fact most of the classical literature on logification of natural language has this `problem'.
-In this paper I supply such an experiment, combining technologies introduced over the last four decades, creating a fully functioning `English to logic' framework.
-
-The transformation rules of the supported English fragment are defined as a Montague Grammar, the semantics of which is discussed and defined in the first part of the paper.
+In fact most of the classical literature on logification of natural language exhibits this `problem': Not testing computationally makes it easy to miss obvious details.
+With the present paper I supply such an experiment.
+
+I build a rewriting system with a well defined set rules, discussing possible semantics.
 I apply the framework to a large collection of sentences, showing the conversion process step by step.
@@
 \smallskip\smallskip
-\noindent \small{\textbf{Keywords:} Montague Grammar, Dynamic Predicate Logic, Natural Language, Meaning, Donkey, Monoid, Computational, Haskell}
+\noindent \small{\textbf{Keywords:} Natural Language Semantics, Meaning, Dynamic Predicate Logic, Donkey, Monoid, Montague Grammar, Computational, Haskell}
 
@@
 
-\section{Introduction (5)}
+\section{Introduction}
 % 5 pagish Donkey example to follow through.
@@
 For a computer scientist, the biggest advantage is that we can compute with logic.
-There is a horde of algorithms for doing querying, question answering, reasoning, unification and resolution.
-They key to this powerhouse is obtaining a first order logic form of your information.
+There is a horde of algorithms for doing querying, question answering, reasoning, unification and resolution - to all of which they key is a formal logic definition of your information.
 
@@
 \item[Robots] Since the beginning of Artificial Intelligence, logic based approaches have done well.\cite{russell2010artificial} Robots need to store relationships they learn about the world and react to it.
-For modern `robots' like apples Siri to be able to communicate with people, a translation layer between natural language and logic is a plausible path.
+For modern `robots' like Apple's Siri to be able to communicate with people, a translation layer between natural language and logic is a plausible path.
 \end{description}
@@
 The reason logics like DPL and Visser Monoids exist is a famous sentence noted by Peter Thomas Geach in 1962\cite{geach1962reference}: `Every farmer who owns a donkey beats it'.
-As we will see, this sentence illustrates that even if we restrict ourselves to factual sentences, formal first order logic still runs into a lot of problems when representing many natural sentences.
+As we will see, this sentence illustrates that even if we restrict ourselves to factual sentences, formal first-order logic still runs into a lot of problems when representing many natural sentences.
 In fact this sentence broke early automatic approaches to logification, and led to the construction of Discourse Representation Theory (DRT)\cite{kamp1993discourse}.
@@
 
-\item `All his donkeys are treated well': This sentence somewhat willingly let's itself translate into: $\forall y(his(x,y)\wedge donkey(y) \rightarrow treated\_well(y))$, which has a free variable $x$ referring to some (male?) entity in the context of the discourse.
-
-\item `Every farmer beats his donkey': This sentence clearly has two possible meanings: $\forall x(farmer(x)\rightarrow \exists y (beats(x,y)\wedge his(x,y)\wedge donkey(y))$ if `his' refers to the farmers or $\exists y(his(x,y)\wedge donkey(y)\wedge \forall z(farmer(z) \rightarrow beats(z,y))$ if it refers to somebody else.
+\item `All his donkeys are treated well': This sentence somewhat willingly lets itself translate into: $\forall y(his(x,y)\wedge donkey(y) \rightarrow treated\_well(y))$, which has a free variable $x$ referring to some (male?) entity in the context of the discourse.
+
+\item `Every farmer beats his donkey': This sentence clearly has two possible meanings: $\forall x(farmer(x)\rightarrow \exists y (beats(x,y)\wedge his(x,y)\wedge donkey(y))$, if `his' refers to the farmers, or $\exists y(his(x,y)\wedge donkey(y)\wedge \forall z(farmer(z) \rightarrow beats(z,y))$ if it refers to somebody else.
 There is no way we can know without more context information.
@@
 
-These examples should hopefully have shown that some uses of English, especially example 4 and 5, are well beyond what we can hope to automatically logify.
-However you might think the sentences of `factual nature' like 1 and 2 shouldn't be too hard?
-Unfortunately, as sentences get longer and more complex we will see that they are, and this is where Donkey Monads, Dynamic Predicate Logic (DPL) and Discourse Representation Theory (DRT) have a role to play.
+Hopefully those examples make a convincing argument that some uses of English, especially example 4 and 5, are well beyond what we can hope to automatically logify.
+However you might think the sentences of `factual nature' like 1 and 2 look pretty trivial?
+In this paper we will see that they are certainly possible, but as sentences get longer and more complex, certainly tricky, and this is where Donkey Monoids, Dynamic Predicate Logic (DPL) and Discourse Representation Theory (DRT) come into play.
 Let's take a close look at that `Donkey Sentence': `If a farmer owns a donkey, he beats it'.
@@
 A far cry from our original `guess'.
-Moreover the above version even have the oddity that the interpretation of `all' is `the logical one': The formula \eqref{correct_donkey} is, in addition to the intended situations, also considered true if no farmers exist in the domain, or none of them own donkeys.
+Moreover the above version even has the oddity that the interpretation of `all' is `the logical one': The formula \eqref{correct_donkey} is, in addition to the intuitive situations\footnote{For non logicians at least.}, also considered true if no farmers exist in the domain, or none of them own donkeys.
 
@@
 Wasn't some other strategy more obvious?
-Before we go into details of such logification strategies, let's first see how easy it could have been, if we instead of first order logic had used Visser's Donkey Monoids:
+Before we go into details of such logification strategies, let's first see how easy it could have been, if we instead of first-order logic had used Visser's Donkey Monoids:
 %
@@
 
-Neat as the Donkey Monoids are, interested as we are in automation, we should be very concerned about those magical $\Bowtie$s and $\triangle$s that have popped up in our formula.
-What do they mean?
-How do they work?
-And how do we know when to use them?
-Visser himself, while answering question 1 and 2 in detail, doesn't give many hints on question 3.
-Only that they are probably controlled by adverbs and similar `helping words'.
+Neat as the Donkey Monoids are, and interested as we are in their automation, we should be very concerned about those magical $\Bowtie$s and $\triangle$s which have popped up in our formula.
+We'll want to ask questions such ask (1) What do they mean?
+(2) How do they work?
+And (3) how do we know when to use them?
+Visser himself, while answering question 1 and 2 in great detail, doesn't give many hints on 3.
+Only that they `switches' are probably controlled by adverbs and other support words.
 
@@
 
-In the early sixties the American logician, Richard Montague, was a teaching a class in which the students had to do logification of simple sentences.
+In the early sixties the American logician, Richard Montague, was teaching a class in which the students had to do logification of simple sentences.
 He noticed that there were many patterns in the way they did it, and conjectured that it would be possible to formalize the entire process.
@@
 
-The Montague Grammar, as Montague's approach came to be know, works amazingly well, it's simplicity taken into account.
-It is not hard to see how one might convince oneself that it could be used to describe all of language.
+The Montague Grammar, as Montague's approach came to be know, has survived in computational linguistics for decades.
+Perhaps its main charm is in its rich promises for a simple, consistent way to describe all of language.
 The basic idea is to define logical formulas recursively over the so called constituent tree of a sentence.
 The tree is based on the idea of context free grammars for natural language developed mainly by Noam Chomsky\cite{chomsky2002syntactic}.
-Figure \ref{fig:const_tree} is an example of what it might look like.
+Figure \ref{fig:const_tree} is an example of what one might look like.
 
@@
 The main thing to notice now is that for Montague's approach to work, it is crucial that we can always get the meaning of a composed phrase from a composition of the meaning of its subphrases.
-This compositional property is exactly what we saw earlier that first order logic doesn't posses, and Montague goes through many hoops to simmulate it.
-
-I do however avoid many of those problems, since my goal here is not to test the validity of Montague's proposed rules, but to experiment with how Donkey Monoids and to an extent DPL behaves when put in the framework described.
-To draw a full circle I am however then going to translate the Donkey Monoids into first order logic so we can easily check if our formulas have the meanings we expect.
+This compositional property is exactly what we saw earlier that first-order logic doesn't posses, and Montague goes through many hoops to simulate it.
+
+I do however avoid many of those problems, since my choices of logic behave much more compositionally than the FOL in Montague's proposed rules.
+To draw a full circle I am however going to translate the obtained Donkey Monoid representations into first-order logic so we can easily check if our formulas have the meanings we expect.
 
@@
 \item[2 POS tagger] This is the first of three Stanford NLP programs my Haskell script is going to pass the text to.
-The POS tagger tokenizes the sentences and statstically tries to figures out part of speach for each word.
-This is our first source of error.
+Then the POS tagger tokenizes the sentences and statstically tries to figures out part of speech for each word.
+This is also my first source of error.
 
@@
 
-\item[4 Correference Resolution] A lot have been said about how anaphora might be resolved using formal Montague Grammar like approaches.
+\item[4 Correference Resolution] A lot has been said about how anaphora might be resolved using formal Montague Grammar like approaches.
 DRT is especially focused on this topic.
-None the less the current champeon of anaphora competitions\cite{lee2011stanford}\cite{raghunathan2010multi} is the heuristical Stanford Deterministic Coreference Resolution System, the workings of which I will explain in detail.
+None the less the current champion of anaphora competitions\cite{lee2011stanford}\cite{raghunathan2010multi} is the heuristical Stanford Deterministic Coreference Resolution System, the workings of which I will explain in detail.
 
@@
 
-\item[6 Dynamic Predicate Logic] Since the Donkey Monoids are variants of DPL, they are easy to convert into it. It is much more surprising that DPL lets itself convert into first order logic.
-
-\item[7 First Order Logic] We end up translating our dynamic logics into this old friend.
-This is mainly done to test the conversion algorithm, and because the first order logic is what we are used to read.
+\item[6 Dynamic Predicate Logic] Since the Donkey Monoids are variants of DPL, they are easy to convert into it. It is much more surprising that DPL lets itself convert into first-order logic.
+
+\item[7 First-Order Logic] We end up translating our dynamic logics into this old friend.
+This is mainly done to test the conversion algorithm, and because the first-order logic is what we are used to read.
 The results can be surprising.
@@
 
-\section{Formal Stuff (10)}
+\section{Dynamic Semantics and Montague}
 
@@
 
-In the previous chapter I showed how First Order Logic lack certain compositional properties.
+In the previous chapter I showed how first-order logic lacks certain compositional properties.
 A possible solution to this problem is to use a so called `update logic' (a logic with update semantics).
@@
 
-The idea of those logics is a `growth of information in time'.
-Starting from no information, and with each statement adding more.
-Information is here seen in the information theoretic way, where no information is when everything is possible ($\top$) and total information is when nothing is possible ($\bot$).
+The idea of those logics is a `growth of information in time', starting from no information, and with each statement adding more.
+Information is here seen in the semantic way, where no information is when everything is possible ($\top$) and total information is when nothing is possible ($\bot$).
 
@@
 
-Notice here in particular the interesting the `dependent' arrow.
+Notice here in particular the interesting `dependent' arrow.
 This arrow communicates that there is a one-many relationship between farmer and donkey, and hence that we are talking about a specific donkey for each farmer, not a single poor donkey beaten by everyone.
@@
 Just like giving meaning to the actual concepts, this must be handled outside of the logic.
-We will see later in this paper that this stance makes sense, since the current state of the art anaphora resolvers are simply based on simple heuristics over noun phrases.
-
-In the end of this chapter we will see how a certain monadic construction by Visser can give us as much flexibility in DPL as we would get with a more involved system such as DRT or FCS.
+We will see later in this paper that this stance makes sense, since the current state of the art anaphora resolvers are based on simple heuristics over noun phrases.
+
+In the end of this chapter we will see how a certain monoid construction by Visser can give us as much flexibility in DPL as we would get with a more involved system such as DRT or FCS.
 \end{itemize}
@@
 Dynamic Predicate Logic has been developed by a wide group of people: Jan van Eijck and Fer-Jan de Vries\cite{eijck1992dynamic}, Hans Kamp, Veltman, Groenendijk and Stokhof.
-It is the linguist version of Dynamic Logic invented by Pratt\cite{pratt1976semantical} which is again build on top of Floyd-Hoare Logic for reasoning about computer programs.
+It is the linguist version of Dynamic Logic invented by Pratt\cite{pratt1976semantical} which is again built on top of Floyd-Hoare Logic for reasoning about computer programs.
 
@@
 $\psi_1$ and $\psi_2$ cannot bind variables in each other, but in $(\psi_1\cup\psi_2)\cdot\psi_3$ they can both bind variables in $\psi_3$.
-This is going to introduce a lot of weird situations later.
-In the next section we will look closer at these problems, and also consider the alternative definition of `or': $\neg(\neg\psi_1\cdot\neg\psi_2)$.
-
-\subsubsection{The problem with disjunction}
-
-We won't actually use `or' in the implementation.
-Example of the problems we still have in the logic.
-Can't tell what variables are available where.
-
-A major problem with our use of `relation union' as disjunction is the way it allows seemingly sound sentences to combine into syntactically invalid ones.
-Consider the sentence `Either a farmer owns a donkey or he doesn't'.
-We would write that as
-\begin{align}
-\exists x\cdot farmer(x)\cdot(\exists y\cdot owns(x,y)\cdot donkey(y)\vee \bot) \label{owns_donkey_or_not}
-\end{align}
-Now concatenating this sentence with `he beats it': $beat(x,y)$ would give us a problem.
-In fact this problem would be more than a syntactic problem, because before we know the model and the existence of the donkey, we don't know if the sentence is valid or not.
-
-Visser\cite{visser1999donkey} suggests that we should build our logic on sets of relations read disjunctively instead of just relations.
-This `possible world' scenario would transform \eqref{owns_donkey_or_not}$\cdot beats(x,y)$ into $\{\exists x\cdot farmer(x)\cdot\exists y\cdot donkey(y)\cdot owns(x,y)\cdot beats(x,y), \exists x\cdot farmer(x)\cdot\bot\cdot beats(x,y)\} = \{\exists x\cdot farmer(x)\cdot\exists y\cdot donkey(y)\cdot owns(x,y)\cdot beats(x,y), \bot\}$.
-This clearly is always well defined.
-
-Visser's idea doesn't fully save us though. If we now consider `a farmer either owns a donkey or owns a horse':
-\begin{align}
-\exists x\cdot farmer(x)\cdot(&\exists y\cdot owns(x,y)\cdot donkey(y)\nonumber\\
-                              &\vee \exists z\cdot owns(x,z)\cdot horse(z))
-\end{align}
-And combines that with `He beats the donkey': $beat(x,y)$, we get, in Visser's model, $\{\exists x\cdot farmer(x)\cdot\exists y\cdot owns(x,y)\cdot donkey(y)\cdot beats(x,y), \exists x\cdot farmer(x)\cdot\exists z\cdot owns(x,z)\cdot horse(z)\cdot beats(x,y)\}$.
-The second one is clearly not valid (though we wouldn't get a problem if there is no horse or farmer) which we might catch `at compile time' and replace it with something like `$\bot$' or `error'.
-This sort of error handling is discussed and implemented in the implementation section, but is not terribly interesting from a formal point of view.
-
-Much more interesting are the following two questions: How could we have translated \label{owns_donkey_or_horse} so it would actually work?
-After all it seemed like a reasonable sentence until variable naming got us.
-And does the way variable binding works with our definition actually make sense?
-
-Clearly just renaming $z$ to $y$ in the horse term would have solved the first problem.
-However to work in a compositional way, that would either require great luck, or renaming, and we don't have a renaming operator.
-
-A trick we could consider is to declare an `object' variable outside of the disjunction:
-\begin{align}
-\exists s\cdot farmer(s)\cdot\exists o\cdot(
-      &\exists y\cdot owns(s,y)\cdot donkey(y)\cdot o=y \nonumber\\
-      &\vee \exists z\cdot owns(s,z)\cdot horse(z)\cdot o=z)\cdot beats(s,o)
-\end{align}
-%
-This even inspires on idea that we might be able to build everything up around subjects, objects, indirect objects etc.
-This is explored in the implementation, but it does suffer from many problems.
-E.g.
-we might well have a sentence very similar to the above: `A farmer owns a donkey and owns a horse.
-He beat them'.
-Here we suddenly have two objects in the same sentence.
-
-Sentences like the above are easy to make in English because the absence of noun genders make it easy to use the same pronoun to refer to a lot of things at once.
-By choosing the right actors however, it is always possible to run into problems.
-
-The problem is perhaps the non-linearity in the way variables pass through the disjunction. Somehow we lose track of their history when the `results' are merged together.
-
-This leads on to the other question we mentioned: Is the binding flow through our disjunction operator really any good?
-We are is inspired by the sentence `Either a farmer or his scapegoat must go to court.
-In any case he will be picked up by sunset'.
-In terms of DPL formulas, this would be modelled something like $(\psi_1\cup\psi_2)\cdot\psi_3$.
-Clearly we have a problem here: Variables bound in $\psi_1$ are not accessible in $\psi_2$.
-
-This sentence suggests that we might want to define a disjunction operator with a linear flow of variables: $\psi_1\to\psi_2\to\psi_3$ instead of $\psi_1\to\psi_3\text{ and }\psi_2\to\psi_3$.
-However there is a problem: We currently define `false' as the absence of possible assignments.
-Hence if $\psi_1$ is false, there is no way we can define disjunction to push it's variables into $\psi_2$.
-If any variables were left undefined, $\psi_1$ wouldn't be false.
-
-There is a cure to this problem.
-When we extend our logic with more power in the monad chapter, we get a second chance.
-However Let's first have a look at an obvious alternative definition for disjunction considered by Groenendijk\cite{groenendijk1991dynamic}.
-We'll call it `static disjunction':
-\begin{align}
-\alpha[\psi_1\vee\psi_2]\beta
- :\equiv&\ \alpha[\neg(\neg\psi_1\wedge\neg\psi_2)]\beta \nonumber\\
- \leftrightarrow&\ \alpha[\neg\psi_1\rightarrow\psi_2]\beta \nonumber\\
- \leftrightarrow&\ \alpha = \beta \wedge \forall\gamma(\alpha[\neg\psi_1]\gamma\rightarrow\gamma\models[\psi_2]) \nonumber\\
- \leftrightarrow&\ \alpha = \beta \wedge \forall\gamma((\alpha=\gamma\wedge\gamma\not\models\psi_1)\rightarrow\gamma\models[\psi_2]) \nonumber\\
- \leftrightarrow&\ \alpha = \beta \wedge \forall\gamma(\alpha=\gamma\rightarrow(\gamma\models\psi_1\vee\gamma\models\psi_2)) \nonumber\\
- \leftrightarrow&\ \alpha = \beta \wedge (\alpha\models\psi_1\vee\alpha\models\psi_2) \label{sem_or}
-\end{align}
-
-The idea to this follows natural from De Morgan's laws of classical logic.
-Clearly the semantics are different from the ones we defined for $\cup$.
-However not in a good way.
-This static disjunction doesn't allow any variables to escape from inside it's reach.
-
-So does this fix our problem of binding between clauses?
-No.
-In fact this definition doesn't even allow binding from a clause and out.
-The villain is $\neg\psi$.
-From the first mention we defined it to be static so it wouldn't allow unsafe, arbitrary assignments as long as they didn't satisfy $\psi$.
-We wanted our logic to be reasonably monotonic, and we got what we deserved.
-$\neg$ not only closed down this version of disjunction, but our implication definition as well.
-
-We can imagine defining $\neg$ dynamically: $\neg\psi_1\cdot\psi_2 = \neg(\psi_1\cdot\psi_2)$.
-However not in our current model, and it won't solve any of our problems with disjunction.
-It does however allow this sentence: `It is not the case that a farmer doesn't own a donkey.
-He beats it'.
-While it made sense not to allow variables to escape from a single $\neg$, we want them to escape from a double.
-
-\subsubsection{Connection with first order logic}
+We won't actually use `or' in the main implementation\footnote{Turns out the Stanford tools find it weird as well.}, but in the appendix~\ref{sec:problem_disj} I look more closely at the weird situations that can arrise.
+I also consider the alternative definition of `or': $\neg(\neg\psi_1\cdot\neg\psi_2)$.
+
+\subsubsection{Connection with First-Order Logic}
 
@@
 
-And combining the two monads we can do our original sentence as
+And combining the two monoids we can do our original sentence as
 %
@@
 
-For now we define the monad as follows:
+For now we define the monoid as follows:
 %
@@
 While the above remains a hot topic in the philosophy of language, we are going to take a more practical stand.
-In our quest to automate the translation of language into Visser's monadic DPL, it is interesting that Montague was able to define such a large fragment of English using familiar tools of compositionality and lambda calculus.
+In our quest to automate the translation of language into Visser's DPL monoid, it is interesting that Montague was able to define such a large fragment of English using familiar tools of compositionality and lambda calculus.
 
@@
 
-\subsubsection{Weak and strong readings}
+\subsubsection{Weak and Strong Readings}
 
@@
 
-The Stanford procedure above is currently the strongest competitor in various coreference competitions.
-
-\section{Implementation (8)}
+The Stanford procedure above is currently the strongest competitor in various coreference competitions.\cite{lee2011stanford}\cite{raghunathan2010multi}
+
+\section{Implementation}
 
@@
 \begin{haskell}
-data Fol = T | F | ,,,
+data Fol = T | F | ...
 type IFol = Interpretation -> Bool
 
-data Dpl = DT | DF | ,,,
+data Dpl = DT | DF | ...
 type IDpl = Interpretation -> Assignment -> Assignment -> Bool
@@
 
-data Visser1 = VT | VF | ,,,
+data Visser1 = VT | VF | ...
 type IVisser1 = (Dpl, Dpl, Integer)
-data Visser2 = VT | VF | ,,,
+data Visser2 = VT | VF | ...
 type IVisser2 = (Dpl, Dpl, Dpl, Dpl, Integer, Integer)
@@
 
-\subsection{The grammar}
+\subsection{The Grammar}
 
@@
 
-\section{Conclusions and Perspectives (2)}
+\section{Conclusions}
 
@@
 
-\subsection{Personal Report*}
-
-What the...
-
-\begin{itemize}
-\item Studied computational linguistics and learned about state of the art technologies
-\item Found out what was going on with dynamic logics and 
-similar discourse algebras
-\item Ran into many issues some of which have been discussed in great deal before (like strong vs. weak readings) and some that haven't had as much attention (like semantics of `and' and `or').
-\item Combined the technologies and theory in a way not tried before, which yielded very good results
-\end{itemize}
-
 \section{Acknowledgments}
@@
 \section{Appendix}
-\subsection{Derivation of DPL preconditions}
+
+\subsection{Derivation of DPL Preconditions}
 Remember that we have defined, for assignments $\alpha$ and $\beta$, DPL formula $\psi$ and FOL formula $\phi$: $\alpha\models\langle\psi\rangle\phi$ to mean $\exists\beta(\alpha\llbracket\psi\rrbracket\beta\wedge\beta\models\phi)$.
@@
 
-\subsection{Source Code*}
-
-Should I include this?
+\subsection{The Problem with Disjunction}\label{sec:problem_disj}
+
+%We won't actually use `or' in the implementation.
+%Example of the problems we still have in the logic.
+%Can't tell what variables are available where.
+
+A major problem with our use of the union of relations as disjunction is the way it allows seemingly sound sentences to combine into syntactically invalid ones.
+Consider the sentence `Either a farmer owns a donkey or he doesn't'.
+We would write that as
+\begin{align}
+\exists x\cdot farmer(x)\cdot(\exists y\cdot owns(x,y)\cdot donkey(y)\vee \bot) \label{owns_donkey_or_not}
+\end{align}
+Now concatenating this sentence with `he beats it': $beat(x,y)$ would give us a problem.
+In fact this problem would be more than a syntactic problem, because before we know the model and the existence of the donkey, we don't know if the sentence is valid or not.
+
+Visser\cite{visser1999donkey} suggests that we should build our logic on sets of relations read disjunctively instead of just relations.
+This `possible world' scenario would transform \eqref{owns_donkey_or_not}$\cdot beats(x,y)$ into $\{\exists x\cdot farmer(x)\cdot\exists y\cdot donkey(y)\cdot owns(x,y)\cdot beats(x,y), \exists x\cdot farmer(x)\cdot\bot\cdot beats(x,y)\} = \{\exists x\cdot farmer(x)\cdot\exists y\cdot donkey(y)\cdot owns(x,y)\cdot beats(x,y), \bot\}$.
+This clearly is always well defined.
+
+Visser's idea doesn't fully save us though. If we now consider `a farmer either owns a donkey or owns a horse':
+\begin{align}
+\exists x\cdot farmer(x)\cdot(&\exists y\cdot owns(x,y)\cdot donkey(y)\nonumber\\
+                              &\vee \exists z\cdot owns(x,z)\cdot horse(z))
+\end{align}
+And combines that with `He beats the donkey': $beat(x,y)$, we get, in Visser's model, $\{\exists x\cdot farmer(x)\cdot\exists y\cdot owns(x,y)\cdot donkey(y)\cdot beats(x,y), \exists x\cdot farmer(x)\cdot\exists z\cdot owns(x,z)\cdot horse(z)\cdot beats(x,y)\}$.
+The second one is clearly not valid (though we wouldn't get a problem if there is no horse or farmer) which we might catch `at compile time' and replace it with something like `$\bot$' or `error'.
+This sort of error handling is discussed and implemented in the implementation section, but is not terribly interesting from a formal point of view.
+
+Much more interesting are the following two questions: How could we have translated \label{owns_donkey_or_horse} so it would actually work?
+After all it seemed like a reasonable sentence until variable naming got us.
+And does the way variable binding works with our definition actually make sense?
+
+Clearly just renaming $z$ to $y$ in the horse term would have solved the first problem.
+However to work in a compositional way, that would either require great luck, or renaming, and we don't have a renaming operator.
+
+A trick we could consider is to declare an `object' variable outside of the disjunction:
+\begin{align}
+\exists s\cdot farmer(s)\cdot\exists o\cdot(
+      &\exists y\cdot owns(s,y)\cdot donkey(y)\cdot o=y \nonumber\\
+      &\vee \exists z\cdot owns(s,z)\cdot horse(z)\cdot o=z)\cdot beats(s,o)
+\end{align}
+%
+This even inspires on idea that we might be able to build everything up around subjects, objects, indirect objects etc.
+This is explored in the implementation, but it does suffer from many problems.
+E.g.
+we might well have a sentence very similar to the above: `A farmer owns a donkey and owns a horse.
+He beat them'.
+Here we suddenly have two objects in the same sentence.
+
+Sentences like the above are easy to make in English because the absence of noun genders make it easy to use the same pronoun to refer to a lot of things at once.
+By choosing the right actors however, it is always possible to run into problems.
+
+The problem is perhaps the non-linearity in the way variables pass through the disjunction. Somehow we lose track of their history when the `results' are merged together.
+
+This leads on to the other question we mentioned: Is the binding flow through our disjunction operator really any good?
+We are is inspired by the sentence `Either a farmer or his scapegoat must go to court.
+In any case he will be picked up by sunset'.
+In terms of DPL formulas, this would be modelled something like $(\psi_1\cup\psi_2)\cdot\psi_3$.
+Clearly we have a problem here: Variables bound in $\psi_1$ are not accessible in $\psi_2$.
+
+This sentence suggests that we might want to define a disjunction operator with a linear flow of variables: $\psi_1\to\psi_2\to\psi_3$ instead of $\psi_1\to\psi_3\text{ and }\psi_2\to\psi_3$.
+However there is a problem: We currently define `false' as the absence of possible assignments.
+Hence if $\psi_1$ is false, there is no way we can define disjunction to push it's variables into $\psi_2$.
+If any variables were left undefined, $\psi_1$ wouldn't be false.
+
+There is a cure to this problem.
+When we extend our logic with more power in the monoid chapter, we get a second chance.
+However Let's first have a look at an obvious alternative definition for disjunction considered by Groenendijk\cite{groenendijk1991dynamic}.
+We'll call it `static disjunction':
+\begin{align}
+\alpha[\psi_1\vee\psi_2]\beta
+ :\equiv&\ \alpha[\neg(\neg\psi_1\wedge\neg\psi_2)]\beta \nonumber\\
+ \leftrightarrow&\ \alpha[\neg\psi_1\rightarrow\psi_2]\beta \nonumber\\
+ \leftrightarrow&\ \alpha = \beta \wedge \forall\gamma(\alpha[\neg\psi_1]\gamma\rightarrow\gamma\models[\psi_2]) \nonumber\\
+ \leftrightarrow&\ \alpha = \beta \wedge \forall\gamma((\alpha=\gamma\wedge\gamma\not\models\psi_1)\rightarrow\gamma\models[\psi_2]) \nonumber\\
+ \leftrightarrow&\ \alpha = \beta \wedge \forall\gamma(\alpha=\gamma\rightarrow(\gamma\models\psi_1\vee\gamma\models\psi_2)) \nonumber\\
+ \leftrightarrow&\ \alpha = \beta \wedge (\alpha\models\psi_1\vee\alpha\models\psi_2) \label{sem_or}
+\end{align}
+
+The idea to this follows natural from De Morgan's laws of classical logic.
+Clearly the semantics are different from the ones we defined for $\cup$.
+However not in a good way.
+This static disjunction doesn't allow any variables to escape from inside it's reach.
+
+So does this fix our problem of binding between clauses?
+No.
+In fact this definition doesn't even allow binding from a clause and out.
+The villain is $\neg\psi$.
+From the first mention we defined it to be static so it wouldn't allow unsafe, arbitrary assignments as long as they didn't satisfy $\psi$.
+We wanted our logic to be reasonably monotonic, and we got what we deserved.
+$\neg$ not only closed down this version of disjunction, but our implication definition as well.
+
+We can imagine defining $\neg$ dynamically: $\neg\psi_1\cdot\psi_2 = \neg(\psi_1\cdot\psi_2)$.
+However not in our current model, and it won't solve any of our problems with disjunction.
+It does however allow this sentence: `It is not the case that a farmer doesn't own a donkey.
+He beats it'.
+While it made sense not to allow variables to escape from a single $\neg$, we want them to escape from a double.
+
+\subsection{Sample Output}
+
+\lstinputlisting[style=bashstyle]{sample_output}
+
+\subsection{Source Code}
+
+\lstinputlisting[style=haskellstyle,caption=\lstname]{src/Main.hs}
+\lstinputlisting[style=haskellstyle,caption=\lstname]{src/Stanford.hs}
+\lstinputlisting[style=haskellstyle,caption=\lstname]{src/ToLogic.hs}
+\lstinputlisting[style=haskellstyle,caption=\lstname]{src/Dpl.hs}
+\lstinputlisting[style=haskellstyle,caption=\lstname]{src/Visser.hs}
+\lstinputlisting[style=haskellstyle,caption=\lstname]{src/Tests.hs}
+\lstinputlisting[style=haskellstyle,caption=\lstname]{src/Utils.hs}
 

### 2013-05-19 run.sh

@@ -1,4 +1,4 @@
 #!/bin/bash
 cd "$(dirname $0)/src"
-runghc Main.hs
+runghc Main.hs inputs $1
 

### 2013-05-19 src/Main.hs

@@ -6,117 +6,38 @@ import qualified Visser as V
 import qualified Dpl as D
 import qualified Fol as F
 
-import Data.List
-import Text.XML.Light
 import Control.Monad
+import Data.List
 import Data.Maybe
-import System.Process
+import System.Environment
+import System.IO
 
-super :: [String] -> IO [V.Visser]
---super sentences = run sentences >>= (return . map montague)
-super sentences = runDry sentences >>= (return . map montague)
+readSentences :: FilePath -> IO [String]
+readSentences path = do
+		handle <- openFile path ReadMode
+		contents <- hGetContents handle
+		return [trimComments l | l <- lines contents, (not.null) (trimComments l)]
 
-sentences = [
-	"If a man eats a sausage, he is happy.",
-	"Every person has a pet.",
-	"Every man supports every woman.",
-	"Every man supports no woman.",
-	"Every farmer beats some donkey.",
-	"If a farmer owns a donkey, he beats it",
-
-	-- sequentials
-	"A man comes in. He sees a donkey. He smiles.", -- very sequential
-	"A farmer. A Donkey. If he owns it, he beats it.", -- sequential
-
-	-- ifs
-	"A dog barks. If it is beaten, it whines.", -- a donkey brays, but Stanford doesn't like it
-	"If a farmer owns a donkey, he beats it. If he rewards it, he doesn't own it.",
-	"If a farmer owns a donkey, he owns a horse. If he doesn't own it, he owns a cow.",
-	"If a woman is American, she loves a soldier. If she is Dutch, she loves a bike-rider.",
-
-	-- stringifying
-	"A farmer owns a donkey. It is brown and he beats it every day.",
-
-	-- doubles
-	"A farmer starves a horse and beats a donkey.", -- double vp
-	"All donkeys and a horse sing a song.", -- double np
-	"Some farmer and entrepreneur beats a donkey.", -- double description
-	"A man and a woman owned a horse and a donkey. If the donkey did not walk, and the man was a farmer, he beat it.", -- cross product
-
-	-- free variables
-	"She is a mother of five", -- mother becomes the root, because is is not a 'real' verb
-	"If a farmer owns it and he beats it, it is a donkey",
-	"No man walks in the park. He yodels.", -- coref checker screws this up, but dpl fixes it
-
-------------- I Can do until this line
-
[…]

### 2013-05-19 src/Utils.hs

@@ -13,3 +13,4 @@ replace pat repl target
 wrap :: Int -> (String, Int) -> String
 wrap n (p, k) | k > n     = "(" ++ p ++ ")"
               | otherwise = p
+

### 2013-05-19 src/Visser.hs

@@ -50,7 +50,7 @@ simp (VC p VT) = simp p
 simp (VC VT p) = simp p
 --simp (VC p VF) = VF
 --simp (VC VF p) = VF
--- TODO: Get rid of conjunctive, equal stream devices △△, ⋈⋈ etc.
+-- TODO: Get rid of conjunctive, equal stream devices
 -- Fall throughs
 simp (VC p q) = simp p `VC` simp q
 simp (VD p q) = simp p `VD` simp q

### 2013-05-19 paper.tex
Spell check

@@
 I apply the framework to a large collection of sentences, showing the conversion process step by step.
-Finally I discuss ways to enlarge the supported fragment of English, how we can combat and interprete the inherent ambiguity we meet, and how the flexibility of Donkey Monoids relate to the classical Montague lambda structures.
+Finally I discuss ways to enlarge the supported fragment of English, how we can combat and interpret the inherent ambiguity we meet, and how the flexibility of Donkey Monoids relate to the classical Montague lambda structures.
 
@@
 \begin{description}
-\item[Knowledge bases] In the field of Knowledge Representation and Reasoning, they long ago discovered that formal logic provides a good language for buliding ontologies.
+\item[Knowledge bases] In the field of Knowledge Representation and Reasoning, they long ago discovered that formal logic provides a good language for building ontologies.
 Typical users are crime fighters or life scientists that need to draw conclusions based on complex systems\cite{malone2010modeling} where most of the known facts are only described in natural language.
-Today those people use mainly have to manually build large models in gui tools, but an automated approach is much wanted.
+Today those people use mainly have to manually build large models in GUI tools, but an automated approach is much wanted.
 
@@
 \item `The donkeys are being stubborn!': This is an often used example of an `implied action' sentence and thought to have the same meaning as `Please go out and beat them!'.
-The best we can do here seems to be $please\_beat\_the\_doneys()$ or perhaps $\forall x(donkey(x)\rightarrow please\_beat(x))$ or even $please(\forall x(donkey(x)\rightarrow beat(x)))$.
+The best we can do here seems to be $please\_beat\_the\_donkeys()$ or perhaps $\forall x(donkey(x)\rightarrow please\_beat(x))$ or even $please(\forall x(donkey(x)\rightarrow beat(x)))$.
 
@@
 
-From the examples before, and perhaps from common intuition, we might find it natural to think of the sentence as two subsentences combined by `if': `if (a farmer owns a donkey) (he beats it)'.
+From the examples before, and perhaps from common intuition, we might find it natural to think of the sentence as two sub-sentences combined by `if': `if (a farmer owns a donkey) (he beats it)'.
 Translating the parentheses first: `A farmer owns a donkey' logifies to $\exists x (farmer(x) \wedge \exists y (donkey(y)\wedge owns(x,y))$, and `he beats it' translates to $beats(x,y)$.
@@
 %
-Which is totally wrong. The $beats(x,y)$ clause ends up refering to the variables outside of their scope, and even if we try to push it inside the parentheses:
+Which is totally wrong. The $beats(x,y)$ clause ends up referring to the variables outside of their scope, and even if we try to push it inside the parentheses:
 %
@@
 If we read $\exists$ as `a' this is basically identical to the English sentence.
-The donkey monoid even supports placing the transitive verb between the two referants.
+The donkey monoid even supports placing the transitive verb between the two referents.
 Clearly some magic is going on, that we will have to investigate further.
@@
 Basically there are two well described ways to go: statistical vector space methods, which is very much the most popular alternative right now, and logical methods.
-As you might have guessed I will lean in the dirrection of the later.
+As you might have guessed I will lean in the direction of the later.
 The most well described methods are Montague Grammars and Discourse Representation Theory (invented after the donkey sentence).
@@
 
-We will see later how exactly the Montague Grammar obtains a full sentence formula from its reqursive process, but in any case for a computer scientist the introduction of trees is always a promising step.
-
-The main thing to notice now is that for Montague's approach to work, it is crucial that we can always get the meaning of a composed phrase from a composition of the meaning of its subphrases.
+We will see later how exactly the Montague Grammar obtains a full sentence formula from its recursive process, but in any case for a computer scientist the introduction of trees is always a promising step.
+
+The main thing to notice now is that for Montague's approach to work, it is crucial that we can always get the meaning of a composed phrase from a composition of the meaning of its sub-phrases.
 This compositional property is exactly what we saw earlier that first-order logic doesn't posses, and Montague goes through many hoops to simulate it.
@@
 \item[2 POS tagger] This is the first of three Stanford NLP programs my Haskell script is going to pass the text to.
-Then the POS tagger tokenizes the sentences and statstically tries to figures out part of speech for each word.
+Then the POS tagger tokenizes the sentences and statistically tries to figures out part of speech for each word.
 This is also my first source of error.
@@
 
-\item[4 Correference Resolution] A lot has been said about how anaphora might be resolved using formal Montague Grammar like approaches.
+\item[4 Coreference Resolution] A lot has been said about how anaphora might be resolved using formal Montague Grammar like approaches.
 DRT is especially focused on this topic.
@@
 
-In these chapters I'm going to describe the logical theories by Visser and earlier dynamic semantists.
-I will introduce some addtions of my own and prove interesting properties about them.
+In these chapters I'm going to describe the logical theories by Visser and earlier dynamic semanticists.
+I will introduce some additions of my own and prove interesting properties about them.
 Finally, but most importantly, I will go into depth with the tools and ideas we need to write a strong logification framework.
@@
 The basic idea is very simple.
-For each postag we define a type.
+For each POS-tag we define a type.
 E.g.
@@
 
-Later in the implementation section we will look at more issues regarding to using Montague Grammars in practise. First we need to look at a very important source of ambiguity in sentences.
+Later in the implementation section we will look at more issues regarding to using Montague Grammars in practice. First we need to look at a very important source of ambiguity in sentences.
 
@@
 
-`Weak' and `strong' are quite accepted terms in linguistics for describing different ways to interprete quantifiers.
+`Weak' and `strong' are quite accepted terms in linguistics for describing different ways to interpret quantifiers.
 In line with information theory, we can say that weak readings are about weak information and strong readings about strong information.
@@
 
-We have equally many reaings for the first `A', creating nine possible interpretations.
-This is because `a' is a very ambigious quantifier.
+We have equally many readings for the first `A', creating nine possible interpretations.
+This is because `a' is a very ambiguous quantifier.
 Other quantifiers such as `every' and `all' are more clearly defined.
-`Some' is another example of an ambigious quantifier.
+`Some' is another example of an ambiguous quantifier.
 
@@
 In our example `he beats it' becomes an `adverbial clause modifier' of the main verb.
-This means it modifies the verb, here in the sence of making the main verb a condition.
+This means it modifies the verb, here in the sense of making the main verb a condition.
 
 Dependency graphs naturally seems to offer more semantic information than constituent trees, and hence that they would be a good intermediate step towards our goal of logification.
-However the first disappointment arrise as we start to look at how they are created.
+However the first disappointment arise as we start to look at how they are created.
 
@@
 
-I've worked hard on trying to define a Montague like set of equations for Enlish, but over the dependency graph.
+I've worked hard on trying to define a Montague like set of equations for English, but over the dependency graph.
 This has however failed, mostly due to one simple reason: In a Montague grammar we expect every type of phrase to have a specific logical type.
@@
 In conclusion, Montague like grammars for dependency graphs could be very interesting to make.
-It would however require some welldefined rules for what we can expect to find under each type of arrow.
+It would however require some well defined rules for what we can expect to find under each type of arrow.
 This was a blind road.
@@
 Coreference tagging is one part of computational linguistics where machine learning algorithms still haven't been able to outperform large iterative applications of linguistic heuristics.
-Of course the procedure assumes (most likely) machine learned postagging and named entity recognition, but the main algorithm is transparent and deterministic.
+Of course the procedure assumes (most likely) machine learned POS-tagging and named entity recognition, but the main algorithm is transparent and deterministic.
 
@@
 
-Since this program is going to be used for research by computational lingusts, easy access to modify the code is more useful than big user interfaces and parameterization.
+Since this program is going to be used for research by computational linguists, easy access to modify the code is more useful than big user interfaces and parametrization.
 Our interface is going to be simple text based, with an input file of text you want translated and a command you run on said file to start producing logifications.
@@
 The Stanford Parser tools provide a rich Java library for interfacing directly with the system.
-They have classes for things like constituent trees, dependencies, correferences and more.
+They have classes for things like constituent trees, dependencies, coreferences and more.
 Obviously I couldn't use any of that.
@@
 Haskell has a nice XML library in Text.XML.Light.
-However the constituent found by the parser were not actually in xml, but in a simple plain text format.
+However the constituent found by the parser were not actually in XML, but in a simple plain text format.
 To parse this I wrote my own parser using the great monadic Parsec library.
@@
 The next step for my code was to tag the relevant tree leafs with correct variable names, so it could go into the Montague code.
-As stated before the Stanford toolchain provided me with correference tagging, but unfortunately not in a great format for my needs.
+As stated before the Stanford tool chain provided me with coreference tagging, but unfortunately not in a great format for my needs.
 If a sentence said `A man and a dog, they had a shower', it would assign the same `mention' to `a man' and `a dog', but nothing to `a shower'.
@@
 Types such as `Interpretation $\rightarrow$ Bool' forgets to allow for basic `errors' such as references to unbound variables.
-For FOL we can check this statically and not allow the syntax, but for a DPL sentence $(\exists x\cdot P(x) \cup \exists y)\cdot Q(x)$ the existance of $x$ in $Q(x)$ depends on the value of $P(x)$.
+For FOL we can check this statically and not allow the syntax, but for a DPL sentence $(\exists x\cdot P(x) \cup \exists y)\cdot Q(x)$ the existence of $x$ in $Q(x)$ depends on the value of $P(x)$.
 
@@
 It doesn't make the formulas any nicer though.
-If we decide to even implement Visser's `many world' semantics as discussed earlier, we get absolutely terrible formulas that don't honour the simplicity of the logic in any way.
+If we decide to even implement Visser's `many world' semantics as discussed earlier, we get absolutely terrible formulas that don't honor the simplicity of the logic in any way.
 
@@
 Once I have sculpted the input text into a nice tree shape\footnote{I use the type `PosTree IWord' where:\\
-\usebox{\LstBox}}, I parse it to my `montague' function. As I described in the early chapters, this applies a set of rules recursively to create a formal interpretation.
+\usebox{\LstBox}}, I parse it to my `Montague' function. As I described in the early chapters, this applies a set of rules recursively to create a formal interpretation.
 
@@
 
-The mentioned `montague' function basically passes the tree to $\llbracket\cdot\rrbracket_\text{ROOT}$ and does a bit of cleaning of the result.
+The mentioned `Montague' function basically passes the tree to $\llbracket\cdot\rrbracket_\text{ROOT}$ and does a bit of cleaning of the result.
 
@@
 The $\Bowtie$ and $\triangle$ monoid makes logic look close enough to natural language, that it starts to become tempting utilizing machine learning methods from translation.
-Unfortunately this idea runs into the same barrier as the above: that for superviced learning, we first need a large base of already translated sentences.
+Unfortunately this idea runs into the same barrier as the above: that for supervised learning, we first need a large base of already translated sentences.
 
@@
 Until now we have been very focused on the same one or two simple sentences, or whatever sentences were relevant to the features we were discussing, but now we are going to introduce a much larger set of complicated test cases.
-Ideally we want test cases for testing every combination of grammar rules immaginable.
+Ideally we want test cases for testing every combination of grammar rules imaginable.
 
 In the test of this program, we used a branch of 100 sentences.
-Some copied from the papers referenced, others invented based on features considered or the daily life of farmers and donkeys as immagined by the author.
+Some copied from the papers referenced, others invented based on features considered or the daily life of farmers and donkeys as imagined by the author.
 The sentences were converted to Visser, DPL and FOL, and the meanings were checked manually to make sense.
@@
 \item
-Let's start up with a very sequential example to showcase the strength of DPL. Notice how the sentences are easily concattenated with no need for scope handling.
+Let's start up with a very sequential example to showcase the strength of DPL. Notice how the sentences are easily concatenated with no need for scope handling.
 \begin{align*}
@@
 \item
-We can also handle cases where `a farmer' is introduced inside the anticedent of an `if'.
+We can also handle cases where `a farmer' is introduced inside the antecedent of an `if'.
 \begin{align*}
@@
 A big consequence of/decision in the grammar created, is that no new entities can be introduced in the `then part' of an `if'.
-The reason for this is that such objects are not guarenteed to exist on the ground of the sentence.
+The reason for this is that such objects are not guaranteed to exist on the ground of the sentence.
 We enforce this rule using the $?_1$ function.
@@
 The first one is a case of a conjunctive VP.
-Notice that we treat `starves' as a transitive verb, even though the sentence is clearly ambigious. In this case the phrase parser makes the decision for us.
+Notice that we treat `starves' as a transitive verb, even though the sentence is clearly ambiguous. In this case the phrase parser makes the decision for us.
 \begin{align*}
@@
 In Montague's PTQ\cite{montague1973proper} he doesn't give rules for conjunction of NPs, but only sentences and VP's.
-In \cite{rooth1982conjunction} and \cite{partee1983generalized} Mats Rooth and Barbara Partee discuss different issues arrising, but never reach any usable construction.
+In \cite{rooth1982conjunction} and \cite{partee1983generalized} Mats Rooth and Barbara Partee discuss different issues arising, but never reach any usable construction.
 Personally I've found the below to be very useful for conjunctive NPs that act as a union.
@@
 A final example that really shows how DPL scope and free variables all play along together is this the case of a `non existent' antecedent.
-The stanford coref tagger mistakingly tags the two men in this sentence as the same, but because negation creates a closed scope, the final sentence still comes out right.
+The Stanford Coreference tagger mistakenly tags the two men in this sentence as the same, but because negation creates a closed scope, the final sentence still comes out right.
 \begin{align*}
@@
 
-A few examples of sentences that my framework already supports, but which fail due to short commings in the Stanford toolkit are `Every dog sees a cat.
+A few examples of sentences that my framework already supports, but which fail due to short comings in the Stanford toolkit are `Every dog sees a cat.
 It chases it.' where the anaphora doesn't get resolved, and `Buffalo buffalo Buffalo buffalo buffalo buffalo Buffalo buffalo' that manages to slightly confuse the parser.
@@
 We have seen how weak and strong readings of the actors in the sentences can give very different results.
-Also conjunction of transitive verbs like `A farmer starves and beats a donkey' are inherently ambigious.
-Perhaps if we returned many possible logifications for each sentence, we could take the interferance to get the real meaning.
+Also conjunction of transitive verbs like `A farmer starves and beats a donkey' are inherently ambiguous.
+Perhaps if we returned many possible logifications for each sentence, we could take the intersection to get the real meaning.
 
@@
 Support in more cases for more negatives like `no' and `nobody' hasn't been explored well.
-Possessive pronouns is something I've struggeled with, but if we accept that `He took his jacket' translates to $\exists y\cdot owns(x,y)\cdot jacket(y)\cdot took(x,y)$ it shouldn't be hard.
+Possessive pronouns is something I've struggled with, but if we accept that `He took his jacket' translates to $\exists y\cdot owns(x,y)\cdot jacket(y)\cdot took(x,y)$ it shouldn't be hard.
 Also there are things like `existential there' and more kinds of SBARs like started by `then', `whom' etc.
@@
 Even higher up are lots of classical linguistic problems.
-Sentences such as `The golden mountain does not exist', `The ancient greeks thought the morning star was the evening star' have had entire papers written on them, without any definive answers.
+Sentences such as `The golden mountain does not exist', `The ancient Greeks thought the morning star was the evening star' have had entire papers written on them, without any definite answers.
 Probably we'll have to consider the use case to know how a well defined computational logic should work.
@@
 The possibility for pushing control elements into the semantic of the logic is really the beauty of Visser's approach and unfortunately a road I haven't studied much in this paper.
-It is not hard to immagine $(Dpl,Dpl,Dpl,Dpl,Integer,Integer)$ exchanged for something like $(Dpl,...,PosTag)$ and coresponding composition rules.
+It is not hard to imagine $(Dpl,Dpl,Dpl,Dpl,Integer,Integer)$ exchanged for something like $(Dpl,...,PosTag)$ and corresponding composition rules.
 
 \section{Acknowledgments}
-I would like to thank Samson Abrahamski for introducing me to the world of Natural Language parsing and semantics and for helpful and stimulating discussions.
+I would like to thank Samson Abramsky for introducing me to the world of Natural Language parsing and semantics and for helpful and stimulating discussions.
 
@@
 
-Finnaly I wish to thanks to The Donkey Sanctuary\cite{donkey2013sanctuary} for taking donkey's welfare seriously and rescuing neglected donkeys in the UK and all over the world.
+Finally I wish to thanks to The Donkey Sanctuary\cite{donkey2013sanctuary} for taking donkey's welfare seriously and rescuing neglected donkeys in the UK and all over the world.
 
@@
 %
-\intertext{For predicates it's mostly a matter of juggling variables, values and assigments:}
+\intertext{For predicates it's mostly a matter of juggling variables, values and assignments:}
 %
@@
 
+\lstinputlisting[style=bashstyle,caption=\lstname]{run.sh}
 \lstinputlisting[style=haskellstyle,caption=\lstname]{src/Main.hs}

### 2013-05-19 run.sh

@@ -1,4 +1,4 @@
 #!/bin/bash
 cd "$(dirname $0)/src"
-runghc Main.hs inputs $1
+runghc Main.hs $1 $2
 

### 2013-05-19 paper.tex
Fixed DPL part

@@
 A possible solution to this problem is to use a so called `update logic' (a logic with update semantics).
-Update semantics are logics inspired on the semantics of programming languages where a variable, once introduced, will stick around for a period not know at the time of it's introduction.
+Update semantics are logics inspired on the semantics of programming languages where a variable, once introduced, will stick around for a period not know at the time of its introduction.
 
@@
 
-The logic cleverly uses these advances in logical understanding to create an update logic in a pragmatic way, without any external constructs and a minimum of new syntax, while still getting all the benefits of compositionality etc.
-
-To model a simple sentence `A man comes in. He sees a donkey. He smiles.', we write:
+DPL cleverly uses this basic logical research to create a very simple, self-contained logic, while still getting all the benefits of compositionality etc.
+To model a simple sentence `A man comes in.
+He sees a donkey.
+He smiles.', we write:
 %
@@
 
-Too understand the semantics of the above, let's start by only considering a dynamic propositional logic with three propositions: $P$, $Q$ and $R$.
-We'll write $p$ to indicate a predicate is true and $\overline{p}$ if it's false.
-We talked about earlier, that the update semantics modeled `growth of information over time', so we'll start with case that every $2^3$ state is possible:
-%
-\begin{align}
-\{\overline{PQR}, \overline{PQ}R, \overline{P}Q\overline{R}, \overline{P}QR, P\overline{QR}, P\overline{Q}R, PQ\overline{R}, PQR\} \label{dpl_all_states}
-\end{align}
-
-Now we'll consider a statement in our predicate logic as a relation between possible states.
-For example $P\overline{Q}R[P\cdot Q]P\overline{Q}R$ is false, while $P\overline{Q}R[P\cup Q]PQR$ is true.
-If we take the image of \eqref{dpl_all_states} under $P\cdot\neg Q$ we get
-%
-\begin{align}
-\{P\overline{QR}, P\overline{Q}R\} \label{dpl_PnQ_states}
-\end{align}
-
-If we then take the image of \eqref{dpl_PnQ_states} under $\overline{P}$ we end up with the empty set of states.
-That is there is no possible assignment satisfying our formula, and we consider it a falsum.
-
-The most interesting question is how we work with quantifiers.
-Which on earth relation could $\exists S$ be?
-We are going to take a hint from computer programming and consider the variable declaration: $S:=\bot$.
-We want to write this in DPL as $\exists S \cdot \neg S$, such that e.g.
-$P\overline{Q}[\exists S \cdot \neg S]P\overline{QS}$ is true.
-Now since $\cdot\neg S$ acts as a filter, removing all assignments that doesn't have $\overline{S}$, it must be that $\exists S$ is a `random reset, introducing $S$ and allowing it every value, such that $P\overline{Q}[\exists S]P\overline{QS}$ and $P\overline{Q}[\exists S]P\overline{Q}S$ are both true.
-The image of \eqref{dpl_all_states} under $\exists S$ has 16 possible states.
+Too understand the semantics of the above, let's start by considering a simple example with two variables ($a$ and $b$) and a domain with three entities ($0$, $1$ and $2$).
+I said earlier, that the update semantics modeled `growth of information over time', so we'll start with the case that every $3^2$ assignments are possible:
+%
+\begin{align}
+\{&\langle a\mapsto0,b\mapsto0\rangle, \langle a\mapsto0,b\mapsto1\rangle, \langle a\mapsto0,b\mapsto2\rangle,\nonumber\\
+&\langle a\mapsto1,b\mapsto0\rangle, \langle a\mapsto1,b\mapsto1\rangle, \langle a\mapsto1,b\mapsto2\rangle,\nonumber\\
+&\langle a\mapsto2,b\mapsto0\rangle, \langle a\mapsto2,b\mapsto1\rangle, \langle a\mapsto2,b\mapsto2\rangle\}\label{dpl_all_states}
+\end{align}
+
+A DPL formula is defined as a relation between assignments, and $\llbracket\cdot\rrbracket$ is the function that takes a formula to its relation.
+For example $\langle a\mapsto1,b\mapsto1\rangle\llbracket a=1\rrbracket\langle a\mapsto1,b\mapsto1\rangle$ is true because the $a=1$ `program' executes successfully.
+We also require that simple tests like this one doesn't change the state, so $\langle a\mapsto 1,b\mapsto2\rangle$ would not have been a valid assignment on the right hand side.
+
+Just like with programs, we can combine multiple formulas into bigger ones: $\langle a\mapsto0,b\mapsto2\rangle\llbracket a=0\cdot b=a\rrbracket\psi$ is false for all $\psi$, and $\langle a\mapsto2,b\mapsto1\rangle\llbracket a=0\cup b<2\rrbracket\langle a\mapsto2,b\mapsto1\rangle$ is true.
+If we take the image of \eqref{dpl_all_states} under $a=0\cdot\neg (b=2)$ we get
+%
+\begin{align}
+\{&\langle a\mapsto0,b\mapsto0\rangle, \langle a\mapsto0,b\mapsto1\rangle\} \label{dpl_PnQ_states}
+\end{align}
+
+If we take the image of \eqref{dpl_PnQ_states} under $a=1$, we end up with the empty set of states.
+That is because there are no possible assignments satisfying our formula.
+Hence we consider $a=0\cdot\neg (b=2)\cdot a=1$ a falsum, also defined as $\bot$.
+
+So why do we define all of this using relations, and not just functions from states to truth values?
+The reason is that we want to introduce a special relation $\exists x$ (for some variable name $x$).
+We want it to analogous to $\exists x(...)$ from first-order logic.
+It's not immediately obvious what the correct definition should be.
+
+The trick is to consider the conjunctive DPL relation $\exists x\cdot x=1$ - we want this to work such that $\langle a\mapsto0\rangle\llbracket\exists x\cdot x=1\rrbracket\langle a\mapsto0,x\mapsto1\rangle$.
+Now considering $x=1$ is a filter, it must be that $\exists x$ is a relation that allows any value for $x$ on its second argument.
+For example we see that the image of \eqref{dpl_all_states} under $\exists x$ has 27 possible states.
 
 Let's have a look at the model theoretic definition of the semantics.
-Visser builds this up in two steps: First a relational algebra and then DPL on top of it.
-But since DPL is basically identical to the relational algebra, we are going to try and do it in one step.
+This is sometimes done in two steps: First defining a relational algebra and then builid DPL on top of it.
+However DPL is basically identical to the relational algebra, so we are going to try and do everything in one step:
 
@@
 Reset_\Sigma &= \{\exists x \mid x\in Var\}\\
-Atom_\Sigma &= Var \cup Prop_\Sigma \cup Reset_\Sigma
-\end{align}
-
-A model $\mathcal{M}$ of a signature $\Sigma$ is a tuple $\langle\mathcal{D},\mathcal{I}\rangle$ where $\mathcal{D}$ is a domain which we will always set as $\{\top,\bot\}$, and $\mathcal{I}$ is the interpretation function $Prop_\Sigma\to\mathcal{D}$.
+Atom_\Sigma &= Prop_\Sigma \cup Reset_\Sigma
+\end{align}
 
@@
 
-We are also going to say that an assignment can satisfy a formula, defined as:
-\begin{align}
-\alpha\models\psi :\equiv \exists\beta(\alpha[\psi]\beta)
-\end{align}
-
-Let $[\cdot]$ be the interpretation function from syntax to relation. Then we define the semantics of our logic pointwise for all assignments $\alpha$ and $\beta$:
-\begin{align}
-\alpha[\psi_1\cdot\psi_2]\beta :\equiv&\ \exists\gamma (\alpha[\psi_1]\gamma \wedge \gamma[\psi_2]\beta) \label{sem_and} \\
-\alpha[\psi_1\cup\psi_2]\beta :\equiv&\ \alpha[\psi_1]\beta \vee \alpha[\psi_2]\beta \label{sem_union} \\
-\alpha[P(x_1,\dots,x_n)]\beta :\equiv&\ \alpha = \beta \wedge \mathcal{I}(P(\alpha_{x_1},\dots,\alpha_{x_n})) \label{sem_pred} \\
-\alpha[\neg\psi]\beta :\equiv&\ \alpha = \beta \wedge \alpha\not\models[\psi] \label{sem_neg}\\
-\alpha[\exists x]\beta :\equiv&\ \forall\omega (\omega \neq x \rightarrow \alpha_{\omega} = \beta_{\omega}) \label{sem_exists} \\
-\intertext{Using those basic axioms we can derive for our built in propositions:}
-\alpha[\bot]\beta \leftrightarrow&\ \alpha = \beta \wedge \bot \nonumber\\
-                  \leftrightarrow&\ \bot \label{sem_bot}\\
-\alpha[\top]\beta \leftrightarrow&\ \alpha = \beta \wedge \top \nonumber\\
-                  \leftrightarrow&\ \alpha = \beta \label{sem_top}\\
-\intertext{Taking $\psi_1\rightarrow\psi_2$ to mean $\neg(\psi_1\wedge\neg\psi_2)$ we also get the following neat equation, originally by Groenendijk and Stokhof\cite{groenendijk1991dynamic}:}
-\alpha[\psi_1\rightarrow\psi_2]\beta
- :\equiv&\ \alpha[\neg(\psi_1\wedge\neg\psi_2)]\beta \nonumber\\
- \leftrightarrow&\ \alpha = \beta \wedge \alpha\not\models[\psi_1\wedge\neg\psi_2] \nonumber\\
- \leftrightarrow&\ \alpha = \beta \wedge \neg\exists\gamma(\alpha[\psi_1\wedge\neg\psi_2]\gamma) \nonumber\\
- \leftrightarrow&\ \alpha = \beta \wedge \neg\exists\gamma(\exists\delta(\alpha[\psi_1]\delta\wedge\delta[\neg\psi_2]\gamma)) \nonumber\\
- \leftrightarrow&\ \alpha = \beta \wedge \neg\exists\gamma(\exists\delta(\alpha[\psi_1]\delta\wedge\delta=\gamma\wedge\delta\not\models[\psi_2])) \nonumber\\
- \leftrightarrow&\ \alpha = \beta \wedge \forall\gamma(\forall\delta(\delta=\gamma\rightarrow(\alpha[\psi_1]\delta\rightarrow\delta\models[\psi_2]))) \nonumber\\
- \leftrightarrow&\ \alpha = \beta \wedge \forall\gamma(\alpha[\psi_1]\gamma\rightarrow\gamma\models[\psi_2]) \label{sem_impl}
+A model $\mathcal{M}$ of a signature $\Sigma$ is a tuple $\langle\mathcal{D},\mathcal{I}\rangle$.
+The model is very similar to first-order logic in that $\mathcal{D}$ is the domain, a set of entities, and $\mathcal{I}$ is the interpretation function $Prop_\Sigma\to(Asgmt\to\mathbb{B})$, ($Asgmt$ being the set of possible assignments, $\mathbb{B}$ the set of boolean values).
+
+Just like in first-order logic we want to denote whether a formula is true or false, and just like in first-order logic we do this by saying that an assignment can satisfy a formula, defined as below.
+If a formula is satisfied by all assignments (equivalent to being satisfied by the empty assignment), it is a tautology.
+\begin{align}
+\alpha\models\psi :\equiv \exists\beta(\alpha\llbracket\psi\rrbracket\beta)
+\end{align}
+
+We then define our relations in the set theoretic way, as sets of pairs:
+\begin{align}
+\llbracket\psi_1\cdot\psi_2\rrbracket :\equiv&\ \{\langle\alpha,\beta\rangle \in Asgmt^2 \mid \exists\gamma(\langle\alpha,\gamma\rangle\in\llbracket\psi_1\rrbracket\wedge\langle\gamma,\beta\rangle\in\llbracket\psi_2\rrbracket)\} \label{sem_set_and} \\
+\llbracket P(x_1,\dots,x_n)\rrbracket :\equiv&\ \{\langle\alpha,\alpha\rangle \in Asgmt^2 \mid \mathcal{I}(P)(\alpha)\} \label{sem_set_pred}
+\end{align}
+%
+Or we can define them more succinctly, pointwise for all assignments $\alpha$, $\beta$:
+\begin{align}
+\alpha\llbracket \psi_1\cdot\psi_2\rrbracket \beta :\equiv&\ \exists\gamma (\alpha\llbracket \psi_1\rrbracket \gamma \wedge \gamma\llbracket \psi_2\rrbracket \beta) \label{sem_and} \\
+\alpha\llbracket \psi_1\cup\psi_2\rrbracket \beta :\equiv&\ \alpha\llbracket \psi_1\rrbracket \beta \vee \alpha\llbracket \psi_2\rrbracket \beta \label{sem_union} \\
+\alpha\llbracket P(x_1,\dots,x_n)\rrbracket \beta :\equiv&\ \alpha = \beta \wedge \mathcal{I}(P)(\alpha) \label{sem_pred} \\
+\alpha\llbracket \neg\psi\rrbracket \beta :\equiv&\ \alpha = \beta \wedge \alpha\not\models\llbracket \psi\rrbracket  \label{sem_neg}\\
+\alpha\llbracket \exists x\rrbracket \beta :\equiv&\ \forall\omega (\omega \neq x \rightarrow \alpha_{\omega} = \beta_{\omega}) \label{sem_exists}
+\end{align}
+%
+Using those basic axioms we can derive $\top$ and $\bot$ given our usual interpretations:
+\begin{align}
+\alpha\llbracket \bot\rrbracket \beta \equiv&\ \alpha = \beta \wedge \mathcal{I}(\bot)(\alpha) \equiv \bot \label{sem_bot}\\
+\alpha\llbracket \top\rrbracket \beta \equiv&\ \alpha = \beta \wedge \mathcal{I}(\top)(\alpha) \equiv \alpha = \beta \label{sem_top}\\
+\end{align}
+%
+Taking $\psi_1\rightarrow\psi_2$ to mean $\neg(\psi_1\wedge\neg\psi_2)$ we also get the following neat equation, originally by Groenendijk and Stokhof\cite{groenendijk1991dynamic}
+\begin{align}
+\alpha\llbracket \psi_1\rightarrow\psi_2\rrbracket \beta
+ :\equiv&\ \alpha\llbracket \neg(\psi_1\wedge\neg\psi_2)\rrbracket \beta \nonumber\\
+ \leftrightarrow&\ \alpha = \beta \wedge \alpha\not\models\llbracket \psi_1\wedge\neg\psi_2\rrbracket  \nonumber\\
+ \leftrightarrow&\ \alpha = \beta \wedge \neg\exists\gamma(\alpha\llbracket \psi_1\wedge\neg\psi_2\rrbracket \gamma) \nonumber\\
+ \leftrightarrow&\ \alpha = \beta \wedge \neg\exists\gamma(\exists\delta(\alpha\llbracket \psi_1\rrbracket \delta\wedge\delta\llbracket \neg\psi_2\rrbracket \gamma)) \nonumber\\
+ \leftrightarrow&\ \alpha = \beta \wedge \neg\exists\gamma(\exists\delta(\alpha\llbracket \psi_1\rrbracket \delta\wedge\delta=\gamma\wedge\delta\not\models\llbracket \psi_2\rrbracket )) \nonumber\\
+ \leftrightarrow&\ \alpha = \beta \wedge \forall\gamma(\forall\delta(\delta=\gamma\rightarrow(\alpha\llbracket \psi_1\rrbracket \delta\rightarrow\delta\models\llbracket \psi_2\rrbracket ))) \nonumber\\
+ \leftrightarrow&\ \alpha = \beta \wedge \forall\gamma(\alpha\llbracket \psi_1\rrbracket \gamma\rightarrow\gamma\models\llbracket \psi_2\rrbracket ) \label{sem_impl}
 \end{align}
@@
 In my implementation section, I describe how the above can be efficiently implemented in Haskell using an isomorphism between relations $\mathcal{P}(A\times B)$ and functions $A\to\mathcal{P}(B)$.
-I also show how the kleiski combinator takes the role of $\cdot$.
+I also show how the Kleisli combinator takes the role of $\cdot$.
 
@@
 These equations we can think of intuitively as having a closed scope.
-We saw earlier how DRT nicely determined that the `it' in `A farmer doesn't own a donkey.
-He beats it' couldn't refer to `a donkey' because the negation didn't let it's variable out.
+We saw earlier how DRT nicely determined that the `it' in `Some farmer owns no donkey.
+He beats it' couldn't refer to `a donkey' because the negation didn't let its variable out.
 In DPL we get the same effect.
@@
 This should be a good sentence, but we can't handle it.
-In the next chapter we will look at how this might be handled.
+In section~\ref{sec:problem_disj} I will look at how this might be handled.
 
@@
 This gives an alternative way to define and understand the semantics of DPL, and it relaxes us that nothing `weird' is going on.
-We might also be a bit disappointed that after all this, we don't possess any extra power, but let's forget about that for now, and see how the conversion works.
-
-Let $\phi$ be a formula in negation normal form, or something. Hmm, come back here once the disjunction chapter is done.
-\begin{align}
-&[\exists x(\phi)] \equiv \neg(\neg(\exists x \cdot [\phi]))\\
-&[\forall x(\phi)] \equiv \exists x \rightarrow [\phi]\\
-&[\phi_1\wedge\phi_2] \equiv [\phi_1]\cdot[\phi_2]\\
-&[\phi_1\vee\phi_2] \equiv \dots\\
-&[\neg(\exists x)] \equiv \dots\\
-&[\neg(\phi_1\wedge\phi_2)] \equiv \neg[\phi_1]\cdot\neg[\phi_2]\\
-&[\neg(\phi_1\vee\phi_2)] \equiv \dots
-\end{align}
-
-Similarly we may create PL formulas with the same meaning as any DPL formula.
-But perhaps we should be a bit concerned on how to interpret the meaning of a PL formula and a DPL formula being equivalent.
-After all DPL formulas are relations where PL formulas a propositions.
+We might also be a bit disappointed that after all this work, we don't possess any extra power, but let's forget about that for now, and see how the conversion works.
+
+We can define a conversion function $\llbracket\cdot\rrbracket$ from FOL to DPL very easily.
+The only trick is that we need a way in DPL to create the closed scopes used in FOL.
+$\neg(\neg(\cdot))$ does nicely:
+\begin{align}
+\llbracket P(x_1,\dots,x_n)\rrbracket &\equiv P(x_1,\dots,x_n)\\
+\llbracket \neg(\phi)\rrbracket &\equiv \neg\llbracket \phi\rrbracket\\
+\llbracket \phi_1\wedge\phi_2\rrbracket &\equiv \llbracket \phi_1\rrbracket\cdot\llbracket \phi_2\rrbracket\\
+\llbracket \phi_1\vee\phi_2\rrbracket &\equiv \llbracket \phi_1\rrbracket\cup\llbracket \phi_2\rrbracket\\
+\llbracket \exists x(\phi)\rrbracket &\equiv \neg(\neg(\exists x \cdot \llbracket \phi\rrbracket))\\
+\llbracket \forall x(\phi)\rrbracket &\equiv \exists x \rightarrow \llbracket \phi\rrbracket
+\end{align}
+
+Similarly we can create a function from DPL to FOL, but perhaps we should be a bit concerned on how to interpret the meaning of an FOL formula and a DPL formula being equivalent.
+After all DPL formulas are relations where PL formulas are propositions.
 In the above it was all easy, but let's think again about what we are doing.
 
-Remember that we said an assignment could model a DPL relation.
-If we consider the truth value of a DPL relation to be whether there exists any relation that models it $\exists\alpha(\alpha\models\psi)$, we can use a trick.
-The trick is to think of the weakest precondition such an assignment must satisfy, and model that in PL.
-In fact the precondition is not just for our assignment, but for the entire model.
-For example to model $\exists x\cdot farmer(x)$ the starting assignment might well be empty ($\bot$), but our model must contain an entity for which $farmer$ is true.
-
-To smoothen the following derivations, we will introduce a special syntax `$\langle\psi\rangle\phi$' where $\psi$ is a DRA formula, $\phi$ is a PL formula and the whole thing is a PL formula.
-For some assignment $\alpha$, we define `$\alpha\models\langle\psi\rangle\phi$' to mean $\exists\beta(\alpha[\psi]\beta \wedge \beta\models\phi)$.
-%
-\begin{align}
-\alpha\models\langle\exists x\rangle\phi
- & \leftrightarrow \alpha\models\exists x (\phi) \label{conv_exists}\\
-\alpha\models\langle\psi_1\cdot\psi_2\rangle\phi
- & \leftrightarrow \alpha\models\langle\psi_1\rangle\langle\psi_2\rangle\phi \label{conv_and}\\
-\alpha\models\langle\psi_1 \cup \psi_2\rangle\phi
- & \leftrightarrow \alpha\models\langle\psi_1\rangle\phi \vee \langle\psi_2\rangle\phi \label{conv_or}\\
- \alpha\models\langle\psi_1 \vee \psi_2\rangle\phi
-\alpha\models\langle P(x_1,\dots,x_n)\rangle\phi
- & \leftrightarrow \alpha\models P(x_1,\dots,x_n) \wedge \phi \label{conv_pred}\\
-\alpha\models\langle\neg\psi\rangle\phi
- & \leftrightarrow \alpha\models \neg\langle\psi\rangle\top \wedge \phi \label{conv_neg}
-\end{align}
-
-See the appendix for my full derivations of every equation.
-Also notice that $\alpha\models\langle\bot\rangle\phi\leftrightarrow a\models\bot$ and $\alpha\models\langle\top\rangle\phi\leftrightarrow a\models\phi$ follow trivially from \eqref{conv_pred}.
-Equations for `convenience' operators such as $\rightarrow$ and $\vee$ can be derived too:
-\begin{align}
-\langle\psi_1\vee\psi_2\rangle\phi
-& \leftrightarrow \langle\neg(\neg\psi_1\cdot\neg\psi_2)\rangle\phi \nonumber\\
-& \leftrightarrow \neg\langle\neg\psi_1\cdot\neg\psi_2\rangle\top \wedge\phi \nonumber\\
-& \leftrightarrow \neg\langle\neg\psi_1\rangle\langle\neg\psi_2\rangle\top \wedge\phi \nonumber\\
-& \leftrightarrow \neg\langle\neg\psi_1\rangle(\neg\langle\psi_2\rangle\top) \wedge\phi \nonumber\\
-& \leftrightarrow \neg(\neg\langle\psi_1\rangle\top \wedge \neg\langle\psi_2\rangle\top) \wedge\phi \nonumber\\
-& \leftrightarrow (\langle\psi_1\rangle\top \vee \langle\psi_2\rangle\top) \wedge\phi \label{conv_vee}
-\end{align}
-
-In the implementation we will use the above equations to recursively convert our DPL formulas to PL. The $\langle\rangle$ notation easily lends itself to calculation by a stack manipulation function.
-
-The rules will look like:
-\begin{haskell}
-addDpl :: Prop -> [DPL] -> Prop
-addDpl p ((DNot d):ds) = addDpl (And (Not (addDpl T [d])) p) ds
-\end{haskell}
+We are going to use our definition of $\alpha\models\phi$ for DPL formulas.
+An FOL formula $\phi$ and a DPL formula $\psi$ are defined to be equivalent exactly when $\forall\alpha(\alpha\models\phi\leftrightarrow\alpha\models\psi)$.
+Perhaps the best way to see why this is a good definition (other than its simplicity) is to think programs and weakest preconditions.
+Every initial state for the `computation' $farmer(y)\cdot\exists x\cdot x=y$ must model $farmer(y)\wedge\exists x(x=y)$ in order to succeed.
+
+To smoothen the following derivations, we will define the function $\langle\psi\rangle\phi:DPL\times FOL\to FOL$ by
+\begin{align}
+\alpha\models\langle\psi\rangle\phi\ \leftrightarrow\ \exists\beta(\alpha\llbracket\psi\rrbracket\beta \wedge \beta\models\phi)\label{foldpl}
+\end{align}
+Notice from this definition that $\langle\psi\rangle\top$ is the conversion function from DPL to FOL.
+Using definition \eqref{foldpl} and formulas \eqref{sem_and} to \eqref{sem_exists} we can derive the following equivalences. See appendix~\ref{sec:derivation_dpl} for my full derivations.
+%
+\begin{align}
+\langle\exists x\rangle\phi & \equiv \exists x (\phi) \label{conv_exists}\\
+\langle\psi_1\cdot\psi_2\rangle\phi & \equiv \langle\psi_1\rangle\langle\psi_2\rangle\phi \label{conv_and}\\
+\langle\psi_1 \cup \psi_2\rangle\phi & \equiv \langle\psi_1\rangle\phi \vee \langle\psi_2\rangle\phi \label{conv_or}\\
+\langle P(x_1,\dots,x_n)\rangle\phi & \equiv  P(x_1,\dots,x_n) \wedge \phi \label{conv_pred}\\
+\langle\neg\psi\rangle\phi & \equiv  \neg\langle\psi\rangle\top \wedge \phi \label{conv_neg}\\
+\langle\psi_1\vee\psi_2\rangle\phi & \equiv (\langle\psi_1\rangle\top \vee \langle\psi_2\rangle\top) \wedge\phi \label{conv_vee}
+\end{align}
+
+The simple predicates $\langle\bot\rangle\phi\equiv\bot$ and $\langle\top\rangle\phi\equiv\phi$ follow trivially from \eqref{conv_pred}.
+Using the first 5 equations, equivalences for derives operators such as $\rightarrow$ and $\vee$ can be derived syntactically as also shown in appendix~\ref{sec:derivation_dpl}.
+
+An alternative way to the $\langle\psi\rangle\phi$ function for looking at the above equations is to think of the $\langle\cdot\rangle$s as elements on a stack.
+This is indeed how I have implemented the DPL to FOL routine in appendix~\ref{sec:source_code}.
 
@@
 
-\subsubsection{Examples}
+\subsubsection{A Bunch of Examples}
 
@@
 
-A few examples of sentences that my framework already supports, but which fail due to short comings in the Stanford toolkit are `Every dog sees a cat.
+A few examples of sentences that my framework already supports, but which fail due to shortcomings in the Stanford toolkit are `Every dog sees a cat.
 It chases it.' where the anaphora doesn't get resolved, and `Buffalo buffalo Buffalo buffalo buffalo buffalo Buffalo buffalo' that manages to slightly confuse the parser.
@@
 
-Finally I wish to thanks to The Donkey Sanctuary\cite{donkey2013sanctuary} for taking donkey's welfare seriously and rescuing neglected donkeys in the UK and all over the world.
+Finally I wish to thank The Donkey Sanctuary\cite{donkey2013sanctuary} for taking donkey's welfare seriously and rescuing neglected donkeys in the UK and all over the world.
 
@@
 
-\subsection{Derivation of DPL Preconditions}
+\subsection{The Problem with Disjunction}\label{sec:problem_disj}
+
+%We won't actually use `or' in the implementation.
+%Example of the problems we still have in the logic.
+%Can't tell what variables are available where.
+
+A major problem with our use of the union of relations as disjunction is the way it allows seemingly sound sentences to combine into syntactically invalid ones.
+Consider the sentence `Either a farmer owns a donkey or he doesn't'.
+We would write that as
+\begin{align}
+\exists x\cdot farmer(x)\cdot(\exists y\cdot owns(x,y)\cdot donkey(y)\vee \bot) \label{owns_donkey_or_not}
+\end{align}
+Now concatenating this sentence with `he beats it': $beat(x,y)$ would give us a problem.
+In fact this problem would be more than a syntactic problem, because before we know the model and the existence of the donkey, we don't know if the sentence is valid or not.
+
+Visser\cite{visser1999donkey} suggests that we should build our logic on sets of relations read disjunctively instead of just relations.
+This `possible world' scenario would transform \eqref{owns_donkey_or_not}$\cdot beats(x,y)$ into $\{\exists x\cdot farmer(x)\cdot\exists y\cdot donkey(y)\cdot owns(x,y)\cdot beats(x,y), \exists x\cdot farmer(x)\cdot\bot\cdot beats(x,y)\} = \{\exists x\cdot farmer(x)\cdot\exists y\cdot donkey(y)\cdot owns(x,y)\cdot beats(x,y), \bot\}$.
+This clearly is always well defined.
+
+Visser's idea doesn't fully save us though. If we now consider `a farmer either owns a donkey or owns a horse':
+\begin{align}
+\exists x\cdot farmer(x)\cdot(&\exists y\cdot owns(x,y)\cdot donkey(y)\nonumber\\
+                              &\vee \exists z\cdot owns(x,z)\cdot horse(z))
+\end{align}
+And combines that with `He beats the donkey': $beat(x,y)$, we get, in Visser's model, $\{\exists x\cdot farmer(x)\cdot\exists y\cdot owns(x,y)\cdot donkey(y)\cdot beats(x,y), \exists x\cdot farmer(x)\cdot\exists z\cdot owns(x,z)\cdot horse(z)\cdot beats(x,y)\}$.
+The second one is clearly not valid (though we wouldn't get a problem if there is no horse or farmer) which we might catch `at compile time' and replace it with something like `$\bot$' or `error'.
+This sort of error handling is discussed and implemented in the implementation section, but is not terribly interesting from a formal point of view.
+
+Much more interesting are the following two questions: How could we have translated \label{owns_donkey_or_horse} so it would actually work?
+After all it seemed like a reasonable sentence until variable naming got us.
+And does the way variable binding works with our definition actually make sense?
+
+Clearly just renaming $z$ to $y$ in the horse term would have solved the first problem.
+However to work in a compositional way, that would either require great luck, or renaming, and we don't have a renaming operator.
+
+A trick we could consider is to declare an `object' variable outside of the disjunction:
+\begin{align}
+\exists s\cdot farmer(s)\cdot\exists o\cdot(
+      &\exists y\cdot owns(s,y)\cdot donkey(y)\cdot o=y \nonumber\\
+      &\vee \exists z\cdot owns(s,z)\cdot horse(z)\cdot o=z)\cdot beats(s,o)
+\end{align}
+%
+This even inspires on idea that we might be able to build everything up around subjects, objects, indirect objects etc.
+This is explored in the implementation, but it does suffer from many problems.
+E.g.
+we might well have a sentence very similar to the above: `A farmer owns a donkey and owns a horse.
+He beat them'.
+Here we suddenly have two objects in the same sentence.
+
+Sentences like the above are easy to make in English because the absence of noun genders make it easy to use the same pronoun to refer to a lot of things at once.
+By choosing the right actors however, it is always possible to run into problems.
+
+The problem is perhaps the non-linearity in the way variables pass through the disjunction. Somehow we lose track of their history when the `results' are merged together.
+
+This leads on to the other question we mentioned: Is the binding flow through our disjunction operator really any good?
+We are is inspired by the sentence `Either a farmer or his scapegoat must go to court.
+In any case he will be picked up by sunset'.
+In terms of DPL formulas, this would be modelled something like $(\psi_1\cup\psi_2)\cdot\psi_3$.
+Clearly we have a problem here: Variables bound in $\psi_1$ are not accessible in $\psi_2$.
+
+This sentence suggests that we might want to define a disjunction operator with a linear flow of variables: $\psi_1\to\psi_2\to\psi_3$ instead of $\psi_1\to\psi_3\text{ and }\psi_2\to\psi_3$.
+However there is a problem: We currently define `false' as the absence of possible assignments.
+Hence if $\psi_1$ is false, there is no way we can define disjunction to push its variables into $\psi_2$.
+If any variables were left undefined, $\psi_1$ wouldn't be false.
+
+There is a cure to this problem.
+When we extend our logic with more power in the monoid chapter, we get a second chance.
+However let's first have a look at an obvious alternative definition for disjunction considered by Groenendijk\cite{groenendijk1991dynamic}.
+We'll call it `static disjunction':
+\begin{align}
+\alpha[\psi_1\vee\psi_2]\beta
+ :\equiv&\ \alpha[\neg(\neg\psi_1\wedge\neg\psi_2)]\beta \label{sem_or_def}\\
+ \leftrightarrow&\ \alpha[\neg\psi_1\rightarrow\psi_2]\beta \nonumber\\
+ \leftrightarrow&\ \alpha = \beta \wedge \forall\gamma(\alpha[\neg\psi_1]\gamma\rightarrow\gamma\models[\psi_2]) \nonumber\\
+ \leftrightarrow&\ \alpha = \beta \wedge \forall\gamma((\alpha=\gamma\wedge\gamma\not\models\psi_1)\rightarrow\gamma\models[\psi_2]) \nonumber\\
+ \leftrightarrow&\ \alpha = \beta \wedge \forall\gamma(\alpha=\gamma\rightarrow(\gamma\models\psi_1\vee\gamma\models\psi_2)) \nonumber\\
+ \leftrightarrow&\ \alpha = \beta \wedge (\alpha\models\psi_1\vee\alpha\models\psi_2) \label{sem_or}
+\end{align}
+
+The idea to this follows naturally from De Morgan's laws of classical logic.
+Clearly the semantics are different from the ones we defined for $\cup$.
+However not in a good way.
+This static disjunction doesn't allow any variables to escape from inside its reach.
+
+So does this fix our problem of binding between clauses?
+No.
+In fact this definition doesn't even allow binding from a clause and out.
+The villain is $\neg\psi$.
+From the first mention we defined it to be static so it wouldn't allow unsafe, arbitrary assignments as long as they didn't satisfy $\psi$.
+We wanted our logic to be reasonably monotonic, and we got what we deserved.
+$\neg$ not only closed down this version of disjunction, but our implication definition as well.
+
+We can imagine defining $\neg$ dynamically: $\neg\psi_1\cdot\psi_2 = \neg(\psi_1\cdot\psi_2)$.
+However not in our current model, and it won't solve any of our problems with disjunction.
+It does however allow this sentence: `It is not the case that a farmer doesn't own a donkey.
+He beats it'.
+While it made sense not to allow variables to escape from a single $\neg$, we want them to escape from a double.
+
+\subsection{Derivation of DPL Preconditions}\label{sec:derivation_dpl}
 Remember that we have defined, for assignments $\alpha$ and $\beta$, DPL formula $\psi$ and FOL formula $\phi$: $\alpha\models\langle\psi\rangle\phi$ to mean $\exists\beta(\alpha\llbracket\psi\rrbracket\beta\wedge\beta\models\phi)$.
@@
  & \leftrightarrow \exists\nu (\alpha[x:=\nu]\models\phi) \nonumber\\
- & \leftrightarrow \alpha\models\exists x (\phi) \\
+ & \leftrightarrow \alpha\models\exists x (\phi) \tag{$\ref{conv_exists}$} \\
 %
@@
  & \leftrightarrow \exists\gamma (\alpha\llbracket\psi_1\rrbracket\gamma \wedge \gamma\models\langle\psi_2\rangle\phi) \nonumber\\
- & \leftrightarrow \alpha\models\langle\psi_1\rangle\langle\psi_2\rangle\phi
+ & \leftrightarrow \alpha\models\langle\psi_1\rangle\langle\psi_2\rangle\phi \tag{$\ref{conv_and}$}
 %
@@
  & \leftrightarrow (\alpha\models\langle\psi_1\rangle\phi) \vee (\alpha\models\langle\psi_2\rangle\phi) \nonumber\\
- & \leftrightarrow \alpha\models\langle\psi_1\rangle\phi \vee \langle\psi_2\rangle\phi
+ & \leftrightarrow \alpha\models\langle\psi_1\rangle\phi \vee \langle\psi_2\rangle\phi \tag{$\ref{conv_or}$}
 %
@@
  & \leftrightarrow P(\alpha_{x_1},\dots,\alpha_{x_n}) \wedge \alpha\models\phi \nonumber\\
- & \leftrightarrow \alpha\models P(x_1,\dots,x_n) \wedge \phi
+ & \leftrightarrow \alpha\models P(x_1,\dots,x_n) \wedge \phi \tag{$\ref{conv_pred}$}
 %
@@
  & \leftrightarrow \alpha\not\models\langle\psi\rangle\top \wedge \alpha\models\phi \nonumber\\
- & \leftrightarrow \alpha\models \neg\langle\psi\rangle\top \wedge \phi
-%
-\intertext{Because $\psi$ is this way `evaluated on $\top$ only', we know that no variables introduced in $\psi$ can affect $\phi$.}\nonumber
-\end{align}
-
-\subsection{The Problem with Disjunction}\label{sec:problem_disj}
-
-%We won't actually use `or' in the implementation.
-%Example of the problems we still have in the logic.
-%Can't tell what variables are available where.
-
-A major problem with our use of the union of relations as disjunction is the way it allows seemingly sound sentences to combine into syntactically invalid ones.
-Consider the sentence `Either a farmer owns a donkey or he doesn't'.
-We would write that as
-\begin{align}
-\exists x\cdot farmer(x)\cdot(\exists y\cdot owns(x,y)\cdot donkey(y)\vee \bot) \label{owns_donkey_or_not}
-\end{align}
-Now concatenating this sentence with `he beats it': $beat(x,y)$ would give us a problem.
-In fact this problem would be more than a syntactic problem, because before we know the model and the existence of the donkey, we don't know if the sentence is valid or not.
-
-Visser\cite{visser1999donkey} suggests that we should build our logic on sets of relations read disjunctively instead of just relations.
-This `possible world' scenario would transform \eqref{owns_donkey_or_not}$\cdot beats(x,y)$ into $\{\exists x\cdot farmer(x)\cdot\exists y\cdot donkey(y)\cdot owns(x,y)\cdot beats(x,y), \exists x\cdot farmer(x)\cdot\bot\cdot beats(x,y)\} = \{\exists x\cdot farmer(x)\cdot\exists y\cdot donkey(y)\cdot owns(x,y)\cdot beats(x,y), \bot\}$.
-This clearly is always well defined.
-
-Visser's idea doesn't fully save us though. If we now consider `a farmer either owns a donkey or owns a horse':
-\begin{align}
-\exists x\cdot farmer(x)\cdot(&\exists y\cdot owns(x,y)\cdot donkey(y)\nonumber\\
-                              &\vee \exists z\cdot owns(x,z)\cdot horse(z))
-\end{align}
-And combines that with `He beats the donkey': $beat(x,y)$, we get, in Visser's model, $\{\exists x\cdot farmer(x)\cdot\exists y\cdot owns(x,y)\cdot donkey(y)\cdot beats(x,y), \exists x\cdot farmer(x)\cdot\exists z\cdot owns(x,z)\cdot horse(z)\cdot beats(x,y)\}$.
-The second one is clearly not valid (though we wouldn't get a problem if there is no horse or farmer) which we might catch `at compile time' and replace it with something like `$\bot$' or `error'.
-This sort of error handling is discussed and implemented in the implementation section, but is not terribly interesting from a formal point of view.
-
-Much more interesting are the following two questions: How could we have translated \label{owns_donkey_or_horse} so it would actually work?
-After all it seemed like a reasonable sentence until variable naming got us.
-And does the way variable binding works with our definition actually make sense?
-
-Clearly just renaming $z$ to $y$ in the horse term would have solved the first problem.
-However to work in a compositional way, that would either require great luck, or renaming, and we don't have a renaming operator.
-
-A trick we could consider is to declare an `object' variable outside of the disjunction:
-\begin{align}
-\exists s\cdot farmer(s)\cdot\exists o\cdot(
-      &\exists y\cdot owns(s,y)\cdot donkey(y)\cdot o=y \nonumber\\
-      &\vee \exists z\cdot owns(s,z)\cdot horse(z)\cdot o=z)\cdot beats(s,o)
-\end{align}
-%
-This even inspires on idea that we might be able to build everything up around subjects, objects, indirect objects etc.
-This is explored in the implementation, but it does suffer from many problems.
-E.g.
-we might well have a sentence very similar to the above: `A farmer owns a donkey and owns a horse.
-He beat them'.
-Here we suddenly have two objects in the same sentence.
-
-Sentences like the above are easy to make in English because the absence of noun genders make it easy to use the same pronoun to refer to a lot of things at once.
-By choosing the right actors however, it is always possible to run into problems.
-
-The problem is perhaps the non-linearity in the way variables pass through the disjunction. Somehow we lose track of their history when the `results' are merged together.
-
-This leads on to the other question we mentioned: Is the binding flow through our disjunction operator really any good?
-We are is inspired by the sentence `Either a farmer or his scapegoat must go to court.
-In any case he will be picked up by sunset'.
-In terms of DPL formulas, this would be modelled something like $(\psi_1\cup\psi_2)\cdot\psi_3$.
-Clearly we have a problem here: Variables bound in $\psi_1$ are not accessible in $\psi_2$.
-
-This sentence suggests that we might want to define a disjunction operator with a linear flow of variables: $\psi_1\to\psi_2\to\psi_3$ instead of $\psi_1\to\psi_3\text{ and }\psi_2\to\psi_3$.
-However there is a problem: We currently define `false' as the absence of possible assignments.
-Hence if $\psi_1$ is false, there is no way we can define disjunction to push it's variables into $\psi_2$.
-If any variables were left undefined, $\psi_1$ wouldn't be false.
-
-There is a cure to this problem.
-When we extend our logic with more power in the monoid chapter, we get a second chance.
-However Let's first have a look at an obvious alternative definition for disjunction considered by Groenendijk\cite{groenendijk1991dynamic}.
-We'll call it `static disjunction':
-\begin{align}
-\alpha[\psi_1\vee\psi_2]\beta
- :\equiv&\ \alpha[\neg(\neg\psi_1\wedge\neg\psi_2)]\beta \nonumber\\
- \leftrightarrow&\ \alpha[\neg\psi_1\rightarrow\psi_2]\beta \nonumber\\
- \leftrightarrow&\ \alpha = \beta \wedge \forall\gamma(\alpha[\neg\psi_1]\gamma\rightarrow\gamma\models[\psi_2]) \nonumber\\
- \leftrightarrow&\ \alpha = \beta \wedge \forall\gamma((\alpha=\gamma\wedge\gamma\not\models\psi_1)\rightarrow\gamma\models[\psi_2]) \nonumber\\
- \leftrightarrow&\ \alpha = \beta \wedge \forall\gamma(\alpha=\gamma\rightarrow(\gamma\models\psi_1\vee\gamma\models\psi_2)) \nonumber\\
- \leftrightarrow&\ \alpha = \beta \wedge (\alpha\models\psi_1\vee\alpha\models\psi_2) \label{sem_or}
-\end{align}
-
-The idea to this follows natural from De Morgan's laws of classical logic.
-Clearly the semantics are different from the ones we defined for $\cup$.
-However not in a good way.
-This static disjunction doesn't allow any variables to escape from inside it's reach.
-
-So does this fix our problem of binding between clauses?
-No.
-In fact this definition doesn't even allow binding from a clause and out.
-The villain is $\neg\psi$.
-From the first mention we defined it to be static so it wouldn't allow unsafe, arbitrary assignments as long as they didn't satisfy $\psi$.
-We wanted our logic to be reasonably monotonic, and we got what we deserved.
-$\neg$ not only closed down this version of disjunction, but our implication definition as well.
-
-We can imagine defining $\neg$ dynamically: $\neg\psi_1\cdot\psi_2 = \neg(\psi_1\cdot\psi_2)$.
-However not in our current model, and it won't solve any of our problems with disjunction.
-It does however allow this sentence: `It is not the case that a farmer doesn't own a donkey.
-He beats it'.
-While it made sense not to allow variables to escape from a single $\neg$, we want them to escape from a double.
+ & \leftrightarrow \alpha\models \neg\langle\psi\rangle\top \wedge \phi \tag{$\ref{conv_neg}$}
+%
+\intertext{Because $\psi$ is this way `evaluated on $\top$ only', we know that no variables introduced in $\psi$ can affect $\phi$.
+The derivation of $\vee$ can be done from the previous results alone:}
+\langle\psi_1\vee\psi_2\rangle\phi
+& \equiv \langle\neg(\neg\psi_1\cdot\neg\psi_2)\rangle\phi & \text{by \eqref{sem_or_def}}\nonumber\\
+& \equiv \neg\langle\neg\psi_1\cdot\neg\psi_2\rangle\top \wedge\phi & \text{by \eqref{conv_neg}}\nonumber\\
+& \equiv \neg\langle\neg\psi_1\rangle\langle\neg\psi_2\rangle\top \wedge\phi & \text{by \eqref{conv_and}}\nonumber\\
+& \equiv \neg\langle\neg\psi_1\rangle(\neg\langle\psi_2\rangle\top) \wedge\phi & \text{by \eqref{conv_neg}}\nonumber\\
+& \equiv \neg(\neg\langle\psi_1\rangle\top \wedge \neg\langle\psi_2\rangle\top) \wedge\phi & \text{by \eqref{conv_neg}}\nonumber\\
+& \equiv (\langle\psi_1\rangle\top \vee \langle\psi_2\rangle\top) \wedge\phi &\text{by \eqref{sem_or_def}} \tag{$\ref{conv_vee}$}
+\end{align}
 
@@
 
-\subsection{Source Code}
+\subsection{Source Code}\label{sec:source_code}
 

### 2013-05-20 paper.tex
Addressed main points

@@
 \subsection{Donkey Monoids}
-
-Imagine translating this sentence to DPL:
+With the introduction of DPL we came a long way in terms of a compositional logic. However looking once more at the donkey sentence variation:
 %
 \begin{quotation}
-A farmer owns a donkey, if he beats it
+A farmer owns a donkey, if he beats it.
 \end{quotation}
 %
-We would have to write something along the lines of:
+And its DPL formula:
 %
 \begin{equation}
-\exists x \cdot farmer(x) \cdot \exists y \cdot donkey(y) \cdot beats(x,y) \rightarrow owns(x,y)
+\exists x \cdot farmer(x) \cdot \exists y \cdot donkey(y) \cdot beats(x,y) \rightarrow owns(x,y)\label{eq:dpl_beats_if_owns}
 \end{equation}
 %
-The problem here is that reorderings like this work against our sought property of linearity: of being able to add new information as we go.
-We cannot compose `A farmer owns a donkey' and `if he beats it' in a meaningful way.
-
-To rescue us from this problem Albert Visser suggests constructing a special e-monoid over the dynamic relational algebra.
-The idea is that we are going to simultaneously work on two `streams' of dpl, and switch easily between them with a special element $\Bowtie$.
-
-Our sentence from above will then be expressible as
+We see that we still haven't reached the compositional Nirvana.
+Equation \eqref{eq:dpl_beats_if_owns} is clearly not based on the composition of `A farmer owns a donkey' and `if he beats it' in a direct way.
+
+\subsubsection{Polarity Switcher}
+To help our course, Albert Visser suggests the construction of a special monoid over DPL.
+The idea is that we are going to simultaneously work on two `streams' of DPL, and switch easily between them with a special constant $\Bowtie$.
+
+The inverse donkey sentence can now be expressed as:
 %
@@
 %
-which can be written in `stream form' as
+which `internally' is represented by the following streams of information:
 %
@@
 %
-The meaning of the above is $(-) \rightarrow (+)$
-
-Another example from Visser is:
-\begin{quotation}
-A man keeps his word. He is honest
-\end{quotation}
-which roughly translates to:
+Visser interprets the above as the DPL formula $(-) \rightarrow (+)$.
+
+Another example from Visser is: `A man keeps his word.
+He is honest' which roughly translates into: $\Bowtie\cdot\exists x\cdot man(x)\cdot\Bowtie\cdot keepword(x)\cdot honest(x)$ for the general reading.
+Visser has a whole lot more examples in his paper\cite{visser1999donkey}.
+
+So how do we define this nifty $\Bowtie$? The trick is to move the entire dynamic predicate logic from before into an monoid that will help us keep track of some extra details.
+
+A monoid is an simple algebraic structure with a one associative binary operation and an identity element.
+Our monoid $\langle\mathcal{D},\cdot\rangle$ is going to have the domain $\mathcal{D}=DPL\times DPL\times\{+,-\}$, which is populated by:
 \begin{equation}
-\Bowtie\cdot\exists x\cdot man(x)\cdot\Bowtie\cdot keepword(x)\cdot honest(x)
+\langle\top,\phi,+\rangle \in \mathcal{D} \quad \text{for all}\enspace \phi\in\text{DPL}
 \end{equation}
-If you are interested in a lot more examples, see Visser's paper\cite{visser1999donkey}.
-
-So how do we define this nifty $\Bowtie$? The trick is to move the entire dynamic relation algebra from before into an e-monoid that will help us keep track of some details.
-
-If $\phi$ from before is the basic algebra, we specify the construction of our monoid $\Phi$ as follows.
-%
-\begin{flalign}
-&\top := \langle \top, \top, + \rangle & \\
-&\bot := \langle \top, \bot, + \rangle & \\
-&\Bowtie := \langle \top, \top, - \rangle & \\
-&\langle q_-, q_+, \alpha \rangle \cdot \langle r_-, r_+, \beta \rangle := \langle q_- \cdot r_{-\alpha}, q_+ \cdot r_{+\alpha}, \alpha\beta \rangle&
-\end{flalign}
+With no chance of confusion, we will simply denote these elements by their DPL representation.
+For example $\mathcal{D}$ contains the constants $\top$ and $\bot$ and the $\Bowtie$ constant here defined:
+\begin{align}
+\top &:= \langle \top, \top, + \rangle \\
+\bot &:= \langle \top, \bot, + \rangle \\
+\Bowtie &:= \langle \top, \top, - \rangle
+\end{align}
+The associative operation $\cdot$ is as below. To make the definition slicker, we give the set $\{+,-\}$ a multiplication operation $++\mapsto+$, $+-\mapsto-$, $-+\mapsto-$, $--\mapsto+$.
+\begin{align}
+&\langle q_-, q_+, \alpha \rangle \cdot \langle r_-, r_+, \beta \rangle := \langle q_- \cdot r_{-\alpha},\enspace q_+ \cdot r_{+\alpha},\enspace  \alpha\beta \rangle
+\end{align}
 %
@@
 
-An even more interesting element than $\Bowtie$ is $\triangle$. $\triangle$ is a scope modifier, and it allows us to write:
+\subsubsection{Scope Modifier}
+We won't spend much time with the above monoid though, since we are immediately going to introduce a generalization! We are going to introduce the scope modifier $\triangle$, which allows us to write:
 %
@@
 \begin{equation}
-\triangle \cdot \exists x \cdot farmer(x) \cdot \triangle \cdot owns(x,y) \cdot \triangle \cdot \exists y \cdot donkey(y) \cdot \triangle \cdot beats(x,y)
+\exists x \cdot farmer(x) \cdot owns(x,y) \cdot \triangle \cdot \exists y \cdot donkey(y) \cdot \triangle \cdot beats(x,y)
 \end{equation}
@@
 
-And combining the two monoids we can do our original sentence as
-%
-\begin{quote}
-If a farmer owns a donkey, he beats it
-\end{quote}
-%
-as
+And combining the scope modifier with the polarity switched we can finally, as promised, write our original sentence `If a farmer owns a donkey, he beats it' as:
 \begin{equation}
-\Bowtie \cdot \triangle \cdot \exists x \cdot farmer(x) \cdot \triangle \cdot owns(x,y) \cdot \triangle \cdot \exists y \cdot donkey(y) \cdot \triangle \cdot \Bowtie \cdot beats(x,y)
+\Bowtie \cdot \exists x \cdot farmer(x) \cdot owns(x,y) \cdot \triangle \cdot \exists y \cdot donkey(y) \cdot \triangle \cdot \Bowtie \cdot beats(x,y)
 \end{equation}
@@
     \hline
-    $(-,1)$ & $\exists x \cdot farmer(x)$ & ~ & ~ $ \exists y \cdot donkey(y)$ & ~ \\\hline
-    $(-,0)$ & ~ & $owns(x,y)$ & ~ & ~ \\\hline
+    $(-,1)$ & ~ & ~ & ~ $ \exists y \cdot donkey(y)$ & ~ \\\hline
+    $(-,0)$ & $\exists x \cdot farmer(x)$ & $owns(x,y)$ & ~ & ~ \\\hline
     $(+,1)$ & ~ & ~ & ~ & ~ \\\hline
@@
 %
-With the meaning being $(-,1) \cdot (-,0) \rightarrow (+,1) \cdot (+,0)$.
+With the DPL interpretation being $(-,1) \cdot (-,0) \rightarrow (+,1) \cdot (+,0)$.
 Notice that the negative relations are again put on the left side of the implication, and the relations with larger scope value are put before those of lower value.
-
-Notice $\Bowtie$ and $\triangle$ are commutative, since they work entirely on their own indicator.
+We see $\Bowtie$ and $\triangle$ are commutative, since they work entirely on their own indicator.
+
+As expected from the above diagram, we are going to define our new monoid $\langle\mathcal{D},\cdot\rangle$ as having the domain\footnote{We have chosen to model our scopes in $\mathbb{Z}_2$.
 It's an interesting question to consider what sentences might require three or more levels of scoping.
-We could define $\triangle$ to go up in scope and $\bigtriangledown$ to go down.
-Instead of ${0,1}$ we would work on $\mathbb{Z}$.
-
-For now we define the monoid as follows:
-%
-\begin{flalign}
-&\top := \langle \top, \top, \top, \top, +, 0 \rangle & \\
-&\bot := \langle \top, \top, \top, \bot, +, 0 \rangle & \\
-&\Bowtie := \langle \top, \top, \top, \top -, 0 \rangle & \\
-&\triangle := \langle \top, \top, \top, \top +, 1 \rangle & \\
+If we chose to work on the entire $\mathbb{Z}$, we could define $\triangle$ to go up in scope and $\triangledown$ to go down.} $DPL^4\times\{+,-\}\times\mathbb{Z}_2$ with the elements:
+\begin{equation}
+\langle \top, \top, \top, \phi, +, 0 \rangle \in \mathcal{D} \quad \text{for all}\enspace \phi\in\text{DPL}
+\end{equation}
+In addition to $\Bowtie$ and $\triangle$ defined as:
+\begin{align}
+\top &:= \langle \top, \top, \top, \top, +, 0 \rangle \\
+\bot &:= \langle \top, \top, \top, \bot, +, 0 \rangle \\
+\Bowtie &:= \langle \top, \top, \top, \top -, 0 \rangle \\
+\triangle &:= \langle \top, \top, \top, \top +, 1 \rangle
+\end{align}
+For $\cdot$ the definition works just like in the previous case, though slightly more complicated:
+\begin{align}
 &\langle q_{-,1}, q_{-,0}, q_{+,1}, q_{+,0}, \alpha, i \rangle \cdot \langle r_{-,1}, r_{-,0}, r_{+,1}, r_{+,0}, \beta, j \rangle := &\\
-&\hspace{1cm}\langle q_{-,1} \cdot r_{-\alpha,1+i}, q_{-,0} \cdot r_{-\alpha,i}, q_{+,1} \cdot r_{\alpha,1+i}, q_{+,0} \cdot r_{\alpha,i}, \alpha\beta, i+j \rangle& \nonumber
-\end{flalign}
-
-In his paper Visser now goes on to define $\triangleleft$ and $\triangleright$ which work similarly to $\Bowtie$ except they work in different directions. 
-%
+&\hspace{1cm}\langle q_{-,1} \cdot r_{-\alpha,1+i},\enspace q_{-,0} \cdot r_{-\alpha,i},\enspace q_{+,1} \cdot r_{\alpha,1+i},\enspace q_{+,0} \cdot r_{\alpha,i},\enspace \alpha\beta,\enspace i+j \rangle& \nonumber
+\end{align}
+
+In his paper Visser now goes on to define $\triangleleft$ and $\triangleright$ which work similarly to $\Bowtie$ except they work `backwards and forward in time'.
+This allows us to model `if he beats it' entirely on its own as $\triangleleft\cdot beats(x,y)$ in
 \begin{equation}
-\Bowtie \cdot \triangle \cdot \exists x \cdot farmer(x) \cdot \triangle \cdot owns(x,y) \cdot \triangle \cdot \exists y \cdot donkey(y) \cdot \triangle \cdot \triangleleft \cdot beats(x,y)
+  \exists x \cdot farmer(x) \cdot owns(x,y) \cdot \triangle \cdot \exists y \cdot donkey(y) \cdot \triangle \cdot \triangleleft \cdot beats(x,y)
 \end{equation}
 %
-This is quite interesting for certain sentences containing `surprises', like
-%
+If we immagined a system that tried to build meaning from its input as soon as it came in, $\triangleleft$ would allow us to account for `surprises' like in the sentence:
 \begin{quotation}
@@
 %
-In this case we would compositionally assume to start out with
-$sees(x,y)$ and finish with $\cdot \exists y \cdot donkey(y)\triangle$ hence we want to write something like:
+In this case we would compositionally like to start out with $sees(x,y)$ expecting a $\triangle\cdot\exists y\cdot\triangle$ somehow, but when we instead see the negation, we end up with the formula:
 %
@@
 %
-Notice that that means the word `no'/`not' must be $\triangleleft\cdot\bot\cdot\triangleright$.
-Unfortunately these retrospective triangles tend to mess sentence enough up, that they require lots of `stream unification brackets', and don't generally give very nice formulas.
-In the semantic parsing in the next section, they wont be used at all.
+Visser suggests $\triangleleft\cdot\bot\cdot\triangleright$ often is a good candidate for words like `no' and `nobody'.
+Unfortunately these retrospective triangles tend to mess sentences enough up, that I had to include brackets everywhere to keep things under control.
+More brackets are seldom the way to beautiful formulas, so in the automatic framework in the next section, we will settle with the $\Bowtie$ and $\triangle$ monoid.
 
@@
 Makoto Kanazawa has written an entire paper on issues with weak and strong readings of donkey sentences\cite{kanazawa1994weak}.
-For the donkey sentence, we may have at least the following readings of the second `a':
+For our main donkey sentence, we may have at least the following readings of the second `a':
 \begin{enumerate}
@@
 It would however require some well defined rules for what we can expect to find under each type of arrow.
-This was a blind road.
+This was a blind alley.
 
@@
 Because these rules are tested in practice on a large set of English sentences, they will hopefully avoid some simple mistakes and oversights that are easy to make in purely theoretical settings.
-It seams like a good approach to start from a small working fragment, and slowly add more phrase rules and sentences.
+It seems like a good approach to start from a small working fragment, and slowly add more phrase rules and sentences.
 The opposite of this approach is what is often observed when logicians and philosophers go astray in very abstract corners of English, without having covered the basic ground.

### 2013-05-20 paper.tex
Final version? 9781 words in main text.

@@
 Notice from this definition that $\langle\psi\rangle\top$ is the conversion function from DPL to FOL.
-Using definition \eqref{foldpl} and formulas \eqref{sem_and} to \eqref{sem_exists} we can derive the following equivalences. See appendix~\ref{sec:derivation_dpl} for my full derivations.
+Using definition \eqref{foldpl} and formulas \eqref{sem_and} to \eqref{sem_exists} we can derive the following equivalences.
+See appendix~\ref{sec:derivation_dpl} for my full proofs of meaning preservation.
 %
@@
 
-In the test of this program, we used a branch of 100 sentences.
+In the test of this program, we used a branch of 53 sentences.
 Some copied from the papers referenced, others invented based on features considered or the daily life of farmers and donkeys as imagined by the author.
 The sentences were converted to Visser, DPL and FOL, and the meanings were checked manually to make sense.
+See appendix~\ref{sec:sample_output} for the raw program output.
 
@@
 The possibility for pushing control elements into the semantic of the logic is really the beauty of Visser's approach and unfortunately a road I haven't studied much in this paper.
-It is not hard to imagine $(Dpl,Dpl,Dpl,Dpl,Integer,Integer)$ exchanged for something like $(Dpl,...,PosTag)$ and corresponding composition rules.
+It is not hard to imagine $(Dpl,Dpl,Dpl,Dpl,Integer,Integer)$ exchanged for something like $(Dpl,\dots,PosTag)$ and corresponding composition rules.
 
@@
 
-\subsection{Sample Output}
+\subsection{Sample Output}\label{sec:sample_output}
 
@@
 \subsection{Source Code}\label{sec:source_code}
+
+An online repository of the source code can be found at\cite{ahle2013donkey}.
 

### 2013-05-20 src/Dpl.hs
More comments

@@ -19,8 +19,7 @@ data Dpl =
   | DP Ref [Ref]
   deriving (Eq, Show)
 
-
-
+-- Printing functions
 
 tex :: Dpl -> String
 tex = fst . tex' where
@@ -45,16 +44,14 @@ pretty = fst . pretty' where
   pretty' (DP f ks) = (f ++ "(" ++ concat (intersperse "," ks) ++ ")", 0)
 
 
+simplify :: Dpl -> Dpl
+simplify p = iterate simp p !! 10
+
 simp :: Dpl -> Dpl
 simp (DC p DT) = simp p
 simp (DC DT p) = simp p
 simp (DC p DF) = DF
 simp (DC DF p) = DF
--- These disjunction rules are a bit doubtful
--- simp (DD p DT) = DT
--- simp (DD DT p) = DT
--- simp (DD p DF) = simp p
--- simp (DD DF p) = simp p
 simp (DI p DF) = DN p
 -- scope related
 simp (DI DT (DI DT p)) = DT `DI` simp p
@@ -65,8 +62,11 @@ simp (DD p q) = simp p `DD` simp q
 simp (DI p q) = simp p `DI` simp q
 simp (DN p) = DN (simp p)
 simp p = p
-simplify :: Dpl -> Dpl
-simplify p = iterate simp p !! 10
+-- See the paper on why the following disjunction rules are invalid
+-- simp (DD p DT) = DT
+-- simp (DD DT p) = DT
+-- simp (DD p DF) = simp p
+-- simp (DD DF p) = simp p
 
 -----------------------------------------------------------
 -- Interpretation
@@ -86,6 +86,8 @@ intDpl (DI p q) i x y                = intDpl (DN (p `DC` (DN q))) i x y
 intDpl (DE var) i x y                = hide var x == hide var y where hide key = filter (\(x,y) -> x /= key)
 intDpl (DP name args) (_,_,intp) x y = x == y && intp name (map (lookup' x) args)
 
+-- Like `lookup', but throws an error in case the entity was not available.
+-- We can use this version since our formulas are guarenteed to be valid.
 lookup' :: Assignment -> Ref -> Entity
 lookup' xys key | Just e <- lookup key xys = e
                 | otherwise                = error ("Variable not in scope: " ++ key)

### 2013-05-20 src/Main.hs

@@ -12,6 +12,7 @@ import Data.Maybe
 import System.Environment
 import System.IO
 
+-- Reads and returns every line from a file, modulo comments and blanks
 readSentences :: FilePath -> IO [String]
 readSentences path = do
 		handle <- openFile path ReadMode

### 2013-05-20 src/Stanford.hs

@@ -34,6 +34,7 @@ lemmas doc = map lemma (findElements (unqual "sentence") sentences)
 
 ------------------
 
+-- Extracts the constituent tree from a document.
 postrees :: Element -> [PosTree Word]
 postrees doc = map rootfilter roots
 	where
@@ -66,10 +67,12 @@ posLeaf =
 		word <- many (noneOf " ()")
 		return (L tag word)
 
-
+-- Takes a postree and a lemmaized list of words for each sentence
+-- and produces a cleaned up and index tagged tree
 cleanTrees :: [PosTree Word] -> [Sentence] -> [PosTree IWord]
 cleanTrees trees ls = map cleanElements $ map cleanTags $ improveTrees trees ls
 
+-- Tags leafs with indicies
 improveTrees :: [PosTree Word] -> [Sentence] -> [PosTree IWord]
 improveTrees trees ls = [improve t l s | (t, l, s) <- zip3 trees ls [0..]]
 	where improve t l s = substitute t (zip l (zip (repeat s) [0..]))
@@ -127,6 +130,7 @@ substitute tree list = snd (substitute_ tree list)
 
 ------------------
 
+-- Extract the deptree for each sentence
 deps :: Element -> String -> [DepTree]
 deps doc name = map (buildTree deps) roots
 	where
@@ -135,15 +139,19 @@ deps doc name = map (buildTree deps) roots
 		rights = nub [dep | (typ, gov, dep) <- deps]
 		roots = lefts \\ rights
 
+-- The dependency information comes in tuples of (label, arrow_from, arrow_to)
+-- This function transforms a set of tuples into an algebraic tree structure
 buildTree :: [(String, Int, Int)] -> Int -> DepTree
 buildTree deps root = Dep root [(name, buildTree deps dep) | (name, gov, dep) <- deps, gov == root]
 
+-- Extract dependecy tuples form xml
 listDeps :: Element -> String -> [(String, Int, Int)]
 listDeps doc name = map parseDep deps
 	where
 		Just dep = findElement (unqual name) doc
 		deps = findElements (unqual "dep") dep
 
+-- Extract single tuple from `dep' element
 parseDep :: Element -> (String, Int, Int)
 parseDep e = (typ, read gov - 1, read dep - 1)
 	where
@@ -153,6 +161,16 @@ parseDep e = (typ, read gov - 1, read dep - 1)
 
 ------------------
 
+-- Calls `corefs' to extract a map from word-index to variable name.
[…]

### 2013-05-20 src/ToLogic.hs

@@ -11,8 +11,12 @@ import Data.Maybe
 type Predicate = Ref -> Visser
 type Predicate2 = Ref -> Ref -> Visser
 
+-- Runs a montague grammar transformation on a constituent tree and a
+-- word-index to variable name map. The produced logic is a
+-- polarity-switch + scope-modifier donkey monoid
 montague :: (Store, [PosTree IWord]) -> Visser
 montague (refs, tree) = cleanVariables $ intRoot tree refs
+
 
 intRoot :: [PosTree IWord] -> Store -> Visser
 intRoot ((P "S" s):ss) r = intS s r `VC` intRoot ss r

### 2013-05-20 src/Visser.hs

@@ -61,6 +61,7 @@ simp p = p
 simplify :: Visser -> Visser
 simplify p = iterate simp p !! 2
 
+-- Renames variables to [a..], removing gaps.
 cleanVariables :: Visser -> Visser
 cleanVariables p = foldl rename p (zip used im ++ zip im variables)
   where

### 2013-05-20 paper.tex
A slightly longer conclusions section wc 9842

@@
 My program can correctly logify entire paragraphs of text containing advanced anaphora, conjunctive nouns, verbs and adjectives, implications and scoping.
-It handles sentences with free variables as well as entirely factual sentences.
-
-A few examples of sentences that my framework already supports, but which fail due to shortcomings in the Stanford toolkit are `Every dog sees a cat.
-It chases it.' where the anaphora doesn't get resolved, and `Buffalo buffalo Buffalo buffalo buffalo buffalo Buffalo buffalo' that manages to slightly confuse the parser.
+In some cases the my defined rules create quite creative FOL formulas that would likely have taken a person a few tries to get right.
+In addition to factual sentences my program also handles query type sentences with free variables.
+
+My fragment of English is however quite limited to the extent that most sentences require some adaption to pass through.
+Word classes such as adverbs and plurals are not supported, and in some cases my framework supports a sentence, but logification still fails due to shortcomings in the Stanford toolkit.
+For example in `Every dog sees a cat.
+It chases it.' where the anaphora doesn't get resolved, and `Buffalo buffalo Buffalo buffalo buffalo buffalo Buffalo buffalo'\cite{rapaport2006september} that also manages to slightly confuse the parser.
 
